### 2014-01-22 brains-conclusion.tex
Add all files, initial import

% !TEX root = brainscopypaste.tex
\section{Conclusion}\label{sec:conclusion}

Main contributions:
\begin{itemize}
\item large \emph{in vivo} psycholinguistics experiment, emphasizing the importance of the semantic network structure (Wordnet: Pagerank, degree, clustering[, distance?]), finely describing the impact of classically-influent psycholinguistic variables (\emph{aoa}, number of phonemes, etc.); 
\item new in the sense that it does not focus on ease of recall but rather bias of substitution, and that in this respect it not only provides a finer description of the bias but also corresponds to an  ``input-output'' reformulation couple describing the joint properties of (substituted$\rightarrow$substituting) terms.
\item beyond that, provide the first bricks of an empirical \emph{fitness landscape} for the epidemiology of representations
\end{itemize}

### 2014-01-22 brains-introduction.tex
Add all files, initial import

% !TEX root = /Users/camille/Shared/Dropbox/BrainCopyPaste/brainscopypaste.tex
% ============================
\section{Introduction} % =====
% ============================

%On assiste aujourd'hui à une convergence progressive entre des disciplines étudiant la cognition humaine, d'un côté, et des disciplines étudiant la culture et les \emph{choses sociales}, de l'autre. Ces deux familles de disciplines ont en effet bien des frontières en commun ; celles-ci se rendent visibles par l'intermédiaire des thèmes pour lesquels il devient incontournable de s'intéresser à l'humain à la fois dans sa globalité et dans ses fonctionnements particuliers.

The understanding of the origin of cultural similarity has triggered a sizable literature in the recent past, spanning over a vast area of research fields ranging from cultural anthropology to social network analysis and complex system modeling --- and diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'', for the most part.  Broadly, this type of research program relies on the understanding of phenomena pertaining to both cognitive science and social science, in order to jointly appraise the processing and transmission of information at the individual and social levels. % information propagation phenomena -- and where we witness the development of studies relying on both individual and social cognition.

Several theories have been proposed and debated within these research fields, mixing social and individual cognition --- notably from the side of cultural anthropology.  There is, first, the debate around the ``memetic'' program initiated by \citet{Dawkins76}, for which the collection of works by \citet{Aunger00} provides a solid overview. A second field relates to the development of evolutionary models of norms (see for instance \citealp{Ehrlich05}), following the seminal work of  \citet{Boyd85}.

A third approach, ``cultural epidemiology'' initiated by \citet{sper:expl} recently attracted significant attention. Works such as \citet{Atran03} suggest that this stance could be anthropologically more satisfying than memetics; see \citet{Kuper00}, and \citet{Bloch00}, for further details on the main issues in this debate.
It focuses on the notion of \emph{representation}, linking the cognitive science concept of \emph{mental representation} to the concept of \emph{public representation}, the latter being the equivalent of the former outside of the brain (\hbox{i.e.} in cultural artefacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
%\begin{quotation}
{%Une représentation peut exister à l'intérieur même de l'utilisateur : il s'agit alors d'une \emph{représentation mentale}. Un souvenir, une hypothèse, une intention sont des exemples de représentations mentales. L'utilisateur et le producteur d'une représentation mentale ne font qu'un. Une représentation peut aussi exister dans l'environnement de l'utilisateur comme par exemple le texte qui est sous vos yeux. Il s'agit alors d'une \emph{représentation publique}. Une représentation publique est généralement un moyen de communication entre un producteur et un utilisateur distincts l'un de l'autre.
``%
A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention. The producer and the user of a mental representation are one and the same person.  A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.}%~(\citealp[p.~49]{sper:expl}.)
''
%\end{quotation}
}
The key point here is that representations are not being replicated through a hi-fidelity copying process, but are being reformulated through an interpretation process, and thus evolve. Cultural epidemiology postulates this type of conceptual mutation and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domains of an underlying semantic dynamical system.
Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not been empirically validated.

%Bien qu'ayant connu des développements récents en modélisation (par exemple \citealp{Claidiere07}, pour la notion d'attracteur culturel), la difficulté à effectuer des mesures quantitatives sur les notions clés du domaine a fait que les hypothèses à la base de l'édifice théorique développé n'ont pas encore pu être testées empiriquement.

The last decade has however witnessed an avalanche of observable \emph{in vivo} data stemming from online interactions. These large amounts of textual traces of information production by humans are plausibly not records of strictly ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), yet they actually constitute a wealth of observations on the dynamics of public representations. As such, they are prone to radically improve the prospects of empirical validation of the individual-level processes of cultural evolution.

We aim to empirically describe the transformation of a specific type of public representations by focusing on the possible cognitive alterations performed by individuals during reformulation.  % On s'est proposé d'étudier de façon \emph{empirique} la transformation des représentations publiques, en se focalisant sur la mise en évidence d'un biais cognitif dans le traitement  par le cerveau de certaines représentations culturelles. 
To deal with robust and simple cultural representations, we paid attention to the evolution of quotations. Admittedly, these verbatim public representations should not suffer any reformulation alteration (as opposed to more sophisticated expressions and opinions), but still do.  We will in particular exhibit a non-trivial cognitive alteration by which some portions of the quotation, namely individual words, may be modified; and will wonder which are generally the semantic and structural characteristics of these words and their substitutions. 
%On verra que par cette analyse de la reproduction des citations par les auteurs on observe un biais cognitif non trivial ; on observera ainsi un lien entre l'attractivité de certains mots (en termes d'épidémiologie culturelle) et les propriétés sémantiques de ces mêmes mots, calculées sur la base du réseau formé par ces mots.
%On s'est proposé d'étudier de façon \emph{empirique} la transformation des représentations publiques, c'est-à-dire d'explorer à l'aide de données obtenues in vivo l'effet du cerveau sur l'évolution des représentations ; une telle étude permettrait d'observer empiriquement l'effet d'un attracteur culturel tel que défini par Sperber. La question est donc la suivante : peut-on observer un biais cognitif dans le traitement de certaines représentations culturelles par le cerveau, et si oui quel est-il ?On verra que par cette analyse de la reproduction des citations par les auteurs on observe un biais cognitif non trivial ; on observera ainsi un lien entre l'attractivité de certains mots (en termes d'épidémiologie culturelle) et les propriétés sémantiques de ces mêmes mots, calculées sur la base du réseau formé par ces mots.
Broadly, we contend that using this type of data is both equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards the design of empirically realistic models of cultural evolution.

% --------------------------------
%\subsection*{Problématique} % -----
% --------------------------------
\label{subsec:problematique}

%\paragraph*{Plan}

The next section (Sec~\ref{sec:related}) describes the relevant recent state-of-the-art on this matter. In Sec.~\ref{sec:protocol}, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material. Section~\ref{sec:results} describes the significant psycholinguistic cues and biases observed during \emph{in vivo} quotation reformulation, followed by a discussion and general guidelines for further work in Sec.~\ref{sec:conclusion}.
%Le travail de recherche effectué étant essentiellement empirique, on le présente en deux grandes étapes : en premier lieu, on détaille le protocole d'expérimentation utilisé ainsi que les différentes tentatives et pistes explorées pendant son élaboration. Pour cela, on commencera par analyser la question ci-dessus plus à fond afin d'en distiller une question opérationnelle, puis on expliquera en détail la façon dont on a construit le protocole expérimental permettant de faire les mesures pour répondre à cette question%\footnote{Le terme "protocole expérimental", utilisé ici, est bien le terme approprié pour ce dont il s'agit : on verra qu'on part d'un corpus de données pour lequel on construit une façon de faire les mesures qu'on considère pertinentes, et cette démarche correspond à la construction d'un protocole expérimental comme on peut le faire pour une expérience de psychologie cognitive ; c'est à ce moment qu'on définit ce qu'on va mesurer et comment cette mesure est faite.}
%. La deuxième grande partie concerne les résultats : on montre tout d'abord la façon dont les mesures sont utilisées pour construire les observables qu'on cherche, ce qui permettra ensuite d'interpréter les résultats obtenus au regard de la question posée.
%Une dernière partie concernera les perspectives que ce travail ouvre, et les façons dont il pourrait être poursuivi et étendu.

\section{Related work}\label{sec:related}

The relevant literature on \emph{public representation dynamics} features two main streams. On one hand, we find studies aiming at understanding the macroscopic phenomena of \emph{social diffusion} of public representations, describing for instance the propagation of cultural items across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.     
%L'état de l'art en ce qui concerne l'épidémiologie culturelle peut être résumé ainsi : d'un côté, des études %s'intégrant au deuxième groupe de disciplines
% permettent une bonne compréhension des phénomènes macroscopiques de \emph{diffusion sociale} des représentations publiques, tels que le chemin emprunté par une information atomique donnée, les temps caractéristiques et les cycles de diffusion dans ces réseaux, la relation entre blogs et sites de médias, ou encore l'autorité et l'influence exercée par certains n\oe{}uds du réseau. 
These studies are relatively independent from anthropology and cognition and are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis). Without necessarily relying on specific social science theories, %these studies are interesting in that they aim at describing cultural dynamics by using large social media corpuses. 
this research stream is of interest for its aim of cultural dynamics description by using large social media corpora.
% ,    Ces études sont relativement indépendantes de l'anthropologie et de la cognition et  se situent à l'interface entre le \emph{data mining}, l'étude des systèmes complexes, et la sociologie quantitative (e.g. sociologie des réseaux), en exploitant de grands corpus de données produits par des collectes sur Internet. Sans nécessairement s'appuyer sur des cadres théoriques précis ou intégrés, ces études présentent néanmoins la particularité de vouloir décrire les dynamiques culturelles en exploitant les possibilités ouvertes par les larges corpus de données issus d'Internet. %, allant de l'étude de l'impact des réseaux sociaux sur la diffusion des informations sur ces mêmes réseaux \citep{Bakshy12} à l'analyse du rôle de la presse en ligne dans la vie politique locale \citep{Parasie12}, en passant par le détail des interactions entre éditeurs sur Wikipedia \citep{Jurgens12} ou l'évolution du vocabulaire des langues au cours du temps \citep{Lieberman07}. 
%On peut citer ici l'étude de la propagation et de la diffusion de l'information dans la blogosphère faite par \citet{Gruhl04}, l'étude du \emph{news cycle} faite par \citet{Leskovec09} qui fait intervenir la plupart de ces aspects, ou encore l'analyse des influences réciproques entre un réseau social et le réseau sémantique formé par les contenus diffusés sur ce premier réseau, faite par \citet{Cointet09} ; %on l'a déjà mentionné, 
 %, ces thèmes étant explorés principalement par le premier groupe de disciplines.

On the other hand, the study of the \emph{transformation} of public representations has emerged only recently. For one, the notion of ``cultural attractor'' has been the focus of models rendering the evolution of some representations and behaviors not until a few years ago \citep{Claidiere07}. Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist of a series of papers investigation \emph{quotation} transformations from a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}. They generally show several types of regularities and propose models of diffusion-transformation of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.  One of the main conclusions of these works lies in the fact that even for the supposedly most stable type of public representation, namely quotations, it is indeed still possible to observe and measure significant transformations.  However, these studies address transformations by focusing on the properties of the author (e.g. news vs. blogs) of the quotation, or the surrounding public space (e.g. quote frequency in the corpus), rather than the very cognitive features which may trigger or, at least, bias these transformations. % Cependant ces auteurs n'étudient les transformations observées qu'en fonction du type d'auteur qui reproduit les citations, et pas en fonction des citations elles-mêmes (en particulier les questions d'ordre cognitif concernant la raison pour laquelle ces transformations ont lieu, ou la raison pour laquelle elles ont lieu dans tel sens et pas dans tel autre, ne sont pas abordées). %Les possibilités offertes par cette étude étant vastes, on a décidé de travailler sur les mêmes données que \citeauthor{Simmons11} pour traiter notre question.

At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations of public representation evolution by studying the influence of word features on the ease of recall. This state-of-the-art is relatively developed and shows the positive impact of classical psycholinguistic variables, such as word frequency (see \citet{Yonelinas02} for a review), age-of-acquisition \citep{Zevin02}, number of phonemes and number of syllables (see for instance \citet{Rey98} and \citet{nick-diss}).

\TB{The study of network-based properties, while heavily discussed around the notion of connectionism and its normative processual models, appeared much more recently as an empirical investigation field \CN} \rk{C'est sûr ? e.g. http://tiny.cc/bl7usw}. Let us mention two interesting and recent studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors in the cognitive processes and reproduction of those words. %Les deux autres travaux viennent de la psycholinguistique, et posent les fondations pour l'étude des propriétés des représentations (qui seront ici des citations). 
First, \citet{Griffiths07} analyze a task where patients are asked to speak out the first word which comes to their mind when they are presented with a random alphabet letter. The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (PageRank) in a language-wide semantic network built from external word association data. % property concernant une tâche où l'on présente aux sujets une lettre de l'alphabet en leur demandant de dire le premier mot commençant par cette lettre qui leur vient à l'esprit. Les auteurs montrent qu'une certaine propriété des mots (le PageRank calculé sur un réseau sémantique construit à partir de données d'associations de mots) est un bon facteur prédictif de quel mot sera rappelé parmi les mots commençant par la lettre demandée ; ils montrent ainsi un lien entre une caractéristique sémantique quantitative des mots et la facilité de rappel de ces mots dans cette tâche. 
A second psycholinguistic study by \citet{Chan10} shows, in a task of picture naming, that words come quicker when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data). %, étudie une tâche où l'on demande aux sujets de nommer oralement une image. Les auteurs montrent qu'une autre propriété des mots, phonologique cette fois (le coefficient de regroupement des mots calculé sur le réseau phonologique formé par ces mots), a un effet sur la rapidité à nommer l'image présentée : c'est un autre cas de lien entre une caractéristique quantitative (phonologique) d'un mot et la facilité à produire ce mot.

On the whole, the current psycholinguistic state-of-the-art seems to hint at two antagonistic types of results.  On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal \CN. On the other hand, when the task consists in remembering a specific list of items, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernable \CN. The jury is still out as to whether reformulation alteration, i.e. spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.  Our paper additionally sheds light on this debate.

\smallskip
More broadly, we thus observe a gap between, on one side, macro-level empirical studies on diffusion dynamics in a social system and, on the other side, studies essentially focused on micro-level transformations of representations --- these latter studies being either strongly normative, or yielding information whose integration into realistic cultural epidemiology models remains difficult. %On remarque de plus que le volet cognitif, central pour les phénomènes étudiés, est relativement peu développé dans les modèles et peu exploré de façon empirique.
%Les deux dernières études, issues de la psycholinguistique et dans un cadre strictement \emph{in vitro} et strictly at the vocabulary level, montrent donc que des propriétés des mots du vocabulaire, calculées sur un réseau formé par ces mêmes mots, sont des facteurs importants dans la façon dont on les traite cognitivement et dans la façon dont on les produit. C'est justement l'élément qui nous manquait : en se basant sur ces trois travaux, on va pouvoir utiliser des propriétés ainsi calculées sur les mots des citations pour mesurer quantitativement les transformations subies par ces citations lorsqu'elles sont reproduites. Pour rendre l'analyse possible, on se concentrera sur des cas de \emph{substitution} d'un mot par un autre lorsqu'un auteur reproduit une citation : à chaque substitution observée, on assiste à un échange entre deux mots pour lesquels on va pouvoir comparer les propriétés sémantiques.

%%% GARBAGE

%Cependant, depuis plus d'une décennie le paysage des choses observables dans le domaine des représentations publiques est en train de changer radicalement, du fait du déluge de données provenant des interactions en lignes, permettant l'étude empirique des représentations dans de cadre de l'épidémiologie culturelle.

%Nous vivons aujourd'hui non pas avec un manque de données concernant les représentations publiques, mais au contraire sous un véritable déluge de telles données. Quel est ce déluge ? Il est le produit d'une collecte qui est née avec l'utilisation commerciale et massive d'Internet. Cette collecte n'est pas un phénomène mineur, ni même un phénomène d'ampleur moyenne : les services sur Internet enregistrent aujourd'hui la quasi-totalité de ce qui est mesurable des interactions entre les personnes connectées ainsi que des interactions des personnes avec les machines. Vu l'importance que ce réseau a pris dans nos vies de tous les jours, on assiste bien à une explosion de ces données ; à tel point que certains observateurs d'Internet estiment que 90\% des données stockées aujourd'hui par les services sur Internet ont étés générées il y a moins de deux ans.
%Sans être des enregistrements des interactions \emph{physiques} entre les personnes (``real life''), il s'agit bien d'observations sur le terrain nouveau que constituent l'action, la perception, et l'interaction connectées, au travers des machines et d'Internet. En un mot, elles sont en fait des observations en grandes quantités de \emph{représentations publiques}. %\footnote{Bien sûr, une bonne partie des données collectées sur Internet appartient à des entreprises qui n'ont aucun intérêt à les partager ni à les exploiter autrement que financièrement. Mais les données publiquement accessibles forment tout de même une base considérable, qu'il est tout à fait possible d'exploiter comme telle.}.
%L'approche qu'on a esquissée ici n'est qu'une des façons d'aborder l'étude de la culture et de la cognition de façon empirique. En effet, les courants de recherche qui gravitent autour de ce thème sont nombreux et variés ; voici une rapide présentation des champs concernés et de la façon dont ils interagissent.

%Un groupe de disciplines est particulièrement pertinent ici, à l'interface entre le \emph{data mining}, l'étude des systèmes complexes, et la sociologie quantitative (e.g. sociologie des réseaux). Un grand nombre de travaux apparaissent exploitant de grands corpus de données produits par des collectes sur Internet, allant de l'étude de l'impact des réseaux sociaux sur la diffusion des informations sur ces mêmes réseaux \citep{Bakshy12} à l'analyse du rôle de la presse en ligne dans la vie politique locale \citep{Parasie12}, en passant par le détail des interactions entre éditeurs sur Wikipedia \citep{Jurgens12} ou l'évolution du vocabulaire des langues au cours du temps \citep{Lieberman07}. Une partie de ces travaux n'inscrivent leurs résultats que dans des cadres théoriques naïfs (autour de thèmes pour lesquels il existe pourtant des théories réfléchies développées par des champs des sciences sociales), et relèvent parfois plus des sciences de l'informatique que de l'étude de la culture (d'aucuns parlent d'ailleurs de \emph{social computing}). Ce groupe de disciplines présente néanmoins la particularité de vouloir étudier la culture en exploitant les possibilités ouvertes par les larges corpus de données issus d'Internet.

%% ----------------------------------------
%\subsection{Relevant fields} % -----
%% ----------------------------------------
%
%L'idée d'étudier la culture en lien avec la cognition apparaît comme une extension naturelle pour plusieurs disciplines historiquement séparées qui, chacune de leur côté, s'en sont approchées en partant de leur propre espace de concepts et en étendant leurs outils. Les dialogues entre ces disciplines ne sont pas rares, mais les traditions et les épistémologies différentes ont rendu les unifications plus ardues. On peut distinguer au moins deux grands groupes de disciplines débattant autour de thématiques reliées à ce sujet.%\footnote{On regroupe ici les disciplines par les thèmes qu'elles débattent, plutôt que par les similarités entre leurs façons de traiter ces thèmes.}.
%
%Le premier groupe est issu des débats entre la biologie évolutionniste, la psychologie cognitive et les sciences sociales (et plus particulièrement l'anthropologie), desquels ont émergé les domaines de la psychologie évolutionniste et de l'anthropologie cognitive. %Plusieurs théories ont été avancées et sont débattues en parallèle par ces disciplines. On se doit de mentionner ici en premier lieu le débat autour du programme mémétique initié par \citet{Dawkins76} et pour lequel l'ouvrage collectif de \citet{Aunger00} est un bon exemple. Un deuxième terrain exploré par cet ensemble de disciplines est le développement de modèles évolutionnistes basés sur les normes (voir par exemple \citealp{Ehrlich05}), avec comme travail fondamental celui de \citet{Boyd85}. Le troisième champ, particulièrement important aujourd'hui, est précisément celui de l'épidémiologie culturelle, initié par \citet{Sperber96} et défendu face aux autres courants dans des travaux comme celui de \citet{Atran03}.
%
%Le deuxième groupe, plus diffus, provient des échanges entre le \emph{data mining}, l'étude des systèmes complexes, et la sociologie quantitative (e.g. sociologie des réseaux). Un grand nombre de travaux apparaissent exploitant de grands corpus de données produits par des collectes sur Internet, allant de l'étude de l'impact des réseaux sociaux sur la diffusion des informations sur ces mêmes réseaux \citep{Bakshy12} à l'analyse du rôle de la presse en ligne dans la vie politique locale \citep{Parasie12}, en passant par le détail des interactions entre éditeurs sur Wikipedia \citep{Jurgens12} ou l'évolution du vocabulaire des langues au cours du temps \citep{Lieberman07}. Une partie de ces travaux n'inscrivent leurs résultats que dans des cadres théoriques naïfs (autour de thèmes pour lesquels il existe pourtant des théories réfléchies développées par des champs des sciences sociales), et relèvent parfois plus des sciences de l'informatique que de l'étude de la culture (d'aucuns parlent d'ailleurs de \emph{social computing}). Ce groupe de disciplines présente néanmoins la particularité de vouloir étudier la culture en exploitant les possibilités ouvertes par les larges corpus de données issus d'Internet.

### 2014-01-22 brains-protocol.tex
Add all files, initial import

% !TEX root = brainscopypaste.tex
\section{Protocol}\label{sec:protocol}

%\tb{On a choisi de prendre une forme de représentation publique en apparence basique, mais qui entre bien dans la définition que \citeauthor{sper:expl} fait des représentations publiques~\citep[p.~49]{sper:expl} : les citations. C'est une forme de représentation d'autant plus instructive à analyser que, si l'on s'en tient aux règles de l'écrit, les citations ne devraient pas changer : lorsqu'on cite quelqu'un il est attendu qu'on reproduise précisément les mots utilisés par l'auteur, éventuellement en n'en prenant qu'une partie pour mieux l'intégrer à son propos, mais certainement pas en en changeant les termes. On peut faire l'hypothèse que les auteurs qui citent \emph{cherchent à le faire} en suivant cette règle. De ce fait, si l'on observe en effet une évolution des citations au-delà du "rognage" elle doit être l'effet d'un biais cognitif involontaire (de l'ordre de l'automatique). L'analyse de l'évolution des citations sur Internet est donc un terrain idéal pour observer un biais cognitif à l'origine de transformations des représentations publiques.}
%
%\tb{Revenons brièvement sur l'hypothèse faite ci-dessus car elle mérite un peu plus d'attention pour être valide. Elle est justifiée pour la simple raison qu'il est souvent facile d'aller vérifier les propos d'origine, et parce que cette transparence fait partie de la raison d'être et de la légitimité mêmes des citations. En effet, les auteurs qui en citent d'autres peuvent tout à fait intégrer une citation en en déformant l'intention d'origine, par exemple en la rognant à souhait ou en extrayant la citation de son contexte, ou encore en la juxtaposant à d'autres citations qui renforcent le propos de l'auteur citant, mais il est rarement à l'avantage de l'auteur de changer les mots utilisés. Ceci pour deux raisons : d'une part, parce que la légitimité et la force de l'information transmise ou de l'argumentaire déployé reposent en partie sur l'honnêteté intellectuelle qu'on peut prêter à l'auteur, d'autre part parce que l'exactitude des citations est justement le premier des indicateurs à utiliser lorsqu'on cherche à mesurer cette honnêteté intellectuelle.}
%
%\tb{Enfin, et pour terminer de justifier cette hypothèse, pour que ce garde-fou fonctionne il n'est pas nécessaire que les citations faites par un auteur soient vérifiées systématiquement, ni même qu'elles soient vérifiées souvent : il suffit que toute l'information nécessaire à la vérification soit publique, pour que le risque encouru par la déformation d'une citation soit bien réel. C'est bien souvent le cas sur Internet\footnote{De la même manière qu'exiger que les comptes d'une entreprise soient publics garantit un certain niveau d'"honnêteté comptable" (sans pour autant que des vérifications soient faites systématiquement), le fait que les données nécessaires à la vérification d'une citation soient publiques est un bon garant du fait que les auteurs qui en citent d'autres cherchent à rester fidèles aux mots d'origine.}. Il n'est pas non plus nécessaire que la vérification soit facile à faire pour que cette garantie tienne, mais il suffit que le risque de chute de légitimité soit assez grand pour que peu de monde s'y risque sérieusement\footnote{S'il n'est pas toujours facile de vérifier qu'une citation est exacte, le jour où quelqu'un montre qu'elle est inexacte il sera bien compliqué pour l'auteur citant de se défendre. Ce raisonnement s'applique moins à la déformation du contexte de la citation, cas dans lequel il est plus facile de se défendre en arguant de la divergence d'interprétation. La \emph{citation inexacte}, en revanche, relève plus de la \emph{donnée fausse} que de l'\emph{interprétation divergente}.}.}

To measure cognitive bias during reformulation, we chose to focus on public representations which could be expected to undergo variations which could be described easily and automatically.
Quotations appeared to be a perfect candidate. First, they are generally cleanly delimited by quotes, which greatly facilitates their detection in text corpora. Second, they stem from a unique ``original'' version, and could ideally be traceable back to it.
Third, and perhaps most importantly, their duplication should \emph{a priori} be highly faithful, apart from cropping issues: not only should variations be of a moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quote.

\subsection{Dataset} % .....

We thus relied on a large and reliable quotation dataset collected by \citet{Leskovec09} which 
%matched these essential criteria: made of public representations, subject to variations which could be expected to be of a relatively moderate magnitude, be ideally described by reference to an ideal original representation, and 
is of a sufficiently large size to lend itself to statistical analysis.
%\tb{On a donc choisi de s'appuyer sur un corpus déjà disponible au lieu de construire le nôtre, ce qui est une des raisons pour lesquelles l'analyse de la question est si tributaire du choix des données.  Le jeu de données utilisé par  et \citet{Simmons11} se prête bien à la notion de représentation publique qu'on a choisie ci-dessus. Construit et publié en 2009 par \citeauthor{Leskovec09}, c'est un jeu de données reconnu et utilisé dans la littérature, largement pré-traité pour nos besoins comme on le verra ci-dessous. Il est d'une dimension bien suffisante pour atteindre des niveaux de significativité intéressants, tout en restant manipulable sur un serveur de calcul modeste ou même sur un bon ordinateur portable.}
This dataset consists of the daily collection of news stories or blog posts from around a million online sources, with the approximate publication rate of 900k texts per day, over a nine-month period ranging from August 2008 to April 2009 \cite{Leskovec09-url}.\footnote{The original article~\citep{Leskovec09} does not provide additional details on the method of source selection.}
Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the cited person. Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words). A community detection algorithm is applied on that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).  Eventually, the data contains around $70,000$ sets of quotations. Each set admittedly contains all variations of a same parent utterance, together with the respective URLs and timestamps of publication of each individual variation. % We have the list of occurrences, URL and publication time of the quotations of a given such set. %  which are possibly slightly differing from a parent utterance.

%\tb{Un des travaux des auteurs a été de reconstituer ces ensembles de citations par une méthode de catégorisation (i.e. construction de \emph{clusters})\footnote{Se référer à \citet{Leskovec09} pour plus de détails. Ce traitement des données comporte aussi quelques étapes de filtrage, décrites dans l'article.}, et c'est à ce stade de traitement qu'on obtient les données : elles sont constituées d'environs 70~000 ensembles de citations, à l'intérieur desquels toutes les citations proviennent d'un même énoncé parent (elles en sont des sous-parties transformées). Pour chaque citation à l'intérieur de chaque ensemble, on a une liste de triplets \emph{(URL, date d'apparition, nombre d'occur\-rences à cette URL)} décrivant les adresses et les dates auxquelles cette citation est apparue.}

\subsection{Word-level measures}

\subsubsection{Standard psycholinguistic indices}

\RK{introduce the classical psycholinguistic measures -- kind of like ``demographic'' usual suspects}

We also considered grammatical types within quotations by detecting of so-called \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology \citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and 
adverbs.  The results were however extremely similar across the various categories, exhibiting no specific behavior of words belonging to different POS categories. %. \tb{Enfin, on ajoute à ces mesures la catégorie grammaticale, dans les citations, des mots échangés lors des substitutions. On s'est restreint à quatre des catégories les plus fréquentes, en utilisant la catégorisation et la notation des \emph{Part-of-Speech} (abrégé POS)\footnote{Le terme \emph{Part-of-Speech}, ou "partie du discours" en français, désigne les catégories grammaticales des mots (ou même des parties de mots) utilisées en linguistique.} du Penn TreeBank Project\footnote{Le Penn TreeBank Project définit une typologie très complète des POS, ainsi qu'une façon de les noter : à chaque POS correspond un code de deux ou trois lettres majuscules. Par exemple, tous les POS tenant du verbe commencent par la lettre $V$ (e.g. $VBN$ pour participe passé, $VBG$ pour participe présent, etc.).}~\citep{Santorini90} : verbes (POS dont le code commence par $V$), noms (POS dont le code commence par $N$), adjectifs (POS dont le code commence par $J$), et adverbes (POS dont le code commence par $R$). Cette mesure supplémentaire affine les observables définies ci-dessus, et permet d'observer un éventuel effet des catégories grammaticales sur les résultats.}

\subsubsection{Network-based definitions}
We rely on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citep{Nelson04}.

The WordNet (WN) database gathers more than 117~000 concepts and about 147~000 words attached to these concepts. WordNet entries are disambuised words, or ``\emph{lemmas}'', which are rattached to concepts called ``synsets'', each of them gathering the set of lemmas corresponding to the same single meaning. Note that the choice of a synset for a lemma immediately determines the grammatical category of that lemma. Lemmas make the nodes of the semantic network (homonyms are aggregated into a single node), whose neighbors are synonymous lemmas belonging to the same synset(s) as the lemma(s) represented by that node.  In other words, synsets induce cliques in the semantic network. Links connecting two synonymous lemmas are being assigned a weight corresponding to the number of times these lemmas are mentioned by WordNet as synonyms, i.e. in the same synset.

\rk{Explain Free Association}

\TB{The original WN network is a weighted, undirected network; the FA network is a directed network where weights have a different meaning. Both semantic networks describe synonymy relationships corresponding to qualitatively distinct forms of association: either from a linguistic viewpoint (WN) or from the perspective of actual information retrieval by humans (FA).  In  order to work in a unified framework, we consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighted networks.}

\bigskip
We introduce three standard network-based features, which may also be relevant psycholinguistically:
\begin{itemize}
	\item \textbf{Distance}, a classical dyadic network measure corresponding to the shortest path length between two nodes. This measure may be used as a proxy of the remoteness between two given lemmas, which is likely to have an influence on the likelihood of being a relevant substitution candidate.
	\item \textbf{Centrality} $k$, measured by the number of neighbors of a given node, and interpretable as a proxy for the polysemy of words. For the directed network FA we may also define the incoming centrality $k'$ (number of neighbors pointing \emph{to} that node).  Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank} \cite{Page99}, which had already been used in \cite{Griffiths07} and may be interpreted as a generalized and recursive measure of lemma polysemy --- central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on, recursively.\footnote{{Betweenness coefficient}, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network \cite{free:set}. More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher, the more important is that node in ensuring the connectedness of the rest of the network. Correlation between FA-BC and FA-PR is $0.74$; between WN-BC and WN-PR it is $0.83$.}
	\item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs. possible links between a node's neighbors \cite{watt-coll}. A lemma is likely to have a higher clustering coefficient whenever the synsets which it belongs to are themselves semantically linked. This measure had been used by \citet{Chan10}.
\end{itemize}

\subsubsection{Variable correlations}

It would be difficult to discuss possible correlations between distance ($d$), on one side, and centrality ($k$ or $k'$) or clustering coefficient ($c$) on the other side, as the former is a dyadic measure while the latter are monadic measures.  We may however note that centrality and clustering are relatively weakly (and negatively) correlated, with a Pearson coefficient of $-0.26$.

An important question here relates to the various possible correlations between those variables.  The age of acquisition is a key variable, primarily as a usual suspect in psycholinguistic studies, and also appears to be generally relatively correlated to all the other variables.  This coincides with an ongoing debate suggesting that age of acquisition encodes a variety of phenomena which are difficult to disentangle from more specific phenomena, which could be captured by more independent variables such as \CN.
In effect, the minimal linear correlation of $aoa$ ($-0.31$) with other variables has a higher magnitude than the maximal linear correlation of other variables with each other ($-0.26$).

Additionally, phonemes and syllables exhibit a strong linear correlation ($0.85$).  We see a better prediction effect of phonemes over syllables, which is consistent with \cite{nick-diss}, and choose to focus on phonemes only. \rk{check that!}

\bigskip
Network properties, finally, appear to be relatively weakly correlated with each other, across one network or between both networks (i.e. nodes with high index values are generally distinct between FA and WN), and may be used . We rely on FA as there is a direct possible psychological interpretation (that human copy-pasting may or may not show the same type of trend as human free association), noting that prediction results will be matched against WN as well. \rk{comment peut-on dire que les nœuds importants dans FA ne sont pas nécessairement importants dans WN mais que, pourtant, la forme de la prédiction est la même? cela me semble illogique en première lecture}

\begin{figure}[!th]
\includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
\end{figure}

\subsection{Detection of substitutions}

Nonetheless, these sets bear no explicit mention of which is the authoritative, original quotation.  Defining substitutions from the dataset therefore requires additional assumptions.  
%\tb{Il manque en effet certaines informations dans le corpus pour pouvoir lire ces substitutions de façon immédiate dans les données : les métadonnées pour chaque citation ne concernant que les URLs, les dates d'apparitions et les nombres d'occurrences à chaque URL, on n'a aucune information sur les liens entre les citations. Or, on a besoin de savoir sur quelles sources un auteur s'est appuyé pour reproduire une citation, c'est-à-dire qu'on a besoin de savoir quelle citation a été générée à partir de quelle autre. Cette information n'est pas présente dans les données et on va donc devoir l'inférer, à l'aide de plusieurs hypothèses.}

Indeed, an author writing a new post and reproducing (possibly altering) a quotation bases himself on original quotations we have no access to. We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quote for each occurrence of each quote. We therefore model the underlying quote selection process by making a few additional assumptions which let us define substitutions from the available data.

The main question in this problem concerns the way we consider later occurrences of a quote which, when it first appeared, was identified to be an alteration from an original quote. Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on Jan 18th, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January. If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the author of $q_b$ on the 19th based himself on $q_a$. But what about the following occurrences of $q_b$? Should we consider them to be substitutions based on $q_a$ (i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?

To settle this question we model the process as follows: we assume that when a quote $q$ appears at time $t$, if not original (i.e. if not stemming from a source external to the dataset, e.g. initiating a new set of quotations), it is based solely on the most frequent quote $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$. The length $\Delta t$ of that period of time is fixed as a fraction of the total duration of the considered set of quotations; we took one fifth in the implementation (i.e. one month for a set of quotations spanning five months, or one day for a set spanning only five days; this takes the dynamism of the set of quotes into account). If $q$ differs from $q_{max}$ by only a word, it is counted as a substitution from $q_{max}$ to $q$. In any other case, i.e. if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.

An example of this detection in a situation akin to the one described above can be seen on figure \ref{fig:model-slidetimebags}: $q_a$ and $q_b$ differ by only a word, and $q_a$ appears first and stays the most frequent in the beginning. This is why the occurrences of $q_b$ at $t_1$ and $t_2$ are detected as substitutions stemming from $q_a$. After a time the situation is reversed: $q_b$ becomes more frequent than $q_a$. This entails that occurrences of $q_b$ are seen as reproductions of itself ($t_3$), and that occurrences of $q_a$ are detected as substitutions stemming from $q_b$ ($t_4$), i.e. \emph{re-creations} of $q_a$.

\begin{figure*}[h]
	\centering
	\def\svgwidth{\textwidth}
	\small
	\input{images/model-slidetimebags-b.pdf_tex}
	\caption{Detection of substitutions according to model \rk{update text in figure}}
	\label{fig:model-slidetimebags}
\end{figure*}

The assumptions embedded in this model are only a subset of a wider set of possibilities, all leading to alternative models. We identified and implemented five other such possible models, which all yielded essentially the same results. These models differ in the definition of the source quote from which new occurrences stem, essentially modifying the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance. For example, the time-windows considered can have different lengths, can include all occurrences from the beginning of the set of quotations, or can have fixed positions. The source quote of a potential substitution can be chosen among those time-windows, or otherwise (e.g. among \emph{all} quotes having appeared before $t$ and differing by a word from the considered arrival quote; this detection process would include all possible substitutions detected by other models, but would also include many false-positives).

\subsection{Characterization of substitutions}

We then measure the cognitive alteration of human copy-pasting by comparing the relative properties of substituted and substituting words in quotations.  We only consider altered quotations, \hbox{i.e.} quotations where at least one word has been modified; in other words, we measure the properties of alteration \emph{knowing that} there has been a variation (which indicates a certain reformulation) while we do not take invariant quotations into account, as it is impossible to know, in that case, whether there has been perfect individual reformulation, or just digital copy-pasting of a source (``{\sf Ctrl-C}/{\sf Ctrl-V}'').

We build two main observables for each word property. First, we measure the variation of word features over a substitution, looking at the difference of the value of a given feature between arrival and start words. %This address the question ``how do word features vary over a subs  \tb{la variation de caractéristique sémantique lors du passage du mot remplacé au mot remplaçant dans une substitution. Ceci répond à la question "comment fait-on varier la caractéristique sémantique lorsqu'on substitue un mot ?".} 
Second, we measure the susceptibility for words to be the target of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to be the target of a substitution. % Second, \tb{la susceptibilité que les mots ont à être remplacés, en fonction de leur caractéristique sémantique. Cette mesure répond à la question "quelles caractéristiques sémantiques a-t-on plus tendance à substituer ?"%\footnote{Attention, cette mesure ne suffit pas en elle-même pour parler d'\emph{influence} de la caractéristique sémantique d'un mot sur le fait qu'il soit substitué ou non, parce qu'on n'a pas fait d'analyse des causes impliquées dans le fait qu'un mot soit substitué ; en effet, d'autres facteurs pourraient être plus importants que ces caractéristiques sémantiques dans les causes d'une substitution. On ne répond donc \emph{pas} à la question "quelle est l'influence de la caractéristique sémantique d'un mot sur sa propension à être substitué ?".}.

\subsubsection{Alteration}
%The above protocol enables us to observe in detail how word features vary upon substitution. 
For a given feature $\mathfrak{F}$, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$. Let us denote this quantity:
$$\Delta (\wstart,\warrival) \defeq \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart)$$

Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster, we first average substitutions over each cluster, by considering the average of arrival word features for a given starting word. Indeed, substitutions occurring in the same cluster are likely not statistically independent.} This 
quantity can be written:
$$\left< \Delta(\wstart,\warrival) \right>_f = \left< \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from % , let us denote the corresponding quantity This value is to be compared to the variation the word's feature would have undergone had the arrival word been chosen randomly in 
the pool of \TB{WN/FA} words. The corresponding average quantity over all start words may be written as:\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
%$$\Delta_{\mathcal{H}_0} (\wstart,\warrival) \defeq f\left(\warrival_{\mathcal{H}_0}\right) - f\left(\wstart\right)$$
%and, similarly, we compute the average over all start words 
$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f
= \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
%Averaging over all start words with a given feature value $f$, we get $$\left< \Delta_{\mathcal{H}_0} \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace \wstart | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word} (neighbors), i.e. an arrival word chosen among semantically plausible though still random words (in this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$). %The values for $d$ were decided upon according to the distribution of distances travelled on the synonym graph upon substitution, normalized by the number of synonyms at that distance. 
%  brought all variation curves shown below nearer to the origin, but did not change the results qualitatively.
%The variation of feature compared to $\mathcal{H}_0$ is then $\left< \Delta \right>_f - \left< \Delta_{\mathcal{H}_0} \right>_f$.

Using this method we obtain the mean variation of feature for each start feature value, and can compare the variations to a situation where arrival words are chosen randomly. This gives us a fine-grained view of how word features evolve upon substitution.

\subsubsection{Susceptibility}

Furthermore, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$. 
We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another. 
Word substitution susceptibility (denoted $\mathfrak{S}_{\mathfrak{F}}(w)$) is computed as the ratio of the number of times $n_s(w)$ a word is substituted to the number of times $n_p(w)$ that word appears in a substitutable position.  We have:
$$\mathfrak{S}_{\mathfrak{F}}(w) \defeq \frac{n_s(w)}{n_p(w)}$$

Now averaging over all words having a given feature value $f$, we obtain the mean susceptibility for the feature value $f$:
$$\left< \mathfrak{S}_{\mathfrak{F}} \right>_f = \left< \frac{n_s(w)}{n_p(w)} \right>_{\left\lbrace w | \mathfrak{F}(w) = f \right\rbrace}$$

This measure focuses on the selection of start word instead of the selection of the arrival word. Indeed, the features have an effect not only on the choice of a new word when a substitution takes place, but also at the preceding moment when it is not yet known which word in the quote --~if any~-- will be substituted.

### 2014-01-22 brains-results.tex
Add all files, initial import

% !TEX root = brainscopypaste.tex
\section{Results}\label{sec:results}

\TB{We may see that $\mathcal{H}_0$ and $\mathcal{H}_{00}$ are slightly translated yet not qualitatively different.}

\subsection{Location of arrival words}

Where in the network are arrival words, are they related to the origin word, when relying on a given semantic network?  (In other words, would a semantic word be a good predictor of the arrival of words?). 
\rk{tout un blabla sur les distances dans les substitutions.}

\subsection{Known psycholinguistic effects}

We first looked at well-known (and well-studied) psycholinguistic features, and doing so we were able to determine which of the two alternative hypotheses presented in the introduction is valid; the features we examined were age of acquisition norms\footnote{The Age-of-Acquisition data is obtained from \citet{Kuperman12}.} and number of phonemes\footnote{The number of phonemes are obtained from the CMU Pronouncing Dictionary for U.S. English \citep{Weide98}, included in the NLTK package \citep{Bird09}. We note that these two measures are likely to be linked, since words with larger numbers of phonemes are likely to be learned later in development.}. First, by looking at substitution susceptibilities for those features, we can see that words with lower age of acquisition (figure \ref{fig:aoa-suscept}) and words with lower number of phonemes (figure \ref{fig:MNphonemes-suscept}) are more likely to be substituted than words with higher values for either of those features. This shows us that, although words learned later in development (as well as words with larger number of phonemes) may be cognitively harder to recall, they seem to be substituted according to hypothesis [X]: their relevance and specificness makes them less susceptible to change, whereas words learned earlier or with smaller number of phonemes are more easily substituted.

Now looking at what type of arrival words are selected upon substitution, we can see that both words learned earlier in development (figure \ref{fig:aoa-absolute}) as well as words with lower number of phonemes (figure \ref{fig:MNphonemes-absolute}) are substituted for words with roughly the same feature values (they tend to augment a little, but stay way below the values expected under $\mathcal{H}_0$ or $\mathcal{H}_{0,n}$). Conversely, words learned later in development and words with higher number of phonemes, when substituted, go towards words with lower feature values, but no lower than what the null hypotheses predict. This shows again that words learned early on are easily interchangeable for other words learned at about the same age. On the other hand, it seems the cognitive system puts little constraint on the arrival word when substituting a word learned later in development (that is, those words are rarely substituted, but when a substitution does occur the arrival word does not show constraints related to the features we examined). We also tested these results for dependence on the grammatical category of words, i.e. POS tags, and found no effect.

Though not originating from the better-known image-naming or controlled lexical-retrieving tasks, these results shed a new light on the strength of the effect of these features depending on their exact value, and on how these features behave in the context of \emph{substitution} and not only free recall.

\begin{figure*}[!th]
	\centering
	\subfloat[][Age-of-Acquisition]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_aoa-Kuperman-all_suscept.png}
		\label{fig:aoa-suscept}
	}
	~
	\subfloat[][Number of phonemes]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_cmu-MNphonemes-all_suscept.png}
		\label{fig:MNphonemes-suscept}
	}
	\caption{Susceptibility values in color over the histograms of feature values}
\end{figure*}

\begin{figure*}[!th]
	\centering
	\subfloat[][Age-of-Acquisition]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_aoa-Kuperman-all_absolute.png}
		\label{fig:aoa-absolute}
	}
	~
	\subfloat[][Number of phonemes]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_cmu-MNphonemes-all_absolute.png}
		\label{fig:MNphonemes-absolute}
	}
	\caption{Variations of features, with corresponding values under $\mathcal{H}_0$ and $\mathcal{H}_{0,n}$}
\end{figure*}

\subsection{Epidemiological setting}

Secondly, and coming back to our primary goal of providing a first empirical testing means for epidemiological models of culture, we added to these two features a number of more abstract properties computed from the Free Association norms used by \citet{Griffiths07}: namely, the PageRank, betweenness coefficient and clustering coefficient of the words.

These new features are classical measures for network-related matters, but --~to our knowledge~-- have seldom been applied to word characterization. They are only a few among many other features that can be used to characterize words based on the network they form (be it the Free Association norms network, another semantic network, or even a phonological network). We consider the complete set of these features (which includes the two mentioned in the previous section, the new network-based ones, as well as any other network-based feature one can compute) as traits characterizing words in an empirical epidemiological evolution of quotes. The benefit we get from measuring the evolution of such features is in how we can use this information to falsify epidemiological culture models. Indeed, in such a setting the information on how a feature is modified upon substitution (detailed to the individual feature value) is in fact a \emph{fitness landscape} for that particular feature. This fitness landscape can in turn be re-injected into epidemiological models to see if those models account for the empirical distribution of quotes, their life-span and relative success. The data for PageRank, betwenness coefficient and clustering coefficient are shown in figures \ref{fig:PR_scores-absolute}, \ref{fig:BCs-absolute} and \ref{fig:CCs-absolute}.

\rk{some talk about what this confirms in Griffith's work}

\begin{figure*}[!th]
	\centering
	\subfloat[][Log PageRank]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-PR_scores-all_absolute.png}
		\label{fig:PR_scores-absolute}
	}
	~
	\subfloat[][Log Betweenness coefficient]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-BCs-all_absolute.png}
		\label{fig:BCs-absolute}
	}
	~
	\subfloat[][Log Clustering coefficient]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-CCs-all_absolute.png}
		\label{fig:CCs-absolute}
	}
	\caption{Variations of features, with corresponding values under $\mathcal{H}_0$ and $\mathcal{H}_{0,n}$}
\end{figure*}

Moreover, the substitution susceptibility corresponds to a \emph{mutation probability} in this setting (shown in figures \ref{fig:PR_scores-suscept} for PageRank, \ref{fig:BCs-suscept} for betweennes coefficients, and \ref{fig:CCs-suscept} for clustering coefficient). To sum up, the data obtained here are empirical measures of (some of) the parameters of epidemiological models of cultural evolution, i.e. essential information to allow empirical testing of these models. If quotes, as a first example of cultural representation, do indeed follow epidemiological rules in their evolution, we may be able to empirically prove the validity -- or falsify -- the existing models of cultural evolution.

\rk{this is very rough and brute-like. It needs to be better presented and further developed.}

\begin{figure*}[!th]
	\centering
	\subfloat[][Log PageRank]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-PR_scores-all_suscept.png}
		\label{fig:PR_scores-suscept}
	}
	~
	\subfloat[][Log Betweenness coefficient]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-BCs-all_suscept.png}
		\label{fig:BCs-suscept}
	}
	~
	\subfloat[][Log Clustering coefficient]{
		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-CCs-all_suscept.png}
		\label{fig:CCs-suscept}
	}
	\caption{Susceptibility values in color over the histograms of feature values}
\end{figure*}

%\begin{figure}[!th]
%\includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
%\end{figure}

### 2014-01-22 algorithms/my.plotcorr.R

@@ -1,0 +1,104 @@
+my.plotcorr <- function (corr, outline = FALSE, col = "grey", upper.panel = c("ellipse", "number", "none"), lower.panel = c("ellipse", "number", "none"), diag = c("none", "ellipse", "number"), digits = 2, bty = "n", axes = FALSE, xlab = "", ylab = "", asp = 1, cex.lab = par("cex.lab"), cex = 0.75 * par("cex"), mar = 0.1 + c(2, 2, 4, 2), ...)
+{
+# this is a modified version of the plotcorr function from the ellipse package
+# this prints numbers and ellipses on the same plot but upper.panel and lower.panel changes what is displayed
+# diag now specifies what to put in the diagonal (numbers, ellipses, nothing)
+# digits specifies the number of digits after the . to round to
+# unlike the original, this function will always print x_i by x_i correlation rather than being able to drop it
+# modified by Esteban Buz
+  if (!require('ellipse', quietly = TRUE, character = TRUE)) {
+    stop("Need the ellipse library")
+  }
+  savepar <- par(pty = "s", mar = mar)
+  on.exit(par(savepar))
+  if (is.null(corr))
+    return(invisible())
+  if ((!is.matrix(corr)) || (round(min(corr, na.rm = TRUE), 6) < -1) || (round(max(corr, na.rm = TRUE), 6) > 1))
+    stop("Need a correlation matrix")
+  plot.new()
+  par(new = TRUE)
+  rowdim <- dim(corr)[1]
+  coldim <- dim(corr)[2]
+  rowlabs <- dimnames(corr)[[1]]
+  collabs <- dimnames(corr)[[2]]
+  if (is.null(rowlabs))
+    rowlabs <- 1:rowdim
+  if (is.null(collabs))
+    collabs <- 1:coldim
+  rowlabs <- as.character(rowlabs)
+  collabs <- as.character(collabs)
+  col <- rep(col, length = length(corr))
+  dim(col) <- dim(corr)
+  upper.panel <- match.arg(upper.panel)
+  lower.panel <- match.arg(lower.panel)
+  diag <- match.arg(diag)
+  cols <- 1:coldim
+  rows <- 1:rowdim
+  maxdim <- max(length(rows), length(cols))
+  plt <- par("plt")
[…]

### 2014-01-22 brains-introduction.tex
Phrasings and obvious mistakes

@@
 A third approach, ``cultural epidemiology'' initiated by \citet{sper:expl} recently attracted significant attention.
-Works such as \citet{Atran03} suggest that this stance could be anthropologically more satisfying than memetics; see \citet{Kuper00}, and \citet{Bloch00}, for further details on the main issues in this debate.
+Works such as \citet{Atran03} suggest that this program is anthropologically more accurate than memetics; see \citet{Kuper00}, and \citet{Bloch00}, for further details on the main issues in this debate.
 It focuses on the notion of \emph{representation}, linking the cognitive science concept of \emph{mental representation} to the concept of \emph{public representation}, the latter being the equivalent of the former outside of the brain (\hbox{i.e.} in cultural artefacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
@@
 The last decade has however witnessed an avalanche of observable \emph{in vivo} data stemming from online interactions.
-These large amounts of textual traces of information production by humans are plausibly not records of strictly ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), yet they actually constitute a wealth of observations on the dynamics of public representations.
+While they are not records of strictly ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), these information trails and productions actually constitute a wealth of observations on the dynamics of public representations.
 As such, they are prone to radically improve the prospects of empirical validation of the individual-level processes of cultural evolution.
@@
 To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
-Admittedly, these verbatim public representations should not suffer any reformulation alteration (as opposed to more sophisticated expressions and opinions), but still do.
+Although these verbatim public representations should not suffer any alterations as they are repeated (as opposed to more elaborate expressions and opinions), in practice they are oftentimes transformed.
 We will in particular exhibit a non-trivial cognitive alteration by which some portions of the quotation, namely individual words, may be modified; and will wonder which are generally the semantic and structural characteristics of these words and their substitutions.
@@
 For one, the notion of ``cultural attractor'' has been the focus of models rendering the evolution of some representations and behaviors not until a few years ago \citep{Claidiere07}.
-Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist of a series of papers investigation \emph{quotation} transformations from a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
+Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist of a series of papers investigating \emph{quotation} transformations from a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
 They generally show several types of regularities and propose models of diffusion-transformation of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.
@@
 
-\TB{The study of network-based properties, while heavily discussed around the notion of connectionism and its normative processual models, appeared much more recently as an empirical investigation field \CN} \rk{C'est sûr ?
-e.g.
-http://tiny.cc/bl7usw}.
+The study of network-based properties, while heavily discussed around the notion of connectionism and its normative processual models, appeared much more recently as an empirical investigation field \CN.
 Let us mention two interesting and recent studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors in the cognitive processes and reproduction of those words.
@@
 Les auteurs montrent qu'une certaine propriété des mots (le PageRank calculé sur un réseau sémantique construit à partir de données d'associations de mots) est un bon facteur prédictif de quel mot sera rappelé parmi les mots commençant par la lettre demandée ; ils montrent ainsi un lien entre une caractéristique sémantique quantitative des mots et la facilité de rappel de ces mots dans cette tâche.
-A second psycholinguistic study by \citet{Chan10} shows, in a task of picture naming, that words come quicker when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
+A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words come quicker when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
 %, étudie une tâche où l'on demande aux sujets de nommer oralement une image.

### 2014-01-22 brains-protocol.tex
Phrasings and obvious mistakes

@@
 Quotations appeared to be a perfect candidate.
-First, they are generally cleanly delimited by quotes, which greatly facilitates their detection in text corpora.
+First, they are generally cleanly delimited by quotation marks (or by HTML markup in web pages), which greatly facilitates their detection in text corpora.
 Second, they stem from a unique ``original'' version, and could ideally be traceable back to it.
@@
 Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
-A community detection algorithm is applied on that quotation graph to detect aggregates of tightly connected, i.e.
+A community detection algorithm is applied to that quotation graph to detect aggregates of tightly connected, i.e.
 sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
-Eventually, the data contains around $70,000$ sets of quotations.
-Each set admittedly contains all variations of a same parent utterance, together with the respective URLs and timestamps of publication of each individual variation.
+The final data contains around $70,000$ sets of quotations.
+Each set allegedly contains all variations of a same parent utterance, annotated with the respective URLs and timestamps of publication of each individual variation.
 % We have the list of occurrences, URL and publication time of the quotations of a given such set.
@@
 
-We also considered grammatical types within quotations by detecting of so-called \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology \citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and
+We also considered grammatical types within quotations by detection of \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology \citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and
 adverbs.
@@
 \subsubsection{Network-based definitions}
-We rely on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citep{Nelson04}.
+We rely on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citet{Nelson04}.
 
 The WordNet (WN) database gathers more than 117~000 concepts and about 147~000 words attached to these concepts.
-WordNet entries are disambuised words, or ``\emph{lemmas}'', which are rattached to concepts called ``synsets'', each of them gathering the set of lemmas corresponding to the same single meaning.
+WordNet entries are disambiguated words, or ``\emph{lemmas}'', which are attached to concepts called ``synsets'', each of them gathering the set of lemmas corresponding to the same single meaning.
 Note that the choice of a synset for a lemma immediately determines the grammatical category of that lemma.
@@
 	\item \textbf{Centrality} $k$, measured by the number of neighbors of a given node, and interpretable as a proxy for the polysemy of words.
-For the directed network FA we may also define the incoming centrality $k'$ (number of neighbors pointing \emph{to} that node).
-Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank} \cite{Page99}, which had already been used in \cite{Griffiths07} and may be interpreted as a generalized and recursive measure of lemma polysemy --- central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on, recursively.\footnote{{Betweenness coefficient}, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network \cite{free:set}.
+For the directed network FA we also define the incoming centrality $k'$ (number of neighbors pointing \emph{to} that node).
+Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank} \cite{Page99}, which had already been used in \cite{Griffiths07} and may be interpreted as a generalized and recursive measure of lemma polysemy --- central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on, recursively.\footnote{\TB{{Betweenness coefficient}, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network \cite{free:set}.
 More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher, the more important is that node in ensuring the connectedness of the rest of the network.
-Correlation between FA-BC and FA-PR is $0.74$; between WN-BC and WN-PR it is $0.83$.}
+Correlation between FA-BC and FA-PR is $0.74$; between WN-BC and WN-PR it is $0.83$.}}
 	\item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs.
 possible links between a node's neighbors \cite{watt-coll}.
-A lemma is likely to have a higher clustering coefficient whenever the synsets which it belongs to are themselves semantically linked.
+A lemma is likely to have a higher clustering coefficient whenever the synsets to which it belongs are themselves semantically linked.
 This measure had been used by \citet{Chan10}.
@@
 
-It would be difficult to discuss possible correlations between distance ($d$), on one side, and centrality ($k$ or $k'$) or clustering coefficient ($c$) on the other side, as the former is a dyadic measure while the latter are monadic measures.
-We may however note that centrality and clustering are relatively weakly (and negatively) correlated, with a Pearson coefficient of $-0.26$.
+\TB{It would be difficult to discuss possible correlations between distance ($d$), on one side, and centrality ($k$ or $k'$) or clustering coefficient ($c$) on the other side, as the former is a dyadic measure while the latter are monadic measures.
+We may however note that centrality and clustering are relatively weakly (and negatively) correlated, with a Pearson coefficient of $-0.26$.} \rk{remove, redundant with below}
 
-An important question here relates to the various possible correlations between those variables.
-The age of acquisition is a key variable, primarily as a usual suspect in psycholinguistic studies, and also appears to be generally relatively correlated to all the other variables.
+An important question here relates to the various possible correlations between all the variables we consider.
+The age of acquisition is a key variable, primarily as a usual suspect in psycholinguistic studies, and also appears to be generally correlated to all the other variables.
 This coincides with an ongoing debate suggesting that age of acquisition encodes a variety of phenomena which are difficult to disentangle from more specific phenomena, which could be captured by more independent variables such as \CN.
-In effect, the minimal linear correlation of $aoa$ ($-0.31$) with other variables has a higher magnitude than the maximal linear correlation of other variables with each other ($-0.26$).
+\TB{Indeed, the minimal linear correlation of $aoa$ ($-0.31$) with other variables has a higher magnitude than the maximal linear correlation of other variables with each other ($-0.26$).} \rk{how do we get to that conclusion?}
 
@@
 Indeed, an author writing a new post and reproducing (possibly altering) a quotation bases himself on original quotations we have no access to.
-We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quote for each occurrence of each quote.
+We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quote for each instance of each quote.
 We therefore model the underlying quote selection process by making a few additional assumptions which let us define substitutions from the available data.

### 2014-01-22 brains-results.tex
Phrasings and obvious mistakes

@@
 
-We first looked at well-known (and well-studied) psycholinguistic features, and doing so we were able to determine which of the two alternative hypotheses presented in the introduction is valid; the features we examined were age of acquisition norms\footnote{The Age-of-Acquisition data is obtained from \citet{Kuperman12}.} and number of phonemes\footnote{The number of phonemes are obtained from the CMU Pronouncing Dictionary for U.S.
+\TB{We first looked at well-known (and well-studied) psycholinguistic features, and doing so we were able to determine which of the two alternative hypotheses presented in the introduction is valid; the features we examined were age of acquisition norms\footnote{The Age-of-Acquisition data is obtained from \citet{Kuperman12}.} and number of phonemes\footnote{The number of phonemes are obtained from the CMU Pronouncing Dictionary for U.S.
 English \citep{Weide98}, included in the NLTK package \citep{Bird09}.
-We note that these two measures are likely to be linked, since words with larger numbers of phonemes are likely to be learned later in development.}.
-First, by looking at substitution susceptibilities for those features, we can see that words with lower age of acquisition (figure \ref{fig:aoa-suscept}) and words with lower number of phonemes (figure \ref{fig:MNphonemes-suscept}) are more likely to be substituted than words with higher values for either of those features.
+We note that these two measures are likely to be linked, since words with larger numbers of phonemes are likely to be learned later in development.}.} \rk{repetition} First, by looking at substitution susceptibilities for those features, we can see that words with lower age of acquisition (figure \ref{fig:aoa-suscept}) and words with lower number of phonemes (figure \ref{fig:MNphonemes-suscept}) are more likely to be substituted than words with higher values for either of those features.
 This shows us that, although words learned later in development (as well as words with larger number of phonemes) may be cognitively harder to recall, they seem to be substituted according to hypothesis [X]: their relevance and specificness makes them less susceptible to change, whereas words learned earlier or with smaller number of phonemes are more easily substituted.
 
-Now looking at what type of arrival words are selected upon substitution, we can see that both words learned earlier in development (figure \ref{fig:aoa-absolute}) as well as words with lower number of phonemes (figure \ref{fig:MNphonemes-absolute}) are substituted for words with roughly the same feature values (they tend to augment a little, but stay way below the values expected under $\mathcal{H}_0$ or $\mathcal{H}_{0,n}$).
+Now looking at what type of arrival words are selected upon substitution, we can see that both words learned earlier in development (figure \ref{fig:aoa-absolute}) as well as words with lower number of phonemes (figure \ref{fig:MNphonemes-absolute}) are substituted for words with roughly the same feature values (they tend to augment a little, but stay way below the values expected under $\mathcal{H}_0$ or $\mathcal{H}_{00}$).
 Conversely, words learned later in development and words with higher number of phonemes, when substituted, go towards words with lower feature values, but no lower than what the null hypotheses predict.
@@
 
-Secondly, and coming back to our primary goal of providing a first empirical testing means for epidemiological models of culture, we added to these two features a number of more abstract properties computed from the Free Association norms used by \citet{Griffiths07}: namely, the PageRank, betweenness coefficient and clustering coefficient of the words.
+\TB{Secondly, and coming back to our primary goal of providing a first empirical testing means for epidemiological models of culture, we added to these two features a number of more abstract properties computed from the Free Association norms used by \citet{Griffiths07}: namely, the PageRank, betweenness coefficient and clustering coefficient of the words.} \rk{repetition}
 

### 2014-01-24 brains-introduction.tex
Rework introduction

@@
 
-%On assiste aujourd'hui à une convergence progressive entre des disciplines étudiant la cognition humaine, d'un côté, et des disciplines étudiant la culture et les \emph{choses sociales}, de l'autre.
-Ces deux familles de disciplines ont en effet bien des frontières en commun ; celles-ci se rendent visibles par l'intermédiaire des thèmes pour lesquels il devient incontournable de s'intéresser à l'humain à la fois dans sa globalité et dans ses fonctionnements particuliers.
+The understanding of the mechanisms behind cultural similarity and diversity has led to a sizeable literature in the recent past, spanning over a vast area of research fields ranging from cultural anthropology to social network analysis and complex systems modelling, all diversely labelled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'', for the most part.
+Broadly, this type of research program investigates phenomena pertaining to both cognitive science and social science, and aims to understand the processing, transmission, and evolution of information both at the individual and social levels.
 
-The understanding of the origin of cultural similarity has triggered a sizable literature in the recent past, spanning over a vast area of research fields ranging from cultural anthropology to social network analysis and complex system modeling --- and diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'', for the most part.
-Broadly, this type of research program relies on the understanding of phenomena pertaining to both cognitive science and social science, in order to jointly appraise the processing and transmission of information at the individual and social levels.
-% information propagation phenomena -- and where we witness the development of studies relying on both individual and social cognition.
+Several theories have been proposed and debated within these research fields, mixing social and individual cognition, notably from the side of cultural anthropology.
+First, the debate around the ``memetic'' program initiated by \citet{Dawkins76}, for which the collection of works by \citet{Aunger00} provides a solid overview.
+A second field has focused on the development of evolutionary models of norms (see for instance \citealp{Ehrlich05}), following the seminal work of \citet{Boyd85}.
 
-Several theories have been proposed and debated within these research fields, mixing social and individual cognition --- notably from the side of cultural anthropology.
-There is, first, the debate around the ``memetic'' program initiated by \citet{Dawkins76}, for which the collection of works by \citet{Aunger00} provides a solid overview.
-A second field relates to the development of evolutionary models of norms (see for instance \citealp{Ehrlich05}), following the seminal work of  \citet{Boyd85}.
-
-A third approach, ``cultural epidemiology'' initiated by \citet{sper:expl} recently attracted significant attention.
-Works such as \citet{Atran03} suggest that this program is anthropologically more accurate than memetics; see \citet{Kuper00}, and \citet{Bloch00}, for further details on the main issues in this debate.
-It focuses on the notion of \emph{representation}, linking the cognitive science concept of \emph{mental representation} to the concept of \emph{public representation}, the latter being the equivalent of the former outside of the brain (\hbox{i.e.} in cultural artefacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
-%\begin{quotation}
-{%Une représentation peut exister à l'intérieur même de l'utilisateur : il s'agit alors d'une \emph{représentation mentale}.
-Un souvenir, une hypothèse, une intention sont des exemples de représentations mentales.
-L'utilisateur et le producteur d'une représentation mentale ne font qu'un.
-Une représentation peut aussi exister dans l'environnement de l'utilisateur comme par exemple le texte qui est sous vos yeux.
-Il s'agit alors d'une \emph{représentation publique}.
-Une représentation publique est généralement un moyen de communication entre un producteur et un utilisateur distincts l'un de l'autre.
-``%
+A third approach, ``cultural epidemiology'', initiated by \citet{sper:expl} recently attracted significant attention.
+Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.
+The cultural epidemiology program focuses on the notion of \emph{representation}, linking the cognitive science concept of \emph{mental representation} to the concept of \emph{public representation}, the latter being the former's counterpart outside the brain (i.e.
+in cultural artefacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
+\begin{quotation}
 A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
 The producer and the user of a mental representation are one and the same person.
-A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.}%~(\citealp[p.~49]{sper:expl}.)
-''
-%\end{quotation}
-}
-The key point here is that representations are not being replicated through a hi-fidelity copying process, but are being reformulated through an interpretation process, and thus evolve.
-Cultural epidemiology postulates this type of conceptual mutation and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domains of an underlying semantic dynamical system.
-Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not been empirically validated.
+A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.
+\end{quotation}}
+The key point here is that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
+Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.
+Despite some recent modelling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analysed.
 
-%Bien qu'ayant connu des développements récents en modélisation (par exemple \citealp{Claidiere07}, pour la notion d'attracteur culturel), la difficulté à effectuer des mesures quantitatives sur les notions clés du domaine a fait que les hypothèses à la base de l'édifice théorique développé n'ont pas encore pu être testées empiriquement.
+However, the last decade has witnessed an avalanche of observable \emph{in vivo} data in the form of online interactions.
+While they are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), these productions and information trails still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
+Given the already significant --~and rapidly growing~-- importance of our online interactions, these records can dramatically improve the prospects of empirical study of the individual-level processes of cultural evolution.
 
-The last decade has however witnessed an avalanche of observable \emph{in vivo} data stemming from online interactions.
-While they are not records of strictly ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), these information trails and productions actually constitute a wealth of observations on the dynamics of public representations.
-As such, they are prone to radically improve the prospects of empirical validation of the individual-level processes of cultural evolution.
+We aim to empirically describe the transformation of a specific type of public representation, by focusing on the possible alterations introduced by individuals when newly producing a representation.
+To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
+While these verbatim public representations should in theory not suffer any alterations as they are produced anew (as opposed to more elaborate expressions and opinions, not identified as quoted utterances), empirical observation shows that they are in fact quite often transformed.
+We will in particular exhibit a non-trivial process by which individual words in quotations are replaced.
+We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
+More generally, we contend that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards building empirically realistic models of cultural evolution.
 
-We aim to empirically describe the transformation of a specific type of public representations by focusing on the possible cognitive alterations performed by individuals during reformulation.
-% On s'est proposé d'étudier de façon \emph{empirique} la transformation des représentations publiques, en se focalisant sur la mise en évidence d'un biais cognitif dans le traitement  par le cerveau de certaines représentations culturelles.
-To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
-Although these verbatim public representations should not suffer any alterations as they are repeated (as opposed to more elaborate expressions and opinions), in practice they are oftentimes transformed.
-We will in particular exhibit a non-trivial cognitive alteration by which some portions of the quotation, namely individual words, may be modified; and will wonder which are generally the semantic and structural characteristics of these words and their substitutions.
-%On verra que par cette analyse de la reproduction des citations par les auteurs on observe un biais cognitif non trivial ; on observera ainsi un lien entre l'attractivité de certains mots (en termes d'épidémiologie culturelle) et les propriétés sémantiques de ces mêmes mots, calculées sur la base du réseau formé par ces mots.
-%On s'est proposé d'étudier de façon \emph{empirique} la transformation des représentations publiques, c'est-à-dire d'explorer à l'aide de données obtenues in vivo l'effet du cerveau sur l'évolution des représentations ; une telle étude permettrait d'observer empiriquement l'effet d'un attracteur culturel tel que défini par Sperber.
-La question est donc la suivante : peut-on observer un biais cognitif dans le traitement de certaines représentations culturelles par le cerveau, et si oui quel est-il ?On verra que par cette analyse de la reproduction des citations par les auteurs on observe un biais cognitif non trivial ; on observera ainsi un lien entre l'attractivité de certains mots (en termes d'épidémiologie culturelle) et les propriétés sémantiques de ces mêmes mots, calculées sur la base du réseau formé par ces mots.
-Broadly, we contend that using this type of data is both equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards the design of empirically realistic models of cultural evolution.
-
-% --------------------------------
-%\subsection*{Problématique} % -----
-% --------------------------------
-\label{subsec:problematique}
-
-%\paragraph*{Plan}
-
-The next section (Sec~\ref{sec:related}) describes the relevant recent state-of-the-art on this matter.
+The next section (Sec~\ref{sec:related}) describes the state-of-the-art on this matter.
 In Sec.~\ref{sec:protocol}, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
 Section~\ref{sec:results} describes the significant psycholinguistic cues and biases observed during \emph{in vivo} quotation reformulation, followed by a discussion and general guidelines for further work in Sec.~\ref{sec:conclusion}.
-%Le travail de recherche effectué étant essentiellement empirique, on le présente en deux grandes étapes : en premier lieu, on détaille le protocole d'expérimentation utilisé ainsi que les différentes tentatives et pistes explorées pendant son élaboration.
-Pour cela, on commencera par analyser la question ci-dessus plus à fond afin d'en distiller une question opérationnelle, puis on expliquera en détail la façon dont on a construit le protocole expérimental permettant de faire les mesures pour répondre à cette question%\footnote{Le terme "protocole expérimental", utilisé ici, est bien le terme approprié pour ce dont il s'agit : on verra qu'on part d'un corpus de données pour lequel on construit une façon de faire les mesures qu'on considère pertinentes, et cette démarche correspond à la construction d'un protocole expérimental comme on peut le faire pour une expérience de psychologie cognitive ; c'est à ce moment qu'on définit ce qu'on va mesurer et comment cette mesure est faite.}
-%.
-La deuxième grande partie concerne les résultats : on montre tout d'abord la façon dont les mesures sont utilisées pour construire les observables qu'on cherche, ce qui permettra ensuite d'interpréter les résultats obtenus au regard de la question posée.
-%Une dernière partie concernera les perspectives que ce travail ouvre, et les façons dont il pourrait être poursuivi et étendu.
 
-\section{Related work}\label{sec:related}
+% ============================
+\section{Related work} % =====
+% ============================
+\label{sec:related}
 

### 2014-01-24 brains-introduction.tex
Rework "Related work" a little

@@
 The relevant literature on \emph{public representation dynamics} features two main streams.
-On one hand, we find studies aiming at understanding the macroscopic phenomena of \emph{social diffusion} of public representations, describing for instance the propagation of cultural items across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
-%L'état de l'art en ce qui concerne l'épidémiologie culturelle peut être résumé ainsi : d'un côté, des études %s'intégrant au deuxième groupe de disciplines
-% permettent une bonne compréhension des phénomènes macroscopiques de \emph{diffusion sociale} des représentations publiques, tels que le chemin emprunté par une information atomique donnée, les temps caractéristiques et les cycles de diffusion dans ces réseaux, la relation entre blogs et sites de médias, ou encore l'autorité et l'influence exercée par certains n\oe{}uds du réseau.
+On one hand, we find studies of the macroscopic \emph{social diffusion} of public representations, describing for instance the propagation of cultural artefacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
 These studies are relatively independent from anthropology and cognition and are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
-Without necessarily relying on specific social science theories, %these studies are interesting in that they aim at describing cultural dynamics by using large social media corpuses.
-this research stream is of interest for its aim of cultural dynamics description by using large social media corpora.
-% ,    Ces études sont relativement indépendantes de l'anthropologie et de la cognition et  se situent à l'interface entre le \emph{data mining}, l'étude des systèmes complexes, et la sociologie quantitative (e.g.
-sociologie des réseaux), en exploitant de grands corpus de données produits par des collectes sur Internet.
-Sans nécessairement s'appuyer sur des cadres théoriques précis ou intégrés, ces études présentent néanmoins la particularité de vouloir décrire les dynamiques culturelles en exploitant les possibilités ouvertes par les larges corpus de données issus d'Internet.
-%, allant de l'étude de l'impact des réseaux sociaux sur la diffusion des informations sur ces mêmes réseaux \citep{Bakshy12} à l'analyse du rôle de la presse en ligne dans la vie politique locale \citep{Parasie12}, en passant par le détail des interactions entre éditeurs sur Wikipedia \citep{Jurgens12} ou l'évolution du vocabulaire des langues au cours du temps \citep{Lieberman07}.
-%On peut citer ici l'étude de la propagation et de la diffusion de l'information dans la blogosphère faite par \citet{Gruhl04}, l'étude du \emph{news cycle} faite par \citet{Leskovec09} qui fait intervenir la plupart de ces aspects, ou encore l'analyse des influences réciproques entre un réseau social et le réseau sémantique formé par les contenus diffusés sur ce premier réseau, faite par \citet{Cointet09} ; %on l'a déjà mentionné,
- %, ces thèmes étant explorés principalement par le premier groupe de disciplines.
+Without necessarily relying on specific social science theories, this research stream is of interest for its use of large social media corpora in studying cultural dynamics.
 
 On the other hand, the study of the \emph{transformation} of public representations has emerged only recently.
-For one, the notion of ``cultural attractor'' has been the focus of models rendering the evolution of some representations and behaviors not until a few years ago \citep{Claidiere07}.
-Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist of a series of papers investigating \emph{quotation} transformations from a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
-They generally show several types of regularities and propose models of diffusion-transformation of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.
-One of the main conclusions of these works lies in the fact that even for the supposedly most stable type of public representation, namely quotations, it is indeed still possible to observe and measure significant transformations.
-However, these studies address transformations by focusing on the properties of the author (e.g.
-news vs.
-blogs) of the quotation, or the surrounding public space (e.g.
-quote frequency in the corpus), rather than the very cognitive features which may trigger or, at least, bias these transformations.
-% Cependant ces auteurs n'étudient les transformations observées qu'en fonction du type d'auteur qui reproduit les citations, et pas en fonction des citations elles-mêmes (en particulier les questions d'ordre cognitif concernant la raison pour laquelle ces transformations ont lieu, ou la raison pour laquelle elles ont lieu dans tel sens et pas dans tel autre, ne sont pas abordées).
-%Les possibilités offertes par cette étude étant vastes, on a décidé de travailler sur les mêmes données que \citeauthor{Simmons11} pour traiter notre question.
+For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07}.
+Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
+They show several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.
+One of the main conclusions of these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to observe and measure significant transformations.
+However, these studies address transformations by focusing on the properties of the source of the quotation (e.g.
+news outlet v.
+blog), or the surrounding public space (e.g.
+quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
 
-At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations of public representation evolution by studying the influence of word features on the ease of recall.
-This state-of-the-art is relatively developed and shows the positive impact of classical psycholinguistic variables, such as word frequency (see \citet{Yonelinas02} for a review), age-of-acquisition \citep{Zevin02}, number of phonemes and number of syllables (see for instance \citet{Rey98} and \citet{nick-diss}).
+At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations for public representation evolution by studying the influence of word features on the ease of recall.
+This field is well developed and details the impact that classical psycholinguistic variables such as word frequency (see \citet{Yonelinas02} for a review), age-of-acquisition \citep{Zevin02}, number of phonemes or number of syllables (see for instance \citet{Rey98} and \citet{nick-diss}), have in this type of task.
 
-The study of network-based properties, while heavily discussed around the notion of connectionism and its normative processual models, appeared much more recently as an empirical investigation field \CN.
-Let us mention two interesting and recent studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors in the cognitive processes and reproduction of those words.
-%Les deux autres travaux viennent de la psycholinguistique, et posent les fondations pour l'étude des propriétés des représentations (qui seront ici des citations).
-First, \citet{Griffiths07} analyze a task where patients are asked to speak out the first word which comes to their mind when they are presented with a random alphabet letter.
+Less classical linguistic variables, based on the study of semantic network properties, have recently appeared as an empirical investigation field, after having been heavily discussed around the notion of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
+Let us mention two interesting and recent studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
+First, \citet{Griffiths07} analyse a task where patients are asked to name the first word which comes to their mind when they are presented with a random letter from the Latin alphabet.
 The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (PageRank) in a language-wide semantic network built from external word association data.
-% property concernant une tâche où l'on présente aux sujets une lettre de l'alphabet en leur demandant de dire le premier mot commençant par cette lettre qui leur vient à l'esprit.
-Les auteurs montrent qu'une certaine propriété des mots (le PageRank calculé sur un réseau sémantique construit à partir de données d'associations de mots) est un bon facteur prédictif de quel mot sera rappelé parmi les mots commençant par la lettre demandée ; ils montrent ainsi un lien entre une caractéristique sémantique quantitative des mots et la facilité de rappel de ces mots dans cette tâche.
 A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words come quicker when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
-%, étudie une tâche où l'on demande aux sujets de nommer oralement une image.
-Les auteurs montrent qu'une autre propriété des mots, phonologique cette fois (le coefficient de regroupement des mots calculé sur le réseau phonologique formé par ces mots), a un effet sur la rapidité à nommer l'image présentée : c'est un autre cas de lien entre une caractéristique quantitative (phonologique) d'un mot et la facilité à produire ce mot.
 
-On the whole, the current psycholinguistic state-of-the-art seems to hint at two antagonistic types of results.
+\rk{add \url{https://www.sciencedirect.com/science/article/pii/S0893608012000330}, \url{http://research.clps.brown.edu/austerweil/pdfs/papers/randomWalkNips2012.pdf}, \url{http://link.springer.com/article/10.3758/s13421-013-0312-y} to that picture}
+
+\TB{On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
 On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal \CN.
-On the other hand, when the task consists in remembering a specific list of items, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernable \CN.
-The jury is still out as to whether reformulation alteration, i.e.
-spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
-Our paper additionally sheds light on this debate.
+On the other hand, when the task consists in remembering a specific list of items, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible \CN.
+The jury is still out as to whether reformulation alteration, i.e. spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
+Our paper additionally sheds light on this debate.}
+\rk{But do we really care about this?}
 
 \smallskip
-More broadly, we thus observe a gap between, on one side, macro-level empirical studies on diffusion dynamics in a social system and, on the other side, studies essentially focused on micro-level transformations of representations --- these latter studies being either strongly normative, or yielding information whose integration into realistic cultural epidemiology models remains difficult.
-%On remarque de plus que le volet cognitif, central pour les phénomènes étudiés, est relativement peu développé dans les modèles et peu exploré de façon empirique.
-%Les deux dernières études, issues de la psycholinguistique et dans un cadre strictement \emph{in vitro} et strictly at the vocabulary level, montrent donc que des propriétés des mots du vocabulaire, calculées sur un réseau formé par ces mêmes mots, sont des facteurs importants dans la façon dont on les traite cognitivement et dans la façon dont on les produit.
-C'est justement l'élément qui nous manquait : en se basant sur ces trois travaux, on va pouvoir utiliser des propriétés ainsi calculées sur les mots des citations pour mesurer quantitativement les transformations subies par ces citations lorsqu'elles sont reproduites.
-Pour rendre l'analyse possible, on se concentrera sur des cas de \emph{substitution} d'un mot par un autre lorsqu'un auteur reproduit une citation : à chaque substitution observée, on assiste à un échange entre deux mots pour lesquels on va pouvoir comparer les propriétés sémantiques.
-
-%%% GARBAGE
-
-%Cependant, depuis plus d'une décennie le paysage des choses observables dans le domaine des représentations publiques est en train de changer radicalement, du fait du déluge de données provenant des interactions en lignes, permettant l'étude empirique des représentations dans de cadre de l'épidémiologie culturelle.
-
-%Nous vivons aujourd'hui non pas avec un manque de données concernant les représentations publiques, mais au contraire sous un véritable déluge de telles données.
-Quel est ce déluge ?
-Il est le produit d'une collecte qui est née avec l'utilisation commerciale et massive d'Internet.
-Cette collecte n'est pas un phénomène mineur, ni même un phénomène d'ampleur moyenne : les services sur Internet enregistrent aujourd'hui la quasi-totalité de ce qui est mesurable des interactions entre les personnes connectées ainsi que des interactions des personnes avec les machines.
-Vu l'importance que ce réseau a pris dans nos vies de tous les jours, on assiste bien à une explosion de ces données ; à tel point que certains observateurs d'Internet estiment que 90\% des données stockées aujourd'hui par les services sur Internet ont étés générées il y a moins de deux ans.
-%Sans être des enregistrements des interactions \emph{physiques} entre les personnes (``real life''), il s'agit bien d'observations sur le terrain nouveau que constituent l'action, la perception, et l'interaction connectées, au travers des machines et d'Internet.
-En un mot, elles sont en fait des observations en grandes quantités de \emph{représentations publiques}.
-%\footnote{Bien sûr, une bonne partie des données collectées sur Internet appartient à des entreprises qui n'ont aucun intérêt à les partager ni à les exploiter autrement que financièrement.
-Mais les données publiquement accessibles forment tout de même une base considérable, qu'il est tout à fait possible d'exploiter comme telle.}.
-%L'approche qu'on a esquissée ici n'est qu'une des façons d'aborder l'étude de la culture et de la cognition de façon empirique.
-En effet, les courants de recherche qui gravitent autour de ce thème sont nombreux et variés ; voici une rapide présentation des champs concernés et de la façon dont ils interagissent.
-
-%Un groupe de disciplines est particulièrement pertinent ici, à l'interface entre le \emph{data mining}, l'étude des systèmes complexes, et la sociologie quantitative (e.g.
-sociologie des réseaux).
-Un grand nombre de travaux apparaissent exploitant de grands corpus de données produits par des collectes sur Internet, allant de l'étude de l'impact des réseaux sociaux sur la diffusion des informations sur ces mêmes réseaux \citep{Bakshy12} à l'analyse du rôle de la presse en ligne dans la vie politique locale \citep{Parasie12}, en passant par le détail des interactions entre éditeurs sur Wikipedia \citep{Jurgens12} ou l'évolution du vocabulaire des langues au cours du temps \citep{Lieberman07}.
-Une partie de ces travaux n'inscrivent leurs résultats que dans des cadres théoriques naïfs (autour de thèmes pour lesquels il existe pourtant des théories réfléchies développées par des champs des sciences sociales), et relèvent parfois plus des sciences de l'informatique que de l'étude de la culture (d'aucuns parlent d'ailleurs de \emph{social computing}).
-Ce groupe de disciplines présente néanmoins la particularité de vouloir étudier la culture en exploitant les possibilités ouvertes par les larges corpus de données issus d'Internet.
-
-%% ----------------------------------------
-%\subsection{Relevant fields} % -----
-%% ----------------------------------------
-%
-%L'idée d'étudier la culture en lien avec la cognition apparaît comme une extension naturelle pour plusieurs disciplines historiquement séparées qui, chacune de leur côté, s'en sont approchées en partant de leur propre espace de concepts et en étendant leurs outils.
-Les dialogues entre ces disciplines ne sont pas rares, mais les traditions et les épistémologies différentes ont rendu les unifications plus ardues.
-On peut distinguer au moins deux grands groupes de disciplines débattant autour de thématiques reliées à ce sujet.%\footnote{On regroupe ici les disciplines par les thèmes qu'elles débattent, plutôt que par les similarités entre leurs façons de traiter ces thèmes.}.
-%
-%Le premier groupe est issu des débats entre la biologie évolutionniste, la psychologie cognitive et les sciences sociales (et plus particulièrement l'anthropologie), desquels ont émergé les domaines de la psychologie évolutionniste et de l'anthropologie cognitive.
-%Plusieurs théories ont été avancées et sont débattues en parallèle par ces disciplines.
-On se doit de mentionner ici en premier lieu le débat autour du programme mémétique initié par \citet{Dawkins76} et pour lequel l'ouvrage collectif de \citet{Aunger00} est un bon exemple.
-Un deuxième terrain exploré par cet ensemble de disciplines est le développement de modèles évolutionnistes basés sur les normes (voir par exemple \citealp{Ehrlich05}), avec comme travail fondamental celui de \citet{Boyd85}.
-Le troisième champ, particulièrement important aujourd'hui, est précisément celui de l'épidémiologie culturelle, initié par \citet{Sperber96} et défendu face aux autres courants dans des travaux comme celui de \citet{Atran03}.
-%
-%Le deuxième groupe, plus diffus, provient des échanges entre le \emph{data mining}, l'étude des systèmes complexes, et la sociologie quantitative (e.g.
-sociologie des réseaux).
-Un grand nombre de travaux apparaissent exploitant de grands corpus de données produits par des collectes sur Internet, allant de l'étude de l'impact des réseaux sociaux sur la diffusion des informations sur ces mêmes réseaux \citep{Bakshy12} à l'analyse du rôle de la presse en ligne dans la vie politique locale \citep{Parasie12}, en passant par le détail des interactions entre éditeurs sur Wikipedia \citep{Jurgens12} ou l'évolution du vocabulaire des langues au cours du temps \citep{Lieberman07}.
-Une partie de ces travaux n'inscrivent leurs résultats que dans des cadres théoriques naïfs (autour de thèmes pour lesquels il existe pourtant des théories réfléchies développées par des champs des sciences sociales), et relèvent parfois plus des sciences de l'informatique que de l'étude de la culture (d'aucuns parlent d'ailleurs de \emph{social computing}).
-Ce groupe de disciplines présente néanmoins la particularité de vouloir étudier la culture en exploitant les possibilités ouvertes par les larges corpus de données issus d'Internet.
+Stepping back, we observe a gap between, on one side, macro-level empirical studies of the diffusion dynamics in a social system and, on the other side, studies focused on micro-level transformations of representations --- these latter studies being either strongly normative, or whose results are difficult to integrate with realistic cultural epidemiology models.

### 2014-01-27 brains-introduction.tex
Change master document meta-data. brains-introduction is now completely clean.
Minor phrasing

@@
-% !TEX root = /Users/camille/Shared/Dropbox/BrainCopyPaste/brainscopypaste.tex
+% !TEX root = brainscopypaste.tex
 % ============================
@@
 \smallskip
-Stepping back, we observe a gap between, on one side, macro-level empirical studies of the diffusion dynamics in a social system and, on the other side, studies focused on micro-level transformations of representations --- these latter studies being either strongly normative, or whose results are difficult to integrate with realistic cultural epidemiology models.
+Stepping back, we observe a gap between, on one side, macro-level empirical studies of the diffusion dynamics in a social system and, on the other side, studies focused on micro-level transformations of representations --- these latter studies being either strongly normative, or whose results are difficult to articulate with realistic cultural epidemiology models.

### 2014-01-27 brains-protocol.tex
Rework foreword

@@
 % !TEX root = brainscopypaste.tex
-\section{Protocol}\label{sec:protocol}
+% ============================
+\section{Protocol} % =========
+% ============================
+\label{sec:protocol}
 
-%\tb{On a choisi de prendre une forme de représentation publique en apparence basique, mais qui entre bien dans la définition que \citeauthor{sper:expl} fait des représentations publiques~\citep[p.~49]{sper:expl} : les citations.
-C'est une forme de représentation d'autant plus instructive à analyser que, si l'on s'en tient aux règles de l'écrit, les citations ne devraient pas changer : lorsqu'on cite quelqu'un il est attendu qu'on reproduise précisément les mots utilisés par l'auteur, éventuellement en n'en prenant qu'une partie pour mieux l'intégrer à son propos, mais certainement pas en en changeant les termes.
-On peut faire l'hypothèse que les auteurs qui citent \emph{cherchent à le faire} en suivant cette règle.
-De ce fait, si l'on observe en effet une évolution des citations au-delà du "rognage" elle doit être l'effet d'un biais cognitif involontaire (de l'ordre de l'automatique).
-L'analyse de l'évolution des citations sur Internet est donc un terrain idéal pour observer un biais cognitif à l'origine de transformations des représentations publiques.}
-%
-%\tb{Revenons brièvement sur l'hypothèse faite ci-dessus car elle mérite un peu plus d'attention pour être valide.
-Elle est justifiée pour la simple raison qu'il est souvent facile d'aller vérifier les propos d'origine, et parce que cette transparence fait partie de la raison d'être et de la légitimité mêmes des citations.
-En effet, les auteurs qui en citent d'autres peuvent tout à fait intégrer une citation en en déformant l'intention d'origine, par exemple en la rognant à souhait ou en extrayant la citation de son contexte, ou encore en la juxtaposant à d'autres citations qui renforcent le propos de l'auteur citant, mais il est rarement à l'avantage de l'auteur de changer les mots utilisés.
-Ceci pour deux raisons : d'une part, parce que la légitimité et la force de l'information transmise ou de l'argumentaire déployé reposent en partie sur l'honnêteté intellectuelle qu'on peut prêter à l'auteur, d'autre part parce que l'exactitude des citations est justement le premier des indicateurs à utiliser lorsqu'on cherche à mesurer cette honnêteté intellectuelle.}
-%
-%\tb{Enfin, et pour terminer de justifier cette hypothèse, pour que ce garde-fou fonctionne il n'est pas nécessaire que les citations faites par un auteur soient vérifiées systématiquement, ni même qu'elles soient vérifiées souvent : il suffit que toute l'information nécessaire à la vérification soit publique, pour que le risque encouru par la déformation d'une citation soit bien réel.
-C'est bien souvent le cas sur Internet\footnote{De la même manière qu'exiger que les comptes d'une entreprise soient publics garantit un certain niveau d'"honnêteté comptable" (sans pour autant que des vérifications soient faites systématiquement), le fait que les données nécessaires à la vérification d'une citation soient publiques est un bon garant du fait que les auteurs qui en citent d'autres cherchent à rester fidèles aux mots d'origine.}.
-Il n'est pas non plus nécessaire que la vérification soit facile à faire pour que cette garantie tienne, mais il suffit que le risque de chute de légitimité soit assez grand pour que peu de monde s'y risque sérieusement\footnote{S'il n'est pas toujours facile de vérifier qu'une citation est exacte, le jour où quelqu'un montre qu'elle est inexacte il sera bien compliqué pour l'auteur citant de se défendre.
-Ce raisonnement s'applique moins à la déformation du contexte de la citation, cas dans lequel il est plus facile de se défendre en arguant de la divergence d'interprétation.
-La \emph{citation inexacte}, en revanche, relève plus de la \emph{donnée fausse} que de l'\emph{interprétation divergente}.}.}
+In order to start bridging this gap, we set out to \emph{empirically} study public representation transformations at the microscopic level, aiming to stay compatible with macroscopic-level studies of these public representations.
+Quotations appeared to be a perfect candidate as public representations.
+First, they are usually cleanly delimited by quotation marks (and often with HTML markup in web pages), which greatly facilitates their detection in text corpora.
+Second, they stem from a unique ``original'' version, and could ideally be traceable back to that version.
+Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quote.
+Quotation evolution is therefore a perfect environment to measure cognition-induced transformations and relate those findings to macroscopic social dynamics.
 
-To measure cognitive bias during reformulation, we chose to focus on public representations which could be expected to undergo variations which could be described easily and automatically.
-Quotations appeared to be a perfect candidate.
-First, they are generally cleanly delimited by quotation marks (or by HTML markup in web pages), which greatly facilitates their detection in text corpora.
-Second, they stem from a unique ``original'' version, and could ideally be traceable back to it.
-Third, and perhaps most importantly, their duplication should \emph{a priori} be highly faithful, apart from cropping issues: not only should variations be of a moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quote.
-
-\subsection{Dataset} % .....
+\subsection{Dataset}
 

### 2014-01-27 brains-protocol.tex
Rephrasing in 3.1 dataset

@@
 
-We thus relied on a large and reliable quotation dataset collected by \citet{Leskovec09} which 
-%matched these essential criteria: made of public representations, subject to variations which could be expected to be of a relatively moderate magnitude, be ideally described by reference to an ideal original representation, and
-is of a sufficiently large size to lend itself to statistical analysis.
-%\tb{On a donc choisi de s'appuyer sur un corpus déjà disponible au lieu de construire le nôtre, ce qui est une des raisons pour lesquelles l'analyse de la question est si tributaire du choix des données.
-Le jeu de données utilisé par  et \citet{Simmons11} se prête bien à la notion de représentation publique qu'on a choisie ci-dessus.
-Construit et publié en 2009 par \citeauthor{Leskovec09}, c'est un jeu de données reconnu et utilisé dans la littérature, largement pré-traité pour nos besoins comme on le verra ci-dessous.
-Il est d'une dimension bien suffisante pour atteindre des niveaux de significativité intéressants, tout en restant manipulable sur un serveur de calcul modeste ou même sur un bon ordinateur portable.}
-This dataset consists of the daily collection of news stories or blog posts from around a million online sources, with the approximate publication rate of 900k texts per day, over a nine-month period ranging from August 2008 to April 2009 \cite{Leskovec09-url}.\footnote{The original article~\citep{Leskovec09} does not provide additional details on the method of source selection.}
-Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the cited person.
+We relied on a reliable quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
+This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009) \cite{Leskovec09-url}.\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
+Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person.
 Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
-A community detection algorithm is applied to that quotation graph to detect aggregates of tightly connected, i.e.
-sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
-The final data contains around $70,000$ sets of quotations.
-Each set allegedly contains all variations of a same parent utterance, annotated with the respective URLs and timestamps of publication of each individual variation.
-% We have the list of occurrences, URL and publication time of the quotations of a given such set.
-%  which are possibly slightly differing from a parent utterance.
-
-%\tb{Un des travaux des auteurs a été de reconstituer ces ensembles de citations par une méthode de catégorisation (i.e.
-construction de \emph{clusters})\footnote{Se référer à \citet{Leskovec09} pour plus de détails.
-Ce traitement des données comporte aussi quelques étapes de filtrage, décrites dans l'article.}, et c'est à ce stade de traitement qu'on obtient les données : elles sont constituées d'environs 70~000 ensembles de citations, à l'intérieur desquels toutes les citations proviennent d'un même énoncé parent (elles en sont des sous-parties transformées).
-Pour chaque citation à l'intérieur de chaque ensemble, on a une liste de triplets \emph{(URL, date d'apparition, nombre d'occur\-rences à cette URL)} décrivant les adresses et les dates auxquelles cette citation est apparue.}
+A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
+This analysis yielded the final data, with a total of about $70,000$ sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, annotated with the respective publication URLs and timestamps of each individual variation.
 

### 2014-01-27 brains-protocol.tex
Small rephrasing

@@
 A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
-This analysis yielded the final data, with a total of about $70,000$ sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, annotated with the respective publication URLs and timestamps of each individual variation.
+This analysis yielded the final data we had access to, with a total of about $70,000$ sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
 

### 2014-01-27 brains-protocol.tex
Flesh out classical measures, add an open question to be solved

@@
 
+To keep the analysis palatable, we restricted the analysis to quotation transformations which consisted in the \emph{substitution} of a word by another word (and only those cases).
+To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
+The following sections detail the features we used.
+
 \subsubsection{Standard psycholinguistic indices}
 
-\RK{introduce the classical psycholinguistic measures -- kind of like ``demographic'' usual suspects}
+We first introduced some of the most classical psycholinguistic measures on words:
 
-We also considered grammatical types within quotations by detection of \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology \citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and
-adverbs.
-The results were however extremely similar across the various categories, exhibiting no specific behavior of words belonging to different POS categories.
-%.
-\tb{Enfin, on ajoute à ces mesures la catégorie grammaticale, dans les citations, des mots échangés lors des substitutions.
-On s'est restreint à quatre des catégories les plus fréquentes, en utilisant la catégorisation et la notation des \emph{Part-of-Speech} (abrégé POS)\footnote{Le terme \emph{Part-of-Speech}, ou "partie du discours" en français, désigne les catégories grammaticales des mots (ou même des parties de mots) utilisées en linguistique.} du Penn TreeBank Project\footnote{Le Penn TreeBank Project définit une typologie très complète des POS, ainsi qu'une façon de les noter : à chaque POS correspond un code de deux ou trois lettres majuscules.
-Par exemple, tous les POS tenant du verbe commencent par la lettre $V$ (e.g.
-$VBN$ pour participe passé, $VBG$ pour participe présent, etc.).}~\citep{Santorini90} : verbes (POS dont le code commence par $V$), noms (POS dont le code commence par $N$), adjectifs (POS dont le code commence par $J$), et adverbes (POS dont le code commence par $R$).
-Cette mesure supplémentaire affine les observables définies ci-dessus, et permet d'observer un éventuel effet des catégories grammaticales sur les résultats.}
+\begin{itemize}
+    \item \textbf{Age of Acquisition}: the average age at which words are learned, obtained from~\citet{kuperman12},
+    \item The average \textbf{Number of Phonemes} for all pronunciations of a word, obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98},\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.}
+    \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary.
+    \item \rk{Why exactly don't we include word frequencies as a measure?}
+\end{itemize}
+
+We also considered grammatical types within quotations by detection of \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology~\citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and adverbs.
+The results were however extremely similar across the various categories, exhibiting no specific effect of words belonging to different POS categories.
 
 \subsubsection{Network-based definitions}
+
 We rely on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citet{Nelson04}.

### 2014-01-27 brains-protocol.tex
Rework WN description

@@
 A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
-This analysis yielded the final data we had access to, with a total of about $70,000$ sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
+This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
 
@@
 
-\subsubsection{Network-based definitions}
+\subsubsection{Network-based indices}
 
-We rely on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citet{Nelson04}.
+Additional to classical psycholinguistic measures, we also considered more recently studied variables based on semantic network properties.
+We relied on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citet{Nelson04}.
 
-The WordNet (WN) database gathers more than 117~000 concepts and about 147~000 words attached to these concepts.
-WordNet entries are disambiguated words, or ``\emph{lemmas}'', which are attached to concepts called ``synsets'', each of them gathering the set of lemmas corresponding to the same single meaning.
-Note that the choice of a synset for a lemma immediately determines the grammatical category of that lemma.
-Lemmas make the nodes of the semantic network (homonyms are aggregated into a single node), whose neighbors are synonymous lemmas belonging to the same synset(s) as the lemma(s) represented by that node.
-In other words, synsets induce cliques in the semantic network.
-Links connecting two synonymous lemmas are being assigned a weight corresponding to the number of times these lemmas are mentioned by WordNet as synonyms, i.e.
-in the same synset.
+The WordNet (WN) database gathers more than \num{117000} concepts and about \num{147000} words attached to these concepts.
+WordNet entries are disambiguated words, called ``lemmas'', which are grouped into concepts called ``synsets'': all the lemmas in a synset are synonyms for the same single meaning.
+Note that synsets also encode grammatical category, so the choice of a synset for a word (thus creating a lemma) immediately determines the grammatical category of that word.
+Non-disambiguated words make the nodes of the semantic network we build (i.e.
+homograph lemmas are aggregated into a single node), and a node's neighbours are synonymous lemmas belonging to the same synset(s) as the lemma(s) represented by that node.
+In other words, synsets induce cliques in the semantic network, and a word with several meanings belongs to as many cliques as it has meanings.
+\TB{Links connecting two synonymous lemmas are assigned a weight corresponding to the number of times these lemmas are mentioned by WordNet as synonyms, i.e. in the same synset.}
+\rk{Didn't we decide to drop that last point, the weighing in the graph?}
 
@@
 Both semantic networks describe synonymy relationships corresponding to qualitatively distinct forms of association: either from a linguistic viewpoint (WN) or from the perspective of actual information retrieval by humans (FA).
-In  order to work in a unified framework, we consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighted networks.}
+In  order to work in a unified framework, we consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighed networks.} \rk{Check that's the case in the analysis}
 

### 2014-01-27 brains-protocol.tex
Add FA description

@@
 Additional to classical psycholinguistic measures, we also considered more recently studied variables based on semantic network properties.
-We relied on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' relationships collected by~\citet{Nelson04}.
+We relied on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' norms collected by~\citet{Nelson04}.
 
@@
 
-\rk{Explain Free Association}
+Free Association (FA) norms record the words that come to mind when people are presented with a given cue (that is the ``free association'' task). As \citeauthor{Nelson04} explain,
+\begin{quote}
+free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.~\citep{Nelson04}
+\end{quote}
+Following \citet{Griffiths07}, we consider the directed weighed network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
 
-\TB{The original WN network is a weighted, undirected network; the FA network is a directed network where weights have a different meaning.
+While the original WN network is a weighted, undirected network, the FA network is a directed network where weights have a different meaning.
 Both semantic networks describe synonymy relationships corresponding to qualitatively distinct forms of association: either from a linguistic viewpoint (WN) or from the perspective of actual information retrieval by humans (FA).
-In  order to work in a unified framework, we consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighed networks.} \rk{Check that's the case in the analysis}
+In  order to work in a unified framework, we shall consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighed networks.
+\rk{Check that's the case in the analysis}
 

### 2014-01-27 brains-protocol.tex
Remove trailing whitespace up to where we are
Reread teh network measures a little

@@
 \TB{Links connecting two synonymous lemmas are assigned a weight corresponding to the number of times these lemmas are mentioned by WordNet as synonyms, i.e. in the same synset.}
-\rk{Didn't we decide to drop that last point, the weighing in the graph?}
+\rk{We might spare ourselves that last sentence since we drop that aspect further down.}
 
-Free Association (FA) norms record the words that come to mind when people are presented with a given cue (that is the ``free association'' task). As \citeauthor{Nelson04} explain,
+Free Association (FA) norms record the words that come to mind when people are presented with a given cue (that is the ``free association'' task).
+As \citeauthor{Nelson04} explain,
 \begin{quote}
@@
 \bigskip
-We introduce three standard network-based features, which may also be relevant psycholinguistically:
+We introduce three standard network-based features to be used on both the WN and FA networks:
+
 \begin{itemize}
     \item \textbf{Distance}, a classical dyadic network measure corresponding to the shortest path length between two nodes.
-This measure may be used as a proxy of the remoteness between two given lemmas, which is likely to have an influence on the likelihood of being a relevant substitution candidate.
-    \item \textbf{Centrality} $k$, measured by the number of neighbors of a given node, and interpretable as a proxy for the polysemy of words.
-For the directed network FA we also define the incoming centrality $k'$ (number of neighbors pointing \emph{to} that node).
-Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank} \cite{Page99}, which had already been used in \cite{Griffiths07} and may be interpreted as a generalized and recursive measure of lemma polysemy --- central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on, recursively.\footnote{\TB{{Betweenness coefficient}, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network \cite{free:set}.
-More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher, the more important is that node in ensuring the connectedness of the rest of the network.
-Correlation between FA-BC and FA-PR is $0.74$; between WN-BC and WN-PR it is $0.83$.}}
-    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs.
-possible links between a node's neighbors \cite{watt-coll}.
-A lemma is likely to have a higher clustering coefficient whenever the synsets to which it belongs are themselves semantically linked.
+    This measure may be used as a proxy of the remoteness between two given words, which is likely to influence the likelihood of a word being a relevant substitution candidate for another word.
+    \item \textbf{Centrality} $k$, measured by the number of neighbours of a given node, which can be interpreted as a proxy for the polysemy of words.
+    \TB{For the directed network FA we also define the incoming centrality $k'$ (number of neighbours pointing \emph{to} that node)}.
+    \rk{Remove that, since we restrict ourselves to undirected networks}
+    Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank}~\citep{Page99}, which had already been used in~\citet{Griffiths07} and may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on recursively.
+    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual v.
+possible edges between a node's neighbours \cite{watt-coll}.
+In the WN network, a word is likely to have a higher clustering coefficient whenever the synsets to which it belongs are themselves semantically linked.
     This measure had been used by \citet{Chan10}.
 \end{itemize}
+
+\TB{Betweenness coefficient, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network~\citep{free:set}.
+More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher the coefficient, the more important that node is in ensuring the connectedness of the rest of the network.
+Correlation between FA-BC and FA-PR is $0.74$; between WN-BC and WN-PR it is $0.83$.}
+\rk{We dropped this, right?}
 

### 2014-01-27 brains-protocol.tex
3.2 is reworked, but needs fact-checking and talk with Camille

@@
 
-We first introduced some of the most classical psycholinguistic measures on words:
+We first introduce some of the most classical psycholinguistic measures on words:
 
@@
     \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary.
-    \item \rk{Why exactly don't we include word frequencies as a measure?}
+    \item \rk{Why exactly don't we include word frequencies as a measure?
+The alleged self-correlation loop doesn't determine at which end of a substitution a given high-frequency word would be, so it doesn't really force correlation}
 \end{itemize}
@@
 \bigskip
-We introduce three standard network-based features to be used on both the WN and FA networks:
+We introduce three standard network-based measures to be used on both the WN and FA networks:
 
 \begin{itemize}
-    \item \textbf{Distance}, a classical dyadic network measure corresponding to the shortest path length between two nodes.
+    \item \textbf{Distance} between two words, a classical dyadic network measure corresponding to the shortest path length between two nodes.
     This measure may be used as a proxy of the remoteness between two given words, which is likely to influence the likelihood of a word being a relevant substitution candidate for another word.
@@
 
-\TB{It would be difficult to discuss possible correlations between distance ($d$), on one side, and centrality ($k$ or $k'$) or clustering coefficient ($c$) on the other side, as the former is a dyadic measure while the latter are monadic measures.
-We may however note that centrality and clustering are relatively weakly (and negatively) correlated, with a Pearson coefficient of $-0.26$.} \rk{remove, redundant with below}
+An important question here relates to the various possible correlations between all the variables we consider.
 
-An important question here relates to the various possible correlations between all the variables we consider.
-The age of acquisition is a key variable, primarily as a usual suspect in psycholinguistic studies, and also appears to be generally correlated to all the other variables.
-This coincides with an ongoing debate suggesting that age of acquisition encodes a variety of phenomena which are difficult to disentangle from more specific phenomena, which could be captured by more independent variables such as \CN.
-\TB{Indeed, the minimal linear correlation of $aoa$ ($-0.31$) with other variables has a higher magnitude than the maximal linear correlation of other variables with each other ($-0.26$).} \rk{how do we get to that conclusion?}
+The age of acquisition is a key variable, primarily as a usual suspect in psycholinguistic studies, and also appears to be generally correlated to many of the other variables.
+This relates to an ongoing debate suggesting that age of acquisition encodes a variety of phenomena which are difficult to disentangle from more specific phenomena, which could be captured by more independent variables such as \CN.
+\rk{What do we conclude of that? Nuance results with AoA?}
+\TB{Indeed, the minimal linear correlation of $aoa$ ($-0.31$) with other variables has a higher magnitude than the maximal linear correlation of other variables with each other ($-0.26$).}
+\rk{how do we get to that conclusion?}
 
-Additionally, phonemes and syllables exhibit a strong linear correlation ($0.85$).
-We see a better prediction effect of phonemes over syllables, which is consistent with \cite{nick-diss}, and choose to focus on phonemes only.
-\rk{check that!}
+Additionally, number of phonemes and number of syllables exhibit a strong linear correlation ($0.85$).
+The analysis showed a better prediction effect of number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to focus the presented results on the former only.
+\rk{Check that!}
 
@@
 Network properties, finally, appear to be relatively weakly correlated with each other, across one network or between both networks (i.e.
-nodes with high index values are generally distinct between FA and WN), and may be used .
-We rely on FA as there is a direct possible psychological interpretation (that human copy-pasting may or may not show the same type of trend as human free association), noting that prediction results will be matched against WN as well.
+nodes with high index values are generally distinct between FA and WN).
+We chose to focus the results on FA as this network presents a direct psychological interpretation (that human copy-pasting may or may not show the same type of trend as human free association), noting that prediction results are matched against WN as well.
 \rk{comment peut-on dire que les nœuds importants dans FA ne sont pas nécessairement importants dans WN mais que, pourtant, la forme de la prédiction est la même?
 cela me semble illogique en première lecture}
+\rk{Check those assertions}
 

### 2014-01-27 brains-protocol.tex
Code style
3.3 is reread and good.

@@
 Second, they stem from a unique ``original'' version, and could ideally be traceable back to that version.
-Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quote.
+Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.
 Quotation evolution is therefore a perfect environment to measure cognition-induced transformations and relate those findings to macroscopic social dynamics.
@@
 
-Nonetheless, these sets bear no explicit mention of which is the authoritative, original quotation.  Defining substitutions from the dataset therefore requires additional assumptions.  
-%\tb{Il manque en effet certaines informations dans le corpus pour pouvoir lire ces substitutions de façon immédiate dans les données : les métadonnées pour chaque citation ne concernant que les URLs, les dates d'apparitions et les nombres d'occurrences à chaque URL, on n'a aucune information sur les liens entre les citations.
-Or, on a besoin de savoir sur quelles sources un auteur s'est appuyé pour reproduire une citation, c'est-à-dire qu'on a besoin de savoir quelle citation a été générée à partir de quelle autre.
-Cette information n'est pas présente dans les données et on va donc devoir l'inférer, à l'aide de plusieurs hypothèses.}
+The data we use presents an additional challenge: each set of quotations bears no explicit information either about the authoritative original quotation, or about the source quotation(s) each author inspired himself from when creating a new post and reproducing (possibly altering) those sources.
+Substitutions are therefore not explicitly encoded in the dataset.
 
-Indeed, an author writing a new post and reproducing (possibly altering) a quotation bases himself on original quotations we have no access to.
-We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quote for each instance of each quote.
-We therefore model the underlying quote selection process by making a few additional assumptions which let us define substitutions from the available data.
+We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
+We therefore model the underlying quotation selection process by making a few additional assumptions which let us define substitutions from the available data.
 
-The main question in this problem concerns the way we consider later occurrences of a quote which, when it first appeared, was identified to be an alteration from an original quote.
-Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on Jan 18th, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.
+The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
+Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.
 If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the author of $q_b$ on the 19th based himself on $q_a$.
@@
 
-To settle this question we model the process as follows: we assume that when a quote $q$ appears at time $t$, if not original (i.e.
+To settle this question we model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e.
 if not stemming from a source external to the dataset, e.g.
-initiating a new set of quotations), it is based solely on the most frequent quote $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$.
-The length $\Delta t$ of that period of time is fixed as a fraction of the total duration of the considered set of quotations; we took one fifth in the implementation (i.e.
-one month for a set of quotations spanning five months, or one day for a set spanning only five days; this takes the dynamism of the set of quotes into account).
+initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$.
+The length $\Delta t$ of that period of time is fixed as a fraction of the total duration of the set of quotations; we took one fifth in the implementation (i.e.
+one month for a set of quotations spanning five months, or one day for a set spanning only five days; this takes the dynamism of the set of quotations into account).
 If $q$ differs from $q_{max}$ by only a word, it is counted as a substitution from $q_{max}$ to $q$.
@@
 
-The assumptions embedded in this model are only a subset of a wider set of possibilities, all leading to alternative models.
-We identified and implemented five other such possible models, which all yielded essentially the same results.
+The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
+We identified and implemented five other such models, and they all yielded essentially the same results.
 These models differ in the definition of the source quote from which new occurrences stem, essentially modifying the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
 For example, the time-windows considered can have different lengths, can include all occurrences from the beginning of the set of quotations, or can have fixed positions.
-The source quote of a potential substitution can be chosen among those time-windows, or otherwise (e.g.
-among \emph{all} quotes having appeared before $t$ and differing by a word from the considered arrival quote; this detection process would include all possible substitutions detected by other models, but would also include many false-positives).
+The source quotation of a potential substitution can be chosen among those time-windows, or otherwise (e.g.
+among \emph{all} quotations having appeared before $t$ and differing by a word from the arrival quotation; this detection process would include all possible substitutions detected by other models, but would also include many false positives).
 

### 2014-01-27 brains-protocol.tex
3.4 is reread and good

@@
 
-\subsubsection{Network-based indices}
+\subsubsection{Network-based measures}
 
@@
     \includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
+    \caption{Variable correlations \rk{rebuild with the variables we talk about and keep}}
 \end{figure}
@@
 
-We then measure the cognitive alteration of human copy-pasting by comparing the relative properties of substituted and substituting words in quotations.
-We only consider altered quotations, \hbox{i.e.} quotations where at least one word has been modified; in other words, we measure the properties of alteration \emph{knowing that} there has been a variation (which indicates a certain reformulation) while we do not take invariant quotations into account, as it is impossible to know, in that case, whether there has been perfect individual reformulation, or just digital copy-pasting of a source (``{\sf Ctrl-C}/{\sf Ctrl-V}'').
+We then measure the alterations introduced when authors reproduce quotations by comparing the relative features of substituted and substituting words.
+Note that since we only consider substitutions, we measure the features of an alteration \emph{knowing that there has been a alteration}, and we do not take invariant quotations into account.
+Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
 
-We build two main observables for each word property.
-First, we measure the variation of word features over a substitution, looking at the difference of the value of a given feature between arrival and start words.
-%This address the question ``how do word features vary over a subs  \tb{la variation de caractéristique sémantique lors du passage du mot remplacé au mot remplaçant dans une substitution.
-Ceci répond à la question "comment fait-on varier la caractéristique sémantique lorsqu'on substitue un mot ?".}
-Second, we measure the susceptibility for words to be the target of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to be the target of a substitution.
-% Second, \tb{la susceptibilité que les mots ont à être remplacés, en fonction de leur caractéristique sémantique.
-Cette mesure répond à la question "quelles caractéristiques sémantiques a-t-on plus tendance à substituer ?"%\footnote{Attention, cette mesure ne suffit pas en elle-même pour parler d'\emph{influence} de la caractéristique sémantique d'un mot sur le fait qu'il soit substitué ou non, parce qu'on n'a pas fait d'analyse des causes impliquées dans le fait qu'un mot soit substitué ; en effet, d'autres facteurs pourraient être plus importants que ces caractéristiques sémantiques dans les causes d'une substitution.
-On ne répond donc \emph{pas} à la question "quelle est l'influence de la caractéristique sémantique d'un mot sur sa propension à être substitué ?".}.
+We build two main observables for each word feature.
+First, we measure the variation of word feature over a substitution, looking at the variation of a given feature between arrival and start words.
+Second, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution.
 
 \subsubsection{Alteration}
-%The above protocol enables us to observe in detail how word features vary upon substitution. 
-For a given feature $\mathfrak{F}$, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$. Let us denote this quantity:
+
+For a given feature $\mathfrak{F}$, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$.
+Let us denote this quantity:
 $$\Delta (\wstart,\warrival) \defeq \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart)$$
 
-Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster, we first average substitutions over each cluster, by considering the average of arrival word features for a given starting word.
-Indeed, substitutions occurring in the same cluster are likely not statistically independent.} This
-quantity can be written:
+Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.
+\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
+Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
+This quantity can be written:
 $$\left< \Delta(\wstart,\warrival) \right>_f = \left< \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
 
-We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from % , let us denote the corresponding quantity This value is to be compared to the variation the word's feature would have undergone had the arrival word been chosen randomly in
-the pool of \TB{WN/FA} words.
-The corresponding average quantity over all start words may be written as:\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
-%$$\Delta_{\mathcal{H}_0} (\wstart,\warrival) \defeq f\left(\warrival_{\mathcal{H}_0}\right) - f\left(\wstart\right)$$
-%and, similarly, we compute the average over all start words 
-$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f
-= \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
-%Averaging over all start words with a given feature value $f$, we get $$\left< \Delta_{\mathcal{H}_0} \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace \wstart | \mathfrak{F}(\wstart) = f \right\rbrace}$$
+We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
+\rk{Check that's true in the code}
+The corresponding average quantity over all start words may be written:
+\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
+$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
 
-We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word} (neighbors), i.e.
-an arrival word chosen among semantically plausible though still random words (in this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$).
-%The values for $d$ were decided upon according to the distribution of distances travelled on the synonym graph upon substitution, normalized by the number of synonyms at that distance.
-%  brought all variation curves shown below nearer to the origin, but did not change the results qualitatively.
-%The variation of feature compared to $\mathcal{H}_0$ is then $\left< \Delta \right>_f - \left< \Delta_{\mathcal{H}_0} \right>_f$.
+We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word} (neighbours in the WN network \rk{We could also do that with neighbours in directed or undirected FA}), i.e.
+an arrival word chosen among semantically plausible though still random words.
+\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.}
 
@@
 This measure focuses on the selection of start word instead of the selection of the arrival word.
-Indeed, the features have an effect not only on the choice of a new word when a substitution takes place, but also at the preceding moment when it is not yet known which word in the quote --~if any~-- will be substituted.
+Indeed, the features have an effect not only on the choice of a new word when a substitution takes place, but also at the preceding moment when it is not yet known which word in the quotation -- if any -- will be substituted.

### 2014-02-14 brains-protocol.tex
Add a todo for classical psycholinguistic features
Some additional answers for fact-checks

@@
 We first introduce some of the most classical psycholinguistic measures on words:
+
+\rk{Add some bibliography about those features' known effects}
 
@@
 \rk{Check that's the case in the analysis}
+\rk{In fact, the graphs are much less interesting for undirected unweighed FA.}
 
@@
 The analysis showed a better prediction effect of number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to focus the presented results on the former only.
-\rk{Check that!}
+\rk{Rather, graphs are much nicer with number of phonemes than number of syllables: clearer tendencies, more data (there's holes with number of syllables). But not a prediction (prediction of what?)}
 

### 2014-02-19 brains-protocol.tex
More info on the fact-checks
Update remarks after today's conversation on current issues

@@
 \begin{itemize}
+    \item \rk{Add word frequency as a feature, and maybe link it to Kolmogorov's 5/3 law in turbulences}
     \item \textbf{Age of Acquisition}: the average age at which words are learned, obtained from~\citet{kuperman12},
@@
     \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary.
-    \item \rk{Why exactly don't we include word frequencies as a measure?
-The alleged self-correlation loop doesn't determine at which end of a substitution a given high-frequency word would be, so it doesn't really force correlation}
 \end{itemize}
@@
 In  order to work in a unified framework, we shall consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighed networks.
-\rk{Check that's the case in the analysis}
-\rk{In fact, the graphs are much less interesting for undirected unweighed FA.}
+\rk{Roll back to directed FA, and explain that it's a link between in-vivo spontaneous substitution and in-vitro forced substitution.
+We could even add a more precise analysis using FA data directly (instead of features on the FA network), or keep that for the processual model.}
 
@@
     \TB{For the directed network FA we also define the incoming centrality $k'$ (number of neighbours pointing \emph{to} that node)}.
-    \rk{Remove that, since we restrict ourselves to undirected networks}
+    \rk{Rephrase once we rolled back FA to directed}
     Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank}~\citep{Page99}, which had already been used in~\citet{Griffiths07} and may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on recursively.
@@
 We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
-\rk{Check that's true in the code}
 The corresponding average quantity over all start words may be written:

### 2014-03-28 brains-feature-cards.tex
Move figures around, and create one page of figures per feature

\clearpage
\onecolumn
\appendix
\section{Feature figures}

% aoa_Kuperman ===============================
\subsection{aoa Kuperman}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/aoa_Kuperman/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/aoa_Kuperman/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/aoa_Kuperman/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/aoa_Kuperman/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% cmu_MNphonemes ===============================
\clearpage
\subsection{cmu MNphonemes}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNphonemes/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/cmu_MNphonemes/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNphonemes/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/cmu_MNphonemes/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% cmu_MNsyllables ===============================
\clearpage
\subsection{cmu MNsyllables}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNsyllables/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/cmu_MNsyllables/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNsyllables/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/cmu_MNsyllables/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_BCs ===============================
\clearpage
\subsection{fa BCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_BCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_BCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_BCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_BCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_CCs ===============================
\clearpage
\subsection{fa CCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_CCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_CCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_CCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_CCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_degrees ===============================
\clearpage
\subsection{fa degrees}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_degrees/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_degrees/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_degrees/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_degrees/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_PR_scores ===============================
\clearpage
\subsection{fa PR scores}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_PR_scores/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_PR_scores/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_PR_scores/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_PR_scores/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% mt_frequencies ===============================
\clearpage
\subsection{mt frequencies}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/mt_frequencies/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/mt_frequencies/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/mt_frequencies/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/mt_frequencies/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% mt_start_frequencies ===============================
\clearpage
\subsection{mt start frequencies}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/mt_start_frequencies/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/mt_start_frequencies/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/mt_start_frequencies/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/mt_start_frequencies/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_BCs ===============================
\clearpage
\subsection{wn BCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_BCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_BCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_BCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_BCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_CCs ===============================
\clearpage
\subsection{wn CCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_CCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_CCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_CCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_CCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_degrees ===============================
\clearpage
\subsection{wn degrees}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_degrees/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_degrees/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_degrees/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_degrees/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_MNSyns ===============================
\clearpage
\subsection{wn MNSyns}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_MNSyns/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_MNSyns/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_MNSyns/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_MNSyns/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_NSigns ===============================
\clearpage
\subsection{wn NSigns}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_NSigns/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_NSigns/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_NSigns/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_NSigns/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_PR_scores ===============================
\clearpage
\subsection{wn PR scores}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_PR_scores/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_PR_scores/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_PR_scores/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_PR_scores/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

### 2014-03-28 brains-results.tex
Add a mising question to be addressed
Move figures around, and create one page of figures per feature

@@
 
-\begin{figure*}[!th]
-	\centering
-	\subfloat[][Age-of-Acquisition]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_aoa-Kuperman-all_suscept.png}
-		\label{fig:aoa-suscept}
-	}
-	~
-	\subfloat[][Number of phonemes]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_cmu-MNphonemes-all_suscept.png}
-		\label{fig:MNphonemes-suscept}
-	}
-	\caption{Susceptibility values in color over the histograms of feature values}
-\end{figure*}
-
-\begin{figure*}[!th]
-	\centering
-	\subfloat[][Age-of-Acquisition]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_aoa-Kuperman-all_absolute.png}
-		\label{fig:aoa-absolute}
-	}
-	~
-	\subfloat[][Number of phonemes]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_cmu-MNphonemes-all_absolute.png}
-		\label{fig:MNphonemes-absolute}
-	}
-	\caption{Variations of features, with corresponding values under $\mathcal{H}_0$ and $\mathcal{H}_{0,n}$}
-\end{figure*}
-
 \subsection{Epidemiological setting}
@@
 
-\begin{figure*}[!th]
-	\centering
-	\subfloat[][Log PageRank]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-PR_scores-all_absolute.png}
-		\label{fig:PR_scores-absolute}
-	}
-	~
-	\subfloat[][Log Betweenness coefficient]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-BCs-all_absolute.png}
-		\label{fig:BCs-absolute}
-	}
-	~
-	\subfloat[][Log Clustering coefficient]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-CCs-all_absolute.png}
-		\label{fig:CCs-absolute}
-	}
-	\caption{Variations of features, with corresponding values under $\mathcal{H}_0$ and $\mathcal{H}_{0,n}$}
-\end{figure*}
-
 Moreover, the substitution susceptibility corresponds to a \emph{mutation probability} in this setting (shown in figures \ref{fig:PR_scores-suscept} for PageRank, \ref{fig:BCs-suscept} for betweennes coefficients, and \ref{fig:CCs-suscept} for clustering coefficient).
@@
 
-\begin{figure*}[!th]
-	\centering
-	\subfloat[][Log PageRank]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-PR_scores-all_suscept.png}
-		\label{fig:PR_scores-suscept}
-	}
-	~
-	\subfloat[][Log Betweenness coefficient]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-BCs-all_suscept.png}
-		\label{fig:BCs-suscept}
-	}
-	~
-	\subfloat[][Log Clustering coefficient]{
-		\includegraphics[width=0.3\textwidth]{images/Fff_Mtbgs_Sno_Pall_N5_fa-CCs-all_suscept.png}
-		\label{fig:CCs-suscept}
-	}
-	\caption{Susceptibility values in color over the histograms of feature values}
-\end{figure*}
-
-%\begin{figure}[!th]
-%\includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
-%\end{figure}
+\rk{A big question to be addressed is the ``why'': what underlying aspect of words with these features make them words to be substituted or not, and why does the brain do that?
+How was that selected for in evolution?}

### 2014-04-08 algorithms/my.plotcorr.R
Also removed now useless algorithms folder

@@ -1,104 +1,0 @@
-my.plotcorr <- function (corr, outline = FALSE, col = "grey", upper.panel = c("ellipse", "number", "none"), lower.panel = c("ellipse", "number", "none"), diag = c("none", "ellipse", "number"), digits = 2, bty = "n", axes = FALSE, xlab = "", ylab = "", asp = 1, cex.lab = par("cex.lab"), cex = 0.75 * par("cex"), mar = 0.1 + c(2, 2, 4, 2), ...)
-{
-# this is a modified version of the plotcorr function from the ellipse package
-# this prints numbers and ellipses on the same plot but upper.panel and lower.panel changes what is displayed
-# diag now specifies what to put in the diagonal (numbers, ellipses, nothing)
-# digits specifies the number of digits after the . to round to
-# unlike the original, this function will always print x_i by x_i correlation rather than being able to drop it
-# modified by Esteban Buz
-  if (!require('ellipse', quietly = TRUE, character = TRUE)) {
-    stop("Need the ellipse library")
-  }
-  savepar <- par(pty = "s", mar = mar)
-  on.exit(par(savepar))
-  if (is.null(corr))
-    return(invisible())
-  if ((!is.matrix(corr)) || (round(min(corr, na.rm = TRUE), 6) < -1) || (round(max(corr, na.rm = TRUE), 6) > 1))
-    stop("Need a correlation matrix")
-  plot.new()
-  par(new = TRUE)
-  rowdim <- dim(corr)[1]
-  coldim <- dim(corr)[2]
-  rowlabs <- dimnames(corr)[[1]]
-  collabs <- dimnames(corr)[[2]]
-  if (is.null(rowlabs))
-    rowlabs <- 1:rowdim
-  if (is.null(collabs))
-    collabs <- 1:coldim
-  rowlabs <- as.character(rowlabs)
-  collabs <- as.character(collabs)
-  col <- rep(col, length = length(corr))
-  dim(col) <- dim(corr)
-  upper.panel <- match.arg(upper.panel)
-  lower.panel <- match.arg(lower.panel)
-  diag <- match.arg(diag)
-  cols <- 1:coldim
-  rows <- 1:rowdim
-  maxdim <- max(length(rows), length(cols))
-  plt <- par("plt")
[…]

### 2014-04-08 brains-protocol.tex
Move figures around, and create one page of figures per feature
General cleanup of figures, adding a link to computed figures from the code
Rework up to 3.2.3 (excluded)

@@
 \begin{itemize}
-    \item \rk{Add word frequency as a feature, and maybe link it to Kolmogorov's 5/3 law in turbulences}
+    \item \textbf{Word frequency}: the frequency at which words appear in our dataset, \rk{add ref saying it's important}
     \item \textbf{Age of Acquisition}: the average age at which words are learned, obtained from~\citet{kuperman12},
     \item The average \textbf{Number of Phonemes} for all pronunciations of a word, obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98},\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.}
-    \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary.
+    \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary,
+    \item The \textbf{Number of Meanings} for a word, obtained from WordNet~\citep{WordNet10}. \rk{add ref on why it's important}
 \end{itemize}
@@
 The results were however extremely similar across the various categories, exhibiting no specific effect of words belonging to different POS categories.
+\rk{See \#8 for this fact-check}
 
@@
 Additional to classical psycholinguistic measures, we also considered more recently studied variables based on semantic network properties.
-We relied on two types of network-based data: curated synonym associations provided by WordNet~\citep{WordNet10}, and ``free association'' norms collected by~\citet{Nelson04}.
-
-The WordNet (WN) database gathers more than \num{117000} concepts and about \num{147000} words attached to these concepts.
-WordNet entries are disambiguated words, called ``lemmas'', which are grouped into concepts called ``synsets'': all the lemmas in a synset are synonyms for the same single meaning.
-Note that synsets also encode grammatical category, so the choice of a synset for a word (thus creating a lemma) immediately determines the grammatical category of that word.
-Non-disambiguated words make the nodes of the semantic network we build (i.e.
-homograph lemmas are aggregated into a single node), and a node's neighbours are synonymous lemmas belonging to the same synset(s) as the lemma(s) represented by that node.
-In other words, synsets induce cliques in the semantic network, and a word with several meanings belongs to as many cliques as it has meanings.
-\TB{Links connecting two synonymous lemmas are assigned a weight corresponding to the number of times these lemmas are mentioned by WordNet as synonyms, i.e. in the same synset.}
-\rk{We might spare ourselves that last sentence since we drop that aspect further down.}
+We relied on the ``free association'' norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
 
@@
 \end{quote}
-Following \citet{Griffiths07}, we consider the directed weighed network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
-
-While the original WN network is a weighted, undirected network, the FA network is a directed network where weights have a different meaning.
-Both semantic networks describe synonymy relationships corresponding to qualitatively distinct forms of association: either from a linguistic viewpoint (WN) or from the perspective of actual information retrieval by humans (FA).
-In  order to work in a unified framework, we shall consider links as an information on the existence of some sort of association between two words and transform both networks into undirected, unweighed networks.
-\rk{Roll back to directed FA, and explain that it's a link between in-vivo spontaneous substitution and in-vitro forced substitution.
-We could even add a more precise analysis using FA data directly (instead of features on the FA network), or keep that for the processual model.}
+Following \citet{Griffiths07}, we first consider the directed weighed network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
+\rk{we're in fact using the unweighed version of the network. Why?}
+This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyse is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
 
 \bigskip
-We introduce three standard network-based measures to be used on both the WN and FA networks:
+We introduce three standard network-based measures to be used on both the FA network \rk{adapt once the weighing question (\#8) is settled}:
 
 \begin{itemize}
-    \item \textbf{Distance} between two words, a classical dyadic network measure corresponding to the shortest path length between two nodes.
-    This measure may be used as a proxy of the remoteness between two given words, which is likely to influence the likelihood of a word being a relevant substitution candidate for another word.
-    \item \textbf{Centrality} $k$, measured by the number of neighbours of a given node, which can be interpreted as a proxy for the polysemy of words.
-    \TB{For the directed network FA we also define the incoming centrality $k'$ (number of neighbours pointing \emph{to} that node)}.
-    \rk{Rephrase once we rolled back FA to directed}
-    Note that in the present case, there is a quasi-perfect correlation between node degree and node \emph{PageRank}~\citep{Page99}, which had already been used in~\citet{Griffiths07} and may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the PageRank sense are those which are linked to many nodes themselves linked to many nodes, and so on recursively.
+    \item \textbf{Centrality} $k$, measured by the number of incoming edges to a given node, i.e.
+the number of cues for which a given word is triggered as an association, which strongly relates to word polysemy.
+    Note that in the present case, there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, a measure which had already been used by~\citet{Griffiths07} and which may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
     \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual v.
 possible edges between a node's neighbours \cite{watt-coll}.
-    In the WN network, a word is likely to have a higher clustering coefficient whenever the synsets to which it belongs are themselves semantically linked.
-    This measure had been used by \citet{Chan10}.
+    We compute the clustering coefficient on the undirected version of the FA network; we thus measure if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
+    \item \textbf{Betweenness coefficient} $b$, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network~\citep{free:set}.
+    More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher the coefficient, the more important that node is in ensuring the connectedness of the rest of the network.
+    This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.
 \end{itemize}
-
-\TB{Betweenness coefficient, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network~\citep{free:set}.
-More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher the coefficient, the more important that node is in ensuring the connectedness of the rest of the network.
-Correlation between FA-BC and FA-PR is $0.74$; between WN-BC and WN-PR it is $0.83$.}
-\rk{We dropped this, right?}
 
@@
 \begin{figure}[!th]
-    \includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
+    \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
     \caption{Variable correlations \rk{rebuild with the variables we talk about and keep}}
@@
 
-\begin{figure*}[h]
-    \centering
-    \def\svgwidth{\textwidth}
-    \small
-    \input{images/model-slidetimebags-b.pdf_tex}
-    \caption{Detection of substitutions according to model \rk{update text in figure}}
-    \label{fig:model-slidetimebags}
-\end{figure*}
+\rk{Add figure describing the chosen model}
 

### 2014-04-09 brains-protocol.tex
Rework variable correlation part

@@
 \begin{itemize}
-    \item \textbf{Centrality} $k$, measured by the number of incoming edges to a given node, i.e.
+    \item \textbf{Centrality} $k$, initially measured by the number of incoming edges to a given node, i.e.
 the number of cues for which a given word is triggered as an association, which strongly relates to word polysemy.
-    Note that in the present case, there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, a measure which had already been used by~\citet{Griffiths07} and which may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
+    However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
+Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
     \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual v.
@@
 
-An important question here relates to the various possible correlations between all the variables we consider.
+There is an important questions here about the possible correlations between all the variables we consider.
 
-The age of acquisition is a key variable, primarily as a usual suspect in psycholinguistic studies, and also appears to be generally correlated to many of the other variables.
-This relates to an ongoing debate suggesting that age of acquisition encodes a variety of phenomena which are difficult to disentangle from more specific phenomena, which could be captured by more independent variables such as \CN.
-\rk{What do we conclude of that? Nuance results with AoA?}
-\TB{Indeed, the minimal linear correlation of $aoa$ ($-0.31$) with other variables has a higher magnitude than the maximal linear correlation of other variables with each other ($-0.26$).}
-\rk{how do we get to that conclusion?}
+\begin{figure}[!th]
+    \centering
+    \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
+    \caption{Correlations in the initial set of features \rk{show non log version?}}
+    \label{fig:feature-corrs-initial}
+\end{figure}
 
-Additionally, number of phonemes and number of syllables exhibit a strong linear correlation ($0.85$).
+Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies and is also usually correlated to many of the other variables.
+This relates to an ongoing debate suggesting that age of acquisition encodes a variety of phenomena, difficult to disentangle from more specific phenomena which could be captured by more independent variables such as~\CN{}.
+Here however, as can be seen in Figure~\ref{fig:feature-corrs-initial}, age of acquisition has a relatively low correlation to the other variables (absolute value not above $0.42$ if we exclude the centrality measures), leading us to keep the variable in the rest of the analysis.
+
+Number of phonemes and number of syllables naturally exhibit a strong linear correlation ($0.83$).
 The analysis showed a better prediction effect of number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to focus the presented results on the former only.
-\rk{Rather, graphs are much nicer with number of phonemes than number of syllables: clearer tendencies, more data (there's holes with number of syllables). But not a prediction (prediction of what?)}
+
+Frequency and number of meanings both have relatively low levels of correlation to the other variables; we therefore also keep them in the rest of the analysis.
 
 \bigskip
-Network properties, finally, appear to be relatively weakly correlated with each other, across one network or between both networks (i.e.
-nodes with high index values are generally distinct between FA and WN).
-We chose to focus the results on FA as this network presents a direct psychological interpretation (that human copy-pasting may or may not show the same type of trend as human free association), noting that prediction results are matched against WN as well.
-\rk{comment peut-on dire que les nœuds importants dans FA ne sont pas nécessairement importants dans WN mais que, pourtant, la forme de la prédiction est la même?
-cela me semble illogique en première lecture}
-\rk{Check those assertions}
+Network properties, on the other hand, are strongly dependant on one another.
+As mentioned earlier, word degree and word pagerank have a very strong correlation ($0.89$) and, degree being generally more correlated to other variables, we chose to remove this variable from the results presented.
+Finally betweenness centrality also exhibits strong correlation levels to the other network properties ($0.62$, $0.64$, and $0.72$ in absolute value), leading us to drop this final feature due to its redundancy.
+
+The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.
 
 \begin{figure}[!th]
-    \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
-    \caption{Variable correlations \rk{rebuild with the variables we talk about and keep}}
+    \centering
+    \includegraphics[width=0.788\linewidth]{images/computed-figures/feature_correlations-filter2.png}
+    \caption{Correlations in the filtered set of features \rk{show non log version?}}
+    \label{fig:feature-corrs-filtered}
 \end{figure}

### 2014-04-09 brains-protocol.tex
Start reworking 3.3

@@
 
-\subsection{Detection of substitutions}
+\subsection{Detection of transformations}
 
 The data we use presents an additional challenge: each set of quotations bears no explicit information either about the authoritative original quotation, or about the source quotation(s) each author inspired himself from when creating a new post and reproducing (possibly altering) those sources.
-Substitutions are therefore not explicitly encoded in the dataset.
+Quote-to-quote transformations, and much less substitutions, are therefore not explicitly encoded in the dataset.
 
-We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
-We therefore model the underlying quotation selection process by making a few additional assumptions which let us define substitutions from the available data.
+We face an inference problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
+We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote transformations and substitutions from the available data.
 
-The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
+\rk{Rework all this with model classification}
+
+\TB{The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
 Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.
@@
 Should we consider them to be substitutions based on $q_a$ (i.e.
-re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?
+re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?}
 
-To settle this question we model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e.
+\TB{To settle this question we model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e.
 if not stemming from a source external to the dataset, e.g.
@@
 In any other case, i.e.
-if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.
+if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.}
 
-An example of this detection in a situation akin to the one described above can be seen on figure \ref{fig:model-slidetimebags}: $q_a$ and $q_b$ differ by only a word, and $q_a$ appears first and stays the most frequent in the beginning.
+\TB{An example of this detection in a situation akin to the one described above can be seen on figure \ref{fig:model-slidetimebags}: $q_a$ and $q_b$ differ by only a word, and $q_a$ appears first and stays the most frequent in the beginning.
 This is why the occurrences of $q_b$ at $t_1$ and $t_2$ are detected as substitutions stemming from $q_a$.
@@
 This entails that occurrences of $q_b$ are seen as reproductions of itself ($t_3$), and that occurrences of $q_a$ are detected as substitutions stemming from $q_b$ ($t_4$), i.e.
-\emph{re-creations} of $q_a$.
+\emph{re-creations} of $q_a$.}
 
@@
 
-The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
+\TB{The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
 We identified and implemented five other such models, and they all yielded essentially the same results.
@@
 The source quotation of a potential substitution can be chosen among those time-windows, or otherwise (e.g.
-among \emph{all} quotations having appeared before $t$ and differing by a word from the arrival quotation; this detection process would include all possible substitutions detected by other models, but would also include many false positives).
+among \emph{all} quotations having appeared before $t$ and differing by a word from the arrival quotation; this detection process would include all possible substitutions detected by other models, but would also include many false positives).}
 
@@
 We then measure the alterations introduced when authors reproduce quotations by comparing the relative features of substituted and substituting words.
-Note that since we only consider substitutions, we measure the features of an alteration \emph{knowing that there has been a alteration}, and we do not take invariant quotations into account.
+Note that since we only consider substitutions, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').

### 2014-04-09 brains-protocol.tex
Reordering and phrasing

@@
 
-\subsection{Characterization of substitutions}
+\subsection{Global evolution of quotation families}
 
-We then measure the alterations introduced when authors reproduce quotations by comparing the relative features of substituted and substituting words.
-Note that since we only consider substitutions, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
+\subsection{Transformation process}
+
+We then focus on the transformation process at work when authors reproduce quotations, by examining the features of substituted and substituting words in each substitution.
+Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
@@
 We build two main observables for each word feature.
-First, we measure the variation of word feature over a substitution, looking at the variation of a given feature between arrival and start words.
-Second, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution.
+First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution.
+Second, we measure the variation of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
+
+\subsubsection{Susceptibility}
+
+For a given feature $\mathfrak{F}$, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
+We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
+Word substitution susceptibility (denoted $\mathfrak{S}_{\mathfrak{F}}(w)$) is computed as the ratio of the number of times $n_s(w)$ a word is substituted to the number of times $n_p(w)$ that word appears in a substitutable position.
+We have:
+$$\mathfrak{S}_{\mathfrak{F}}(w) \defeq \frac{n_s(w)}{n_p(w)}$$
+
+Now averaging over all words having a given feature value $f$, we obtain the mean susceptibility for the feature value $f$:
+\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
+Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
+$$\left< \mathfrak{S}_{\mathfrak{F}} \right>_f = \left< \frac{n_s(w)}{n_p(w)} \right>_{\left\lbrace w | \mathfrak{F}(w) = f \right\rbrace}$$
+
+This measure focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation -- if any -- will be substituted.
 
@@
 
-For a given feature $\mathfrak{F}$, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$.
+Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$.
 Let us denote this quantity:
@@
 Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.
-\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
-Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
 This quantity can be written:
@@
 
-We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word} (neighbours in the WN network \rk{We could also do that with neighbours in directed or undirected FA}), i.e.
+We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word}, i.e.
 an arrival word chosen among semantically plausible though still random words.
-\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.}
+\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}
 
@@
 This gives us a fine-grained view of how word features evolve upon substitution.
-
-\subsubsection{Susceptibility}
-
-Furthermore, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
-We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
-Word substitution susceptibility (denoted $\mathfrak{S}_{\mathfrak{F}}(w)$) is computed as the ratio of the number of times $n_s(w)$ a word is substituted to the number of times $n_p(w)$ that word appears in a substitutable position.
-We have:
-$$\mathfrak{S}_{\mathfrak{F}}(w) \defeq \frac{n_s(w)}{n_p(w)}$$
-
-Now averaging over all words having a given feature value $f$, we obtain the mean susceptibility for the feature value $f$:
-$$\left< \mathfrak{S}_{\mathfrak{F}} \right>_f = \left< \frac{n_s(w)}{n_p(w)} \right>_{\left\lbrace w | \mathfrak{F}(w) = f \right\rbrace}$$
-
-This measure focuses on the selection of start word instead of the selection of the arrival word.
-Indeed, the features have an effect not only on the choice of a new word when a substitution takes place, but also at the preceding moment when it is not yet known which word in the quotation -- if any -- will be substituted.

### 2014-04-09 brains-protocol.tex
Add a section on the global evolution of quotation families

@@
 
-\subsection{Detection of transformations}
+\subsection{Detection of substitutions}
 
@@
 We face an inference problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
-We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote transformations and substitutions from the available data.
+We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote substitutions from the available data.
 
@@
 
-\subsection{Global evolution of quotation families}
+\subsection{Long-term evolution of quotation families}
 
-\subsection{Transformation process}
+We first examine the evolution of quotation families under the repeated action of substitutions.
+Our goal in this step is to identify the long-term effect of cognitive bias over the lifetimes of quotation families and thus over the framing of public information.
 
-We then focus on the transformation process at work when authors reproduce quotations, by examining the features of substituted and substituting words in each substitution.
+\rk{on pourrait en fait modéliser et formaliser ça de façon bien poussée en allant chercher du côté de l'analyse des séries temporelles et des processus stochastiques en général, mais bon...
+c'est pas le but ici, et ça serait une toute autre recherche, et un autre public.
+Oder?}
+
+A discreet time (as explained in the discreet-time models in the previous section) is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 1 to 5 in practice).
+Each quote in a time bag is counted as many times as it occurs in the corresponding time interval.
+
+We then compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
+This yields a distribution of feature values for each $n$, which is the simplest possible view of the state of an average quotation family in its $n$-th time bag, i.e.
+after $n \times \Delta t$ days.\footnote{For consistency, if we do this for $n$ going up to $N$, we only include quotation families that span long enough to have at least $N$ time bags.}
+Such a computation makes no assumptions on the way quotations undergo substitutions over time.
+
+The raw distributions built with this computation are stationary.
+\rk{figure/stats à l'appui ?
+bouger ça dans les résultats ?
+ou le garder comme ça, mentionné en passant pour justifier la suite ?}
+That is, the substitutions on quotations over time have no global effect on quotation families, either because external quotations are continuously fed into the family and compensate for the effect of substitutions, or because substitutions operate on a marginal portion of the quotation families, or both.
+
+\bigskip
+To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: after selecting one of the discreet-time substitution detection models (see previous section), we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
+The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
+We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the action of repeated cognitive bias.
+
+\subsection{Substitution process}
+
+We then focus on the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
 Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.

### 2014-04-09 brains-results.tex
Add a section on the global evolution of quotation families

@@
 \TB{We may see that $\mathcal{H}_0$ and $\mathcal{H}_{00}$ are slightly translated yet not qualitatively different.}
+
+\TB{Results for global evolution of features:}
+
+\TB{1/ no effect on global distribution => either continuous reinjection with base distribution, or substitutions are marginal (we can't separate those two with the information we have), or they are on marginal quotations (but no, because in that case we would see a difference when dropping quote frequency, which we don't).
+So the global distribution is stationary.}
+
+\TB{2/ we do see an effect when looking at chains, i.e. restricting the source quotes to be the product from previous step.}
+
+\TB{3/ also no effect when allowing more complex transformations than 1-word substitutions => they are either reinjections, or the process is not as constrained as 1-word substitutions.}
 

### 2014-04-11 brains-introduction.tex
Update remarks after today's conversation on current issues
Typos and phrasing

@@
 
-The next section (Sec~\ref{sec:related}) describes the state-of-the-art on this matter.
+The next section (Sec.~\ref{sec:related}) describes the state-of-the-art on this matter.
 In Sec.~\ref{sec:protocol}, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
-Section~\ref{sec:results} describes the significant psycholinguistic cues and biases observed during \emph{in vivo} quotation reformulation, followed by a discussion and general guidelines for further work in Sec.~\ref{sec:conclusion}.
+Sec.~\ref{sec:results} describes the significant psycholinguistic biases observed during \emph{in vivo} quotation reformulation as well as their epidemiological setting, followed by a discussion and general guidelines for further work in Sec.~\ref{sec:conclusion}.
 
@@
 This field is well developed and details the impact that classical psycholinguistic variables such as word frequency (see \citet{Yonelinas02} for a review), age-of-acquisition \citep{Zevin02}, number of phonemes or number of syllables (see for instance \citet{Rey98} and \citet{nick-diss}), have in this type of task.
+\rk{add a ref for word frequency}
 
@@
 First, \citet{Griffiths07} analyse a task where patients are asked to name the first word which comes to their mind when they are presented with a random letter from the Latin alphabet.
-The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (PageRank) in a language-wide semantic network built from external word association data.
-A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words come quicker when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
+The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
+A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
 
@@
 Our paper additionally sheds light on this debate.}
-\rk{But do we really care about this?}
+\rk{Show the link with epidemiology, and fitness: oddness is a kind of fitness}
 
 \smallskip
-Stepping back, we observe a gap between, on one side, macro-level empirical studies of the diffusion dynamics in a social system and, on the other side, studies focused on micro-level transformations of representations --- these latter studies being either strongly normative, or whose results are difficult to articulate with realistic cultural epidemiology models.
+Stepping back, we observe a gap between, on one side, macro-level empirical studies of the diffusion dynamics in a social system and, on the other side, studies focused on micro-level transformations of representations --- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.

### 2014-04-11 brains-results.tex
Minial hints for the future

@@
 \section{Results}\label{sec:results}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
+    \caption{XXXX}
+    \label{fig:timebags-evolution}
+\end{figure*}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+    \caption{XXXX}
+    \label{fig:feature-susceptibilities}
+\end{figure*}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
+    \caption{XXXX}
+    \label{fig:feature-variations}
+\end{figure*}
 
@@
 
-Where in the network are arrival words, are they related to the origin word, when relying on a given semantic network?
-(In other words, would a semantic word be a good predictor of the arrival of words?).
+\TB{Where in the network are arrival words, are they related to the origin word, when relying on a given semantic network?
+(In other words, would a semantic word be a good predictor of the arrival of wrds?).}
 \rk{tout un blabla sur les distances dans les substitutions.}
+
+\rk{is this really necessary?}
 

### 2014-04-11 brains-protocol.tex
Continue protocol rework, 3.3 still needs to be redone

@@
 
-We relied on a reliable quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
+We used a reliable quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
 This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009) \cite{Leskovec09-url}.\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
@@
 
-Additional to classical psycholinguistic measures, we also considered more recently studied variables based on semantic network properties.
+Aside from classical psycholinguistic measures, we also considered more recently studied variables based on semantic network properties.
 We relied on the ``free association'' norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
 
-Free Association (FA) norms record the words that come to mind when people are presented with a given cue (that is the ``free association'' task).
+Free association (FA) norms record the words that come to mind when someone is presented with a given cue (that is the ``free association'' task).
 As \citeauthor{Nelson04} explain,
@@
 \bigskip
-We introduce three standard network-based measures to be used on both the FA network \rk{adapt once the weighing question (\#8) is settled}:
+We introduce three standard network-based measures to be used on the FA network \rk{adapt once the weighing question (\#8) is settled}:
 
@@
 
-There is an important questions here about the possible correlations between all the variables we consider.
+An important question arises concerning the possible correlations between all the variables we use.
 

### 2014-04-11 brains-protocol.tex
Rework susbstitution detection models

@@
 
-\rk{Rework all this with model classification}
-
-\TB{The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
+The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
 Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.
@@
 Should we consider them to be substitutions based on $q_a$ (i.e.
-re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?}
+re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?
 
-\TB{To settle this question we model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e.
+To settle this question we can model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e.
 if not stemming from a source external to the dataset, e.g.
 initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$.
-The length $\Delta t$ of that period of time is fixed as a fraction of the total duration of the set of quotations; we took one fifth in the implementation (i.e.
-one month for a set of quotations spanning five months, or one day for a set spanning only five days; this takes the dynamism of the set of quotations into account).
+The length $\Delta t$ of that period of time is chosen as a fixed number of days (2 days in the implementation).
 If $q$ differs from $q_{max}$ by only a word, it is counted as a substitution from $q_{max}$ to $q$.
 In any other case, i.e.
-if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.}
+if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.
 
-\TB{An example of this detection in a situation akin to the one described above can be seen on figure \ref{fig:model-slidetimebags}: $q_a$ and $q_b$ differ by only a word, and $q_a$ appears first and stays the most frequent in the beginning.
-This is why the occurrences of $q_b$ at $t_1$ and $t_2$ are detected as substitutions stemming from $q_a$.
+An example of this detection in a situation akin to the one described above can be seen in Figure \ref{fig:substitution-detection-example}: $q_a$ and $q_b$ differ by only a word, and $q_a$ appears first, once, and so is the most frequent in the beginning.
+The first two occurrences of $q_b$ are therefore detected as substitutions stemming from $q_a$.
 After a time the situation is reversed: $q_b$ becomes more frequent than $q_a$.
-This entails that occurrences of $q_b$ are seen as reproductions of itself ($t_3$), and that occurrences of $q_a$ are detected as substitutions stemming from $q_b$ ($t_4$), i.e.
-\emph{re-creations} of $q_a$.}
+This entails that occurrences of $q_b$ are seen as reproductions of itself, and that occurrences of $q_a$ are detected as substitutions stemming from $q_b$, i.e. \emph{re-creations} of $q_a$.
 
-\rk{Add figure describing the chosen model}
+\begin{figure}[h]
+    \centering
+    \def\svgwidth{\linewidth}
+    \small
+    \input{images/substitution-detection-example.pdf_tex}
+    \caption{Example substitution detection model}
+    \label{fig:substitution-detection-example}
+\end{figure}
 
-\TB{The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
-We identified and implemented five other such models, and they all yielded essentially the same results.
-These models differ in the definition of the source quote from which new occurrences stem, essentially modifying the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
-For example, the time-windows considered can have different lengths, can include all occurrences from the beginning of the set of quotations, or can have fixed positions.
-The source quotation of a potential substitution can be chosen among those time-windows, or otherwise (e.g.
-among \emph{all} quotations having appeared before $t$ and differing by a word from the arrival quotation; this detection process would include all possible substitutions detected by other models, but would also include many false positives).}
+The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
+We identified and implemented eleven other such models, and they all yielded essentially the same results.
+These models differ on four fundamental principles or constraints:
+
+\begin{itemize}
+    \item \textbf{Time} can be kept continuous or made discreet (the model described above has continuous time),
+    \item The \textbf{source time interval} from which a source quote is selected can span a fixed number of days, or can go back to the beginning of the quotation family,
+    \item \textbf{Source quotes} in the source interval for substitutions can be further restricted, e.g. to only the most frequent quote in the interval as in the model described above,
+    \item \textbf{Destination quotes} for substitutions can be further restricted, e.g.
+to only quotes that did not appear in the source time interval (ensuring it is the first occurrence of the destination quote).
+\end{itemize}
+
+These variations essentially modify the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
 
@@
 
-A discreet time (as explained in the discreet-time models in the previous section) is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 1 to 5 in practice).
+A discreet time is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 2 in practice).
 Each quote in a time bag is counted as many times as it occurs in the corresponding time interval.

### 2014-04-11 brains-protocol.tex
Clean up some comments

@@
 \begin{itemize}
-    \item \textbf{Word frequency}: the frequency at which words appear in our dataset, \rk{add ref saying it's important}
+    \item \textbf{Word frequency}: the frequency at which words appear in our dataset, \rk{add ref on why it's important}
     \item \textbf{Age of Acquisition}: the average age at which words are learned, obtained from~\citet{kuperman12},
@@
     \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
-    \caption{Correlations in the initial set of features \rk{show non log version?}}
+    \caption{Correlations in the initial set of features \rk{show non-log version?}}
     \label{fig:feature-corrs-initial}
@@
 Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies and is also usually correlated to many of the other variables.
-This relates to an ongoing debate suggesting that age of acquisition encodes a variety of phenomena, difficult to disentangle from more specific phenomena which could be captured by more independent variables such as~\CN{}.
+This relates to an ongoing debate suggesting that age of acquisition encodes a variety of phenomena, difficult to disentangle from more specific phenomena which could be captured by more independent variables~\CN{}.
 Here however, as can be seen in Figure~\ref{fig:feature-corrs-initial}, age of acquisition has a relatively low correlation to the other variables (absolute value not above $0.42$ if we exclude the centrality measures), leading us to keep the variable in the rest of the analysis.
@@
     \includegraphics[width=0.788\linewidth]{images/computed-figures/feature_correlations-filter2.png}
-    \caption{Correlations in the filtered set of features \rk{show non log version?}}
+    \caption{Correlations in the filtered set of features \rk{show non-log version?}}
     \label{fig:feature-corrs-filtered}
@@
 Our goal in this step is to identify the long-term effect of cognitive bias over the lifetimes of quotation families and thus over the framing of public information.
-
-\rk{on pourrait en fait modéliser et formaliser ça de façon bien poussée en allant chercher du côté de l'analyse des séries temporelles et des processus stochastiques en général, mais bon...
-c'est pas le but ici, et ça serait une toute autre recherche, et un autre public.
-Oder?}
 

### 2014-04-11 brains-protocol.tex
Typos

@@
 Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
-    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual v.
-possible edges between a node's neighbours \cite{watt-coll}.
+    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual \emph{v.} possible edges between a node's neighbours \cite{watt-coll}.
     We compute the clustering coefficient on the undirected version of the FA network; we thus measure if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
@@
 \bigskip
-Network properties, on the other hand, are strongly dependant on one another.
+Network properties, on the other hand, are strongly dependent on one another.
 As mentioned earlier, word degree and word pagerank have a very strong correlation ($0.89$) and, degree being generally more correlated to other variables, we chose to remove this variable from the results presented.

### 2014-04-11 brains-results.tex
Work on figure captions

@@
     \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
-    \caption{XXXX}
+    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 39).
+    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
+    After a period of time, each feature becomes concentrated in a specific range of its own.}
     \label{fig:timebags-evolution}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{XXXX}
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution \emph{v.} average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
     \label{fig:feature-susceptibilities}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{XXXX}
+    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ \emph{v.} average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
+    The fact that all the curves have positive slopes shows the influence of the disappearing word on the appearing word (indeed, the contraposition is that selection of the appearing word independently from the disappearing word would produce horizontal curves).
+    Finally, the fact that all the curves also have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range, which is consistent with the evolution observed in Figure~\ref{fig:timebags-evolution}.}
     \label{fig:feature-variations}

### 2014-04-14 brains-protocol.tex
first camments

@@
 Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person.
+\cam{btw do we have an excerpt of quotations here, could we even feature a few prototypical examples?}
 Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
@@
     \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
-    \caption{Correlations in the initial set of features \rk{show non-log version?}}
+    \caption{Correlations in the initial set of features \rk{show non-log version?} \cam{if we use log variables I guess we should show correlations on log variables -- I'm surprised by the level of correlation b/n (aoa) and (log PR), though}}
     \label{fig:feature-corrs-initial}
@@
 We build two main observables for each word feature.
-First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution.
+First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution \new{under this condition}.
 Second, we measure the variation of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
@@
 an arrival word chosen among semantically plausible though still random words.
-\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}
+\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}
 

### 2014-04-14 brains-protocol.tex
zusammenarbeit 14.04

@@
 
-To keep the analysis palatable, we restricted the analysis to quotation transformations which consisted in the \emph{substitution} of a word by another word (and only those cases).
+To keep the analysis palatable, we restricted the analysis to quotation transformations which consisted in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
@@
 \end{quote}
-Following \citet{Griffiths07}, we first consider the directed weighed network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
-\rk{we're in fact using the unweighed version of the network. Why?}
+%Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
+% \rk{we're in fact using the unweighed version of the network. Why?}
+\new{Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.}
 This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyse is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
@@
 \bigskip
-We introduce three standard network-based measures to be used on the FA network \rk{adapt once the weighing question (\#8) is settled}:
+We introduce three standard network-based measures to be used on the FA network% \rk{adapt once the weighing question (\#8) is settled}
+:
 
@@
 The length $\Delta t$ of that period of time is chosen as a fixed number of days (2 days in the implementation).
-If $q$ differs from $q_{max}$ by only a word, it is counted as a substitution from $q_{max}$ to $q$.
+If $q$ differs from $q_{max}$ by only one word, as noted above, it is counted as a substitution from $q_{max}$ to $q$.
 In any other case, i.e.
@@
 The first two occurrences of $q_b$ are therefore detected as substitutions stemming from $q_a$.
-After a time the situation is reversed: $q_b$ becomes more frequent than $q_a$.
-This entails that occurrences of $q_b$ are seen as reproductions of itself, and that occurrences of $q_a$ are detected as substitutions stemming from $q_b$, i.e. \emph{re-creations} of $q_a$.
+After a time the situation is reversed and $q_b$ becomes more frequent than $q_a$: in this time frame $q_b$ is thus inaccurately considered as the original quote.
+This entails that occurrences of $q_b$ are seen as identical occurrences of the original quote, while occurrences of $q_a$ are detected as substitutions stemming from $q_b$.
 
@@
 
-These variations essentially modify the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
+These various flavors of an ideal substitution detection model %not only
+ have an impact on whether occurrences are considered as substitutions or repetition of the original quote. %, but also essentially modify how many times a given
 
-\subsection{Long-term evolution of quotation families}
+%variations essentially modify the balance between
+
+%considering the second blue quote as a reproduction of the first blue because the time window is too short
+
+%quote or as a mutation
+
+%taking into account the possibility of a
+
+%reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
+
+\section{Long-term evolution of quotation families}
 
@@
 \bigskip
-To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: after selecting one of the discreet-time substitution detection models (see previous section), we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
+To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned discrete-time substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
 The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
-We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the action of repeated cognitive bias.
+\new{As a result, the size of observed words/quotes drastically decreases across time, yet it assuredly focuses on successive mutations.} \cam{ajouter graphique pédagogique?}
+We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.
 
-\subsection{Substitution process}
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
+    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 39).
+    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
+    After a period of time, each feature becomes concentrated in a specific range of its own.}
+    \label{fig:timebags-evolution}
+\end{figure*}
+
+\section{Substitution process}
 
@@
 
-\subsubsection{Susceptibility}
+\subsection{Susceptibility}
 
@@
 
-\subsubsection{Alteration}
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution \emph{v.} average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
+    \label{fig:feature-susceptibilities}
+\end{figure*}
+
+\subsection{Alteration}
 
@@
 This gives us a fine-grained view of how word features evolve upon substitution.
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
+    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ \emph{v.} average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
+    The fact that all the curves have positive slopes shows the influence of the disappearing word on the appearing word (indeed, the contraposition is that selection of the appearing word independently from the disappearing word would produce horizontal curves).
+    Finally, the fact that all the curves also have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range, which is consistent with the evolution observed in Figure~\ref{fig:timebags-evolution}.}
+    \label{fig:feature-variations}
+\end{figure*}

### 2014-04-15 brains-protocol.tex
Rework presentation of measures

@@
 
-For a given feature $\mathfrak{F}$, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
+For a given feature $\phi$, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
-Word substitution susceptibility (denoted $\mathfrak{S}_{\mathfrak{F}}(w)$) is computed as the ratio of the number of times $n_s(w)$ a word is substituted to the number of times $n_p(w)$ that word appears in a substitutable position.
-We have:
-$$\mathfrak{S}_{\mathfrak{F}}(w) \defeq \frac{n_s(w)}{n_p(w)}$$
+Word substitution susceptibility is computed as the ratio of the number $s_w$ of times a word is substituted to the number $p_w$ of times that word appears in a substitutable position, i.e.
+$\sfrac{s_w}{p_w}$.
 
-Now averaging over all words having a given feature value $f$, we obtain the mean susceptibility for the feature value $f$:
+Now averaging over all words such that $\phi(w) = f$, we obtain the mean susceptibility for the feature value $f$:
 \footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
 Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
-$$\left< \mathfrak{S}_{\mathfrak{F}} \right>_f = \left< \frac{n_s(w)}{n_p(w)} \right>_{\left\lbrace w | \mathfrak{F}(w) = f \right\rbrace}$$
+$$\sigma_{\phi}(f) \defeq \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
 
@@
 
-Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$.
-Let us denote this quantity:
-$$\Delta (\wstart,\warrival) \defeq \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart)$$
-
-Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.
-This quantity can be written:
-$$\left< \Delta(\wstart,\warrival) \right>_f = \left< \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
+Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival) - \phi(\wstart)$.
+Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
+$$\Delta_{\phi}(f) \defeq \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \phi(\wstart) = f \right\rbrace}$$
 
 We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
-The corresponding average quantity over all start words may be written:
-\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
-$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
+The new quantity under $\mathcal{H}_0$ is:
+\footnote{Note that $\phi(\warrival_{\mathcal{H}_0})$ is in fact a constant in this averaging, since by definition $\warrival_{\mathcal{H}_0}$ does not depend on $\wstart$.}
+$$\Delta_{\phi, \mathcal{H}_0}(f) \defeq \left< \phi({\warrival_{\mathcal{H}_0}}) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \phi(\wstart) = f \right\rbrace}$$
 
@@
 an arrival word chosen among semantically plausible though still random words.
-\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}
+\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}\rk{S: Then we must say the results are the same.}
 

### 2014-04-15 brains-protocol.tex
Time is always discrete, either binned or sliding

@@
 \begin{itemize}
-    \item \textbf{Time} can be kept continuous or made discreet (the model described above has continuous time),
+    \item \textbf{Time} can be sliced into bins to build time bags (as in the model described above) or kept fine-grained by using sliding time bags,
     \item The \textbf{source time interval} from which a source quote is selected can span a fixed number of days, or can go back to the beginning of the quotation family,
@@
 
-These various flavors of an ideal substitution detection model %not only
- have an impact on whether occurrences are considered as substitutions or repetition of the original quote. %, but also essentially modify how many times a given
-
-%variations essentially modify the balance between
-
-%considering the second blue quote as a reproduction of the first blue because the time window is too short
-
-%quote or as a mutation
-
-%taking into account the possibility of a
-
-%reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
+These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote or repetitions of the original quote.
 
@@
 
-A discreet time is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 2 in practice).
+A discrete time is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 2 in practice).
 Each quote in a time bag is counted as many times as it occurs in the corresponding time interval.

### 2014-04-16 brains-protocol.tex
Add guiding remark for 3.3

@@
 The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
+\rk{it's a balance between repetition, substitution, or injection from outside/other process from inside}
 Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.

### 2014-04-16 brains-protocol.tex
Update correlation figures, maybe needs further filtering, and text needs to be updated

@@
     \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
-    \caption{Correlations in the initial set of features \rk{show non-log version?} \cam{if we use log variables I guess we should show correlations on log variables -- I'm surprised by the level of correlation b/n (aoa) and (log PR), though}}
+    \caption{Spearman correlations in the initial set of features}
     \label{fig:feature-corrs-initial}
 \end{figure}
+
+\rk{Correlation talk needs to be redone for new set of features}
 
@@
     \includegraphics[width=0.788\linewidth]{images/computed-figures/feature_correlations-filter2.png}
-    \caption{Correlations in the filtered set of features \rk{show non-log version?}}
+    \caption{Spearman correlations in the filtered set of features \rk{further remove FA betweenness?}}
     \label{fig:feature-corrs-filtered}

### 2014-04-16 brains-protocol.tex
Start working on model example figure

@@
 
-The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
-\rk{it's a balance between repetition, substitution, or injection from outside/other process from inside}
-Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.
-If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the author of $q_b$ on the 19th based himself on $q_a$.
-But what about the following occurrences of $q_b$?
+The main question in this problem concerns the way we consider later occurrences of a quotation that has already occurred earlier: the author of the later occurrences could have sourced an earlier occurrence and accurately copied it, they could have transformed an earlier occurrence of \emph{another} quote into this one, or they could have relied on information or other quotes from outside the dataset, i.e.
+from an external source we do not measure.
+
+Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs twice on the 19th and once on the 20th of January.
+If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 19th based himself on $q_a$.
+But what about the second occurrence of $q_b$ on the 19th, and the one on the 20th?
 Should we consider them to be substitutions based on $q_a$ (i.e.
 re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?
+
+\begin{figure}[h]
+    \centering
+    \def\svgwidth{\linewidth}
+    \small
+    \input{images/substitution-detection-example1.pdf_tex}
+    \caption{XXXXX}
+    \label{fig:substitution-detection-example1}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \def\svgwidth{\linewidth}
-    \small
-    \input{images/substitution-detection-example.pdf_tex}
-    \caption{Example substitution detection model}
-    \label{fig:substitution-detection-example}
-\end{figure}
-
 The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
@@
 
-These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote or repetitions of the original quote.
+These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
 
@@
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
+\rk{We also restrict to words substituted at least once.}
 Word substitution susceptibility is computed as the ratio of the number $s_w$ of times a word is substituted to the number $p_w$ of times that word appears in a substitutable position, i.e.

### 2014-04-17 brains-protocol.tex
Rework substitution detection paragraph, and figure

@@
 
-Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs twice on the 19th and once on the 20th of January.
-If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 19th based himself on $q_a$.
-But what about the second occurrence of $q_b$ on the 19th, and the one on the 20th?
+Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs on the 21st, 22nd and 23rd of January.
+If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in the top drawing of Figure~\ref{fig:substitution-detection-example1}.
+But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?
 Should we consider them to be substitutions based on $q_a$ (i.e.
-re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?
+re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or accuratee reproductions of the earlier occurrences of $q_b$ (options shown in Figure~\ref{fig:substitution-detection-example1})?
 
@@
 
-To settle this question we can model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e.
+To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution
+We further assume that when a quotation $q$ appears in time bag $i$, if it is not original (i.e.
 if not stemming from a source external to the dataset, e.g.
-initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$.
-The length $\Delta t$ of that period of time is chosen as a fixed number of days (2 days in the implementation).
-If $q$ differs from $q_{max}$ by only one word, as noted above, it is counted as a substitution from $q_{max}$ to $q$.
+initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($i - 1$).
+If $q$ differs from $q_{max}$ by only one word it is counted as a substitution from $q_{max}$ to~$q$.
 In any other case, i.e.
@@
 
-An example of this detection in a situation akin to the one described above can be seen in Figure \ref{fig:substitution-detection-example}: $q_a$ and $q_b$ differ by only a word, and $q_a$ appears first, once, and so is the most frequent in the beginning.
-The first two occurrences of $q_b$ are therefore detected as substitutions stemming from $q_a$.
-After a time the situation is reversed and $q_b$ becomes more frequent than $q_a$: in this time frame $q_b$ is thus inaccurately considered as the original quote.
-This entails that occurrences of $q_b$ are seen as identical occurrences of the original quote, while occurrences of $q_a$ are detected as substitutions stemming from $q_b$.
+This model has two main effects.
+First, the binning influences how many times quote occurrences can be counted as substitutions: in the second drawing on Figure~\ref{fig:substitution-detection-example1}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-detection-example1}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-detection-example1}, this seems the most likely scenario).
 

### 2014-04-17 brains-protocol.tex
Rework introduction of long-term evolution, needs comments on results

@@
 To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution
-We further assume that when a quotation $q$ appears in time bag $i$, if it is not original (i.e.
+We further assume that when a quotation $q$ appears in time bag $n$, if it is not original (i.e.
 if not stemming from a source external to the dataset, e.g.
-initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($i - 1$).
+initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($n - 1$).
 If $q$ differs from $q_{max}$ by only one word it is counted as a substitution from $q_{max}$ to~$q$.
@@
 
-\section{Long-term evolution of quotation families}
+\section{Macroscopic evolution of quotation families}
 
@@
 
-A discrete time is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 2 in practice).
-Each quote in a time bag is counted as many times as it occurs in the corresponding time interval.
+To do so we use a variation of the model presented in the previous section: quote occurrences are binned into 2-day time bags, substitutions can stem from any quote in the time bag preceding the current one, but destination quotes must be absent from that preceding time bag.
+The last two points constitute a variation from the ``majority'' rule which let us detect substitution chains later on.
 
@@
 The raw distributions built with this computation are stationary.
-\rk{figure/stats à l'appui ?
-bouger ça dans les résultats ?
-ou le garder comme ça, mentionné en passant pour justifier la suite ?}
 That is, the substitutions on quotations over time have no global effect on quotation families, either because external quotations are continuously fed into the family and compensate for the effect of substitutions, or because substitutions operate on a marginal portion of the quotation families, or both.
@@
 \bigskip
-To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned discrete-time substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
+To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
 The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
-\new{As a result, the size of observed words/quotes drastically decreases across time, yet it assuredly focuses on successive mutations.} \cam{ajouter graphique pédagogique?}
+As a result, the size of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
+\rk{add pedagogical figure}
 We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.
+
+\rk{comment results, transition to micro process}
 
@@
 
-\section{Substitution process}
+\section{Microscopic evolution: the substitution process}
 

### 2014-04-17 brains-protocol.tex
Add comments to choose figure for 4.

@@
 
+\rk{Adapt this last paragraph once the model is really chosen (see below)}
+
 We then compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
@@
 
-\rk{comment results, transition to micro process}
+\rk{Choose which figure to show (see garbage at the end of file)}
 
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
-    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 39).
-    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
-    After a period of time, each feature becomes concentrated in a specific range of its own.}
-    \label{fig:timebags-evolution}
-\end{figure*}
+\rk{Once chosen, comment results and transition to micro process}
+
+%\begin{figure*}[!th]
+%    \centering
+%    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
+%    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 39).
+%    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
+%    After a period of time, each feature becomes concentrated in a specific range of its own.}
+%    \label{fig:timebags-evolution}
+%\end{figure*}
 

### 2014-04-17 brains-garbage.tex
Acknowledge some changes, set comments for later

--- 2014-04-11 brains-results.tex
@@
 % !TEX root = brainscopypaste.tex
-\section{Results}\label{sec:results}
+\section*{Garbage / Results}\label{sec:results}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
-    \caption{XXXX}
+    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_no-exclusion.png}
+    \caption{Recursive, minimum 20 timebags for each cluster, \textbf{previous quotes not excluded}}
     \label{fig:timebags-evolution}
@@
     \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{XXXX}
-    \label{fig:feature-susceptibilities}
+    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_prevtbg-exclusion.png}
+    \caption{Chains, minimum 20 timebags for each cluster, \textbf{excluding quotes from preceding timebag}}
+    \label{fig:timebags-evolution}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{XXXX}
-    \label{fig:feature-variations}
+    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_prevnew-exclusion.png}
+    \caption{Chains, minimum 20 timebags for each cluster, \textbf{excluding quotes considered new in previous timebag}}
+    \label{fig:timebags-evolution}
 \end{figure*}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_shifting_min20_prevtbg-exclusion.png}
+    \caption{Chains, \textbf{shifting clusters}, minimum 20 timebags for each cluster, \textbf{excluding quotes from preceding timebag}}
+    \label{fig:timebags-evolution}
+\end{figure*}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_shifting_min10_prevtbg-exclusion.png}
+    \caption{Chains, \textbf{shifting clusters}, \textbf{minimum 10 timebags} for each cluster, \textbf{excluding quotes from preceding timebag}}
+    \label{fig:timebags-evolution}
+\end{figure*}
+
+%\begin{figure*}[!th]
+%    \centering
+%    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
+%    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 39).
+%    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
+%    After a period of time, each feature becomes concentrated in a specific range of its own.}
+%    \label{fig:timebags-evolution}
+%\end{figure*}
+%
+%\begin{figure*}[!th]
+%    \centering
+%    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+%    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution \emph{v.} average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+%    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
+%    \label{fig:feature-susceptibilities}
+%\end{figure*}
+%
+%\begin{figure*}[!th]
+%    \centering
+%    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
+%    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ \emph{v.} average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+%    The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
+%    The fact that all the curves have positive slopes shows the influence of the disappearing word on the appearing word (indeed, the contraposition is that selection of the appearing word independently from the disappearing word would produce horizontal curves).
+%    Finally, the fact that all the curves also have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range, which is consistent with the evolution observed in Figure~\ref{fig:timebags-evolution}.}
+%    \label{fig:feature-variations}
+%\end{figure*}
 

### 2014-04-17 brains-protocol.tex
Acknowledge some changes, set comments for later

@@
 We build two main observables for each word feature.
-First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution \new{under this condition}.
+First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution under this condition.
 Second, we measure the variation of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
@@
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
-\rk{We also restrict to words substituted at least once.}
 Word substitution susceptibility is computed as the ratio of the number $s_w$ of times a word is substituted to the number $p_w$ of times that word appears in a substitutable position, i.e.
@@
 
-Now averaging over all words such that $\phi(w) = f$, we obtain the mean susceptibility for the feature value $f$:
+Now averaging over all words such that $\phi(w) = f$ (only taking into account words that are substituted at least once), we obtain the mean susceptibility for the feature value $f$:
 \footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
@@
 This measure focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation -- if any -- will be substituted.
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution \emph{v.} average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
-    \label{fig:feature-susceptibilities}
-\end{figure*}
 
@@
 
+\subsection{Results}
+
+\rk{Comment results}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution \emph{v.} average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
+    \label{fig:feature-susceptibilities}
+\end{figure*}
+
 \begin{figure*}[!th]

### 2014-04-22 brains-protocol.tex
cosmetics 2.0

@@
 
-We face an inference problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
+We face an inference problem where, given all quotations and their occurrence timestamps, we should estimate which was the originating quotation for each instance of each quotation.
 We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote substitutions from the available data.
 
-The main question in this problem concerns the way we consider later occurrences of a quotation that has already occurred earlier: the author of the later occurrences could have sourced an earlier occurrence and accurately copied it, they could have transformed an earlier occurrence of \emph{another} quote into this one, or they could have relied on information or other quotes from outside the dataset, i.e.
+The main question in this problem concerns the manner in which we decide whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
+%consider later occurrences of a quotation that has already occurred earlier: the author of the later occurrences could have sourced an earlier occurrence and accurately copied it, they could have transformed an earlier occurrence of \emph{another} quote into this one, or they could have relied on information or other quotes from outside the dataset, i.e.
 from an external source we do not measure.
@@
 But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?
-Should we consider them to be substitutions based on $q_a$ (i.e.
-re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or accuratee reproductions of the earlier occurrences of $q_b$ (options shown in Figure~\ref{fig:substitution-detection-example1})?
+Should we consider them to be substitutions based on $q_a$ %(i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) 
+or accurate reproductions of the previous occurrences of $q_b$? \cam{là je distinguerais 3a [top] de 3b [middle \& bottom]} (options shown in Figure~\ref{fig:substitution-detection-example1})?
 

### 2014-04-22 brains-protocol.tex
Remove BCs in correlation and variation plots (/!\ but not yet in timebag evolution plots)

@@
     \centering
-    \includegraphics[width=0.788\linewidth]{images/computed-figures/feature_correlations-filter2.png}
-    \caption{Spearman correlations in the filtered set of features \rk{further remove FA betweenness?}}
+    \includegraphics[width=0.6975\linewidth]{images/computed-figures/feature_correlations-filter2.png}
+    \caption{Spearman correlations in the filtered set of features}
     \label{fig:feature-corrs-filtered}

### 2014-04-22 brains-protocol.tex
Further adapt notations in protocol

@@
 Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
-$$\sigma_{\phi}(f) \defeq \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
+$$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
 
@@
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
-$$\Delta_{\phi}(f) \defeq \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \phi(\wstart) = f \right\rbrace}$$
+$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \phi(\wstart) = f \right\rbrace}$$
 
-We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
+We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_0$ had been chosen randomly from the pool of free association words.
 The new quantity under $\mathcal{H}_0$ is:
-\footnote{Note that $\phi(\warrival_{\mathcal{H}_0})$ is in fact a constant in this averaging, since by definition $\warrival_{\mathcal{H}_0}$ does not depend on $\wstart$.}
-$$\Delta_{\phi, \mathcal{H}_0}(f) \defeq \left< \phi({\warrival_{\mathcal{H}_0}}) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \phi(\wstart) = f \right\rbrace}$$
+\footnote{Note that $\phi(\warrival_0)$ is in fact a constant in this averaging, since by definition $\warrival_0$ does not depend on $\wstart$.}
+$$\Delta_{\phi}^0 (f) = \left< \phi({\warrival_0}) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival_0) | \phi(\wstart) = f \right\rbrace}$$
 
@@
 an arrival word chosen among semantically plausible though still random words.
-\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}\rk{S: Then we must say the results are the same.}
+\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}\rk{S: Then we must say the results are the same.}
 

### 2014-04-22 brains-protocol.tex
Simplify legend for fig. 5

@@
     The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
-    The fact that all the curves have positive slopes shows the influence of the disappearing word on the appearing word (indeed, the contraposition is that selection of the appearing word independently from the disappearing word would produce horizontal curves).
-    Finally, the fact that all the curves also have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range, which is consistent with the evolution observed in Figure~\ref{fig:timebags-evolution}.}
+    The fact that all the curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range, which is consistent with the evolution observed in Figure~\ref{fig:timebags-evolution}.}
     \label{fig:feature-variations}

### 2014-05-05 brains-garbage.tex
Typo
Choose timebag evolution model
We use the chains with minimum 20 timebags per cluster, no quotes
excluded from previous timebags. This is the best
number-of-words/explanation-complexity ratio.

@@
 \section*{Garbage / Results}\label{sec:results}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_no-exclusion.png}
-    \caption{Recursive, minimum 20 timebags for each cluster, \textbf{previous quotes not excluded}}
-    \label{fig:timebags-evolution}
-\end{figure*}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_prevtbg-exclusion.png}
-    \caption{Chains, minimum 20 timebags for each cluster, \textbf{excluding quotes from preceding timebag}}
-    \label{fig:timebags-evolution}
-\end{figure*}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_prevnew-exclusion.png}
-    \caption{Chains, minimum 20 timebags for each cluster, \textbf{excluding quotes considered new in previous timebag}}
-    \label{fig:timebags-evolution}
-\end{figure*}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_shifting_min20_prevtbg-exclusion.png}
-    \caption{Chains, \textbf{shifting clusters}, minimum 20 timebags for each cluster, \textbf{excluding quotes from preceding timebag}}
-    \label{fig:timebags-evolution}
-\end{figure*}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=0.95\textwidth]{images/computed-figures/timebags_evolution_recursive_shifting_min10_prevtbg-exclusion.png}
-    \caption{Chains, \textbf{shifting clusters}, \textbf{minimum 10 timebags} for each cluster, \textbf{excluding quotes from preceding timebag}}
-    \label{fig:timebags-evolution}
-\end{figure*}
 

### 2014-05-05 brains-protocol.tex
Fix missing fix of #meanings to #synonyms in protocol
Choose timebag evolution model
We use the chains with minimum 20 timebags per cluster, no quotes
excluded from previous timebags. This is the best
number-of-words/explanation-complexity ratio.

@@
     \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary,
-    \item The \textbf{Number of Meanings} for a word, obtained from WordNet~\citep{WordNet10}. \rk{add ref on why it's important}
+    \item The average \textbf{Number of Synonyms} for all meanings of a word, obtained from WordNet~\citep{WordNet10}. \rk{add ref on why it's important}
 \end{itemize}
@@
 
-\rk{Choose which figure to show (see garbage at the end of file)}
+\rk{Comment results and transition to micro process}
 
-\rk{Once chosen, comment results and transition to micro process}
-
-%\begin{figure*}[!th]
-%    \centering
-%    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution.png}
-%    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 39).
-%    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
-%    After a period of time, each feature becomes concentrated in a specific range of its own.}
-%    \label{fig:timebags-evolution}
-%\end{figure*}
+\begin{figure*}[!tbh]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_no-exclusion.png}
+    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 37).
+    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
+    After a period of time, each feature becomes concentrated in a specific range of its own.}
+    \label{fig:timebags-evolution}
+\end{figure*}
 

### 2014-05-15 brains-protocol.tex
Cut substitution figure in two, rename section 3.3

@@
 
-\subsection{Detection of substitutions}
+\subsection{Temporal binning}
 
@@
 Should we consider them to be substitutions based on $q_a$ %(i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) 
-or accurate reproductions of the previous occurrences of $q_b$? \cam{là je distinguerais 3a [top] de 3b [middle \& bottom]} (options shown in Figure~\ref{fig:substitution-detection-example1})?
+or accurate reproductions of the previous occurrences of $q_b$? \cam{là je distinguerais 3a [top] de 3b [middle \& bottom]} (options shown in Figure~\ref{fig:substitution-temporal-binning})?
 
@@
     \small
-    \input{images/substitution-detection-example1.pdf_tex}
+    \input{images/substitution-temporal-binning.pdf_tex}
     \caption{XXXXX}
-    \label{fig:substitution-detection-example1}
+    \label{fig:substitution-temporal-binning}
 \end{figure}
@@
 This model has two main effects.
-First, the binning influences how many times quote occurrences can be counted as substitutions: in the second drawing on Figure~\ref{fig:substitution-detection-example1}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
-Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-detection-example1}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-detection-example1}, this seems the most likely scenario).
+First, the binning influences how many times quote occurrences can be counted as substitutions: in the second drawing on Figure~\ref{fig:substitution-temporal-binning}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-temporal-binning}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-temporal-binning}, this seems the most likely scenario).
 
@@
 
+\begin{figure}[h]
+    \centering
+    \def\svgwidth{\linewidth}
+    \small
+    \input{images/substitution-q_max.pdf_tex}
+    \caption{XXXXX}
+    \label{fig:substitution-q_max}
+\end{figure}
+
 We build two main observables for each word feature.

### 2014-05-15 brains-protocol.tex
Typo

@@
 Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs on the 21st, 22nd and 23rd of January.
-If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in the top drawing of Figure~\ref{fig:substitution-detection-example1}.
+If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in the top drawing of Figure~\ref{fig:substitution-temporal-binning}.
 But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?

### 2014-05-15 brains-protocol.tex
Work on 3.3 and 4
* Further separate figure 3 into 3a and 3b
* Rework the presentation of binning of quotation families
* Rework 4 to be compatible with the above

@@
 We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote substitutions from the available data.
-
-The main question in this problem concerns the manner in which we decide whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
-%consider later occurrences of a quotation that has already occurred earlier: the author of the later occurrences could have sourced an earlier occurrence and accurately copied it, they could have transformed an earlier occurrence of \emph{another} quote into this one, or they could have relied on information or other quotes from outside the dataset, i.e.
-from an external source we do not measure.
+The main issue at hand is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
 
 Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs on the 21st, 22nd and 23rd of January.
-If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in the top drawing of Figure~\ref{fig:substitution-temporal-binning}.
+If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
 But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?
 Should we consider them to be substitutions based on $q_a$ %(i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) 
-or accurate reproductions of the previous occurrences of $q_b$? \cam{là je distinguerais 3a [top] de 3b [middle \& bottom]} (options shown in Figure~\ref{fig:substitution-temporal-binning})?
+or accurate reproductions of the previous occurrences of $q_b$? (Options shown in Figure~\ref{fig:substitution-temporal-binning-a}.)
 
@@
     \centering
+    \subfloat[Possible paths from occurrence to occurrence]{
 	    \def\svgwidth{\linewidth}
 	    \small
-    \input{images/substitution-temporal-binning.pdf_tex}
-    \caption{XXXXX}
+	    \input{images/substitution-temporal-binning-a.pdf_tex}
+	    \label{fig:substitution-temporal-binning-a}
+	}
+	\hfill \\
+    \subfloat[Binned quotation family]{
+	    \def\svgwidth{\linewidth}
+	    \small
+	    \input{images/substitution-temporal-binning-b.pdf_tex}
+	    \label{fig:substitution-temporal-binning-b}
+	}
+	\caption{Temporal binning of quotation families}
     \label{fig:substitution-temporal-binning}
@@
 
-To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution
-We further assume that when a quotation $q$ appears in time bag $n$, if it is not original (i.e.
-if not stemming from a source external to the dataset, e.g.
-initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($n - 1$).
-If $q$ differs from $q_{max}$ by only one word it is counted as a substitution from $q_{max}$ to~$q$.
-In any other case, i.e.
-if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.
+To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
+Then when a quotation $q$ appears in time bag $n$, it is counted as a substitution from each quote $q^*$ in the preceding time bag ($n - 1$) from which it differs by only one word.
+If no quote in the preceding time bag can qualify as a source in a substitution (i.e.
+$q$ differs from all the quotes in the preceding time bag by more than one word), the occurrence of $q$ is not considered to be an instance of substitution.
+Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
 
-This model has two main effects.
-First, the binning influences how many times quote occurrences can be counted as substitutions: in the second drawing on Figure~\ref{fig:substitution-temporal-binning}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
-Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-temporal-binning}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-temporal-binning}, this seems the most likely scenario).
-
-The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
+The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, time can be sliced into bins to build fixed time bags as is done here, or kept fine-grained by using sliding time bags.}
+These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
 We identified and implemented eleven other such models, and they all yielded essentially the same results.
-These models differ on four fundamental principles or constraints:
-
-\begin{itemize}
-    \item \textbf{Time} can be sliced into bins to build time bags (as in the model described above) or kept fine-grained by using sliding time bags,
-    \item The \textbf{source time interval} from which a source quote is selected can span a fixed number of days, or can go back to the beginning of the quotation family,
-    \item \textbf{Source quotes} in the source interval for substitutions can be further restricted, e.g. to only the most frequent quote in the interval as in the model described above,
-    \item \textbf{Destination quotes} for substitutions can be further restricted, e.g.
-to only quotes that did not appear in the source time interval (ensuring it is the first occurrence of the destination quote).
-\end{itemize}
-
-These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
 
@@
 
-To do so we use a variation of the model presented in the previous section: quote occurrences are binned into 2-day time bags, substitutions can stem from any quote in the time bag preceding the current one, but destination quotes must be absent from that preceding time bag.
-The last two points constitute a variation from the ``majority'' rule which let us detect substitution chains later on.
-
-\rk{Adapt this last paragraph once the model is really chosen (see below)}
-
-We then compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
+To do so we compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
 This yields a distribution of feature values for each $n$, which is the simplest possible view of the state of an average quotation family in its $n$-th time bag, i.e.
 after $n \times \Delta t$ days.\footnote{For consistency, if we do this for $n$ going up to $N$, we only include quotation families that span long enough to have at least $N$ time bags.}
-Such a computation makes no assumptions on the way quotations undergo substitutions over time.
+Such a computation, based solely on the binning of quotation families, makes no assumptions on the way quotations undergo substitutions over time.
 
@@
 \bigskip
-To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
+To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the substitution detection model described in the previous section, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
 The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
-As a result, the size of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
+As a result, the number of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
 \rk{add pedagogical figure}
@@
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+
+\TB{This model has two main effects.
+First, the binning influences how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-temporal-binning}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-temporal-binning}, this seems to be the most likely scenario).}
+
+\TB{\textbf{Source quotes} in the source interval for substitutions can be further restricted, e.g. to only the most frequent quote in the interval as in the model described above,}
 

### 2014-05-15 brains-protocol.tex
Add a pedagogical figure for substitution chains

@@
 To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the substitution detection model described in the previous section, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
-The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
+The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on (see Figure~\ref{fig:substitution-chains} for an illustration of this process).
 As a result, the number of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
-\rk{add pedagogical figure}
+
+\begin{figure}[h]
+    \centering
+    \def\svgwidth{\linewidth}
+    \tiny
+    \input{images/substitution-chains.pdf_tex}
+    \caption{XXXXX}
+    \label{fig:substitution-chains}
+\end{figure}
+
 We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.

### 2014-05-16 brains-protocol.tex
Figure updates
* Replace timebag evolution figure
* Cosmetic changes in substitution chains pedagogical figure
No need for '!' in placing figure
Rework section 4
* It properly explains why the substitution detection model is extended
* It explains the final figure for substitution detection

@@
 \subsection{Temporal binning}
+\label{sec:temporal-binning}
 
@@
 
+\begin{figure*}[th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_no-exclusion.png}
+    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 37).
+    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
+    After a period of time, each feature becomes concentrated in a specific range of its own.}
+    \label{fig:timebags-evolution}
+\end{figure*}
+
 We first examine the evolution of quotation families under the repeated action of substitutions.
@@
 
-\begin{figure*}[!tbh]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_no-exclusion.png}
-    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 37).
-    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
-    After a period of time, each feature becomes concentrated in a specific range of its own.}
-    \label{fig:timebags-evolution}
-\end{figure*}
-
 \section{Microscopic evolution: the substitution process}
@@
 We then focus on the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
-Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
-Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
 
-\TB{This model has two main effects.
-First, the binning influences how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
-Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-temporal-binning}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-temporal-binning}, this seems to be the most likely scenario).}
+To do so we first extend the substitution detection model used in the previous section to reduce possible false positives.
+Indeed, when in the previous section we were forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
+We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding time bag.
+Since it is fairly certain that authors will have come upon the most frequent quote in any given time bag (\emph{v.} all the other quotes in that time bag), restricting substitutions to the most frequent quote makes sure we only detect substitutions that really occurred, greatly reducing the number of false positives.
 
-\TB{\textbf{Source quotes} in the source interval for substitutions can be further restricted, e.g. to only the most frequent quote in the interval as in the model described above,}
+Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
+In time bag~3, $q_b$ now holds the majority and is considered the unique basis for the last occurrence of $q_a$ (in time bag~4).
+This is despite the fact that $q_a$ also appears in time bag~3 alongside $q_b$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
+Conversely, if $q_a$ had been the most frequent quote in time bag 3, the last occurrence of $q_a$ in time bag~4 would have been considered a faithful copy of the occurrence in time bag~3.
 
@@
 
-We build two main observables for each word feature.
+With the substitution detection model thus refined, we build two main observables for each word feature.
 First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution under this condition.
 Second, we measure the variation of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
+
+Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
+Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
 
@@
 We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word}, i.e.
-an arrival word chosen among semantically plausible though still random words.
-\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.} \rk{Do we present these results?}\cam{nope}\rk{S: Then we must say the results are the same.}
+an arrival word chosen among semantically plausible though still random words.\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.}
+The results were not qualitatively changed with this second null hypothesis, so for the sake of clarity we chose to present the results using the simpler~$\mathcal{H}_0$.
 

### 2014-07-09 brains-protocol.tex
Some figures renamed

@@
     \centering
-    \includegraphics[width=\linewidth]{images/computed-figures/feature_correlations-filter1.png}
+    \includegraphics[width=\linewidth]{images/computed-figures/correlations-filter1.png}
     \caption{Spearman correlations in the initial set of features}
@@
     \centering
-    \includegraphics[width=0.6975\linewidth]{images/computed-figures/feature_correlations-filter2.png}
+    \includegraphics[width=0.6975\linewidth]{images/computed-figures/correlations-filter2.png}
     \caption{Spearman correlations in the filtered set of features}
@@
 \section{Macroscopic evolution of quotation families}
-
-\begin{figure*}[th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/timebags_evolution_recursive_min20_no-exclusion.png}
-    \caption{\textbf{Feature distribution evolution:} evolution of the distribution of feature values in substitution chains over successive 2-day time bags (bags 0 to 18, i.e. days 0 to 37).
-    The legends indicate the number of words left in each time bag; these decrease exponentially since only a fraction of the quotes undergo substitution at each step.
-    After a period of time, each feature becomes concentrated in a specific range of its own.}
-    \label{fig:timebags-evolution}
-\end{figure*}
 

### 2014-07-09 brains-protocol.tex
merging

@@
 % ============================
-\section{Protocol} % =========
+\section{Data} % =========
 % ============================
@@
 
-\subsection{Dataset}
+\subsection{In vivo utterances}
 
@@
 
+\section{Observing the substitution process}
+
+We then focus on the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
+
 \subsection{Temporal binning}
@@
 
-The data we use presents an additional challenge: each set of quotations bears no explicit information either about the authoritative original quotation, or about the source quotation(s) each author inspired himself from when creating a new post and reproducing (possibly altering) those sources.
+To do so we first propose a substitution detection model.
+
+The utterance data we use presents a challenge: each set of quotations bears no explicit information either about the authoritative original quotation, or about the source quotation(s) each author inspired himself from when creating a new post and reproducing (possibly altering) those sources.
 Quote-to-quote transformations, and much less substitutions, are therefore not explicitly encoded in the dataset.
@@
 
-\section{Macroscopic evolution of quotation families}
-
-We first examine the evolution of quotation families under the repeated action of substitutions.
-Our goal in this step is to identify the long-term effect of cognitive bias over the lifetimes of quotation families and thus over the framing of public information.
-
-To do so we compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
-This yields a distribution of feature values for each $n$, which is the simplest possible view of the state of an average quotation family in its $n$-th time bag, i.e.
-after $n \times \Delta t$ days.\footnote{For consistency, if we do this for $n$ going up to $N$, we only include quotation families that span long enough to have at least $N$ time bags.}
-Such a computation, based solely on the binning of quotation families, makes no assumptions on the way quotations undergo substitutions over time.
-
-The raw distributions built with this computation are stationary.
-That is, the substitutions on quotations over time have no global effect on quotation families, either because external quotations are continuously fed into the family and compensate for the effect of substitutions, or because substitutions operate on a marginal portion of the quotation families, or both.
-
-\bigskip
-To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the substitution detection model described in the previous section, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
-The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on (see Figure~\ref{fig:substitution-chains} for an illustration of this process).
-As a result, the number of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
-
-\begin{figure}[h]
-    \centering
-    \def\svgwidth{\linewidth}
-    \tiny
-    \input{images/substitution-chains.pdf_tex}
-    \caption{XXXXX}
-    \label{fig:substitution-chains}
-\end{figure}
-
-We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.
-
-\rk{Comment results and transition to micro process}
-
-\section{Microscopic evolution: the substitution process}
-
-We then focus on the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
-
-To do so we first extend the substitution detection model used in the previous section to reduce possible false positives.
-Indeed, when in the previous section we were forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
+We now need to reduce possible false positives.
+Indeed, so far we have been forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
 We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding time bag.
@@
 
-\subsection{Results}
+\section{Results}
 

### 2014-12-12 brainscopycut.tex
enough for tonight

% General document style ------------------------------------------------------------------
\documentclass{acmtog}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% Colors
\usepackage[usenames,dvipsnames]{color}

\acmVolume{VV}
\acmNumber{N}
\acmYear{YYYY}
\acmMonth{Monteh}
\acmArticleNum{XXX}  
\acmdoi{10.1145/XXXXXXX.YYYYYYY}

\markboth{S. Lerique and C. Roth}{How do we copy and paste? The semantic drift of quotations in blogspace}

% Let us have multiline comments
\usepackage{verbatim}

% Footnotes in section headings
\usepackage[stable]{footmisc}

% Formatting bibliography
\usepackage{natbib}
\setcitestyle{authoryear,round,semicolon}

% Mathematics and Code
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[group-separator={,}]{siunitx}
%\usepackage{stmaryrd}
\usepackage{xfrac}
\usepackage{mathtools}
%\newcommand{\defeq}{\overset{\underset{\mathrm{def}}{}}{=}}

\usepackage{listings}
\lstset{
    basicstyle=\ttfamily
}

% Have graphics
\graphicspath{{images/}}
\usepackage{graphicx}
\usepackage{subfig}

% Help in placing the floats
\usepackage{placeins}

% To make comments, corrections, notes
\newcommand{\TB}[1]{\textcolor{blue}{#1}}
\newcommand{\new}[1]{\TB{#1}}
\newcommand{\tb}[1]{\textcolor{blue}{#1}}
\newcommand{\TG}[1]{\textcolor{MidnightBlue}{#1}}
\newcommand{\add}[1]{\TB{TOADD: #1}}
\newcommand{\ra}[1]{$\rightarrow$#1}
\newcommand{\replace}[2]{\sout{#1} \TB{#2}}
\newcommand{\rk}[1]{\TG{{\footnotesize {\bf[\emph{#1}]}}}}
\newcommand{\remark}[1]{\noindent\fbox{\parbox{.99\linewidth}{\TB{#1}}}}
\newcommand{\smallremark}[1]{\noindent\fbox{{\TB{#1}}}}
\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\TB{\scriptsize \sf #1}}}}}
\newcommand{\cam}[1]{\textcolor{blue}{\small{[{\bf C:} }#1{]}}}

% shortcuts
\newcommand{\arrival}{\text{arrival}}
\newcommand{\start}{\text{start}}

\newcommand{\warrival}{w'}
\newcommand{\wstart}{w}

% If citations or references are missing, make it visible
\newcommand{\CN}{\textsuperscript{\TB{[Citation needed]}}}
\newcommand{\CNs}{\textsuperscript{\TB{[Multiple citations needed]}}}
\newcommand{\RN}{\textsuperscript{\TB{[Internal reference needed]}}}

% Start with the real content -------------------------------------------------------------

\title{{How do we copy and paste?\\The semantic drift of quotations in blogspace}}
\date{}
\author{S\'EBASTIEN LERIQUE {\upshape and} CAMILLE ROTH}

\begin{document}
\maketitle

\input{brains-introduction}

\input{brains-protocol}

\input{brains-conclusion}

\subsection*{Acknowledgements}

We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal and Sharon Peperkamp \cam{forgot anyone?} for useful suggestions and comments.

\bibliographystyle{plainnat}
\bibliography{brainscopypaste.bib}

\clearpage 
\input{brains-garbage}

\end{document}

### 2014-12-12 brains-protocol.tex
newest

@@
-% !TEX root = brainscopypaste.tex
+% !TEX root = brainscopycut.tex
 % ============================
-\section{Data} % =========
+\section{Methods} % =========
 % ============================
@@
 
-In order to start bridging this gap, we set out to \emph{empirically} study public representation transformations at the microscopic level, aiming to stay compatible with macroscopic-level studies of these public representations.
-Quotations appeared to be a perfect candidate as public representations.
-First, they are usually cleanly delimited by quotation marks (and often with HTML markup in web pages), which greatly facilitates their detection in text corpora.
+%In order to start bridging this gap, we set out to \emph{empirically} study public representation transformations at the microscopic level; 
+%aiming to stay compatible with macroscopic-level studies of these public representations.
+Quotations appeared to be a perfect candidate to propose a first \emph{in vivo} measure of low-level cognitive bias in a reformulation task. % as public representations.
+First, they are usually cleanly delimited by quotation marks %(and often with HTML markup in web pages), 
+which greatly facilitates their detection in text corpora.
 Second, they stem from a unique ``original'' version, and could ideally be traceable back to that version.
 Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.
-Quotation evolution is therefore a perfect environment to measure cognition-induced transformations and relate those findings to macroscopic social dynamics.
+
+We could therefore examine the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
+To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
+
+The next subsections describe the dataset and measures we used to assess this cognitive bias.
+%Quotation evolution is therefore a perfect environment to measure cognition-induced transformations.%and relate those findings to macroscopic social dynamics.
 
@@
 
-We used a reliable quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
+We used a quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
 This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009) \cite{Leskovec09-url}.\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
-Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person.
-\cam{btw do we have an excerpt of quotations here, could we even feature a few prototypical examples?}
-Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
+Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person. For instance,
+\begin{quote}
+\cam{do we have an excerpt of quotations here, could we even feature a few prototypical examples? Obama said that ``...'' }
+\end{quote}
+Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words): \cam{``...''\ra``...' ''}
 A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
@@
 
-To keep the analysis palatable, we restricted the analysis to quotation transformations which consisted in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
-To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
-The following sections detail the features we used.
-
 \subsubsection{Standard psycholinguistic indices}
 
-We first introduce some of the most classical psycholinguistic measures on words:
-
+We now introduce some of the most classical psycholinguistic measures on words.
 \rk{Add some bibliography about those features' known effects}
@@
 Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
-    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual \emph{v.} possible edges between a node's neighbours \cite{watt-coll}.
+    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs.
+possible edges between a node's neighbours \cite{watt-coll}.
     We compute the clustering coefficient on the undirected version of the FA network; we thus measure if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
@@
 
-\section{Observing the substitution process}
+\subsection{Substitution model}
 
-We then focus on the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
-
-\subsection{Temporal binning}
+%\subsubsection{Temporal binning}
 \label{sec:temporal-binning}
 
-To do so we first propose a substitution detection model.
+We finally need a substitution detection model, for the utterance data we use presents a challenge: quote-to-quote transformations, and much less substitutions, are not explicitly encoded in the dataset.
+More precisely, each set of quotations bears no explicit information about either the authoritative original quotation, or the source quotation(s) each author relied on when creating a new post and reproducing (and possibly altering) that source.
+We thus face an inference problem where, given all quotations and their occurrence timestamps, we should estimate which was the originating quotation for each instance of each quotation.
 
-The utterance data we use presents a challenge: each set of quotations bears no explicit information either about the authoritative original quotation, or about the source quotation(s) each author inspired himself from when creating a new post and reproducing (possibly altering) those sources.
-Quote-to-quote transformations, and much less substitutions, are therefore not explicitly encoded in the dataset.
+We therefore model the underlying quotation selection process by making a few additional assumptions.% which let us define quote-to-quote substitutions from the available data.
+The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
 
-We face an inference problem where, given all quotations and their occurrence timestamps, we should estimate which was the originating quotation for each instance of each quotation.
-We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote substitutions from the available data.
-The main issue at hand is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
-
-Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs on the 21st, 22nd and 23rd of January.
-If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
-But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?
-Should we consider them to be substitutions based on $q_a$ %(i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) 
-or accurate reproductions of the previous occurrences of $q_b$? (Options shown in Figure~\ref{fig:substitution-temporal-binning-a}.)
+Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs on the 21st, 22nd and 23rd of January.
+If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 21st based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
+But what about the second and third occurrences of $q'$, on the 22nd and 23rd?
+Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place) 
+or accurate reproductions of the previous occurrences of $q'$? (Options shown in Figure~\ref{fig:substitution-temporal-binning-a}.)
 
@@
 	}
+	\hfill \\
+	\subfloat[XXXX]{
+	    \def\svgwidth{\linewidth}
+	    \small
+		\input{images/substitution-q_max.pdf_tex}
+	    \label{fig:substitution-q_max}
+	}
 	\caption{Temporal binning of quotation families}
@@
 $q$ differs from all the quotes in the preceding time bag by more than one word), the occurrence of $q$ is not considered to be an instance of substitution.
-Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
 
@@
 We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding time bag.
-Since it is fairly certain that authors will have come upon the most frequent quote in any given time bag (\emph{v.} all the other quotes in that time bag), restricting substitutions to the most frequent quote makes sure we only detect substitutions that really occurred, greatly reducing the number of false positives.
+We assume that authors will have encountered the most frequent quote in a given time bag (\hbox{vs.} all the other quotes in that time bag).
+Thus, restricting substitutions to the most frequent quote ensures that we only detect substitutions that really occurred, greatly reducing the number of false positives.
 
 Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
-In time bag~3, $q_b$ now holds the majority and is considered the unique basis for the last occurrence of $q_a$ (in time bag~4).
-This is despite the fact that $q_a$ also appears in time bag~3 alongside $q_b$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
-Conversely, if $q_a$ had been the most frequent quote in time bag 3, the last occurrence of $q_a$ in time bag~4 would have been considered a faithful copy of the occurrence in time bag~3.
-
-\begin{figure}[h]
-    \centering
-    \def\svgwidth{\linewidth}
-    \small
-    \input{images/substitution-q_max.pdf_tex}
-    \caption{XXXXX}
-    \label{fig:substitution-q_max}
-\end{figure}
+In time bag~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in time bag~4).
+This is despite the fact that $q$ also appears in time bag~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
+Conversely, if $q$ had been the most frequent quote in time bag 3, the last occurrence of $q$ in time bag~4 would have been considered a faithful copy of the occurrence in time bag~3.
 
@@
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+
+\section{Results}\label{sec:results}
 
@@
 
-\section{Results}
-
 \rk{Comment results}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution \emph{v.} average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
     Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ \emph{v.} average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ vs.
+average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
     The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.

### 2014-12-12 brains-introduction.tex
Typos
newest

@@
-% !TEX root = brainscopypaste.tex
+% !TEX root = brainscopycut.tex
 % ============================
@@
 
-The understanding of the mechanisms behind cultural similarity and diversity has led to a sizeable literature in the recent past, spanning over a vast area of research fields ranging from cultural anthropology to social network analysis and complex systems modelling, all diversely labelled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'', for the most part.
-Broadly, this type of research program investigates phenomena pertaining to both cognitive science and social science, and aims to understand the processing, transmission, and evolution of information both at the individual and social levels.
+The understanding of knowledge transmission mechanisms has led to a sizeable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining.
+These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part,
+%Broadly speaking, this type of research program is bound to
+ investigating phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
+ both at the individual and social levels.  Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.
 
-Several theories have been proposed and debated within these research fields, mixing social and individual cognition, notably from the side of cultural anthropology.
-First, the debate around the ``memetic'' program initiated by \citet{Dawkins76}, for which the collection of works by \citet{Aunger00} provides a solid overview.
-A second field has focused on the development of evolutionary models of norms (see for instance \citealp{Ehrlich05}), following the seminal work of \citet{Boyd85}.
+%Broadly speaking, this type of research program is bound to investigate phenomena pertaining to both cognitive science and social science, and aims to understand the processing, transmission, and evolution of information both at the individual and social levels.
+The relevant literature features two main streams, depending on whether the focus lies on social dynamics or cognitive processes.
 
-A third approach, ``cultural epidemiology'', initiated by \citet{sper:expl} recently attracted significant attention.
-Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.
-The cultural epidemiology program focuses on the notion of \emph{representation}, linking the cognitive science concept of \emph{mental representation} to the concept of \emph{public representation}, the latter being the former's counterpart outside the brain (i.e.
-in cultural artefacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
+A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
+It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview) the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the program of ``cultural epidemiology'', initiated by \citet{sper:expl}, which %recently attracted significant attention, and
+% focuses on the notion of \emph{representation}, 
+links %the cognitive science 
+concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e.
+in cultural artifacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
 \begin{quotation}
@@
 \end{quotation}}
-The key point here is that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
-Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.
-Despite some recent modelling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analysed.
 
-However, the last decade has witnessed an avalanche of observable \emph{in vivo} data in the form of online interactions.
-While they are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), these productions and information trails still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
-Given the already significant --~and rapidly growing~-- importance of our online interactions, these records can dramatically improve the prospects of empirical study of the individual-level processes of cultural evolution.
+Here, one of the core claim consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
+Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
+Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
+Despite some recent modelling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been extensively empirically analyzed.
 
-We aim to empirically describe the transformation of a specific type of public representation, by focusing on the possible alterations introduced by individuals when newly producing a representation.
+Another research stream deals with rather macroscopic studies of knowledge diffusion.
+Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace.
+The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by individuals).
+This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, has mainly come from a variety of online interaction contexts.
+While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
+Some could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
+
+These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
+They nonetheless show us the interest of using these records --~given the already significant and rapidly growing importance of our online interactions~--
+%and, %more broadly, large textual corpora \cam{google n-grams??} 
+to dramatically improve the empirical understanding of individual-level processes of cultural evolution.
+
+\bigskip
+Stepping back, we observe a gap between, on one side, empirical studies of the diffusion dynamics in a social system and, on the other side, more theoretical works  focused on knowledge transformation processes.
+%--- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.
+Our research lies at the intersection of these two programs, aiming to shed light on microscopic information transformation processes by leveraging the empirical wealth of (in vivo) social diffusion phenomena.
+We hope to describe some reformulation processes within a large distributed social system such as blogspace.
+More precisely, we aim to empirically describe how some specific type of public representation may possibly be altered by bloggers when they try to reproduce that representation.
+
+We focus on simple linguistic modifications, thereby connecting our research to the broader psycholinguistic literature.
 To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
@@
 
-The next section (Sec.~\ref{sec:related}) describes the state-of-the-art on this matter.
+The next section (Sec.~\ref{sec:related}) describes our hypotheses along with the relevant state-of-the-art on this \cam{psycholinguistic} matter.
 In Sec.~\ref{sec:protocol}, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
@@
 % ============================
-\section{Related work} % =====
+\section{Hypotheses} % =====
 % ============================
@@
 
-The relevant literature on \emph{public representation dynamics} features two main streams.
-On one hand, we find studies of the macroscopic \emph{social diffusion} of public representations, describing for instance the propagation of cultural artefacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
-These studies are relatively independent from anthropology and cognition and are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
-Without necessarily relying on specific social science theories, this research stream is of interest for its use of large social media corpora in studying cultural dynamics.
+%The relevant literature on \emph{public representation dynamics} features two main streams.
+%On one hand, we find studies of the macroscopic \emph{social diffusion} of public representations, describing for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
+%These studies are relatively independent from anthropology and cognition and are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
+%Without necessarily relying on specific social science theories, this research stream is of interest for its use of large textual corpora in studying cultural dynamics.
 
-On the other hand, the study of the \emph{transformation} of public representations has emerged only recently.
+The practical study of the \emph{transformation} of public representations has emerged only recently.
 For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07}.
+
 Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
@@
 One of the main conclusions of these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to observe and measure significant transformations.
-However, these studies address transformations by focusing on the properties of the source of the quotation (e.g.
-news outlet v.
-blog), or the surrounding public space (e.g.
-quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
+However, these studies address transformations by focusing on the properties of the source of the quotation (\hbox{e.g.} news outlet {vs.} blog), or the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
 
@@
 \smallskip
-Stepping back, we observe a gap between, on one side, macro-level empirical studies of the diffusion dynamics in a social system and, on the other side, studies focused on micro-level transformations of representations --- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.

### 2014-12-12 brains-protocol.tex
"time bag" -> "bin"

@@
 
-To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
-Then when a quotation $q$ appears in time bag $n$, it is counted as a substitution from each quote $q^*$ in the preceding time bag ($n - 1$) from which it differs by only one word.
-If no quote in the preceding time bag can qualify as a source in a substitution (i.e.
-$q$ differs from all the quotes in the preceding time bag by more than one word), the occurrence of $q$ is not considered to be an instance of substitution.
+To settle this question we bin the quote occurrences into fixed \emph{bins} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
+Then when a quotation $q$ appears in bin $n$, it is counted as a substitution from each quote $q^*$ in the preceding bin ($n - 1$) from which it differs by only one word.
+If no quote in the preceding bin can qualify as a source in a substitution (i.e.
+$q$ differs from all the quotes in the preceding bin by more than one word), the occurrence of $q$ is not considered to be an instance of substitution.
 Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
 
-The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, time can be sliced into bins to build fixed time bags as is done here, or kept fine-grained by using sliding time bags.}
+The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
 These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
@@
 Indeed, so far we have been forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
-We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding time bag.
-We assume that authors will have encountered the most frequent quote in a given time bag (\hbox{vs.} all the other quotes in that time bag).
+We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding bin.
+We assume that authors will have encountered the most frequent quote in a given bin (\hbox{vs.} all the other quotes in that bin).
 Thus, restricting substitutions to the most frequent quote ensures that we only detect substitutions that really occurred, greatly reducing the number of false positives.
@@
 Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
-In time bag~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in time bag~4).
-This is despite the fact that $q$ also appears in time bag~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
-Conversely, if $q$ had been the most frequent quote in time bag 3, the last occurrence of $q$ in time bag~4 would have been considered a faithful copy of the occurrence in time bag~3.
+In bin~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in bin~4).
+This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
+Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.
 

### 2014-12-12 brains-protocol.tex
morning build

--- 2014-12-12 brains-protocol.tex
@@
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
+
+Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than i expected'', ``did that just happen'', as well as many simple expletive-based sentences).
+Their presence is may largely owe to random variations around a casual expression, while we are interested in transformations of news-related quotes linked to an original, identifiable utterance.
+We thus filter out all clusters lasting more than 80 days, or featuring quotes having less than 5 words \cam{not sure this is what you meant [but then the filter may be a bit asymmetric, oder?]} (or not written in English).
+We eventually keep 19,621 clusters (out of 71,568; i.e.
+27.4\%), 60,246 quotes (out of 310,457; i.e.
+19.4\%) and $\sim$838k occurrences (out of 8.12m, i.e.
+10.3\%).
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose feature more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'', we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
+We could manually check on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
 

### 2014-12-12 brains-introduction.tex
US spelling, and phrasing, for first section (introduction)

@@
 
-The understanding of knowledge transmission mechanisms has led to a sizeable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining.
+The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining.
 These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part,
 %Broadly speaking, this type of research program is bound to
- investigating phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
+ investigate phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
  both at the individual and social levels.  Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.
@@
 A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
-It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview) the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the program of ``cultural epidemiology'', initiated by \citet{sper:expl}, which %recently attracted significant attention, and
+It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program initiated by \citet{sper:expl}, which %recently attracted significant attention, and
 % focuses on the notion of \emph{representation}, 
 links %the cognitive science 
-concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e.
+the concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e.
 in cultural artifacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
@@
 
-Here, one of the core claim consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
+Here, one of the core claims consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
 Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
 Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
-Despite some recent modelling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been extensively empirically analyzed.
+Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been \TB{extensively} empirically analyzed.\marginnote{\seb{Deux adverbes.
+Pas encore trouvé d'autre formulation.}}
 
@@
 Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace.
-The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by individuals).
-This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, has mainly come from a variety of online interaction contexts.
+The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by each individual).
+This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
 While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
@@
 
-These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
+These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
 They nonetheless show us the interest of using these records --~given the already significant and rapidly growing importance of our online interactions~--

### 2014-12-12 brains-introduction.tex

@@
 Let us mention two interesting and recent studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
-First, \citet{Griffiths07} analyse a task where patients are asked to name the first word which comes to their mind when they are presented with a random letter from the Latin alphabet.
+First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the Latin alphabet.
 The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
-A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster when they have a higher clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
+A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
 

### 2014-12-12 brains-introduction.tex
introduced examples + revamped the beginning of sec. 2 + minor corrections

@@
 Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
-Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been \TB{extensively} empirically analyzed.\marginnote{\seb{Deux adverbes.
-Pas encore trouvé d'autre formulation.}}
+Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.
 
@@
 
-The practical study of the \emph{transformation} of public representations has emerged only recently.
+The practical study of the transformation of public representations has emerged only recently.
 For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07}.
 
-Among the empirical approaches on the mutation of representations, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
-They show several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.
-One of the main conclusions of these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to observe and measure significant transformations.
+Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
+%They exhibit several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.
+One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations.
 However, these studies address transformations by focusing on the properties of the source of the quotation (\hbox{e.g.} news outlet {vs.} blog), or the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
+One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations.
+They essentially examine the effect of some properties of the quotation source (\hbox{e.g.} news outlet {vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus).
+Some propose diffusion-transformation models, but the very cognitive features which may determine or, at least, influence these transformations, are nonetheless overlooked; which may appear to be relatively simplistic from a cognitive viewpoint.
 

### 2014-12-12 brains-protocol.tex
introduced examples + revamped the beginning of sec. 2 + minor corrections

@@
 \begin{quote}
-\cam{do we have an excerpt of quotations here, could we even feature a few prototypical examples? Obama said that ``...'' }
+The Bank of England said, ``these operations are designed to address funding pressures over quarter-end.''
 \end{quote}
-Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words): \cam{``...''\ra``...' ''}
+Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
+We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{designed} to address funding pressures over quarter-end.''\end{quote}
 A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).

### 2014-12-12 brains-protocol.tex
Rephrasing of the filtering in 'manual inspection of dataset'

@@
 Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
-We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{designed} to address funding pressures over quarter-end.''\end{quote}
+We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
 A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
@@
 
-Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than i expected'', ``did that just happen'', as well as many simple expletive-based sentences).
-Their presence is may largely owe to random variations around a casual expression, while we are interested in transformations of news-related quotes linked to an original, identifiable utterance.
-We thus filter out all clusters lasting more than 80 days, or featuring quotes having less than 5 words \cam{not sure this is what you meant [but then the filter may be a bit asymmetric, oder?]} (or not written in English).
-We eventually keep 19,621 clusters (out of 71,568; i.e.
-27.4\%), 60,246 quotes (out of 310,457; i.e.
-19.4\%) and $\sim$838k occurrences (out of 8.12m, i.e.
-10.3\%).
+Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
+Their presence is largely due to random variations around a casual expression, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
+We thus filter out all clusters lasting more than 80 days, and all quotes less than 5 words long \seb{it is asymmetrical...} (as well as quotes not written in English).
+We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
 The significantly larger loss in occurrences indicates that, on average, the clusters we lose feature more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'', we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.

### 2014-12-12 brains-introduction.tex
introduction / hypotheses mods

@@
 
-The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining.
-These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part,
-%Broadly speaking, this type of research program is bound to
- investigate phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
- both at the individual and social levels.  Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.
+The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling; from social cognition to data mining.
+These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part, investigate phenomena pertaining to both cognitive science and social science, both at the individual and social levels.
 
-%Broadly speaking, this type of research program is bound to investigate phenomena pertaining to both cognitive science and social science, and aims to understand the processing, transmission, and evolution of information both at the individual and social levels.
-The relevant literature features two main streams, depending on whether the focus lies on social dynamics or cognitive processes.
+Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.
+A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
+It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the former's counterpart outside the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
+\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: \begin{quotation}A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
+The producer and the user of a mental representation are one and the same person.
+A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.\end{quotation}}
 
-A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
-It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program initiated by \citet{sper:expl}, which %recently attracted significant attention, and
-% focuses on the notion of \emph{representation}, 
-links %the cognitive science 
-the concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e.
-in cultural artifacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
-\begin{quotation}
-A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
-The producer and the user of a mental representation are one and the same person.
-A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.
-\end{quotation}}
-
-Here, one of the core claims consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
+Here, one of the core claims consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
 Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
-Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
-Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.
+Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
+Despite some recent modeling attempts (\hbox{e.g.} \citet{Claidiere07}), the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.
 
@@
 These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
-They nonetheless show us the interest of using these records --~given the already significant and rapidly growing importance of our online interactions~--
+They nonetheless show us the interest of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
 %and, %more broadly, large textual corpora \cam{google n-grams??} 
-to dramatically improve the empirical understanding of individual-level processes of cultural evolution.
+in radically improving the empirical understanding of (individual-level) cultural evolution processes.
 
 \bigskip
-Stepping back, we observe a gap between, on one side, empirical studies of the diffusion dynamics in a social system and, on the other side, more theoretical works  focused on knowledge transformation processes.
+Stepping back, we thus observe a gap between, on one side, empirical studies of diffusion dynamics in social systems and, on the other side, more theoretical works  focused on knowledge transformation processes.
 %--- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.
-Our research lies at the intersection of these two programs, aiming to shed light on microscopic information transformation processes by leveraging the empirical wealth of (in vivo) social diffusion phenomena.
-We hope to describe some reformulation processes within a large distributed social system such as blogspace.
-More precisely, we aim to empirically describe how some specific type of public representation may possibly be altered by bloggers when they try to reproduce that representation.
+Our research lies at the intersection of these two programs, aiming to shed light on micro-level information transformation by leveraging the empirical wealth of (in vivo) social diffusion phenomena.
+More precisely, we hope to describe reformulation processes within a large distributed system such as blogspace; showing how some specific types and features of public representations may be altered by bloggers when they freely reproduce that representation.
 

### 2014-12-12 brains-introduction.tex
introduction mods

@@
 We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
-More generally, we contend that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards building empirically realistic models of cultural evolution.
+We moreover contend that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards building empirically realistic models of cultural evolution.
 
-The next section (Sec.~\ref{sec:related}) describes our hypotheses along with the relevant state-of-the-art on this \cam{psycholinguistic} matter.
+The next section (Sec.~\ref{sec:related}) describes our hypotheses along with the relevant state-of-the-art on this psycholinguistic matter.
 In Sec.~\ref{sec:protocol}, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.

### 2014-12-12 brains-introduction.tex
Word frequency is covered

@@
 The practical study of the transformation of public representations has emerged only recently.
-For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07}.
-
+For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07}\marginnote{C: on n'a que C \& S 2007 et c'est la bérézina de ce côté-là, ou il y a d'autres trucs?
+dunno personally}.
 Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
@@
 They essentially examine the effect of some properties of the quotation source (\hbox{e.g.} news outlet {vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus).
-Some propose diffusion-transformation models, but the very cognitive features which may determine or, at least, influence these transformations, are nonetheless overlooked; which may appear to be relatively simplistic from a cognitive viewpoint.
+Some diffusion-transformation models have been proposed, yet the very cognitive features which may determine or, at least, influence these transformations, are overlooked; which may appear to be relatively unsatisfying from a cognitive viewpoint.
 
@@
 This field is well developed and details the impact that classical psycholinguistic variables such as word frequency (see \citet{Yonelinas02} for a review), age-of-acquisition \citep{Zevin02}, number of phonemes or number of syllables (see for instance \citet{Rey98} and \citet{nick-diss}), have in this type of task.
-\rk{add a ref for word frequency}
 
-Less classical linguistic variables, based on the study of semantic network properties, have recently appeared as an empirical investigation field, after having been heavily discussed around the notion of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
-Let us mention two interesting and recent studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
-First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the Latin alphabet.
-The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
+Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
+Let us mention two interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
+First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin
+alphabet.
+The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (PageRank) in a language-wide semantic network built from external word association data.
 A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
-
 \rk{add \url{https://www.sciencedirect.com/science/article/pii/S0893608012000330}, \url{http://research.clps.brown.edu/austerweil/pdfs/papers/randomWalkNips2012.pdf}, \url{http://link.springer.com/article/10.3758/s13421-013-0312-y} to that picture}
 
-\TB{On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
+On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
 On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal \CN.
@@
 The jury is still out as to whether reformulation alteration, i.e. spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
-Our paper additionally sheds light on this debate.}
-\rk{Show the link with epidemiology, and fitness: oddness is a kind of fitness}
-
-\smallskip
+We also aim here at shedding some light on this debate, considering oddness is a kind of fitness.

### 2014-12-12 brains-introduction.tex
related work

@@
 It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the former's counterpart outside the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
-\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: \begin{quotation}A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
+\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
 The producer and the user of a mental representation are one and the same person.
-A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.\end{quotation}}
+A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}''.}
 
@@
 % ============================
-\section{Hypotheses} % =====
+\section{Related work} % =====
 % ============================

### 2014-12-12 brains-protocol.tex
methods [introduction]

@@
 %aiming to stay compatible with macroscopic-level studies of these public representations.
-Quotations appeared to be a perfect candidate to propose a first \emph{in vivo} measure of low-level cognitive bias in a reformulation task. % as public representations.
+In practice, quotations appeared indeed to be a perfect candidate for designing a first \emph{in vivo} measure of low-level cognitive bias in a reformulation task. % as public representations.
 First, they are usually cleanly delimited by quotation marks %(and often with HTML markup in web pages), 
@@
 
-We could therefore examine the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
-To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+We therefore examine transformation processes by comparing the substituted and substituting words.
+To keep the analysis palatable, we focused on quotation transformations consisting of the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
-
 The next subsections describe the dataset and measures we used to assess this cognitive bias.

### 2014-12-12 brains-protocol.tex
methods [substitution]

@@
 
-Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs on the 21st, 22nd and 23rd of January.
-If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 21st based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
-But what about the second and third occurrences of $q'$, on the 22nd and 23rd?
+Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs on Jan 21, 22 and 23.
+If $q$ was sufficiently prominent when $q'$ first appeared, we can reasonably assume that the first author of $q'$ on the 21st based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
+But what about the second and third occurrences of $q'$, on Jan 22 and 23?
 Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place) 
@@
 
-To settle this question we bin the quote occurrences into fixed \emph{bins} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
-Then when a quotation $q$ appears in bin $n$, it is counted as a substitution from each quote $q^*$ in the preceding bin ($n - 1$) from which it differs by only one word.
+To settle this question we group quote occurrences into fixed bins spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
+When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution from each quote $q$ in the preceding bin $t$ from which it differs by only one word.
 If no quote in the preceding bin can qualify as a source in a substitution (i.e.
-$q$ differs from all the quotes in the preceding bin by more than one word), the occurrence of $q$ is not considered to be an instance of substitution.
+$q'$ differs from all the quotes in the preceding bin by more than one word), the occurrence of $q'$ is not considered to be an instance of substitution.
 Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.

### 2014-12-12 brains-protocol.tex
results

@@
 To settle this question we group quote occurrences into fixed bins spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
-When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution from each quote $q$ in the preceding bin $t$ from which it differs by only one word.
-If no quote in the preceding bin can qualify as a source in a substitution (i.e.
-$q'$ differs from all the quotes in the preceding bin by more than one word), the occurrence of $q'$ is not considered to be an instance of substitution.
-Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
-
-The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
+When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote $q$ of the preceding bin $t$ by only one word.
+If not, $q'$ is not considered to be an instance of substitution.
+We assume here that quotes are based on the most frequent quote of the preceding bin (\hbox{vs.} all the other quotes in that bin).
+These assumptions are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, the criterion of the most frequent quote in the preceding bin may be replaced with the most frequent quote overall, or the oldest quote; time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
 These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
@@
 
-We now need to reduce possible false positives.
-Indeed, so far we have been forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
-We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding bin.
-We assume that authors will have encountered the most frequent quote in a given bin (\hbox{vs.} all the other quotes in that bin).
-Thus, restricting substitutions to the most frequent quote ensures that we only detect substitutions that really occurred, greatly reducing the number of false positives.
+Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
 
-Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
+%We now need to reduce possible false positives.
+%Indeed, so far we have been forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
+%We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding bin.
+%We assume that authors will have encountered the most frequent quote in a given bin (\hbox{vs.} all the other quotes in that bin).
+Thus, Restricting substitutions to the most frequent quote ensures that we only detect substitutions that really occurred, greatly reducing the number of false positives.
+
+%Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
 In bin~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in bin~4).
 This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
-Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.
+Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.\marginnote{C: on peut mettre ce paragraphe dans la légende de la figure d'ailleurs}
 
-With the substitution detection model thus refined, we build two main observables for each word feature.
+\section{Results}\label{sec:results}
+
+With the substitution detection model thus defined, we may now define a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.
+To this end, we build two main observables for each word feature.
 First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution under this condition.
@@
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
-
-\section{Results}\label{sec:results}
 

### 2014-12-12 brains-protocol.tex
Redo list of psycholinguitic indices + refs

--- 2014-12-12 brains-protocol.tex
@@
 %aiming to stay compatible with macroscopic-level studies of these public representations.
-In practice, quotations appeared indeed to be a perfect candidate for designing a first \emph{in vivo} measure of low-level cognitive bias in a reformulation task. % as public representations.
+Quotations appeared to be a perfect candidate to propose a first \emph{in vivo} measure of low-level cognitive bias in a reformulation task. % as public representations.
 First, they are usually cleanly delimited by quotation marks %(and often with HTML markup in web pages), 
@@
 
-We therefore examine transformation processes by comparing the substituted and substituting words.
-To keep the analysis palatable, we focused on quotation transformations consisting of the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+We could therefore examine the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
+To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
+
 The next subsections describe the dataset and measures we used to assess this cognitive bias.
@@
 We now introduce some of the most classical psycholinguistic measures on words.
-\rk{Add some bibliography about those features' known effects}
 
 \begin{itemize}
-    \item \textbf{Word frequency}: the frequency at which words appear in our dataset, \rk{add ref on why it's important}
-    \item \textbf{Age of Acquisition}: the average age at which words are learned, obtained from~\citet{kuperman12},
-    \item The average \textbf{Number of Phonemes} for all pronunciations of a word, obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98},\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.}
-    \item The average \textbf{Number of Syllables} for all pronunciations of a word, also obtained from the CMU Pronouncing Dictionary,
-    \item The average \textbf{Number of Synonyms} for all meanings of a word, obtained from WordNet~\citep{WordNet10}. \rk{add ref on why it's important}
+    \item \textbf{Word frequency}: the frequency at which words appear in our dataset, known to be relevant for both recognition and recall~\citep{gregg1976word},
+    \item \textbf{Age of Acquisition}: the average age at which words are learned (obtained from~\citet{kuperman12}), known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
+    \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
+    \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from WordNet~\citep{WordNet10}) as a proxy to word polysemy.
 \end{itemize}

### 2014-12-12 brains-protocol.tex
clarified 4.1

@@
 
-For a given feature $\phi$, the protocol lets us compute substitution \emph{susceptibilities} for each feature value $f$.
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
@@
 $\sfrac{s_w}{p_w}$.
+In other words, it measures how often a subtitutable word $w$ gets actually substituted.
 
-Now averaging over all words such that $\phi(w) = f$ (only taking into account words that are substituted at least once), we obtain the mean susceptibility for the feature value $f$:
+Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$ (only taking into account words that are substituted at least once):
 \footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
@@
 
-This measure focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation -- if any -- will be substituted.
+This measure focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
 

### 2014-12-12 brains-protocol.tex
clarified 4.1b

@@
 
-This measure focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
+Put shortly, susceptibility focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
+
+\medskip
+Results are gathered on Fig.~\ref{fig:feature-susceptibilities}.
+
+\rk{Comment results}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
+    \label{fig:feature-susceptibilities}
+\end{figure*}
 
@@
 
+\medskip
+Results are gathered on Fig.~\ref{fig:feature-variations}.
+
 \rk{Comment results}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
-    \label{fig:feature-susceptibilities}
-\end{figure*}
 

### 2014-12-12 brains-protocol.tex
removed comment 3.1

@@
 Their presence is largely due to random variations around a casual expression, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
-We thus filter out all clusters lasting more than 80 days, and all quotes less than 5 words long \seb{it is asymmetrical...} (as well as quotes not written in English).
+We thus filter out all clusters lasting more than 80 days, and all quotes less than 5 words long %\seb{it is asymmetrical...}
+ (as well as quotes not written in English).
 We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).

### 2014-12-12 brains-protocol.tex
various

@@
     \item \textbf{Word frequency}: the frequency at which words appear in our dataset, known to be relevant for both recognition and recall~\citep{gregg1976word},
-    \item \textbf{Age of Acquisition}: the average age at which words are learned (obtained from~\citet{kuperman12}), known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
+    \item \textbf{Age of Acquisition}: the average age at which words are learned (obtained from~\citet{Kuperman12}), known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
     \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
@@
 To settle this question we group quote occurrences into fixed bins spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
-When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote $q$ of the preceding bin $t$ by only one word.
+When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote $q$ of the preceding bin $t$ (or a substring thereof) by only one word.
 If not, $q'$ is not considered to be an instance of substitution.
@@
 
-Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$ (only taking into account words that are substituted at least once):
+Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$% (only taking into account words that are substituted at least once)
+:
 \footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.

### 2014-12-12 brains-protocol.tex
protocol
consulting features

@@
 We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose feature more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'', we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'', we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
 We could manually check on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
@@
 
-We also considered grammatical types within quotations by detection of \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology~\citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and adverbs.
+%We also considered grammatical types within quotations by detection of \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology~\citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and adverbs.
 The results were however extremely similar across the various categories, exhibiting no specific effect of words belonging to different POS categories.
@@
 	}
-	\caption{Temporal binning of quotation families}
+	\caption{{\bf Temporal binning of quotation families.} In bin~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in bin~4).
+This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
+Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.}
     \label{fig:substitution-temporal-binning}
@@
 If not, $q'$ is not considered to be an instance of substitution.
-We assume here that quotes are based on the most frequent quote of the preceding bin (\hbox{vs.} all the other quotes in that bin).
+%We assume here that quotes are based on the most frequent quote of the preceding bin (\hbox{vs.} all the other quotes in that bin).
 These assumptions are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, the criterion of the most frequent quote in the preceding bin may be replaced with the most frequent quote overall, or the oldest quote; time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
@@
 Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+In practice, from the 838k initial occurrences spread into 19,621 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 2,873 real substitutions obeying to this model.
+From these substitutions we remove those featuring stop words, minor spelling changes (e.g.
+center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 476 valid substitutions.
 
@@
 %Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
-In bin~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in bin~4).
-This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
-Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.\marginnote{C: on peut mettre ce paragraphe dans la légende de la figure d'ailleurs}
 
@@
 
-With the substitution detection model thus defined, we may now define a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.
+With the substitution detection model thus defined, we may now formulate a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.
 To this end, we build two main observables for each word feature.

### 2014-12-14 brains-protocol.tex
rewriting expressions in 4.2, puzzling dependency on f in H_0

@@
 
-We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
-Word substitution susceptibility is computed as the ratio of the number $s_w$ of times a word is substituted to the number $p_w$ of times that word appears in a substitutable position, i.e.
-$\sfrac{s_w}{p_w}$.
+We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on that word or on another one.
+Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, \hbox{i.e.} $\sfrac{s_w}{p_w}$.
 In other words, it measures how often a subtitutable word $w$ gets actually substituted.
@@
 
-\subsection{Alteration}
+\subsection{Variation}
 
@@
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
-$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \phi(\wstart) = f \right\rbrace}$$
+$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
 
 We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_0$ had been chosen randomly from the pool of free association words.
-The new quantity under $\mathcal{H}_0$ is:
-\footnote{Note that $\phi(\warrival_0)$ is in fact a constant in this averaging, since by definition $\warrival_0$ does not depend on $\wstart$.}
-$$\Delta_{\phi}^0 (f) = \left< \phi({\warrival_0}) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival_0) | \phi(\wstart) = f \right\rbrace}$$
+Since $\phi(\warrival_0)$ is in fact a constant in the above averaging (by definition $\warrival_0$ does not depend on $\wstart$) and $\phi(\wstart)=f$, 
+the baseline variation under $\mathcal{H}_0$ may be rewritten:
+$$\Delta_{\phi}^0 (f) = \left<\phi\right>-f$$
 
-We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word}, i.e.
-an arrival word chosen among semantically plausible though still random words.\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.}
-The results were not qualitatively changed with this second null hypothesis, so for the sake of clarity we chose to present the results using the simpler~$\mathcal{H}_0$.
+We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.}
+The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.
 
-Using this method we obtain the mean variation of feature for each start feature value, and can compare the variations to a situation where arrival words are chosen randomly.
+Using this method we obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.
 This gives us a fine-grained view of how word features evolve upon substitution.

### 2014-12-14 brains-protocol.tex
rewriting expressions in 4.2, puzzling dependency on f in H_0 / clearer

@@
 
-We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_0$ had been chosen randomly from the pool of free association words.
-Since $\phi(\warrival_0)$ is in fact a constant in the above averaging (by definition $\warrival_0$ does not depend on $\wstart$) and $\phi(\wstart)=f$, 
-the baseline variation under $\mathcal{H}_0$ may be rewritten:
+We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
+In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore) and $\phi(\wstart)=f$,  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
 $$\Delta_{\phi}^0 (f) = \left<\phi\right>-f$$

### 2014-12-14 brains-protocol.tex
Merge

--- 2014-12-12 brains-protocol.tex
@@
 % \rk{we're in fact using the unweighed version of the network. Why?}
-\new{Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.}
+Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
 This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyse is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
@@
 \begin{itemize}
-    \item \textbf{Centrality} $k$, initially measured by the number of incoming edges to a given node, i.e.
+    \item \textbf{Degree centrality}, measured by the number of incoming edges to a given node, i.e.
 the number of cues for which a given word is triggered as an association, which strongly relates to word polysemy.
-    However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
+    An additional measure for centrality is node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
+    Both measures have 
+    %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
 Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
-    \item \textbf{Clustering coefficient} $c$, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs.
+    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network~\citep{free:set}.
+    This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.
+    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs.
 possible edges between a node's neighbours \cite{watt-coll}.
     We compute the clustering coefficient on the undirected version of the FA network; we thus measure if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
-    \item \textbf{Betweenness coefficient} $b$, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network~\citep{free:set}.
-    More technically, it corresponds to the normalized number of shortest paths connecting dyads which pass through that node; the higher the coefficient, the more important that node is in ensuring the connectedness of the rest of the network.
-    This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.
 \end{itemize}
@@
 
-Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs on Jan 21, 22 and 23.
-If $q$ was sufficiently prominent when $q'$ first appeared, we can reasonably assume that the first author of $q'$ on the 21st based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
-But what about the second and third occurrences of $q'$, on Jan 22 and 23?
+Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs on the 21st, 22nd and 23rd of January.
+If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 21st based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
+But what about the second and third occurrences of $q'$, on the 22nd and 23rd?
 Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place) 
@@
 
-%We now need to reduce possible false positives.
-%Indeed, so far we have been forced to detect substitutions with a very permissive model to make sure chains of substitutions could be extracted (at the expense of possible false positives, the only effect of this being additional noise in the results, making our conclusions more conservative), here we can afford to detect substitutions more precisely.
-%We thus add the constraint that substitutions can only stem from the most frequent quote in the preceding bin.
-%We assume that authors will have encountered the most frequent quote in a given bin (\hbox{vs.} all the other quotes in that bin).
-Thus, Restricting substitutions to the most frequent quote ensures that we only detect substitutions that really occurred, greatly reducing the number of false positives.
-
 %Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.

### 2014-12-14 brains-protocol.tex
Merge

@@
 
-\section{Results}\label{sec:results}
-
 \subsection{Susceptibility}
 
-We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on the considered word or on another.
-Word substitution susceptibility is computed as the ratio of the number $s_w$ of times a word is substituted to the number $p_w$ of times that word appears in a substitutable position, i.e.
-$\sfrac{s_w}{p_w}$.
+We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on that word or on another one.
+Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, \hbox{i.e.} $\sfrac{s_w}{p_w}$.
 In other words, it measures how often a subtitutable word $w$ gets actually substituted.
@@
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
-$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \phi(\wstart) = f \right\rbrace}$$
+$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
 
-We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_0$ had been chosen randomly from the pool of free association words.
-The new quantity under $\mathcal{H}_0$ is:
-\footnote{Note that $\phi(\warrival_0)$ is in fact a constant in this averaging, since by definition $\warrival_0$ does not depend on $\wstart$.}
-$$\Delta_{\phi}^0 (f) = \left< \phi({\warrival_0}) - \phi(\wstart) \right>_{\left\lbrace (\wstart,\warrival_0) | \phi(\wstart) = f \right\rbrace}$$
+We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
+In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore) and $\phi(\wstart)=f$,  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
+$$\Delta_{\phi}^0 (f) = \left<\phi\right>-f$$
 
-We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word}, i.e.
-an arrival word chosen among semantically plausible though still random words.\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.}
-The results were not qualitatively changed with this second null hypothesis, so for the sake of clarity we chose to present the results using the simpler~$\mathcal{H}_0$.
+We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.}
+The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.
 
-Using this method we obtain the mean variation of feature for each start feature value, and can compare the variations to a situation where arrival words are chosen randomly.
+Using this method we obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.
 This gives us a fine-grained view of how word features evolve upon substitution.

### 2014-12-15 brains-protocol.tex
Finish network-features rework. Correlation talk still to be done

@@
 \begin{itemize}
-    \item \textbf{Degree centrality}, measured by the number of incoming edges to a given node, i.e.
-the number of cues for which a given word is triggered as an association, which strongly relates to word polysemy.
-    An additional measure for centrality is node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
-    Both measures have 
+    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as an association, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
+    In the present case the two measures are quasi-perfectly correlated, and measure polysemy by taking into account the structure of the entire network.
     %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
@@
     This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.
-    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes, and defined as the ratio between the number of actual vs.
-possible edges between a node's neighbours \cite{watt-coll}.
-    We compute the clustering coefficient on the undirected version of the FA network; we thus measure if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
+    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual vs.
+possible edges between a node's neighbors.}
+    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
 \end{itemize}

### 2014-12-15 brains-protocol.tex
Fuse figures b and c on subtitution binning, fix figure references in tex

@@
 	\hfill \\
-    \subfloat[Binned quotation family]{
-	    \def\svgwidth{\linewidth}
-	    \small
-	    \input{images/substitution-temporal-binning-b.pdf_tex}
-	    \label{fig:substitution-temporal-binning-b}
-	}
-	\hfill \\
-	\subfloat[XXXX]{
+	\subfloat[Binned quotation family with majority rule]{
 	    \def\svgwidth{\linewidth}
@@
 
-Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
 In practice, from the 838k initial occurrences spread into 19,621 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 2,873 real substitutions obeying to this model.
@@
     The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
-    The fact that all the curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range, which is consistent with the evolution observed in Figure~\ref{fig:timebags-evolution}.}
+    The fact that all the curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.}
     \label{fig:feature-variations}

### 2014-12-15 brains-protocol.tex
Minor phrasings in psycholinguistic and network features

@@
     \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
-    \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from WordNet~\citep{WordNet10}) as a proxy to word polysemy.
+    \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from WordNet~\citep{WordNet10}) as an \emph{a priori} indicator of how easy it would be to replace a word.
 \end{itemize}
@@
 % \rk{we're in fact using the unweighed version of the network. Why?}
-Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
+%Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
+Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
 This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyse is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
@@
 \begin{itemize}
-    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as an association, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
+    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
     In the present case the two measures are quasi-perfectly correlated, and measure polysemy by taking into account the structure of the entire network.
@@
 Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
-    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node tends to connect otherwise remote areas of the network~\citep{free:set}.
+    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{free:set}.
     This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.

### 2014-12-15 brains-protocol.tex
Redo correlation talk

@@
 
-\rk{Correlation talk needs to be redone for new set of features}
-
-Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies and is also usually correlated to many of the other variables.
-This relates to an ongoing debate suggesting that age of acquisition encodes a variety of phenomena, difficult to disentangle from more specific phenomena which could be captured by more independent variables~\CN{}.
-Here however, as can be seen in Figure~\ref{fig:feature-corrs-initial}, age of acquisition has a relatively low correlation to the other variables (absolute value not above $0.42$ if we exclude the centrality measures), leading us to keep the variable in the rest of the analysis.
-
-Number of phonemes and number of syllables naturally exhibit a strong linear correlation ($0.83$).
+Number of phonemes and number of syllables naturally exhibit a strong linear correlation ($0.8$).
 The analysis showed a better prediction effect of number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to focus the presented results on the former only.
 
-Frequency and number of meanings both have relatively low levels of correlation to the other variables; we therefore also keep them in the rest of the analysis.
+Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
+Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see in Figure~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
+
+Frequency and number of synonyms both have relatively low levels of correlation to the other variables (excluding again the network properties); we therefore also keep them in the rest of the analysis.
 
 \bigskip
-Network properties, on the other hand, are strongly dependent on one another.
-As mentioned earlier, word degree and word pagerank have a very strong correlation ($0.89$) and, degree being generally more correlated to other variables, we chose to remove this variable from the results presented.
-Finally betweenness centrality also exhibits strong correlation levels to the other network properties ($0.62$, $0.64$, and $0.72$ in absolute value), leading us to drop this final feature due to its redundancy.
+Network centrality properties, on the other hand, are strongly dependent on one another.
+As mentioned earlier, degree centrality and pagerank have a very strong correlation ($0.85$), and are also redundant with betweenness centrality (with correlation levels at $0.75$ and $0.68$ respectively).
+Furthermore, the three variables are also strongly related to age of acquisition, which leads us to keep the latter as the sole indicator for centrality.
+
+Conversely, clustering coefficient exhibits low correlation levels with all the variables we kept (maximum absolute value $0.38$), leading us to include it in the rest of the analysis.
 

### 2014-12-15 brains-introduction.tex
Add model references to intro

@@
 The practical study of the transformation of public representations has emerged only recently.
-For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07}\marginnote{C: on n'a que C \& S 2007 et c'est la bérézina de ce côté-là, ou il y a d'autres trucs?
-dunno personally}.
+For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07,claidiere2014darwinian}\marginnote{S: il y a beaucoup de meta-études philosophiques où ils reprennent des études empiriques, mais toujours peu de modèles.
+Peut-être que tu veux vérifer les deux que j'ai réjouté?
+il y a aussi éventuellement \citet{claidiere2009imitation} avec un peu de modèle, \citet{maccallum2012evolution} plus empirique, et \citet{claidiere2012effect} qui le commente}.
 Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
@@
 The jury is still out as to whether reformulation alteration, i.e. spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
-We also aim here at shedding some light on this debate, considering oddness is a kind of fitness.
+We also aim here at shedding some light on this debate, considering oddness as a dimension of the purported fitness of utterances.

### 2014-12-15 brains-introduction.tex
Add references for awkward/common jury

@@
 
-Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
+Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading} for an example of such models).
 Let us mention two interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
@@
 On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
-On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal \CN.
-On the other hand, when the task consists in remembering a specific list of items, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible \CN.
+On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
+On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible (see again~\citet{Yonelinas02} for a review).
 The jury is still out as to whether reformulation alteration, i.e. spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.

### 2014-12-15 brains-introduction.tex
Update state of the art on network studies

@@
 Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading} for an example of such models).
-Let us mention two interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
+Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
 First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin
 alphabet.
-The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (PageRank) in a language-wide semantic network built from external word association data.
-A second psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
-\rk{add \url{https://www.sciencedirect.com/science/article/pii/S0893608012000330}, \url{http://research.clps.brown.edu/austerweil/pdfs/papers/randomWalkNips2012.pdf}, \url{http://link.springer.com/article/10.3758/s13421-013-0312-y} to that picture}
+The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
+\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, i.e.
+the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
+A third psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
+\citet{nelson2013activation}, finally, show the importance of clustering coefficient in a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free assocation).
 

### 2014-12-15 brains-introduction.tex
finished correcting 4.2 with new formulas / legend of fig. 5 accordingly / addressed notes / other minor mods

@@
 The practical study of the transformation of public representations has emerged only recently.
-For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07,claidiere2014darwinian}\marginnote{S: il y a beaucoup de meta-études philosophiques où ils reprennent des études empiriques, mais toujours peu de modèles.
+For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citet{Claidiere07,claidiere2014darwinian} as well as a hybrid empirical-theoretical protocol in \citet{maccallum2012evolution}) \seb{il y a beaucoup de meta-études philosophiques où ils reprennent des études empiriques, mais toujours peu de modèles.
 Peut-être que tu veux vérifer les deux que j'ai réjouté?
-il y a aussi éventuellement \citet{claidiere2009imitation} avec un peu de modèle, \citet{maccallum2012evolution} plus empirique, et \citet{claidiere2012effect} qui le commente}.
+il y a aussi éventuellement \citet{claidiere2009imitation} avec un peu de modèle, \citet{maccallum2012evolution} plus empirique, et \citet{claidiere2012effect} qui le commente}\cam{Bien vu MacCallum!
+Mais attends, quelles sont les \emph{2} références ajoutées?
+je vois Claidière 2014, en sus du précédent Claidière 2007; mais les autres sont à choisir dans la parenthèse...
+en tout cas, je penche pour citer Claidière 2007, 2014 et en sandwich MacCallum, avec une précision, cf.
+phrase précédente.
+If it suits you, you may kill these commentaries altogether!}
 Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
@@
 
-Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading} for an example of such models).
+Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
 Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.

### 2014-12-15 brains-protocol.tex
Note error in figures
finished correcting 4.2 with new formulas / legend of fig. 5 accordingly / addressed notes / other minor mods

@@
 
-Number of phonemes and number of syllables naturally exhibit a strong linear correlation ($0.8$).
-The analysis showed a better prediction effect of number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to focus the presented results on the former only.
+The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
+The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
 
 Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
-Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see in Figure~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
+Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Figure~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
 
@@
 \subsection{Substitution model}
+
+\seb{Les figures des résultats sont en fait pour $\Delta t = 1$ jour, pas 2. Sauf contre-indication, je corrige les figures et le texte}
+\cam{Oui très bien, n'hésite pas. Super pour les figures, je sais pas si l'inversion graphique q en haut q' en bas est facile à réaliser?}
 
@@
     \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
+    Each feature exhibits a specific and significant pattern favoring either high- or low-valued words for substitution.}
     \label{fig:feature-susceptibilities}
@@
 
-Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival) - \phi(\wstart)$.
+Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
-$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
+$$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
 
 We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
-In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore) and $\phi(\wstart)=f$,  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
-$$\Delta_{\phi}^0 (f) = \left<\phi\right>-f$$
+In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
+$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ vs.
-average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
-    The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
-    The fact that all the curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.}
+    \caption{\textbf{Feature variation upon substitution:} average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
+    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
+    }
     \label{fig:feature-variations}

### 2014-12-16 brains-introduction.tex
Fix modeling references in intro

--- 2014-12-15 brains-introduction.tex
@@
 The practical study of the transformation of public representations has emerged only recently.
-For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07,claidiere2014darwinian}\marginnote{S: il y a beaucoup de meta-études philosophiques où ils reprennent des études empiriques, mais toujours peu de modèles.
-Peut-être que tu veux vérifer les deux que j'ai réjouté?
-il y a aussi éventuellement \citet{claidiere2009imitation} avec un peu de modèle, \citet{maccallum2012evolution} plus empirique, et \citet{claidiere2012effect} qui le commente}.
+For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citet{Claidiere07,claidiere2014darwinian} as well as a hybrid empirical-theoretical protocol in \citet{maccallum2012evolution}).
 Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.

### 2014-12-16 brains-protocol.tex
Fix number of days for binning in figures and text

@@
 
-\seb{Les figures des résultats sont en fait pour $\Delta t = 1$ jour, pas 2. Sauf contre-indication, je corrige les figures et le texte}
-\cam{Oui très bien, n'hésite pas. Super pour les figures, je sais pas si l'inversion graphique q en haut q' en bas est facile à réaliser?}
-
 %\subsubsection{Temporal binning}
@@
 
-Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs on the 21st, 22nd and 23rd of January.
-If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 21st based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
-But what about the second and third occurrences of $q'$, on the 22nd and 23rd?
+Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs twice on the 20th and once on the 21st of January.
+If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 20th based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
+But what about the second and third occurrences of $q'$, on the 20th and 21st?
 Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place) 
@@
 
-To settle this question we group quote occurrences into fixed bins spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
+To settle this question we group quote occurrences into fixed bins spanning $\Delta t$ days (1 day in the implementation), each one representing a unit of time evolution.
 When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote $q$ of the preceding bin $t$ (or a substring thereof) by only one word.
@@
 These assumptions are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, the criterion of the most frequent quote in the preceding bin may be replaced with the most frequent quote overall, or the oldest quote; time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
-These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
+These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
 We identified and implemented eleven other such models, and they all yielded essentially the same results.
 
-Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
+Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.
 In practice, from the 838k initial occurrences spread into 19,621 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 2,873 real substitutions obeying to this model.

### 2014-12-23 brains-protocol.tex
Minor note to remember to crop figures
First comments on results for susceptibility

@@
 \medskip
-Results are gathered on Fig.~\ref{fig:feature-susceptibilities}.
+The results for this observable, gathered on Fig.~\ref{fig:feature-susceptibilities}, first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
+Indeed, susceptibility goes from nearly 0.4 for low-frequency words down to 0 for very high-frequency words.
 
-\rk{Comment results}
+Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
+These high-valued words also have low occurrence frequencies, which corresponds to the slope change in the frequency curve between extreme low frequency words and mid- to high-frequency words.
+\rk{check this with a plot of freq vs.
+clustering and freq vs.
+AoA, or simply plot of frequency for the high-valued words for AoA and CC.}
+This class of words corresponds to \TB{highly specialized words? domain-specific? a hard limit on too complicated words?}
+
+Finally the results for Clustering coefficient and Number of synonyms, despite some tendencies, do not allow us to conclude either to an effect or no effect of those features on substitution susceptibility.
 

### 2014-12-23 brains-protocol.tex
Start comments on results for feature variation

@@
 
+\rk{Might want to globally add $y = x$? Defines areas above/below $y = x$ and areas above/below $H_0$}
+
+\begin{itemize}
+    \item Clustering: flat, no dependency. But negative bias, which corresponds to literature on word production.
+    \item AoA: global negative bias, also consistent with literature. Question: on one side, effect w.r.t. $H_0$ is weaker for higher AoA, but if you compare to $y = x$, it's not that clear.
+    \item Frequency: very strong positive bias, ok.
+Now extreme high-frequency words have a different pattern (maybe a third category at the opposite of the very low-frequency words?): even if above $H_0$, they're brought quite lower than their original value (complexified?)
+    \item Number of phonemes and Number of synonyms: expanding behavior (but check with $y = x$): lower goes to lower, higher goes to higher.
+\end{itemize}
+
 \begin{figure*}[!th]

### 2014-12-23 brains-protocol.tex
4.1 minor

@@
 \medskip
-The results for this observable, gathered on Fig.~\ref{fig:feature-susceptibilities}, first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
+Results are gathered on Fig.~\ref{fig:feature-susceptibilities}.
+They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
 Indeed, susceptibility goes from nearly 0.4 for low-frequency words down to 0 for very high-frequency words.
+As a reminder, this value of 0.4 means that, this kind of low-frequency words, when present in a quote undergoing a substitution, are indeed the one being substituted 40\% of the time on average.
 
-Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
+Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no significant effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
 These high-valued words also have low occurrence frequencies, which corresponds to the slope change in the frequency curve between extreme low frequency words and mid- to high-frequency words.

### 2014-12-28 brains-protocol.tex
Add notice on human/automatic reproduction

@@
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+\rk{Recognize that this can be a sub-sample of real human manual Ctrl-C/V, since we don't have access to successful manual copy by the brain.
+No need to suppose this sample is representative: we focus on when there is a transformation, and that's the process we study (since we have no access to the difference between manual and automatic).
+We also therefore suppose that manual copy and manual transformation are potentially different processes, and can be studied separately profitably.}
 

### 2014-12-28 brains-protocol.tex
Alteration and perfect copy are different - justify

--- 2014-12-23 brains-protocol.tex
@@
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration --e.g.
+drafting before publishing, double-checking sources, proof-reading--, and may not be representative of the cognitive processes at work during alteration.
+The two situations are different enough to be studied separately, and we focus here on the latter.
 

### 2014-12-30 brains-protocol.tex
4.1 new scenario / 4.2 transition

@@
 
-Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no significant effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
-These high-valued words also have low occurrence frequencies, which corresponds to the slope change in the frequency curve between extreme low frequency words and mid- to high-frequency words.
+The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
+For each of these features, it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
+If these variables have an effect, it is by no means as strong as it is for Word frequency.
+This is remarkably clear for Clustering coefficient and Number of synonyms, where susceptibility values remain within quite small intervals (respectively [0.09-0.13] and [0.11-0.15]).
+We may notice a somewhat marked effect for the highest values of Age of acquisition and Number of phonemes, where the mean susceptibility is almost twice higher than for the rest (respectively 0.19 vs 0.13, and 0.21 vs 0.12).
+Keeping in mind the poor statistical significance of this effect, we could still suggest that the longest and late-learnt words are more susceptible.
+To examine further this phenomenon, we plotted the two-dimensional map of joint susceptibility values for these two features (heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
+\TB{We see that \todo{}}.
+\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
+%Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no significant effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
+%These high-valued words also have low occurrence frequencies, which corresponds to the slope change in the frequency curve between extreme low frequency words and mid- to high-frequency words.
 \rk{check this with a plot of freq vs.
@@
 AoA, or simply plot of frequency for the high-valued words for AoA and CC.}
-This class of words corresponds to \TB{highly specialized words? domain-specific? a hard limit on too complicated words?}
 
-Finally the results for Clustering coefficient and Number of synonyms, despite some tendencies, do not allow us to conclude either to an effect or no effect of those features on substitution susceptibility.
+All in all, apart for Word frequency and despite some tendencies these results do not allow us to conclude either to an effect or no effect of the selected psycholinguistic features on substitution susceptibility.
+We may therefore globally assume that substitution targets are more or less chosen in a uniform way with respect to these features.
+
+% plus ou moins choisies uniformément (modulo les réserves exhibées en 4.1), on peut montrer en quoi elles sont modifiées Finally the results for Clustering coefficient and Number of synonyms, despite some tendencies, do not allow us to conclude either to an effect or no effect of those features on substitution susceptibility.
 
@@
 
-Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
+Next, we can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by the said substitution.
+We therefore measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:

### 2014-12-30 brains-protocol.tex
4.2 reorganized

@@
 We may notice a somewhat marked effect for the highest values of Age of acquisition and Number of phonemes, where the mean susceptibility is almost twice higher than for the rest (respectively 0.19 vs 0.13, and 0.21 vs 0.12).
-Keeping in mind the poor statistical significance of this effect, we could still suggest that the longest and late-learnt words are more susceptible.
+Keeping in mind the poor statistical significance of this effect, we could still suggest that the longest and late-learned words are more susceptible.
 To examine further this phenomenon, we plotted the two-dimensional map of joint susceptibility values for these two features (heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
@@
 
-We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
+Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or repulsion) effect towards some values of each feature.
+In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamic systems.
+
+We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
 In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
 $$\nu_{\phi}^0 (f) = \left<\phi\right>$$
-
-We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.\footnote{In this case $\warrival_{00}$ does depend on $\wstart$.}
-The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.
-
-Using this method we obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.
-This gives us a fine-grained view of how word features evolve upon substitution.
+We thereby obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
+In this case $\warrival_{00}$ does depend on $\wstart$.
+The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
 
 \medskip
+On the whole, this method gives us a fine-grained view of how word features evolve upon substitution with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
 Results are gathered on Fig.~\ref{fig:feature-variations}.
 
-\rk{Comment results}
+We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1.
+In other words, there exists a unique attractor for all features.
+Second, the comparison with $\mathcal{H}_0$ shows that we have two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\mathcal{H}_0$ and $\nu_{\phi}$) or (ii) $\mathcal{H}_0$ remains always above or below $\nu_{\phi}$.
 
-\rk{Might want to globally add $y = x$? Defines areas above/below $y = x$ and areas above/below $H_0$}
+\rk{Comment these results} \todo{}
 
-\begin{itemize}
+\tb{\begin{itemize}
     \item Clustering: flat, no dependency. But negative bias, which corresponds to literature on word production.
@@
     \item Number of phonemes and Number of synonyms: expanding behavior (but check with $y = x$): lower goes to lower, higher goes to higher.
-\end{itemize}
+\end{itemize}}
 

### 2014-12-30 brains-protocol.tex
4.2 reorganized v2

@@
 We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
-In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
-$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
-We thereby obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
+In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
 In this case $\warrival_{00}$ does depend on $\wstart$.
 The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
+$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
+%We thereby obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.
 
-\medskip
-On the whole, this method gives us a fine-grained view of how word features evolve upon substitution with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
-Results are gathered on Fig.~\ref{fig:feature-variations}.
+This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
 
+\medskip Results are gathered on Fig.~\ref{fig:feature-variations}.
  We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1.
 In other words, there exists a unique attractor for all features.
-Second, the comparison with $\mathcal{H}_0$ shows that we have two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\mathcal{H}_0$ and $\nu_{\phi}$) or (ii) $\mathcal{H}_0$ remains always above or below $\nu_{\phi}$.
+Second, the comparison with $\mathcal{H}_0$ shows that we have two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
 

### 2014-12-30 brains-protocol.tex
4.1 comment on heatmap

@@
 To examine further this phenomenon, we plotted the two-dimensional map of joint susceptibility values for these two features (heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-\TB{We see that \todo{}}.
-\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
+\TB{We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).
+If we focus on cells based on at least several dozens of words, values range from 0.11 to 0.22 with little obvious regularity, making it hard to draw any conclusion as regards the direction of an effect and confirming the fact that the slight late increases observed for Age of acquisition and Number of phonemes were rather non-significant.}
+%\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
 %Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no significant effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).

### 2014-12-30 brains-protocol.tex
Add notice for future results writing
Proof-read (rephrasings) of 4.1

--- 2014-12-30 brains-protocol.tex
@@
 Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
-Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration --e.g.
-drafting before publishing, double-checking sources, proof-reading--, and may not be representative of the cognitive processes at work during alteration.
+Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~e.g.
+drafting before publishing, double-checking sources, proof-reading~---, and may not be representative of the cognitive processes at work during alteration.
 The two situations are different enough to be studied separately, and we focus here on the latter.
@@
 Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, \hbox{i.e.} $\sfrac{s_w}{p_w}$.
-In other words, it measures how often a subtitutable word $w$ gets actually substituted.
+In other words, it measures how often a substitutable word $w$ actually gets substituted.
 
@@
 \medskip
-Results are gathered on Fig.~\ref{fig:feature-susceptibilities}.
+Results for this measure are gathered on Fig.~\ref{fig:feature-susceptibilities}.
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
 Indeed, susceptibility goes from nearly 0.4 for low-frequency words down to 0 for very high-frequency words.
-As a reminder, this value of 0.4 means that, this kind of low-frequency words, when present in a quote undergoing a substitution, are indeed the one being substituted 40\% of the time on average.
+To make things clear, this value of 0.4 means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 40\% of the time on average.
 
@@
 This is remarkably clear for Clustering coefficient and Number of synonyms, where susceptibility values remain within quite small intervals (respectively [0.09-0.13] and [0.11-0.15]).
-We may notice a somewhat marked effect for the highest values of Age of acquisition and Number of phonemes, where the mean susceptibility is almost twice higher than for the rest (respectively 0.19 vs 0.13, and 0.21 vs 0.12).
-Keeping in mind the poor statistical significance of this effect, we could still suggest that the longest and late-learned words are more susceptible.
-To examine further this phenomenon, we plotted the two-dimensional map of joint susceptibility values for these two features (heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-\TB{We see that \todo{}}.
-\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
+We may notice a somewhat marked effect for the highest values of Age of acquisition and Number of phonemes, where the mean susceptibility is almost twice as high as in the rest (respectively 0.19 vs 0.13, and 0.21 vs 0.12).
+Keeping in mind the poor statistical significance of this effect, we could still suggest that the longest words and words learned late are more susceptible to substitution.
+To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
+We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).
+If we focus on cells based on at least several dozens of words, values range from 0.11 to 0.22 with little obvious regularity, making it hard to draw any conclusion as regards the direction of an effect.
+This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.
+
+\rk{Add number of words to cells in heatmap}
+
+%\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
 %Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no significant effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
@@
 All in all, apart for Word frequency and despite some tendencies these results do not allow us to conclude either to an effect or no effect of the selected psycholinguistic features on substitution susceptibility.
-We may therefore globally assume that substitution targets are more or less chosen in a uniform way with respect to these features.
+We may therefore globally assume that substitution targets are chosen in a more or less uniform way with respect to these features.
 
@@
 
+\rk{Also, the figures make sense together: they \emph{all} show a single attractor, so mas alla de las differences between each feature, we see attraction.}
+
 \tb{\begin{itemize}

### 2014-12-30 brains-protocol.tex
Proof-read (rephrasings) of 4.2

@@
 
-Next, we can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by the said substitution.
-We therefore measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
+We can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by said substitution.
+We therefore measure how a word $\wstart$'s feature varies as $\wstart$ is substituted for $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
@@
 
-Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or repulsion) effect towards some values of each feature.
-In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamic systems.
+Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or a repulsion) effect towards (resp.
+from) some values of each feature.
+In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamical systems.
 
@@
 
-\medskip Results are gathered on Fig.~\ref{fig:feature-variations}.
+\medskip
+Results are gathered on Fig.~\ref{fig:feature-variations}.
 We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1.
-In other words, there exists a unique attractor for all features.
-Second, the comparison with $\mathcal{H}_0$ shows that we have two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
+In other words, each feature exhibits a unique attractor.  
+Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
 

### 2014-12-30 brains-protocol.tex
Update Fig. 4 caption

@@
 
-All in all, apart for Word frequency and despite some tendencies these results do not allow us to conclude either to an effect or no effect of the selected psycholinguistic features on substitution susceptibility.
+All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to an effect or to no effect of the selected psycholinguistic features on substitution susceptibility.
 We may therefore globally assume that substitution targets are chosen in a more or less uniform way with respect to these features.
@@
     \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    Each feature exhibits a specific and significant pattern favoring either high- or low-valued words for substitution.}
+    \TB{Apart from Word frequency (where low-frequency words are more susceptible than high-frequency words), results do not let us conclude either to an effect or to no effect of the features on susceptibility.
+    The heatmap on the lower-right, detailing the interaction of Age of acquisition and Number of phonemes on susceptibility, also shows no significant effect once outlier values are discarded.}}
     \label{fig:feature-susceptibilities}

### 2014-12-30 brains-protocol.tex
Emphasize that all figures show a single attractor

@@
 Results are gathered on Fig.~\ref{fig:feature-variations}.
-We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1.
-In other words, each feature exhibits a unique attractor.  
+We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
+In other words, each feature exhibits a unique attractor beyond its own particular variation patterns.
 Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
@@
 \rk{Comment these results} \todo{}
-
-\rk{Also, the figures make sense together: they \emph{all} show a single attractor, so mas alla de las differences between each feature, we see attraction.}
 

### 2014-12-31 brains-protocol.tex
Fill up 4.2

@@
 We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
-In other words, each feature exhibits a unique attractor beyond its own particular variation patterns.
+In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibit a unique attractor.
 Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
+The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed
+\TB{(furthermore, for these two features $\mathcal{H}_{00}$ stays inside the confidence intervals of the alteration, confirming that they are irrelevant for the substitution process)}.
+\rk{Ajouter ce texte en bleu si tu décides de mettre les figures avec $\mathcal{H}_{00}$.}
+The second class on the other hand (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias in one direction or the other.
 
-\rk{Comment these results} \todo{}
+Word frequency, with destination words always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words \TB{(this in turn confirms that substitution is more of a recall process than a familiarity process, since common words are favored over awkward ones)}.
+\rk{Mais est-ce que c'est pas une inférence inverse que de dire ceci?
+i.e.
+on sait que "familiarity" implique un biais en faveur des mots awkards, et que "recall" implique un biais en faveur des mots fréquents, mais on peut pas vraiment tirer de conclusions en utilisant la réciproque, si?}.
+The detailed variations of the Word frequency curve, despite an inflexion for high-frequency words, do not seem to follow any significant pattern.
 
-\tb{\begin{itemize}
-    \item Clustering: flat, no dependency. But negative bias, which corresponds to literature on word production.
-    \item AoA: global negative bias, also consistent with literature. Question: on one side, effect w.r.t. $H_0$ is weaker for higher AoA, but if you compare to $y = x$, it's not that clear.
-    \item Frequency: very strong positive bias, ok.
-Now extreme high-frequency words have a different pattern (maybe a third category at the opposite of the very low-frequency words?): even if above $H_0$, they're brought quite lower than their original value (complexified?)
-    \item Number of phonemes and Number of synonyms: expanding behavior (but check with $y = x$): lower goes to lower, higher goes to higher.
-\end{itemize}}
+Age of acquisition and Clustering coefficient, on the other hand, exhibit a clear negative bias for the substitution process.
+Both curves are significantly below their respective $\nu_{\phi}^0$ values, which is consistent with the literature on recall: words learned earlier and words with lower clustering coefficient are easier to produce than average~\citep{nelson2013activation,Zevin02}.
+Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word: a substitution will always produce words with a clustering coefficient near $\exp(-2.4) \simeq 0.1$.
+
+\medskip
+We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
+Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value (i.e.
+$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age of 8, and a Clustering coefficient of 0.1).
 

### 2014-12-31 brains-protocol.tex
Update fig.4 caption for heatmap with numbers of words

@@
     \TB{Apart from Word frequency (where low-frequency words are more susceptible than high-frequency words), results do not let us conclude either to an effect or to no effect of the features on susceptibility.
-    The heatmap on the lower-right, detailing the interaction of Age of acquisition and Number of phonemes on susceptibility, also shows no significant effect once outlier values are discarded.}}
+    The heatmap on the lower-right details the interaction of Age of acquisition and Number of phonemes on susceptibility; susceptibility values are indicated in the cells, and the number of words per cell is indicated in parentheses.
+    Once outlier values with only 11 words are discarded, the heatmap shows no significant interaction effect.}}
     \label{fig:feature-susceptibilities}

### 2014-12-31 brains-protocol.tex
Add an example in 4.2

@@
 
+To make things concrete, here is an example substitution taking place in the dataset.
+At the end of December 2008, many science-related websites reported the following quote,
+
+\begin{quote}
+    ``It could be that our emotions and the systems to regulate them are vestiges of our evolutionary ancestry.''
+\end{quote}
+and a minority of other websites reported the following,
+\begin{quote}
+    ``It could be that our emotions and the systems to regulate them are vestiges of our evolutionary \textbf{history}.''
+\end{quote}
+The word \emph{ancestry} is acquired at an average of 10.11 years old, appears about 12 times in the dataset, and (unfortunately) is not encoded in Free Association norms, meaning we have no Clustering coefficient for this word.
+The word it was replaced with, \emph{history}, is acquired on average at 6.42 years old, appears about 33k times in the dataset, and has a Clustering coefficient of 0.1.
+Such a change, though minor in appearance, is a typical example of alteration along the lines shown by our results.
+
+%ancestry (age 10.11, phonemes 8.0, freq 2.48, cc -, susceptibility 1.0)
+%history  [age 6.42, phonemes 6.0, freq 10.42, cc -2.30, susceptibility 0.0]
+
 \medskip

### 2014-12-31 brainscopycut.tex
US spelling, and phrasing, for first section (introduction)
change seb's color
4.1 new scenario / 4.2 transition
Put back bibliography file deleted in earlier commit
We forgot no one

@@
 \newcommand{\cam}[1]{\textcolor{blue}{\small{[{\bf C:} }#1{]}}}
+\newcommand{\seb}[1]{\textcolor{blue}{[{\bf S:} #1{]}}}
+\newcommand{\todo}{\fcolorbox{blue}{lightblue}{(...)}}
+\definecolor{lightblue}{rgb}{.85,.85,1}
 
@@
 
-We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal and Sharon Peperkamp \cam{forgot anyone?} for useful suggestions and comments.
+We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal and Sharon Peperkamp \cam{forgot anyone?} \seb{nope, c'est bon} for useful suggestions and comments.
 
 \bibliographystyle{plainnat}
-\bibliography{brainscopypaste.bib}
+\bibliography{brainscopycut.bib}
 

### 2014-12-31 brains-protocol.tex
Remove now obsolete note
Minor editorial changes

@@
 
-\rk{Add number of words to cells in heatmap}
-
 %\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
@@
 We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
-In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibit a unique attractor.
+In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
 Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
 The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed
-\TB{(furthermore, for these two features $\mathcal{H}_{00}$ stays inside the confidence intervals of the alteration, confirming that they are irrelevant for the substitution process)}.
+\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
 \rk{Ajouter ce texte en bleu si tu décides de mettre les figures avec $\mathcal{H}_{00}$.}
-The second class on the other hand (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias in one direction or the other.
+On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear positive or negative bias.
 
-Word frequency, with destination words always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words \TB{(this in turn confirms that substitution is more of a recall process than a familiarity process, since common words are favored over awkward ones)}.
+Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words \TB{(this in turn confirms that substitution is more of a recall process than a familiarity process, since common words are favored over awkward ones)}.
 \rk{Mais est-ce que c'est pas une inférence inverse que de dire ceci?
 i.e.
-on sait que "familiarity" implique un biais en faveur des mots awkards, et que "recall" implique un biais en faveur des mots fréquents, mais on peut pas vraiment tirer de conclusions en utilisant la réciproque, si?}.
+on sait qu'une tâche "familiarity/recognition" implique un biais en faveur des mots awkards, et qu'une tâche "recall" implique un biais en faveur des mots fréquents, mais on peut pas vraiment tirer de conclusions en utilisant la réciproque, si?}.
 The detailed variations of the Word frequency curve, despite an inflexion for high-frequency words, do not seem to follow any significant pattern.
@@
 Both curves are significantly below their respective $\nu_{\phi}^0$ values, which is consistent with the literature on recall: words learned earlier and words with lower clustering coefficient are easier to produce than average~\citep{nelson2013activation,Zevin02}.
-Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word: a substitution will always produce words with a clustering coefficient near $\exp(-2.4) \simeq 0.1$.
+Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; i.e.
+on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq 0.1$.
 

### 2014-12-31 brainscopycut.tex
Thank Nico too

@@
 
-We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal and Sharon Peperkamp \cam{forgot anyone?} \seb{nope, c'est bon} for useful suggestions and comments.
+We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard \cam{forgot anyone?} \seb{nope, c'est bon} for useful suggestions and comments.
 

### 2015-01-27 brains-protocol.tex
corrections interpretations 4.2

@@
     \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    \TB{Apart from Word frequency (where low-frequency words are more susceptible than high-frequency words), results do not let us conclude either to an effect or to no effect of the features on susceptibility.
-    The heatmap on the lower-right details the interaction of Age of acquisition and Number of phonemes on susceptibility; susceptibility values are indicated in the cells, and the number of words per cell is indicated in parentheses.
-    Once outlier values with only 11 words are discarded, the heatmap shows no significant interaction effect.}}
+    %\TB{Apart from Word frequency (where low-frequency words are more susceptible than high-frequency words), results do not let us conclude either to an effect or to no effect of the features on susceptibility.
+	%%%%
+    % Cam: J'ai viré la phrase ci-dessus car redondante et même contradictoire avec le texte.
+    %%%%
+    The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
+    Without the rather unsignificant cells (11 words or less), the heatmap shows no specific trend in a direction or another.}
     \label{fig:feature-susceptibilities}
@@
 In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
-Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether (i) there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$) or (ii) $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
+Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether:
+\begin{enumerate}[(i)] 
+\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
+\item or $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
+\end{enumerate}
 The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed
 \TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
-\rk{Ajouter ce texte en bleu si tu décides de mettre les figures avec $\mathcal{H}_{00}$.}
-On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear positive or negative bias.
+\rk{Cam: On vire ou bien?}
 
-Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words \TB{(this in turn confirms that substitution is more of a recall process than a familiarity process, since common words are favored over awkward ones)}.
-\rk{Mais est-ce que c'est pas une inférence inverse que de dire ceci?
+On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative.
+
+Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words.
+This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
+\cam{je crois qu'on peut éviter l'inférence inverse en ne disant pas "confirms" (c'est exagéré en effet) mais en disant juste que c'est cohérent avec un processus mais pas l'autre; j'ai changé le wording}
+% , Mais est-ce que c'est pas une inférence inverse que de dire ceci?
 i.e.
 on sait qu'une tâche "familiarity/recognition" implique un biais en faveur des mots awkards, et qu'une tâche "recall" implique un biais en faveur des mots fréquents, mais on peut pas vraiment tirer de conclusions en utilisant la réciproque, si?}.
-The detailed variations of the Word frequency curve, despite an inflexion for high-frequency words, do not seem to follow any significant pattern.
+However, the detailed variations of the Word frequency curve, despite an inflection for high-frequency words, do not seem to follow any significant pattern.\cam{Je serais pour supprimer cette phrase purement factuelle et pas très clear-cut non plus}
 

### 2015-01-27 brainscopycut.tex
corrections interpretations 4.2

--- 2014-12-31 brainscopycut.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage{enumerate}
 
@@
 
-We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal and Sharon Peperkamp \cam{forgot anyone?} \seb{nope, c'est bon} for useful suggestions and comments.
+We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard for useful suggestions and comments.
 

### 2015-01-28 brains-protocol.tex
Entérine les changements de la discussion par mail

@@
     \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    %\TB{Apart from Word frequency (where low-frequency words are more susceptible than high-frequency words), results do not let us conclude either to an effect or to no effect of the features on susceptibility.
-	%%%%
-    % Cam: J'ai viré la phrase ci-dessus car redondante et même contradictoire avec le texte.
-    %%%%
     The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
-    Without the rather unsignificant cells (11 words or less), the heatmap shows no specific trend in a direction or another.}
+    Without the rather unsignificant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}
     \label{fig:feature-susceptibilities}
@@
 \end{enumerate}
-The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed
-\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
-\rk{Cam: On vire ou bien?}
+The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
+%\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
+%\rk{Cam: On vire ou bien?}
 
@@
 This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
-\cam{je crois qu'on peut éviter l'inférence inverse en ne disant pas "confirms" (c'est exagéré en effet) mais en disant juste que c'est cohérent avec un processus mais pas l'autre; j'ai changé le wording}
+%\cam{je crois qu'on peut éviter l'inférence inverse en ne disant pas "confirms" (c'est exagéré en effet) mais en disant juste que c'est cohérent avec un processus mais pas l'autre; j'ai changé le wording}
 % , Mais est-ce que c'est pas une inférence inverse que de dire ceci?
@@
 on sait qu'une tâche "familiarity/recognition" implique un biais en faveur des mots awkards, et qu'une tâche "recall" implique un biais en faveur des mots fréquents, mais on peut pas vraiment tirer de conclusions en utilisant la réciproque, si?}.
-However, the detailed variations of the Word frequency curve, despite an inflection for high-frequency words, do not seem to follow any significant pattern.\cam{Je serais pour supprimer cette phrase purement factuelle et pas très clear-cut non plus}
+%However, the detailed variations of the Word frequency curve, despite an inflection for high-frequency words, do not seem to follow any significant pattern.\cam{Je serais pour supprimer cette phrase purement factuelle et pas très clear-cut non plus}
 

### 2015-01-28 brains-protocol.tex
Update example substitution with a carton plein

@@
 To make things concrete, here is an example substitution taking place in the dataset.
-At the end of December 2008, many science-related websites reported the following quote,
+At the end of January 2009, many media websites reported the following quote,
 
 \begin{quote}
-    ``It could be that our emotions and the systems to regulate them are vestiges of our evolutionary ancestry.''
+    ``The massive economic upheaval being experienced across the globe is sparing no one in the consumer electronics world.''
 \end{quote}
-and a minority of other websites reported the following,
+and a smaller number of media websites, and blogs, reported the following,
 \begin{quote}
-    ``It could be that our emotions and the systems to regulate them are vestiges of our evolutionary \textbf{history}.''
+    ``The massive economic upheaval being experienced across the \textbf{world} is sparing no one in the consumer electronics world.''
 \end{quote}
-The word \emph{ancestry} is acquired at an average of 10.11 years old, appears about 12 times in the dataset, and (unfortunately) is not encoded in Free Association norms, meaning we have no Clustering coefficient for this word.
-The word it was replaced with, \emph{history}, is acquired on average at 6.42 years old, appears about 33k times in the dataset, and has a Clustering coefficient of 0.1.
+The word \emph{globe} is acquired at an average of 6,5 years old, appears about 3,5k times in the dataset, and has a Clustering coefficient of 0,24.
+The word it was replaced with, \emph{world}, is acquired on average at 5,3 years old, appears about 146k times in the dataset, and has a Clustering coefficient of 0,05.
+(Both words have four phonemes.)
 Such a change, though minor in appearance, is a typical example of alteration along the lines shown by our results.
 
-%ancestry (age 10.11, phonemes 8.0, freq 2.48, cc -, susceptibility 1.0)
-%history  [age 6.42, phonemes 6.0, freq 10.42, cc -2.30, susceptibility 0.0]
+%globe (age 6.5, phonemes 4.0, freq 8.18, cc -1.42, susceptibility 1.0)
+%world  [age 5.32, phonemes 4.0, freq 11.89, cc -2.95, susceptibility 0.272727272727]
 

### 2015-01-28 brains-protocol.tex
Dots for decimal numbers, not commas

@@
 \end{quote}
-The word \emph{globe} is acquired at an average of 6,5 years old, appears about 3,5k times in the dataset, and has a Clustering coefficient of 0,24.
-The word it was replaced with, \emph{world}, is acquired on average at 5,3 years old, appears about 146k times in the dataset, and has a Clustering coefficient of 0,05.
+The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the dataset, and has a Clustering coefficient of 0.24.
+The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the dataset, and has a Clustering coefficient of 0.05.
 (Both words have four phonemes.)

### 2015-01-28 brains-protocol.tex
Note that each feature encodes a different number of words

@@
 
-The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.
+The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.\footnote{Note that since the final features stem from different datasets, they do not encode the same amounts of words (and \emph{a fortiori}, not always the same words).
+Thus, frequency encodes about 22.6k words, age of acquisition 30.1k words, number of phonemes 123.4k words, number of synonyms 111.2k, and clustering coefficient 5.7k words (because of the relatively small size of the Free Association dataset).
+This is not a problem since the analysis is done on a per-feature basis, with no need for words to be encoded in all the features at once.}
 

### 2015-01-28 brains-protocol.tex
Clarify footnote on numbers of words per feature

--- 2015-01-28 brains-protocol.tex
@@
 
-The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.
+The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.\footnote{Note that since the final features stem from different datasets, they do not always encode the same words nor the same number of words.
+Indeed, frequency encodes about 22.6k words, age of acquisition 30.1k words, number of phonemes 123.4k words, number of synonyms 111.2k, and clustering coefficient 5.7k words.
+Quite often then, words from our dataset are encoded in only a subset of the features we use (if any at all); this is not a problem however, since the analysis is done on a per-feature basis and does not need words to be encoded in all the features at once.}
 

### 2015-01-28 brains-introduction.tex
added some room for the upcoming abstract...
massaged the introduction

@@
 % !TEX root = brainscopycut.tex
+
+\subsection*{ABSTRACT}
+
+\cam{TBD.}
+
+\bigskip
+\hrule
+
 % ============================
@@
 Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.
-A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
-It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the former's counterpart outside the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
+A first stream is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
+It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
 \footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
@@
 
-Here, one of the core claims consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
+One of the core claims of this literature consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
 Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
@@
 While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
-Some could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
+Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
 
-These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
-They nonetheless show us the interest of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
+These latter studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
+They nonetheless show us the added value of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
 %and, %more broadly, large textual corpora \cam{google n-grams??} 
-in radically improving the empirical understanding of (individual-level) cultural evolution processes.
+towards radically improving the empirical understanding of (individual-level) cultural evolution processes.
 

### 2015-01-28 brains-conclusion.tex
newest
Add a note for conclusion
revamped the conclusion

@@
-% !TEX root = brainscopypaste.tex
-\section{Conclusion}\label{sec:conclusion}
+% !TEX root = brainscopycut.tex
+\section{Concluding remarks}\label{sec:conclusion}
+We aimed at contributing to the empirical understanding of knowledge transformation processes by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
+To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
+In more detail, we describe the joint properties of the substituted and substituting terms in the reformulation by individuals of a specific type of utterances (quotations).
+%--- in this sense we also diverge from psycholinguistic experiments that focus on ease of recall.
 
-Main contributions:
-\begin{itemize}
-\item large \emph{in vivo} psycholinguistics experiment, emphasizing the importance of the semantic network structure (Wordnet: Pagerank, degree, clustering[, distance?]), finely describing the impact of classically-influent psycholinguistic variables (\emph{aoa}, number of phonemes, etc.);
-\item new in the sense that it does not focus on ease of recall but rather bias of substitution, and that in this respect it not only provides a finer description of the bias but also corresponds to an  ``input-output'' reformulation couple describing the joint properties of (substituted$\rightarrow$substituting) terms.
-\item beyond that, provide the first bricks of an empirical \emph{fitness landscape} for the epidemiology of representations
-\end{itemize}
+For each of the selected psycholinguistic variables, we demonstrate the existence of attractor values in the underlying variable spaces.
+More precisely, beyond the interpretation of our results for each variable, we notice that all variables remarkably exhibit a single attractor and are generally contractile --- as such, even though the observed convergence patterns  explain only partially quote evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
+
+%Rather, we exhibit the bias of substitution, and that in this respect it not only provides a  we provide a fine description of the bias but also corresponds to an  ``input-output'' reformulation couple describing the joint properties of (substituted$\rightarrow$substituting) terms.
+
+%We contend that our results provide some of the first bricks of an empirical \emph{fitness landscape} for the epidemiology of representations
+
+%(age of acquisition, number of phonemes, ), it also emphasizes the importance of the semantic network structure (Wordnet: Pagerank, degree, clustering[, distance?]),
+
+%\begin{itemize}
+%\item new in the sense that it does not focus 
+%From a psycholingustic viewpoint,

### 2015-01-28 brains-protocol.tex
footnote 7 was too masochistic indeed

--- 2015-01-28 brains-protocol.tex
@@
     \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
-    \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from WordNet~\citep{WordNet10}) as an \emph{a priori} indicator of how easy it would be to replace a word.
+    \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from \citet{WordNet10}) as an \emph{a priori} indicator of how easy it would be to replace a word.
 \end{itemize}
@@
 Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
-This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyse is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
+This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyze is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
 
@@
 
-The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.
+The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.\footnote{Note that feature values stem from different datasets which do not always encode the same words.
+Indeed, we have data on frequency for about 22.6k words, on age of acquisition for 30.1k words, on number of phonemes for 123.4k words, number of synonyms 111.2k, and clustering coefficient 5.7k words.
+Quite often then, not all features are available for all words in our dataset; % are encoded in only a subset of the features we use (if any at all); 
+however this is not problematic since the analysis is done on a per-feature basis, all words need not be encoded in all features.}
 
@@
 
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
+    Without the rather unsignificant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}
+    \label{fig:feature-susceptibilities}
+\end{figure*}
 \medskip
@@
 
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
-    Without the rather unsignificant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}
-    \label{fig:feature-susceptibilities}
-\end{figure*}
-
 \subsection{Variation}
@@
 This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
+    \caption{\textbf{Feature variation upon substitution:} average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
+    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
+    }
+    \label{fig:feature-variations}
+\end{figure*}
 
@@
 $\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age of 8, and a Clustering coefficient of 0.1).
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{\textbf{Feature variation upon substitution:} average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
-    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
-    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
-    }
-    \label{fig:feature-variations}
-\end{figure*}

### 2015-01-29 brains-protocol.tex
Minor phrasing

@@
 Quite often then, not all features are available for all words in our dataset; % are encoded in only a subset of the features we use (if any at all); 
-however this is not problematic since the analysis is done on a per-feature basis, all words need not be encoded in all features.}
+however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
 

### 2015-02-09 brains-introduction.tex
abstract

@@
 
-\cam{TBD.}
+We describe reformulation processes within a large distributed system such as blogspace; showing how some specific features of public representations may be altered by bloggers when they freely reproduce them.
+To deal with robust and simple cultural representations, we focus on the evolution of quotations.
+In particular, we uncover some of the semantic and structural characteristics of individual words and the substitutions they undergo.
+Our work amounts to a large in vivo experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
+We show that all variables remarkably exhibit a single attractor and are generally contractile.
+Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
 
@@
 Our research lies at the intersection of these two programs, aiming to shed light on micro-level information transformation by leveraging the empirical wealth of (in vivo) social diffusion phenomena.
-More precisely, we hope to describe reformulation processes within a large distributed system such as blogspace; showing how some specific types and features of public representations may be altered by bloggers when they freely reproduce that representation.
+More precisely, we hope to describe reformulation processes within a large distributed system such as blogspace; showing how some specific types and features of public representations may be altered by bloggers when they freely reproduce them.
 

### 2015-02-09 brains-conclusion.tex
some minor corrections

@@
 For each of the selected psycholinguistic variables, we demonstrate the existence of attractor values in the underlying variable spaces.
-More precisely, beyond the interpretation of our results for each variable, we notice that all variables remarkably exhibit a single attractor and are generally contractile --- as such, even though the observed convergence patterns  explain only partially quote evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
+More precisely, beyond the interpretation of our results for each variable, we notice that all variables remarkably exhibit a single attractor and are generally contractile --- as such, even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
 

### 2015-02-10 brains-protocol.tex
fusion 3.2.1 / 3.2.2 et precisions sur les variables reseaux

@@
 
-\subsubsection{Standard psycholinguistic indices}
+\subsubsection{Psycholinguistic indices}
 
-We now introduce some of the most classical psycholinguistic measures on words.
+We first introduce some of the most classical psycholinguistic measures on words.
 
@@
 
-\subsubsection{Network-based measures}
+The number of synonyms is related to a notion of the connectivity of words in a semantic network.
+To go a bit further in this direction, we appraise the possible role of network-based variables which have received special attention in the recent related literature, following the blooming interest in networks from many disciplines over the last decade.
 
-Aside from classical psycholinguistic measures, we also considered more recently studied variables based on semantic network properties.
+\medskip
+Aside from classical psycholinguistic measures, we thus considered more recently studied variables based on semantic network properties.
 We relied on the ``free association'' norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
-
 Free association (FA) norms record the words that come to mind when someone is presented with a given cue (that is the ``free association'' task).
-As \citeauthor{Nelson04} explain,
-\begin{quote}
-free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.~\citep{Nelson04}
-\end{quote}
+As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
 %Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
@@
 
-\bigskip
-We introduce three standard network-based measures to be used on the FA network% \rk{adapt once the weighing question (\#8) is settled}
+Three standard network-based measures are to be used on the FA network% \rk{adapt once the weighing question (\#8) is settled}
 :
@@
 Furthermore, the three variables are also strongly related to age of acquisition, which leads us to keep the latter as the sole indicator for centrality.
+This may trigger a chicken-and-egg issue where a strong centrality may be due, or be the result, of an early age of acquisition; in any case, the age of acquisition seems to partially capture centrality-based network properties.
 

### 2015-02-10 brains-introduction.tex
corrections generales jusqu'a la section 3

@@
 It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
-\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
+%\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
 The producer and the user of a mental representation are one and the same person.
@@
 One of the core claims of this literature consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
-Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
+Sperber's cultural epidemiology classically opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
 Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
@@
 The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by each individual).
-This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
+This research program nonetheless exhibits a particularly interesting empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
 While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
@@
 
-These latter studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
+These latter studies are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis) and are relatively remote from cognitive science; for a significant part, they rely rather marginally on specific social science theories.
 They nonetheless show us the added value of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
@@
 To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
-While these verbatim public representations should in theory not suffer any alterations as they are produced anew (as opposed to more elaborate expressions and opinions, not identified as quoted utterances), empirical observation shows that they are in fact quite often transformed.
+While these verbatim public representations should in theory not suffer any alterations when they are produced anew (as opposed to more elaborate expressions and opinions, not identified as quoted utterances), empirical observation shows that they are occasionally transformed.
 We will in particular exhibit a non-trivial process by which individual words in quotations are replaced.
 We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
-We moreover contend that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards building empirically realistic models of cultural evolution.
+In some way that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes a first step towards building empirically realistic models of cultural evolution.
 

### 2015-02-10 brains-protocol.tex
corrections generales jusqu'a la section 3

@@
 
-The number of synonyms is related to a notion of the connectivity of words in a semantic network.
+The number of synonyms is related to a notion of the word connectivity in a semantic network.
 To go a bit further in this direction, we appraise the possible role of network-based variables which have received special attention in the recent related literature, following the blooming interest in networks from many disciplines over the last decade.
@@
 \medskip
-Aside from classical psycholinguistic measures, we thus considered more recently studied variables based on semantic network properties.
-We relied on the ``free association'' norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
-Free association (FA) norms record the words that come to mind when someone is presented with a given cue (that is the ``free association'' task).
-As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
+%, we thus considered more recently studied variables based on semantic network properties.
+We relied on the ``free association'' (FA) norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
+FA norms record the words that come to mind when someone is presented with a given cue% (that is the ``free association'' task)
+. As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
 %Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
@@
     \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
-    In the present case the two measures are quasi-perfectly correlated, and measure polysemy by taking into account the structure of the entire network.
+    In the present case these two polysemy-related measures are quasi-perfectly correlated. %, and measure polysemy by taking into account the structure of the entire network.
     %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
@@
     \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{free:set}.
-    This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.
+    This quantity tells us if some words behave like unavoidable waypoints on association chains connecting one word to another.
     \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual vs.
@@
 The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
-The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
+The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.\marginnote{C: which analysis, ours?
+(if yes, the notion of ``prediction effect'' wouldn't appear to be very understandable at this stage of the paper)}
 
@@
 	}
-	\caption{{\bf Temporal binning of quotation families.} In bin~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in bin~4).
+	\caption{{\bf Temporal binning of quotation families.} $q$ and $q'$ are two versions of a quotation belonging to the same cluster.
+In the bottom panel (b), $q'$ holds the majority in the 3rd bin and is considered the unique basis for the last occurrence of $q$ (in the 4th bin).
 This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
-Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.}
+Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence of $q$ in bin~3.}
     \label{fig:substitution-temporal-binning}
@@
 
+\medskip
 Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.

### 2015-02-10 brains-protocol.tex
corrections generales sur la section 4, avec commentaires [le texte peut encore etre simplifie/clarifie]

@@
 
-With the substitution detection model thus defined, we may now formulate a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.
+We may now use this substitution model to formulate a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.
 To this end, we build two main observables for each word feature.
-First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution under this condition.
-Second, we measure the variation of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
+First, we measure the susceptibility for words to be the target of a substitution in a quote, knowing that there has been a variation, in order to show which semantic features are the most likely to ``attract'' a substitution under this condition.
+Second, we measure the evolution of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
 
 Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
-Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
 Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~e.g.
-drafting before publishing, double-checking sources, proof-reading~---, and may not be representative of the cognitive processes at work during alteration.
+drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
 The two situations are different enough to be studied separately, and we focus here on the latter.
@@
 
-Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$% (only taking into account words that are substituted at least once)
+Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$, that is% (only taking into account words that are substituted at least once)
 :
-\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
-Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
+\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
 $$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
@@
 
-All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to an effect or to no effect of the selected psycholinguistic features on substitution susceptibility.
+All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
+of the selected psycholinguistic features on substitution susceptibility.
 We may therefore globally assume that substitution targets are chosen in a more or less uniform way with respect to these features.
@@
 
-Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or a repulsion) effect towards (resp.
-from) some values of each feature.
+Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or a repulsion) effect towards (respectively from) some values of each feature.
 In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamical systems.
 
-We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
+We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the whole pool of free association words.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins?
+la formulation est un peu ambigüe}
 In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
 In this case $\warrival_{00}$ does depend on $\wstart$.
-The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
+Our conclusions hold under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
 $$\nu_{\phi}^0 (f) = \left<\phi\right>$$
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{\textbf{Feature variation upon substitution:} average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    \caption{\textbf{Feature variation upon substitution:} $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
     The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
@@
 \item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
-\item or $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
+\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
 \end{enumerate}
@@
 
-On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative.
+On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative, with respect to the purely random situation ($\mathcal{H}_0$).
 
@@
 %However, the detailed variations of the Word frequency curve, despite an inflection for high-frequency words, do not seem to follow any significant pattern.\cam{Je serais pour supprimer cette phrase purement factuelle et pas très clear-cut non plus}
+
+\marginnote{C: les légendes sur la figure 5 sont un peu trompeuses, on parle de $\nu_\phi$, $\nu_{\phi}^0$ dans le texte et ici de $y=x$ et $\mathcal{H}_0$; peut être préciser que la courbe pleine est $\nu_\phi$, etc.?}
 
@@
 We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
-Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value (i.e.
+Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e.
 $\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age of 8, and a Clustering coefficient of 0.1).

### 2015-02-12 brains-introduction.tex
Typo

@@
 We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
-In some way that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes a first step towards building empirically realistic models of cultural evolution.
+In a way using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes a first step towards building empirically realistic models of cultural evolution.
 
@@
 Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
-Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
+Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
 First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin

### 2015-02-12 brains-protocol.tex
Some phrasing fixes, and a question

@@
 We could manually check on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
+\marginnote{S: dans la footnote: we "could" check, on laisse ça ?}
 
@@
 The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
-The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.\marginnote{C: which analysis, ours?
-(if yes, the notion of ``prediction effect'' wouldn't appear to be very understandable at this stage of the paper)}
+\TB{Our analysis showed clearer results with number of phonemes over number of syllables}, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
 
@@
 
-We therefore model the underlying quotation selection process by making a few additional assumptions.% which let us define quote-to-quote substitutions from the available data.
+We therefore model the underlying quotation selection process by making a few additional assumptions. % which let us define quote-to-quote substitutions from the available data.
 The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
@@
 First, we measure the susceptibility for words to be the target of a substitution in a quote, knowing that there has been a variation, in order to show which semantic features are the most likely to ``attract'' a substitution under this condition.
-Second, we measure the evolution of word feature over a substitution, looking at the variation of a given feature between start and arrival words.
+Second, we measure the change in word feature upon substitution, looking at the variation of a given feature between start and arrival words.
 
@@
 Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, \hbox{i.e.} $\sfrac{s_w}{p_w}$.
-In other words, it measures how often a substitutable word $w$ actually gets substituted.
+\TB{In other words, it measures how often a word $w$ actually gets substituted, compared to how often it could have been substituted (because it appears in quotes undergoing substitution)}.
 
@@
 \medskip
-Results for this measure are gathered on Fig.~\ref{fig:feature-susceptibilities}.
+Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.

### 2015-02-12 brains-protocol.tex
More phrasing and questions

@@
 \footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
+\marginnote{S: en fait cette footnote (8) est valide que pour les variations de features, pas pour les susceptibilités.
+C'est d'ailleurs consistent, puisque sur les susceptibilités cette opération serait pas vraiment possible.
+On la passe au 4.2 ?}
 $$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
@@
     The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
-    Without the rather unsignificant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}
+    Without the rather non-significant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}
     \label{fig:feature-susceptibilities}
@@
 
-We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the whole pool of free association words.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins?
-la formulation est un peu ambigüe}
+We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was \TB{randomly chosen from the whole pool of words in the dataset from which the feature is built (e.g.
+for Clustering coefficient: randomly chosen in the pool of words in the Free Association norms)}.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins?
+la formulation est un peu ambigüe.
+S: oui et non.
+Oui: c'est pas juste les voisins.
+Non: le pool dépend de la feature, c'est pas juste FA.
+Il y avait une erreur, j'ai reformulé pour être plus clair.}
 In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.

### 2015-02-12 brains-conclusion.tex
Last phrasing fixes, also fix figure legends [pull other repo!]

@@
 \section{Concluding remarks}\label{sec:conclusion}
-We aimed at contributing to the empirical understanding of knowledge transformation processes by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
+We aimed to contribute to the empirical understanding of knowledge transformation processes by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
 To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.

### 2015-02-12 brains-protocol.tex
Last phrasing fixes, also fix figure legends [pull other repo!]

@@
 
-This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
+This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\nu_{\phi}^0$).
 \begin{figure*}[!th]
@@
     The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
+    The intersection with $y = x$ marks the attractor value.
     %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
@@
 \medskip
-Results are gathered on Fig.~\ref{fig:feature-variations}.
+Results are gathered in Fig.~\ref{fig:feature-variations}.
 We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
 In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
-Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether:
+Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
 \begin{enumerate}[(i)] 
@@
 %However, the detailed variations of the Word frequency curve, despite an inflection for high-frequency words, do not seem to follow any significant pattern.\cam{Je serais pour supprimer cette phrase purement factuelle et pas très clear-cut non plus}
-
-\marginnote{C: les légendes sur la figure 5 sont un peu trompeuses, on parle de $\nu_\phi$, $\nu_{\phi}^0$ dans le texte et ici de $y=x$ et $\mathcal{H}_0$; peut être préciser que la courbe pleine est $\nu_\phi$, etc.?}
 

### 2015-02-13 brains-protocol.tex
rc-1

@@
 The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'', we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
-We could manually check on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
-\marginnote{S: dans la footnote: we "could" check, on laisse ça ?}
+We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
 
@@
 The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
-\TB{Our analysis showed clearer results with number of phonemes over number of syllables}, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
+Our analysis showed clearer results with number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
 
@@
 Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, \hbox{i.e.} $\sfrac{s_w}{p_w}$.
-\TB{In other words, it measures how often a word $w$ actually gets substituted, compared to how often it could have been substituted (because it appears in quotes undergoing substitution)}.
+{In other words, it measures how often a word $w$ actually gets substituted, compared to how often it could have been substituted (because it appears in quotes undergoing substitution)}.
 
@@
 :
-\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
-\marginnote{S: en fait cette footnote (8) est valide que pour les variations de features, pas pour les susceptibilités.
-C'est d'ailleurs consistent, puisque sur les susceptibilités cette opération serait pas vraiment possible.
-On la passe au 4.2 ?}
+
 $$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
@@
 We therefore measure how a word $\wstart$'s feature varies as $\wstart$ is substituted for $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
-Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
+Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, i.e.:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
 $$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
@@
 
-We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was \TB{randomly chosen from the whole pool of words in the dataset from which the feature is built (e.g.
+We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was randomly chosen from the whole pool of words available in the dataset for that feature.\footnote{For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the dataset of FA norms.}
+%\TB{randomly chosen from the whole pool of words in the dataset from which the feature is built (e.g.
 for Clustering coefficient: randomly chosen in the pool of words in the Free Association norms)}.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins?

### 2015-02-19 brains-conclusion.tex
Add telmo's remarks from 19/02/2015 review in main pdf

@@
 \section{Concluding remarks}\label{sec:conclusion}
-We aimed to contribute to the empirical understanding of knowledge transformation processes by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
+We aimed to contribute to the empirical understanding of \TB{knowledge transformation processes} \rk{T: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?} by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
 To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.

### 2015-02-19 brains-protocol.tex
Add telmo's remarks from 19/02/2015 review in main pdf

@@
 We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'', we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{\TB{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them''} \rk{T: it is not clear why this utterance is lost according to the filtering rules described in sections 3.1}, we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
 We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
@@
 These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
-We identified and implemented eleven other such models, and they all yielded essentially the same results.
+\TB{We identified and implemented eleven other such models, and they all yielded essentially the same results.} \rk{T: I would prefer a bit more detail here.
+Maybe on supp.
+mats.?
+What does “essentially the same” mean?
+This could cause concern to reviewers.}
 
@@
 We can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by said substitution.
-We therefore measure how a word $\wstart$'s feature varies as $\wstart$ is substituted for $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
+We therefore measure how \TB{a word $\wstart$'s feature} \rk{T: This is correct but confusing.
+I had to re-read 3 times.} varies as $\wstart$ is substituted for $\warrival$, i.e.
+$\phi(\warrival)$ as a function of $\phi(\wstart)$.
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, i.e.:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}

### 2015-02-19 brains-conclusion.tex
Correct after some of the remarks from Telmo's review

--- 2015-02-12 brains-conclusion.tex
@@
 \section{Concluding remarks}\label{sec:conclusion}
-We aimed to contribute to the empirical understanding of knowledge transformation processes by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
+We aimed to contribute to the empirical understanding of \TB{representation transformation processes} \marginnote{fixed after Telmo saying: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?} by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
 To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.

### 2015-02-19 brains-protocol.tex
Correct after some of the remarks from Telmo's review

@@
 We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{\TB{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them''} \rk{T: it is not clear why this utterance is lost according to the filtering rules described in sections 3.1}, we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
-We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{\TB{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them'')}, we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
+We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}\marginnote{fixed after Telmo saying: it is not clear why this utterance is lost according to the filtering rules described in sections 3.1}
 
@@
 We can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by said substitution.
-We therefore measure how \TB{a word $\wstart$'s feature} \rk{T: This is correct but confusing.
-I had to re-read 3 times.} varies as $\wstart$ is substituted for $\warrival$, i.e.
-$\phi(\warrival)$ as a function of $\phi(\wstart)$.
+\TB{Looking at a word $\wstart$, we measure how its feature varies as it is substituted for $\warrival$, i.e.
+we look at $\phi(\warrival)$ as a function of $\phi(\wstart)$.} \marginnote{fixed after Telmo saying: This is correct but confusing.
+I had to re-read 3 times.}
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, i.e.:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}

### 2015-02-19 brains-conclusion.tex
ack / updated post-telmo

--- 2015-02-12 brains-conclusion.tex
@@
 \section{Concluding remarks}\label{sec:conclusion}
-We aimed to contribute to the empirical understanding of knowledge transformation processes by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
+We aimed to contribute to the empirical understanding of representation transformation processes %\marginnote{fixed after Telmo saying: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?}
+ by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
 To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.

### 2015-02-19 brains-protocol.tex
ack / updated post-telmo

--- 2015-02-19 brains-protocol.tex
@@
 We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{\TB{While we lose real event-related utterances such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them''} \rk{T: it is not clear why this utterance is lost according to the filtering rules described in sections 3.1}, we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them''), we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
 We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
@@
 We can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by said substitution.
-We therefore measure how \TB{a word $\wstart$'s feature} \rk{T: This is correct but confusing.
-I had to re-read 3 times.} varies as $\wstart$ is substituted for $\warrival$, i.e.
-$\phi(\warrival)$ as a function of $\phi(\wstart)$.
+Considering a word $\wstart$ substituted for $\warrival$, we measure how the feature of $\wstart$ varies when it is replaced with $\warrival$, i.e.
+we look at $\phi(\warrival)$ as a function of $\phi(\wstart)$.
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, i.e.:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}

### 2015-02-19 brainscopycut.tex
Garbage is garbage
ack / updated post-telmo

@@
 
-We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard for useful suggestions and comments.
+We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Telmo Menezes, Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard for useful suggestions and comments.
 
@@
 
-\clearpage 
-\input{brains-garbage}
+%\clearpage 
+%\input{brains-garbage}
 

### 2015-02-24 brains-protocol.tex
options 11 - no other options

@@
 %We assume here that quotes are based on the most frequent quote of the preceding bin (\hbox{vs.} all the other quotes in that bin).
-These assumptions are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, the criterion of the most frequent quote in the preceding bin may be replaced with the most frequent quote overall, or the oldest quote; time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
-These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
-\TB{We identified and implemented eleven other such models, and they all yielded essentially the same results.} \rk{T: I would prefer a bit more detail here.
+Note that these assumptions are admittedly a subset of a much wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, the criterion of the most frequent quote in the preceding bin may be replaced with the most frequent quote overall, or the oldest quote; time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
+It is however not feasible to try them all and, for the sake of simplicity, we decided to go with a sensible set of assumptions, and stick to them without trying alternative options.
+%These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
+%\TB{We identified and implemented eleven other such models, and they all yielded essentially the same results.} \rk{T: I would prefer a bit more detail here.
 Maybe on supp.

### 2015-03-03 brains-protocol.tex
Add marks for where the text must be fixed after new filtering

@@
 
+\rk{fix}
+
 Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
@@
  (as well as quotes not written in English).
+% New filtering: 1051 substitutions out of 6172 mined. 45749 clusters. -- correct text with numbers and for new figures.
 We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
 The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them''), we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
-We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.}
+We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.\rk{fix}}
 
@@
 center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 476 valid substitutions.
+\rk{fix}
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    \caption{\textbf{Substitution susceptibility:} \rk{fix} average susceptibility to substitution vs.
+average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
     The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
@@
 \medskip
-Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
+\rk{fix} Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
@@
 
-The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
+\rk{fix} The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
 For each of these features, it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
@@
 
-All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
+\rk{fix} All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
 of the selected psycholinguistic features on substitution susceptibility.
@@
 Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; i.e.
-on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq 0.1$.
+on average, substitutions will always produce words with a clustering coefficient around \rk{fix} $\exp(-2.4) \simeq 0.1$.
 
@@
 \medskip
-We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
+\rk{fix} We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
 Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e.

### 2015-03-03 brains-protocol.tex
Fix a few numbers after new filtering

@@
 
-\rk{fix}
-
 Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
 Their presence is largely due to random variations around a casual expression, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
-We thus filter out all clusters lasting more than 80 days, and all quotes less than 5 words long %\seb{it is asymmetrical...}
+\TB{We thus filter out all quotes less than 5 words long or lasting more than 80 days} %\seb{it is asymmetrical...}
  (as well as quotes not written in English).
+\TB{If an entire cluster still lasts more than 80 days after this screening (because of short-lived but unrelated quotes far apart in time), we also filter it out.}
 % New filtering: 1051 substitutions out of 6172 mined. 45749 clusters. -- correct text with numbers and for new figures.
-We eventually keep 19,621 clusters (out of 71,568; i.e. 27.4\%), containing a total of 60,246 unique quotes (out of 310,457; i.e. 19.4\%) making up $\sim$838k occurrences (out of 8.12m, i.e. 10.3\%).
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as ``the terrorists have used automatic weapons and in some places grenades have been lobbed the encounters are still going on and we are trying to overpower them''), we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
+We eventually keep 45,749 clusters (out of 71,568; i.e.
+63.9\%), containing a total of 127,778 unique quotes (out of 310,457; i.e.
+41.2\%) making up $\sim$2.43m occurrences (out of 8.16m, i.e.
+29.8\%).
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as \TB{``the city is tired of me and the organization and i have run our course together''}), we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
 We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.\rk{fix}}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} \rk{fix} average susceptibility to substitution vs.
+    \caption{\textbf{Substitution susceptibility:} \rk{fix by replacing "number of synonyms" par "Age of acquisition"} average susceptibility to substitution vs.
 average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
@@
 \medskip
-\rk{fix} Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
+\rk{fix by replacing "number of synonyms" par "Age of acquisition"} Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
@@
 
-\rk{fix} The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
+\rk{fix by replacing "number of synonyms" par "Age of acquisition"} The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
 For each of these features, it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
@@
 
-\rk{fix} All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
+\rk{fix by replacing "number of synonyms" par "Age of acquisition"} All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect
 of the selected psycholinguistic features on substitution susceptibility.

### 2015-03-03 brains-protocol.tex
Do some fixes in text after new filtering

@@
 The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as \TB{``the city is tired of me and the organization and i have run our course together''}), we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
-We manually checked on a subsample of clusters that a large proportion of remaining quotes seem to be relevant, while filtered quotes are essentially irrelevant.\rk{fix}}
+\TB{Manual coding on a subsample of the excluded clusters revealed that XXX\% of them are indeed irrelevant to our analysis.}}\marginnote{Pour la footnote: avec un codage conservateur on a 67\% correctly filtered vs.
+33\% qui pourraient êtres gardés.
+Avec un codage plus normal: 81\% correctly filtered vs.
+19\% qui pourraient être gardés.}
 
@@
 Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.
-In practice, from the 838k initial occurrences spread into 19,621 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 2,873 real substitutions obeying to this model.
+In practice, from the 2.43m initial occurrences spread into 45,749 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 6,172 real substitutions obeying to this model.
 From these substitutions we remove those featuring stop words, minor spelling changes (e.g.
-center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 476 valid substitutions.
-\rk{fix}
+center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 1,051 valid substitutions.
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} \rk{fix by replacing "number of synonyms" par "Age of acquisition"} average susceptibility to substitution vs.
-average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    The heatmap on the lower-right shows the joint effect of Age of acquisition and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
-    Without the rather non-significant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
+    \TB{TO BE FIXED: Without the rather non-significant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}}
     \label{fig:feature-susceptibilities}
@@
 \medskip
-\rk{fix by replacing "number of synonyms" par "Age of acquisition"} Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
+Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
-Indeed, susceptibility goes from nearly 0.4 for low-frequency words down to 0 for very high-frequency words.
-To make things clear, this value of 0.4 means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 40\% of the time on average.
+Indeed, susceptibility goes from 0.33 for low-frequency words down to nearly 0 for very high-frequency words.
+To make things clear, this value of 0.33 means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.
 
-\rk{fix by replacing "number of synonyms" par "Age of acquisition"} The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
-For each of these features, it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
+The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
+\TB{If we set aside the values for low Number of phonemes}, for each of these features it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
 If these variables have an effect, it is by no means as strong as it is for Word frequency.
-This is remarkably clear for Clustering coefficient and Number of synonyms, where susceptibility values remain within quite small intervals (respectively [0.09-0.13] and [0.11-0.15]).
-We may notice a somewhat marked effect for the highest values of Age of acquisition and Number of phonemes, where the mean susceptibility is almost twice as high as in the rest (respectively 0.19 vs 0.13, and 0.21 vs 0.12).
-Keeping in mind the poor statistical significance of this effect, we could still suggest that the longest words and words learned late are more susceptible to substitution.
-To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
+This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.19] and [0.15-0.20]).
+\TB{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as in the rest (respectively 0.9 vs 0.16, and 0.11 vs 0.17).}
+Keeping in mind the poor statistical significance of this effect, we could still suggest that the shortest words and words with few synonyms are less susceptible to substitution.
+\TB{TO BE FIXED: To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
 We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).
 If we focus on cells based on at least several dozens of words, values range from 0.11 to 0.22 with little obvious regularity, making it hard to draw any conclusion as regards the direction of an effect.
-This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.
+This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.}
 
@@
 
-\rk{fix by replacing "number of synonyms" par "Age of acquisition"} All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect
+All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
 of the selected psycholinguistic features on substitution susceptibility.
@@
 Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; i.e.
-on average, substitutions will always produce words with a clustering coefficient around \rk{fix} $\exp(-2.4) \simeq 0.1$.
+on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq 0.1$.
 
@@
 \medskip
-\rk{fix} We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
+We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
 Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e.
-$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age of 8, and a Clustering coefficient of 0.1).
+$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age slightly below 8, and a Clustering coefficient of 0.1).

### 2015-03-04 brains-protocol.tex
remodeling cluster filtering

@@
 Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
-Their presence is largely due to random variations around a casual expression, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
-\TB{We thus filter out all quotes less than 5 words long or lasting more than 80 days} %\seb{it is asymmetrical...}
+Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
+To filter them out, we exclude all quotes having less than 5 words long or lasting more than 80 days %\seb{it is asymmetrical...}
  (as well as quotes not written in English).
-\TB{If an entire cluster still lasts more than 80 days after this screening (because of short-lived but unrelated quotes far apart in time), we also filter it out.}
+If an entire cluster still lasts more than 80 days after this screening (because of short-lived but unrelated quotes far apart in time), we also exclude it.
 % New filtering: 1051 substitutions out of 6172 mined. 45749 clusters. -- correct text with numbers and for new figures.
@@
 41.2\%) making up $\sim$2.43m occurrences (out of 8.16m, i.e.
-29.8\%).
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.\footnote{While we lose some real event-related utterances because of their presence in clusters lasting more than 80 days (such as \TB{``the city is tired of me and the organization and i have run our course together''}), we contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion.
-\TB{Manual coding on a subsample of the excluded clusters revealed that XXX\% of them are indeed irrelevant to our analysis.}}\marginnote{Pour la footnote: avec un codage conservateur on a 67\% correctly filtered vs.
-33\% qui pourraient êtres gardés.
-Avec un codage plus normal: 81\% correctly filtered vs.
-19\% qui pourraient être gardés.}
+29.8\%).\footnote{The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.}
+%We contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion, 
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (such as ``the city is tired of me and the organization and I have run our course together''), we check that our approach essentially fulfills its goals by manually coding a random subsample of 100 excluded clusters: a solid 71\% appear to be entirely irrelevant to our analysis (everyday language rather than quotations), and all but one of the remaining clusters were of relevance to the protocol set out below in Sec.~\ref{sec:model}.
 
@@
 \subsection{Substitution model}
-
 %\subsubsection{Temporal binning}
-\label{sec:temporal-binning}
+\label{sec:temporal-binning}\label{sec:model}
 

### 2015-03-04 brains-protocol.tex
making sense of Fig.4 once again

@@
 The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
-\TB{If we set aside the values for low Number of phonemes}, for each of these features it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
+{If we set aside the values for low Number of phonemes}, for each of these features it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
 If these variables have an effect, it is by no means as strong as it is for Word frequency.
-This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.19] and [0.15-0.20]).
-\TB{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as in the rest (respectively 0.9 vs 0.16, and 0.11 vs 0.17).}
-Keeping in mind the poor statistical significance of this effect, we could still suggest that the shortest words and words with few synonyms are less susceptible to substitution.
-\TB{TO BE FIXED: To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).
-If we focus on cells based on at least several dozens of words, values range from 0.11 to 0.22 with little obvious regularity, making it hard to draw any conclusion as regards the direction of an effect.
-This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.}
+This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.18] and [0.16-0.20]\marginnote{corrigé les intervalles en fonction des arrondis (.5->1), à partir du graphique ceci dit (pas trouvé les valeurs exactes sur le ipython}).
+{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as in the rest \marginnote{the rest c'est la moyenne des other values ou la plus extrême?} (respectively 0.\TB{0}9 vs 0.16, and 0.11 vs 0.17).}
+Keeping in mind the poor statistical significance of this effect, we could still wonder if the shortest words and words with fewest synonyms are significantly less susceptible to substitution.
+To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
+\TB{TBF: [[We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).]]}
+\TB{TBF: [[If we focus on cells based on at least several dozens of words]]} Values tend to navigate around the mean value \TB{[TBF:()]}, \TB{while we observe a few outlier cells with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).} \TB{On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
+\TB{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
 
@@
 
-All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
+All in all, apart from Word frequency and despite some tendencies, \TB{in general} these results do not allow us to conclude either to a marked effect %or to no effect 
 of the selected psycholinguistic features on substitution susceptibility.

### 2015-03-04 brains-protocol.tex
fine polish re: Fig 4 / still stuff to be fixed

@@
     The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
-    \TB{TO BE FIXED: Without the rather non-significant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}}
+    \TG{TBF: Without the rather non-significant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}}
     \label{fig:feature-susceptibilities}
@@
 To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-\TB{TBF: [[We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).]]}
-\TB{TBF: [[If we focus on cells based on at least several dozens of words]]} Values tend to navigate around the mean value \TB{[TBF:()]}, \TB{while we observe a few outlier cells with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).} \TB{On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
-\TB{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
+\TG{TBF: [[We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).]]}
+\TG{TBF: [[If we focus on cells based on at least several dozens of words]]} \TB{Values tend to navigate around the mean value} \TG{[TBF:()]}, \TB{even if there a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).} \TB{On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
+\TG{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
 

### 2015-03-04 brains-protocol.tex
various updates

@@
 If these variables have an effect, it is by no means as strong as it is for Word frequency.
-This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.18] and [0.16-0.20]\marginnote{corrigé les intervalles en fonction des arrondis (.5->1), à partir du graphique ceci dit (pas trouvé les valeurs exactes sur le ipython}).
-{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as in the rest \marginnote{the rest c'est la moyenne des other values ou la plus extrême?} (respectively 0.\TB{0}9 vs 0.16, and 0.11 vs 0.17).}
+This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.18] and [0.16-0.20]).
+{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as the average of the other values (respectively 0.09 vs 0.16, and 0.11 vs 0.17).}
 Keeping in mind the poor statistical significance of this effect, we could still wonder if the shortest words and words with fewest synonyms are significantly less susceptible to substitution.
 To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-\TG{TBF: [[We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).]]}
-\TG{TBF: [[If we focus on cells based on at least several dozens of words]]} \TB{Values tend to navigate around the mean value} \TG{[TBF:()]}, \TB{even if there a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).} \TB{On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
+%\TG{TBF: [[We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).]]}
+%\TG{TBF: [[If we focus on cells based on at least several dozens of words]]} 
+Values tend to navigate around the mean value \TG{[TBF:()]}, even if there are a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).
+{On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
 \TG{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
@@
 
-All in all, apart from Word frequency and despite some tendencies, \TB{in general} these results do not allow us to conclude either to a marked effect %or to no effect 
+All in all, apart from Word frequency and despite some local tendencies, {in general} these results do not allow us to conclude either to a marked effect %or to no effect 
 of the selected psycholinguistic features on substitution susceptibility.

### 2015-03-04 brains-protocol.tex
various updates

@@
     \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
-    \TG{TBF: Without the rather non-significant cells (11 words or less), the heatmap shows no specific trend in one direction or the other.}}
+    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
     \label{fig:feature-susceptibilities}
@@
 {On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
-\TG{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
+%\TG{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
 

### 2015-03-04 brains-protocol.tex
mean value

@@
 %\TG{TBF: [[If we focus on cells based on at least several dozens of words]]} 
-Values tend to navigate around the mean value \TG{[TBF:()]}, even if there are a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).
+Values tend to navigate around the mean value (0.16), even if there are a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).
 {On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}

### 2015-03-05 brains-introduction.tex
Move abstract, comment unnecessary preamble options, rename Figure to Fig.

@@
 % !TEX root = brainscopycut.tex
-
-\subsection*{ABSTRACT}
-
-We describe reformulation processes within a large distributed system such as blogspace; showing how some specific features of public representations may be altered by bloggers when they freely reproduce them.
-To deal with robust and simple cultural representations, we focus on the evolution of quotations.
-In particular, we uncover some of the semantic and structural characteristics of individual words and the substitutions they undergo.
-Our work amounts to a large in vivo experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
-We show that all variables remarkably exhibit a single attractor and are generally contractile.
-Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
-
-\bigskip
-\hrule
 

### 2015-03-05 brains-protocol.tex
Move abstract, comment unnecessary preamble options, rename Figure to Fig.

@@
 Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
-Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Figure~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
+Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Fig.~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
 
@@
 
-The final set of variables we consider, as well as their cross-correlations, can be seen in Figure~\ref{fig:feature-corrs-filtered}.\footnote{Note that feature values stem from different datasets which do not always encode the same words.
+The final set of variables we consider, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.\footnote{Note that feature values stem from different datasets which do not always encode the same words.
 Indeed, we have data on frequency for about 22.6k words, on age of acquisition for 30.1k words, on number of phonemes for 123.4k words, number of synonyms 111.2k, and clustering coefficient 5.7k words.
@@
 Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs twice on the 20th and once on the 21st of January.
-If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 20th based himself on $q$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
+If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 20th based himself on $q$ as is shown in Fig.~\ref{fig:substitution-temporal-binning-a}.
 But what about the second and third occurrences of $q'$, on the 20th and 21st?
 Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place) 
-or accurate reproductions of the previous occurrences of $q'$? (Options shown in Figure~\ref{fig:substitution-temporal-binning-a}.)
+or accurate reproductions of the previous occurrences of $q'$? (Options shown in Fig.~\ref{fig:substitution-temporal-binning-a}.)
 
@@
 In the bottom panel (b), $q'$ holds the majority in the 3rd bin and is considered the unique basis for the last occurrence of $q$ (in the 4th bin).
-This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
+This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Fig.~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
 Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence of $q$ in bin~3.}
@@
 \medskip
-Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.
+Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Fig.~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.
 In practice, from the 2.43m initial occurrences spread into 45,749 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 6,172 real substitutions obeying to this model.
@@
 
-%Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Figure~\ref{fig:substitution-q_max}.
+%Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Fig.~\ref{fig:substitution-q_max}.
 

### 2015-03-05 brainscopycut.tex
Use apa6 documentclass
Move abstract, comment unnecessary preamble options, rename Figure to Fig.

@@
 % General document style ------------------------------------------------------------------
-\documentclass{acmtog}
+\documentclass[man]{apa6}
 \usepackage[utf8]{inputenc}
@@
 
-\acmVolume{VV}
-\acmNumber{N}
-\acmYear{YYYY}
-\acmMonth{Monteh}
-\acmArticleNum{XXX}  
-\acmdoi{10.1145/XXXXXXX.YYYYYYY}
+%\acmVolume{VV}
+%\acmNumber{N}
+%\acmYear{YYYY}
+%\acmMonth{Month}
+%\acmArticleNum{XXX}  
+%\acmdoi{10.1145/XXXXXXX.YYYYYYY}
 
-\markboth{S. Lerique and C. Roth}{How do we copy and paste? The semantic drift of quotations in blogspace}
+%\markboth{S. Lerique and C. Roth}{How do we copy and paste? The semantic drift of quotations in blogspace}
 
@@
 
+% shortcuts
+\newcommand{\arrival}{\text{arrival}}
+\newcommand{\start}{\text{start}}
+
+\newcommand{\warrival}{w'}
+\newcommand{\wstart}{w}
+
 % To make comments, corrections, notes
@@
 
-% shortcuts
-\newcommand{\arrival}{\text{arrival}}
-\newcommand{\start}{\text{start}}
-
-\newcommand{\warrival}{w'}
-\newcommand{\wstart}{w}
-
 % If citations or references are missing, make it visible
@@
 
-\title{{How do we copy and paste?\\The semantic drift of quotations in blogspace}}
+\title{How do we copy and paste? The semantic drift of quotations in blogspace}
+\shorttitle{The semantic drift of quotations in blogspace}
 \date{}
-\author{S\'EBASTIEN LERIQUE {\upshape and} CAMILLE ROTH}
+\author{Sébastien Lerique \& Camille Roth}
+\affiliation{Centre d'Analyse et de Mathématique Sociales () \& Centre Marc Bloch ()}
+\abstract{We describe reformulation processes within a large distributed system such as blogspace;
+showing how some specific features of public representations may be altered by bloggers when they freely reproduce them.
+To deal with robust and simple cultural representations, we focus on the evolution of quotations.
+In particular, we uncover some of the semantic and structural characteristics of individual words and the substitutions they undergo.
+Our work amounts to a large in vivo experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
+We show that all variables remarkably exhibit a single attractor and are generally contractile.
+Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.}
 

### 2015-03-05 brains-introduction.tex
Change \ref's to named references or remove them (apastyle does not number sections)

@@
 
-The next section (Sec.~\ref{sec:related}) describes our hypotheses along with the relevant state-of-the-art on this psycholinguistic matter.
-In Sec.~\ref{sec:protocol}, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
-Sec.~\ref{sec:results} describes the significant psycholinguistic biases observed during \emph{in vivo} quotation reformulation as well as their epidemiological setting, followed by a discussion and general guidelines for further work in Sec.~\ref{sec:conclusion}.
+\TB{The next section %(Sec.~\ref{sec:related}) 
+describes our hypotheses along with the relevant state-of-the-art on this psycholinguistic matter.
+%In Sec.~\ref{sec:protocol}, 
+In the section that follows it, 
+we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
+%Sec.~\ref{sec:results} 
+The results section describes the significant psycholinguistic biases observed during \emph{in vivo} quotation reformulation as well as their epidemiological setting, followed by a discussion and general guidelines for further work in the final section.%Sec.~\ref{sec:conclusion}.
+}
 

### 2015-03-05 brains-protocol.tex
Change \ref's to named references or remove them (apastyle does not number sections)

@@
 %We contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion, 
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (such as ``the city is tired of me and the organization and I have run our course together''), we check that our approach essentially fulfills its goals by manually coding a random subsample of 100 excluded clusters: a solid 71\% appear to be entirely irrelevant to our analysis (everyday language rather than quotations), and all but one of the remaining clusters were of relevance to the protocol set out below in Sec.~\ref{sec:model}.
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (such as ``the city is tired of me and the organization and I have run our course together''), we check that our approach essentially fulfills its goals by manually coding a random subsample of 100 excluded clusters: a solid 71\% appear to be entirely irrelevant to our analysis (everyday language rather than quotations), and all but one of the remaining clusters were of relevance to the protocol \TB{set out below.}% in Sec.~\ref{sec:model}.
 

### 2015-03-05 brains-protocol.tex
APAize enumerations

@@
 
-\begin{itemize}
+\begin{APAitemize}
     \item \textbf{Word frequency}: the frequency at which words appear in our dataset, known to be relevant for both recognition and recall~\citep{gregg1976word},
@@
     \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from \citet{WordNet10}) as an \emph{a priori} indicator of how easy it would be to replace a word.
-\end{itemize}
+\end{APAitemize}
 
@@
 
-\begin{itemize}
+\begin{APAitemize}
     \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
@@
     This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
-\end{itemize}
+\end{APAitemize}
 
@@
 
-This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\nu_{\phi}^0$).
+This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to 
+\begin{seriate}
+\item the original feature (\hbox{vs.} $y=x$) and 
+\item a random arrival (\hbox{vs.} $\nu_{\phi}^0$).
+\end{seriate}
 \begin{figure*}[!th]
@@
 Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
-\begin{enumerate}[(i)] 
+\begin{APAenumerate} 
 \item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
 \item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
-\end{enumerate}
+\end{APAenumerate}
 The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.

### 2015-03-05 brains-introduction.tex
Italics for latin in vivo / in vitro

@@
 %--- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.
-Our research lies at the intersection of these two programs, aiming to shed light on micro-level information transformation by leveraging the empirical wealth of (in vivo) social diffusion phenomena.
+Our research lies at the intersection of these two programs, aiming to shed light on micro-level information transformation by leveraging the empirical wealth of (\emph{in vivo}) social diffusion phenomena.
 More precisely, we hope to describe reformulation processes within a large distributed system such as blogspace; showing how some specific types and features of public representations may be altered by bloggers when they freely reproduce them.

### 2015-03-05 brains-introduction.tex
Latin abbreviations only between parentheses

@@
 The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
-\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, i.e.
-the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
+\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, that is the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
 A third psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
@@
 On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible (see again~\citet{Yonelinas02} for a review).
-The jury is still out as to whether reformulation alteration, i.e. spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
+The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
 We also aim here at shedding some light on this debate, considering oddness as a dimension of the purported fitness of utterances.

### 2015-03-05 brains-protocol.tex
Italics for latin in vivo / in vitro
Latin abbreviations only between parentheses

@@
 
-\subsection{In vivo utterances}
+\subsection{\emph{In vivo} utterances}
 
@@
 We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
-A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, i.e. sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
+A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
@@
     This quantity tells us if some words behave like unavoidable waypoints on association chains connecting one word to another.
-    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual vs.
-possible edges between a node's neighbors.}
+    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual versus possible edges between a node's neighbors.}
     This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
@@
 We therefore model the underlying quotation selection process by making a few additional assumptions. % which let us define quote-to-quote substitutions from the available data.
-The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
+The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, that is from a source external to the data collection perimeter.
 
@@
 Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
-Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~e.g.
-drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
+Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~for instance drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
 The two situations are different enough to be studied separately, and we focus here on the latter.
@@
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on that word or on another one.
-Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, \hbox{i.e.} $\sfrac{s_w}{p_w}$.
+Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, that is $\sfrac{s_w}{p_w}$.
 {In other words, it measures how often a word $w$ actually gets substituted, compared to how often it could have been substituted (because it appears in quotes undergoing substitution)}.
@@
     \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
     The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
@@
 This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.18] and [0.16-0.20]).
-{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as the average of the other values (respectively 0.09 vs 0.16, and 0.11 vs 0.17).}
+{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as the average of the other values (respectively 0.09 vs.
+0.16, and 0.11 vs.
+0.17).}
 Keeping in mind the poor statistical significance of this effect, we could still wonder if the shortest words and words with fewest synonyms are significantly less susceptible to substitution.
@@
 
-We can thus show how words are modified once we know they are substituted, \hbox{i.e.} how their features are modified by said substitution.
-Considering a word $\wstart$ substituted for $\warrival$, we measure how the feature of $\wstart$ varies when it is replaced with $\warrival$, i.e.
-we look at $\phi(\warrival)$ as a function of $\phi(\wstart)$.
-Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, i.e.:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
+We can thus show how words are modified once we know they are substituted, that is how their features are modified by said substitution.
+Considering a word $\wstart$ substituted for $\warrival$, we measure how the feature of $\wstart$ varies when it is replaced with $\warrival$, that is we look at $\phi(\warrival)$ as a function of $\phi(\wstart)$.
+Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
 $$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
@@
 Il y avait une erreur, j'ai reformulé pour être plus clair.}
-In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
+In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, that is an arrival word chosen among semantically plausible though still random words.
 In this case $\warrival_{00}$ does depend on $\wstart$.
@@
 Both curves are significantly below their respective $\nu_{\phi}^0$ values, which is consistent with the literature on recall: words learned earlier and words with lower clustering coefficient are easier to produce than average~\citep{nelson2013activation,Zevin02}.
-Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; i.e.
-on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq 0.1$.
+Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; that is on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq 0.1$.
 

### 2015-03-05 brains-protocol.tex
0.x is .x is the number can't be greater than 1

@@
 
-The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
+The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($.8$).
 Our analysis showed clearer results with number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
@@
 Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
-Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Fig.~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
+Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Fig.~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
 
@@
 Network centrality properties, on the other hand, are strongly dependent on one another.
-As mentioned earlier, degree centrality and pagerank have a very strong correlation ($0.85$), and are also redundant with betweenness centrality (with correlation levels at $0.75$ and $0.68$ respectively).
+As mentioned earlier, degree centrality and pagerank have a very strong correlation ($.85$), and are also redundant with betweenness centrality (with correlation levels at $.75$ and $.68$ respectively).
 Furthermore, the three variables are also strongly related to age of acquisition, which leads us to keep the latter as the sole indicator for centrality.
@@
 
-Conversely, clustering coefficient exhibits low correlation levels with all the variables we kept (maximum absolute value $0.38$), leading us to include it in the rest of the analysis.
+Conversely, clustering coefficient exhibits low correlation levels with all the variables we kept (maximum absolute value $.38$), leading us to include it in the rest of the analysis.
 
@@
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
-Indeed, susceptibility goes from 0.33 for low-frequency words down to nearly 0 for very high-frequency words.
-To make things clear, this value of 0.33 means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.
+Indeed, susceptibility goes from $.33$ for low-frequency words down to nearly 0 for very high-frequency words.
+To make things clear, this value of $.33$ means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.
 
@@
 If these variables have an effect, it is by no means as strong as it is for Word frequency.
-This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively [0.13-0.18] and [0.16-0.20]).
-{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as the average of the other values (respectively 0.09 vs.
-0.16, and 0.11 vs.
-0.17).}
+This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively $[.13-.18]$ and $[.16-.20]$).
+{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as the average of the other values (respectively $.09$ vs.
+$.16$, and $.11$ vs.
+$.17$).}
 Keeping in mind the poor statistical significance of this effect, we could still wonder if the shortest words and words with fewest synonyms are significantly less susceptible to substitution.
@@
 %\TG{TBF: [[If we focus on cells based on at least several dozens of words]]} 
-Values tend to navigate around the mean value (0.16), even if there are a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).
+Values tend to navigate around the mean value ($.16$), even if there are a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).
 {On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
@@
 Both curves are significantly below their respective $\nu_{\phi}^0$ values, which is consistent with the literature on recall: words learned earlier and words with lower clustering coefficient are easier to produce than average~\citep{nelson2013activation,Zevin02}.
-Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; that is on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq 0.1$.
+Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; that is on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq .1$.
 
@@
 \end{quote}
-The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the dataset, and has a Clustering coefficient of 0.24.
-The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the dataset, and has a Clustering coefficient of 0.05.
+The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the dataset, and has a Clustering coefficient of $.24$.
+The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the dataset, and has a Clustering coefficient of $.05$.
 (Both words have four phonemes.)
@@
 Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e.
-$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age slightly below 8, and a Clustering coefficient of 0.1).
+$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age slightly below 8, and a Clustering coefficient of $.1$).

### 2015-03-05 brains-introduction.tex

@@
 A first stream is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
-It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
+It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citealp{Aunger00}, provides a solid overview), the development of evolutionary models of norms~\citep[see for instance][]{Ehrlich05} following the seminal work of \citet{Boyd85}; or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
 %\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
@@
 Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
-Despite some recent modeling attempts (\hbox{e.g.} \citet{Claidiere07}), the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.
+Despite some recent modeling attempts~\citep[e.g.][]{Claidiere07}, the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.
 
@@
 While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
-Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
+Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace~\citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media~\citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues~\citep{Cointet09}.
 
@@
 The practical study of the transformation of public representations has emerged only recently.
-For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citet{Claidiere07,claidiere2014darwinian} as well as a hybrid empirical-theoretical protocol in \citet{maccallum2012evolution}).
+For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citealp{Claidiere07}, and \citealp{claidiere2014darwinian}, as well as a hybrid empirical-theoretical protocol in \citealp{maccallum2012evolution}).
 Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
@@
 At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations for public representation evolution by studying the influence of word features on the ease of recall.
-This field is well developed and details the impact that classical psycholinguistic variables such as word frequency (see \citet{Yonelinas02} for a review), age-of-acquisition \citep{Zevin02}, number of phonemes or number of syllables (see for instance \citet{Rey98} and \citet{nick-diss}), have in this type of task.
+This field is well developed and details the impact that classical psycholinguistic variables such as word frequency~\citep[see][for a review]{Yonelinas02}, age-of-acquisition~\citep{Zevin02}, number of phonemes or number of syllables~\citep[see for instance][]{Rey98,nick-diss}, have in this type of task.
 
-Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
+Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models~\citep[see for instance][]{collins1975spreading}.
 Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
@@
 On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
-On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible (see again~\citet{Yonelinas02} for a review).
+On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{Yonelinas02}.
 The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.

### 2015-03-05 brains-protocol.tex

@@
 We used a quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
-This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009) \cite{Leskovec09-url}.\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
+This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009~---~\citealp{Leskovec09-url}).\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
 Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person. For instance,
@@
 We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
-A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations (see \citet{Leskovec09} for more detail).
+A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see][for more detail]{Leskovec09}.
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
@@
 63.9\%), containing a total of 127,778 unique quotes (out of 310,457; i.e.
-41.2\%) making up $\sim$2.43m occurrences (out of 8.16m, i.e.
+41.2\%) making up about 2.43m occurrences (out of 8.16m, i.e.
 29.8\%).\footnote{The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.}
@@
     \item \textbf{Word frequency}: the frequency at which words appear in our dataset, known to be relevant for both recognition and recall~\citep{gregg1976word},
-    \item \textbf{Age of Acquisition}: the average age at which words are learned (obtained from~\citet{Kuperman12}), known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
-    \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary~\citep{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
-    \item The average \textbf{Number of Synonyms} for all meanings of a word (obtained from \citet{WordNet10}) as an \emph{a priori} indicator of how easy it would be to replace a word.
+    \item \textbf{Age of Acquisition}: the average age at which words are learned~\citep[obtained from][]{Kuperman12}, known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
+    \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary, \citealp{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
+    \item The average \textbf{Number of Synonyms} for all meanings of a word~\citep[obtained from][]{WordNet10} as an \emph{a priori} indicator of how easy it would be to replace a word.
 \end{APAitemize}
@@
 %, we thus considered more recently studied variables based on semantic network properties.
-We relied on the ``free association'' (FA) norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
+We relied on the ``free association'' (FA) norms collected by \citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
 FA norms record the words that come to mind when someone is presented with a given cue% (that is the ``free association'' task)
@@
 \begin{APAitemize}
-    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
+    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by \citet{Griffiths07}.
     In the present case these two polysemy-related measures are quasi-perfectly correlated. %, and measure polysemy by taking into account the structure of the entire network.
@@
     This quantity tells us if some words behave like unavoidable waypoints on association chains connecting one word to another.
-    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual versus possible edges between a node's neighbors.}
+    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\citep{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual versus possible edges between a node's neighbors.}
     This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
@@
 The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($.8$).
-Our analysis showed clearer results with number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.
+Our analysis showed clearer results with number of phonemes over number of syllables, which is consistent with \citet{nick-diss}, and we therefore chose to only present results for the former.
 

### 2015-03-05 brains-introduction.tex
+keywords +authors +validation +minor changes

@@
 
-\TB{The next section %(Sec.~\ref{sec:related}) 
+The next section %(Sec.~\ref{sec:related}) 
 describes our hypotheses along with the relevant state-of-the-art on this psycholinguistic matter.
 %In Sec.~\ref{sec:protocol}, 
-In the section that follows it, 
-we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
+Then, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
 %Sec.~\ref{sec:results} 
-The results section describes the significant psycholinguistic biases observed during \emph{in vivo} quotation reformulation as well as their epidemiological setting, followed by a discussion and general guidelines for further work in the final section.%Sec.~\ref{sec:conclusion}.
-}
+We further describe the significant psycholinguistic biases observed during \emph{in vivo} quotation reformulation as well as their epidemiological setting, followed by a discussion and general guidelines for further work in the final section.%Sec.~\ref{sec:conclusion}.
 

### 2015-03-05 brains-protocol.tex
+keywords +authors +validation +minor changes

@@
 %We contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion, 
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (such as ``the city is tired of me and the organization and I have run our course together''), we check that our approach essentially fulfills its goals by manually coding a random subsample of 100 excluded clusters: a solid 71\% appear to be entirely irrelevant to our analysis (everyday language rather than quotations), and all but one of the remaining clusters were of relevance to the protocol \TB{set out below.}% in Sec.~\ref{sec:model}.
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (such as ``the city is tired of me and the organization and I have run our course together''), we check that our approach essentially fulfills its goals by manually coding a random subsample of 100 excluded clusters: a solid 71\% appear to be entirely irrelevant to our analysis (everyday language rather than quotations), and all but one of the remaining clusters were of relevance to the protocol {set out below.}% in Sec.~\ref{sec:model}.
 

### 2015-03-05 brainscopycut.tex
Add further authornote, keywords, and note fields
Change \ref's to named references or remove them (apastyle does not number sections)
+keywords +authors +validation +minor changes

@@
 % Formatting bibliography
-\usepackage{natbib}
-\setcitestyle{authoryear,round,semicolon}
+%\usepackage{natbib}
+\usepackage[natbibapa]{apacite}
+%\setcitestyle{authoryear,round,semicolon}
 
@@
 \date{}
-\author{Sébastien Lerique \& Camille Roth}
-\affiliation{Centre d'Analyse et de Mathématique Sociales () \& Centre Marc Bloch ()}
+\twoauthors{Sébastien Lerique}{Camille Roth}
+\twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av.
+de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE / USR 3130\\Friedrichstr 191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE / USR 3130\\Friedrichstr 191, D-10117 Berlin}
 \abstract{We describe reformulation processes within a large distributed system such as blogspace;
@@
 Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.}
+\authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
+\keywords{word production; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; in vivo psycholinguistics}
+%\note{\TB{Note XXXX (version? date?)}}
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{apacite}
 \bibliography{brainscopycut.bib}

### 2015-03-06 brainscopycut.tex
Bibliography: +caps, +consistent page numbers, -doi
Italics for latin

@@
 %\usepackage{natbib}
-\usepackage[natbibapa]{apacite}
+\usepackage[natbibapa,nodoi]{apacite}
 %\setcitestyle{authoryear,round,semicolon}
@@
 In particular, we uncover some of the semantic and structural characteristics of individual words and the substitutions they undergo.
-Our work amounts to a large in vivo experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
+Our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
 We show that all variables remarkably exhibit a single attractor and are generally contractile.
@@
 \authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
-\keywords{word production; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; in vivo psycholinguistics}
+\keywords{word production; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; \emph{in vivo} psycholinguistics}
 %\note{\TB{Note XXXX (version? date?)}}

### 2015-03-06 brains-conclusion.tex
Phrasing fixes, blue text to validate

@@
 For each of the selected psycholinguistic variables, we demonstrate the existence of attractor values in the underlying variable spaces.
-More precisely, beyond the interpretation of our results for each variable, we notice that all variables remarkably exhibit a single attractor and are generally contractile --- as such, even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
+More precisely, beyond the interpretation of our results for each variable, we notice that all variables remarkably exhibit a single attractor and are generally contractile --- as such, even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are susceptible to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
 

### 2015-03-06 brains-introduction.tex
Phrasing fixes, blue text to validate

@@
 A first stream is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
-It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citealp{Aunger00}, provides a solid overview), the development of evolutionary models of norms~\citep[see for instance][]{Ehrlich05} following the seminal work of \citet{Boyd85}; or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
+It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citealp{Aunger00}, provides a solid overview), the development of evolutionary models of norms~\citep[see for instance][]{Ehrlich05} following the seminal work of \citet{Boyd85}; or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all kinds of cultural artifacts: texts, utterances, etc.).
 %\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
@@
 At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations for public representation evolution by studying the influence of word features on the ease of recall.
-This field is well developed and details the impact that classical psycholinguistic variables such as word frequency~\citep[see][for a review]{Yonelinas02}, age-of-acquisition~\citep{Zevin02}, number of phonemes or number of syllables~\citep[see for instance][]{Rey98,nick-diss}, have in this type of task.
+This field is well developed and details the impact that classical psycholinguistic variables such as word frequency~\citep[see][for a review]{Yonelinas02}, age-of-acquisition~\citep{Zevin02}, number of phonemes or number of syllables~\citep[see for instance][]{Rey98,nick-diss}, have on this type of task.
 
@@
 On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
-On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central -- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
+On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
 On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{Yonelinas02}.

### 2015-03-06 brains-protocol.tex
Phrasing fixes, blue text to validate

@@
 
-We could therefore examine the individual substitution process at work when authors transform quotations, by examining the features of the substituted and substituting words in each substitution.
-To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+\TB{We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.} To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
@@
 We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
-A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see][for more detail]{Leskovec09}.
+A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see][for more details]{Leskovec09}.
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
@@
 Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
-Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source~(``{\sc Ctrl-C}/{\sc Ctrl-V}'').
 Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~for instance drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
@@
 They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
-Indeed, susceptibility goes from $.33$ for low-frequency words down to nearly 0 for very high-frequency words.
+Indeed, susceptibility goes from $.33$ for low-frequency words down to nearly~0 for very high-frequency words.
 To make things clear, this value of $.33$ means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.
@@
 %\TG{TBF: [[If we focus on cells based on at least several dozens of words]]}
-Values tend to navigate around the mean value ($.16$), even if there are a few outlier cells, yet with little obvious regularity (except for a low number of synonyms, consistently with the unidimensional graph).
+\TB{Even if there are a few outlier cells, values tend to navigate around the mean value ($.16$) with little obvious regularity (except for a low number of synonyms, consistent with the unidimensional graph)}.
 {On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
@@
 
-All in all, apart from Word frequency and despite some local tendencies, {in general} these results do not allow us to conclude either to a marked effect %or to no effect 
+All in all, apart from Word frequency and despite some local tendencies, {in general} these results do not allow us to conclude \TB{\sout{either}} to a marked effect %or to no effect
 of the selected psycholinguistic features on substitution susceptibility.
@@
 We can thus show how words are modified once we know they are substituted, that is how their features are modified by said substitution.
-Considering a word $\wstart$ substituted for $\warrival$, we measure how the feature of $\wstart$ varies when it is replaced with $\warrival$, that is we look at $\phi(\warrival)$ as a function of $\phi(\wstart)$.
+Considering a word~$\wstart$ substituted for~$\warrival$, we measure how the feature of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
 Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}

### 2015-03-09 brains-protocol.tex
thoroughly validated RC1

@@
 
-\TB{We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.} To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+{We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.} To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
@@
 %\TG{TBF: [[If we focus on cells based on at least several dozens of words]]}
-\TB{Even if there are a few outlier cells, values tend to navigate around the mean value ($.16$) with little obvious regularity (except for a low number of synonyms, consistent with the unidimensional graph)}.
+{Even if there are a few outlier cells, values tend to navigate around the mean value ($.16$) with little obvious regularity (except for a low number of synonyms, consistent with the unidimensional graph)}.
 {On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
@@
 
-All in all, apart from Word frequency and despite some local tendencies, {in general} these results do not allow us to conclude \TB{\sout{either}} to a marked effect %or to no effect
+All in all, apart from Word frequency and despite some local tendencies, {in general} these results do not allow us to conclude to a marked effect %or to no effect
 of the selected psycholinguistic features on substitution susceptibility.

### 2015-03-09 brainscopycut.tex
Phrasing fixes, blue text to validate
thoroughly validated RC1

@@
 %\newcommand{\defeq}{\overset{\underset{\mathrm{def}}{}}{=}}
+
+% Strikethrough (with the \sout comment)
+\usepackage[normalem]{ulem}
 
@@
 \twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av.
-de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE / USR 3130\\Friedrichstr 191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE / USR 3130\\Friedrichstr 191, D-10117 Berlin}
+de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE\\Friedrichstr.
+191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE\\Friedrichstr.
+191, D-10117 Berlin}
 \abstract{We describe reformulation processes within a large distributed system such as blogspace;

### 2015-03-10 brains-introduction.tex
Also cite Kleinberg on emails

@@
 While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
-Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace~\citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media~\citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues~\citep{Cointet09}.
+Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace~\citep{Gruhl04} or the email network~\citep{Liben-Nowell25032008}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media~\citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues~\citep{Cointet09}.
 

### 2015-03-10 brainscopycut.tex
thx-jp

@@
 \twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av.
-de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE\\Friedrichstr.
-191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE\\Friedrichstr.
+de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr.
+191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr.
 191, D-10117 Berlin}
@@
 
-We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Telmo Menezes, Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard for useful suggestions and comments.
+We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Telmo Menezes, Jean-Philippe Cointet, Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard for useful suggestions and comments.
 

### 2016-05-24 brains-protocol.tex
chore/build: disable some figures for build to succeed

@@
 
-\begin{figure}[!th]
-    \centering
-    \includegraphics[width=\linewidth]{images/computed-figures/correlations-filter1.png}
-    \caption{Spearman correlations in the initial set of features}
-    \label{fig:feature-corrs-initial}
-\end{figure}
+%\begin{figure}[!th]
+%    \centering
+%    \includegraphics[width=\linewidth]{images/computed-figures/all-feature_correlations.png}
+%    \caption{Spearman correlations in the initial set of features}
+%    \label{fig:feature-corrs-initial}
+%\end{figure}
 
@@
 
-\begin{figure}[!th]
-    \centering
-    \includegraphics[width=0.6975\linewidth]{images/computed-figures/correlations-filter2.png}
-    \caption{Spearman correlations in the filtered set of features}
-    \label{fig:feature-corrs-filtered}
-\end{figure}
+%\begin{figure}[!th]
+%    \centering
+%    \includegraphics[width=0.6412\linewidth]{images/computed-figures/paper-feature_correlations-v2.png}
+%    \caption{Spearman correlations in the filtered set of features}
+%    \label{fig:feature-corrs-filtered}
+%\end{figure}
 
@@
 
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
-    \label{fig:feature-susceptibilities}
-\end{figure*}
+%\begin{figure*}[!th]
+%    \centering
+%    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
+%    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+%    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
+%    \label{fig:feature-susceptibilities}
+%\end{figure*}
 \medskip
@@
 \end{seriate}
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-    \caption{\textbf{Feature variation upon substitution:} $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
-    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
-    The intersection with $y = x$ marks the attractor value.
-    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.
-    }
-    \label{fig:feature-variations}
-\end{figure*}
+%\begin{figure*}[!th]
+%    \centering
+%    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
+%    \caption{\textbf{Feature variation upon substitution:} $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+%    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
+%    The intersection with $y = x$ marks the attractor value.
+%    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.
+%    }
+%    \label{fig:feature-variations}
+%\end{figure*}
 

### 2016-05-24 brainscopycut.tex
Use a readable document style for working
chore/tools: update commands for notes
fix/title: tentative new title
chore/todos: add todo

@@
 % General document style ------------------------------------------------------------------
-\documentclass[man]{apa6}
+\documentclass[jou]{apa6}
+%\documentclass[man]{apa6}
 \usepackage[utf8]{inputenc}
@@
 % To make comments, corrections, notes
-\newcommand{\TB}[1]{\textcolor{blue}{#1}}
-\newcommand{\new}[1]{\TB{#1}}
-\newcommand{\tb}[1]{\textcolor{blue}{#1}}
-\newcommand{\TG}[1]{\textcolor{MidnightBlue}{#1}}
-\newcommand{\add}[1]{\TB{TOADD: #1}}
+\newcommand{\tbb}[1]{\fcolorbox{blue}{lightblue}{#1}}
+\newcommand{\tbg}[1]{\fcolorbox{blue}{lightgreen}{#1}}
+\newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
+\newcommand{\new}[1]{\tg{#1}}
+\newcommand{\add}[1]{\tbb{ADD: #1}}
 \newcommand{\ra}[1]{$\rightarrow$#1}
-\newcommand{\replace}[2]{\sout{#1} \TB{#2}}
-\newcommand{\rk}[1]{\TG{{\footnotesize {\bf[\emph{#1}]}}}}
-\newcommand{\remark}[1]{\noindent\fbox{\parbox{.99\linewidth}{\TB{#1}}}}
-\newcommand{\smallremark}[1]{\noindent\fbox{{\TB{#1}}}}
-\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\TB{\scriptsize \sf #1}}}}}
-\newcommand{\cam}[1]{\textcolor{blue}{\small{[{\bf C:} }#1{]}}}
-\newcommand{\seb}[1]{\textcolor{blue}{[{\bf S:} #1{]}}}
-\newcommand{\todo}{\fcolorbox{blue}{lightblue}{(...)}}
+\newcommand{\replace}[2]{\sout{#1} \tbb{#2}}
+\newcommand{\rk}[1]{\tbg{{\footnotesize {\bf[\emph{#1}]}}}}
+\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\tbb{\scriptsize \sf #1}}}}}
+\newcommand{\cam}[1]{\tbb{\small{[{\bf C:} }#1{]}}}
+\newcommand{\seb}[1]{\tbb{\small{[{\bf S:} #1{]}}}}
+\newcommand{\todo}[1]{\tbb{#1}}
 \definecolor{lightblue}{rgb}{.85,.85,1}
+\definecolor{lightgreen}{rgb}{.85,1,.85}
 
@@
 
-\title{How do we copy and paste? The semantic drift of quotations in blogspace}
+\title{The semantic drift of quotations in blogspace: a case study in short-term cultural evolution}
 \shorttitle{The semantic drift of quotations in blogspace}
@@
 We show that all variables remarkably exhibit a single attractor and are generally contractile.
-Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.}
+Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
+\todo{Update after introduction rework.}}
 \authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}

### 2016-05-26 brains-conclusion.tex
chore/todos: note all github issues in the main text

@@
 \section{Concluding remarks}\label{sec:conclusion}
+
+\todo{\#14: link to introduction discussion, mentioning this can be a model system}
+
+\todo{\#15: relate to missed literature}
+
 We aimed to contribute to the empirical understanding of representation transformation processes %\marginnote{fixed after Telmo saying: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?}
@@
 %--- in this sense we also diverge from psycholinguistic experiments that focus on ease of recall.
+
+\todo{\#18: tone down claims about contractile: it's a possible hypothesis if this were the only process, but not observed with the mix of all other processes}
 

### 2016-05-26 brains-introduction.tex
chore/todos: note all github issues in the main text

@@
 % ============================
+
+\todo{
+\#14: Rework to
+(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
+(2) present current problem of testing theories to incorporate those discoveries,
+(3) present big data literature which can bring measurement tools,
+(4) introduce the question of cultural attractors,
+(5) to be studied through quotes on the web: it's the proposal of a novel approach, we'll be describing and cross-checking with existing literature (vs.
+predicting predict, since there are too many possible things to predict)
+}
 
@@
 
+\todo{
+\#15: add
+(1) sentence recall (Potter \& Lombardi),
+(2) false memories (Deese),
+(3) subjective organization (Tulving, Zaromb),
+(4) working memory and attention (Jefferies),
+(5) iterated learning (Kirby)
+}
+
 %The relevant literature on \emph{public representation dynamics} features two main streams.

### 2016-05-26 brains-protocol.tex
chore/todos: note all github issues in the main text

@@
 
-{We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.} To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+\todo{
+\#19: better explain choice of single-substitution case, referring to further down in (1) substitution detection and (2) susceptibility measure
+}
+
+We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
+To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
@@
 \subsection{\emph{In vivo} utterances}
+
+\todo{\#17: go into more detail on how quotes are selected by Leskovec et al. (or explain there's no info if there's none)}
 
@@
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
+
+\todo{\#17: precision/recall analysis on anti-spam and language}
 
@@
 \subsubsection{Psycholinguistic indices}
+
+\todo{\#16:
+(1) update the set of features,
+(2) better explain feature selection: no PCA or all-feature regression (a predictive model doesn't differentiate correlated features), it's robust feature selection with domain knowledge
+(3) note clustering values have changed because of weights
+}
 
@@
 
+\todo{\#19: show more clearly why this is more robust with single-substitution, justifying our choice of this case}
+
+\todo{\#17: explain all models and the work involved, with graphs in annex}
+
 We finally need a substitution detection model, for the utterance data we use presents a challenge: quote-to-quote transformations, and much less substitutions, are not explicitly encoded in the dataset.
@@
 \medskip
+
+\todo{\#12: fix substitution counts}
+
 Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Fig.~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.
@@
 
+\todo{\#19: explain if we're susceptible to bias from single-substitution}
+
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on that word or on another one.
@@
 \medskip
+
+\todo{\#16:
+(1) show POS don't differ,
+(2) note that stopwords are excluded, which excludes POS open/closed discussion
+}
+
+\todo{\#16: explain we do this on quartiles, rework results discussion}
+
+\todo{\#17: show other feature variations and other models in annex}
+
+\todo{\#15: relate to missed literature}
+
 Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
@@
 \medskip
+
+\todo{\#16:
+(1) add H00 lines to account for semantic similarity (nothing more fancy since it involves a threshold),
+(2) rework results discussion
+}
+
+\todo{\#17: show other feature variations and other models in annex}
+
+\todo{\#17: explain binning}
+
+\todo{\#18: tone down claims about contractile: it's a possible hypothesis if this were the only process, but not observed with the mix of all other processes}
+
+\todo{\#15: relate to missed literature}
+
 Results are gathered in Fig.~\ref{fig:feature-variations}.
@@
 
+\todo{\#16: add the taking-context-into-account parts, potentially moving the non-relative results discussions to here:
+(1) explain and add sentence-relative susceptibility by quartiles, which has a perfectly clear interpretation + results discussion,
+(2) explain and add sentence-relative variations + results discussion
+(3) link those results to non sentence-relative values
+}
+
+\todo{\#17: show other feature variations and other models in annex}
+
+\todo{\#15: relate to missed literature}
+
 \medskip

### 2016-05-30 brains-introduction.tex
chore/todos: update some todos from Gureckis' edited version

--- 2015-03-10 brains-introduction.tex
@@
 % ============================
+
+\todo{
+\#14: Rework to
+(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
+(2) present current problem of testing theories to incorporate those discoveries,
+(3) present big data literature which can bring measurement tools,
+(4) introduce the question of cultural attractors,
+(5) to be studied through quotes on the web: it's the proposal of an online approach, we'll be describing and cross-checking with existing literature (vs.
+predicting, since there are too many possible things to predict)
+}
+
+\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
 
@@
 
+\todo{
+\#15: add
+(1) sentence recall (Potter \& Lombardi),
+(2) false memories (Deese),
+(3) subjective organization (Tulving, Zaromb),
+(4) working memory and attention (Jefferies),
+(5) iterated learning (Kirby)
+}
+
+\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
+
 %The relevant literature on \emph{public representation dynamics} features two main streams.

### 2016-05-30 brains-protocol.tex
chore/todos: update some todos from Gureckis' edited version

@@
 \label{sec:protocol}
+
+\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
 
@@
 \end{quote}
+
+\todo{\#20: make clear we didn't do this, it's Leskovec et al.}
+
 Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
@@
 \medskip
+
+\todo{\#20: FA don't encode reformulation explicitly, they did norms. We hypothesize it's what's at work.}
+
 %, we thus considered more recently studied variables based on semantic network properties.
@@
 
-\todo{\#17: explain all models and the work involved, with graphs in annex}
+\todo{\#17: explain all models and the work involved, with graphs and explanatory figures for each model in annex}
 
@@
 \todo{\#16: explain we do this on quartiles, rework results discussion}
+
+\todo{\#16: explicit all significance claims, referring to the CIs}
 

### 2016-05-30 brains-conclusion.tex
chore/todos: rework some todos after rereading Cog. Sci. letter

@@
 
-\todo{\#14: link to introduction discussion, mentioning this can be a model system}
+\todo{\#14:
+link to introduction discussion:
+(1) this can be a model system,
+(2) convergence can be looked for in any dimension, but that doesn't make a theory, so Epidemiology of Representations is all nice, but:
+(3a) taken simplistically it predicts obviously simplistic stuff (all quotes CV. to a single quote),
+(3b) taken more realistically (many dimensions in life) it gives some ideas, but it's not clear it's a core principle
+(3c) we need more controlled investigation to see how relevant it is.
+}
 

### 2016-06-01 brains-introduction.tex
feat/introduction: first try at a new introduction

--- 2015-03-10 brains-introduction.tex
@@
 % ============================
+
+\todo{
+\#14: Rework to
+(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
+(2) present current problem of testing theories to incorporate those discoveries,
+(3) introduce the question of cultural attractors,
+(4) present big data literature which can bring measurement tools,
+(5) to be studied through quotes on the web: it's the proposal of an online approach, we'll be describing and cross-checking with existing literature (vs.
+predicting, since there are too many possible things to predict)
+}
+
+\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
+
+\begin{new}
+Since the very beginning of psychology, of social sciences, and later that of cognitive science, several authors have proposed frameworks to unify the study of cognition and culture in meaningful ways.
+In social science, the question already existed in Durkheim's initial works \CN, but was only later tackled in earnest by Mauss' \emph{Techniques of the Body} \CN, Giddens' \emph{Structuration Theory} \CN and Bourdieu's \emph{Sens Pratique} \CN.
+
+Today's debate of that question, however, is structured by proponents from cognitive science.
+There is, on one side, a theory allying neo-darwinism and representation-based cognitive science that is best summarized by Sperber in his description of \emph{Epidemiology of Representations} \CN, complementary to Boyd and Richerson's work on gene-culture co-evolution \CN.
+On the other side an enactive proposition which anthropologists like Ingold, in line with Mauss' works, are calling for \CN, is being developed by Froese, Di Paolo, and De Jaegher among others \CNs.
+
+The question is also gaining relevance in other fields, as work in evo-devo and non-genetic inheritance is accumulating evidence not accounted for by the modern synthesis \CN;
+these discoveries are creating demand for new or extended approaches to life evolution that unify its different levels, as well as creative empirical methods to test the predictions these approaches make \CN.
+
+\end{new}
 

### 2016-06-22 brains-conclusion.tex
feat/introduction: much more work on the new introduction

@@
 % !TEX root = brainscopycut.tex
+\section{Discussion}
+
+\todo{Discuss related to introduction. Attractors, lineage with specification, what we couldn't observe, how it fits into Kirby.
+}
+
+\todo{\#15: relate to missed literature}
+
+\begin{new}
+
+ADDTHIS:
+We also chose exploratory vs. predictive to give a detailed view of what happens and because there's too many possible things to predict.
+
+\end{new}
+
 \section{Concluding remarks}\label{sec:conclusion}
@@
 (3b) taken more realistically (many dimensions in life) it gives some ideas, but it's not clear it's a core principle
-(3c) we need more controlled investigation to see how relevant it is.
+(3c) we need more controlled investigation to fuel the discussion and see how relevant it is.
 }
 
-\todo{\#15: relate to missed literature}
+\begin{new}
+
+ADDTHIS:
+On the other side an enactive proposition which anthropologists like Ingold, in line with Mauss' works, are calling for \CN, is being developed by Froese, Di Paolo, and De Jaegher among others \CNs.
+
+The question is also gaining relevance in other fields, as work in evo-devo and non-genetic inheritance is accumulating evidence not accounted for by the modern synthesis \CN;
+these discoveries are creating demand for new or extended approaches to life evolution that unify its different levels, as well as creative empirical methods to test the predictions these approaches make \CN.
+
+\end{new}
 

### 2016-06-22 brains-introduction.tex
feat/introduction: much more work on the new introduction

--- 2016-05-30 brains-introduction.tex
@@
 
-\todo{
-\#14: Rework to
-(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
-(2) present current problem of testing theories to incorporate those discoveries,
-(3) present big data literature which can bring measurement tools,
-(4) introduce the question of cultural attractors,
-(5) to be studied through quotes on the web: it's the proposal of an online approach, we'll be describing and cross-checking with existing literature (vs.
-predicting, since there are too many possible things to predict)
+\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
+
+\begin{new}
+START SENTENCE.
+
+Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
+%the reciprocal influences of cognition and culture in meaningful ways.
+While it has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g.
+Mauss' \emph{Techniques of the Body}, \citeyear{mauss_sociology_1979}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.
+
+These construe culture as an evolutionary process analogous and parallel to biological evolution (and especially the modern synthesis' account of it).
+That analogy can be traced a long way back in the 20th century and earlier, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\citeyearpar{dawkins_selfish_2006}, and later the development of \emph{Dual inheritance theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
+More recently, Dan Sperber has drawn on this principle to explicitly connect anthropology and cognitive science through the theory of \emph{Epidemiology of Representations} \citep{sperber_explaining_1996}, and the study of cultural evolution has been growing steadily since.
+
+The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) have shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
+\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
+\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
+\Citeauthor{kirby_cumulative_2008} \citeyearpar{kirby_cumulative_2008,cornish_systems_2013} have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
+Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.
+
+The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
+these are alternatively located in the mind ("mental representation" in Sperber's terminology), or in the outer world ("public representations") as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
+A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
+Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to e.g.
+memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.
+
+The theory makes an additional strong hypothesis, which this paper focuses on:
+the accumulation of cognitive effects in transmission leads to the emergence of very stable representations (called "cultural representations") that spread throughout an entire society, a process which should manifest itself as attractors (called "cultural attractors") in the dynamical system that models that society.
+
+This hypothesis has been hard to test in concrete situations, as quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
+One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
+This paper exemplifies a second approach, taking advantage of the ever-increasing avalanche of available digital footprints since the 2000's.
+Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly (see for instance \citealp{gruhl_information_2004}, \citealp{liben-nowell_tracing_2008}, \citealp{cointet_socio-semantic_2009}, and \citealp{danescu-niculescu-mizil_you_2012}, for good examples).
+We believe these tools can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.
+
+\bigskip
+
+To do so we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
+These public representations should normally not change when they are copied from one website and published on another (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed \citep{simmons_memes_2011}.\footnote{
+It seems that authors do not always copy-paste, but copy quotes by hand instead.
 }
+Analyzing this process gives us information on (a) the way these public representations are transformed when they are transmitted, and (b) which cognitive biases are responsible for those transformations.
+By using an exploratory approach, our study gives a detailed view of one of the cognitive effects responsible for the evolution of online quotes, and shows that it is compatible with the hypothesis of cultural attractors.
 
-\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
+We chose to restrict our analysis to substitutions (one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
+As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
+This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
+}
+While this restricts the scope of our observations for this particular data set, the methodological point we also make is left intact.
+
+By characterizing substitutions with 6 features on the disappearing and appearing words, we identify what makes a substitution more likely, and how a word changes when it is substituted.
+Consistent with known effects in linguistics, we observe that low-frequency words and words learned later in development are more susceptible to substitution than other words.
+Looking at the context those words appear in, we observe a marked effect for substitution of extreme words in a sentence (either very high-valued or very low-valued features compared to sentence average, except for word frequency).
+Focusing on how words are transformed, we see that the appearing words have significantly higher frequency and lower age-of-acquisition than synonyms of the disappearing word.
+Finally, the patterns we observe are also consistent with an attraction of each of the features towards a (feature-specific) asymptotic value.
+
+While the transformations we describe are not the only ones at work in this data set, our analysis shows that the substitution process is both consistent with known biases and compatible with the existence of feature-specific attractors.
+This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, but with natural language out of the laboratory.
+
+The next section %(Sec.~\ref{sec:related})
+describes our hypotheses along with a review of the psycholinguistics literature.
+%In Sec.~\ref{sec:protocol},
+Then, we describe the data set and detail the various assumptions that were made in order to analyze it.
+%Sec.~\ref{sec:results}
+Next, we describe the measures built to observe the cognitive biases operating in quote transmission.
+Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.
+
+\end{new}
 

### 2016-06-22 brainscopycut.tex
feat/acknowledgements: acknowledge funding
chore/todos: note all github issues in the main text
chore/utils: new is an environment, not a command
fix/utils: \TB is now \tb
feat/introduction: much more work on the new introduction

@@
 % To make comments, corrections, notes
-\newcommand{\tbb}[1]{\fcolorbox{blue}{lightblue}{#1}}
-\newcommand{\tbg}[1]{\fcolorbox{blue}{lightgreen}{#1}}
+\newcommand{\tb}[1]{\textcolor{blue}{#1}}
 \newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
-\newcommand{\new}[1]{\tg{#1}}
-\newcommand{\add}[1]{\tbb{ADD: #1}}
+\newcommand{\newtext}[1]{\tg{#1}}
+\newenvironment{new}{\par\color{MidnightBlue}}{\par}
+\newcommand{\add}[1]{\tb{ADD: #1}}
 \newcommand{\ra}[1]{$\rightarrow$#1}
-\newcommand{\replace}[2]{\sout{#1} \tbb{#2}}
-\newcommand{\rk}[1]{\tbg{{\footnotesize {\bf[\emph{#1}]}}}}
-\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\tbb{\scriptsize \sf #1}}}}}
-\newcommand{\cam}[1]{\tbb{\small{[{\bf C:} }#1{]}}}
-\newcommand{\seb}[1]{\tbb{\small{[{\bf S:} #1{]}}}}
-\newcommand{\todo}[1]{\tbb{#1}}
+\newcommand{\replace}[2]{\sout{#1} \tb{#2}}
+\newcommand{\rk}[1]{\tb{{\footnotesize {\bf[\emph{#1}]}}}}
+\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\tb{\scriptsize \sf #1}}}}}
+\newcommand{\cam}[1]{\tb{\small{[{\bf C:} }#1{]}}}
+\newcommand{\seb}[1]{\tb{\small{[{\bf S:} #1{]}}}}
+\newcommand{\todo}[1]{\tb{#1}}
 \definecolor{lightblue}{rgb}{.85,.85,1}
@@
 % If citations or references are missing, make it visible
-\newcommand{\CN}{\textsuperscript{\TB{[Citation needed]}}}
-\newcommand{\CNs}{\textsuperscript{\TB{[Multiple citations needed]}}}
-\newcommand{\RN}{\textsuperscript{\TB{[Internal reference needed]}}}
+\newcommand{\CN}{\textsuperscript{\tb{[Citation needed]}}}
+\newcommand{\CNs}{\textsuperscript{\tb{[Multiple citations needed]}}}
+\newcommand{\RN}{\textsuperscript{\tb{[Internal reference needed]}}}
 
@@
 
-We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Telmo Menezes, Jean-Philippe Cointet, Jean-Pierre Nadal, Sharon Peperkamp, and Nicolas Baumard for useful suggestions and comments.
+We are warmly grateful to Ana Sofia Morais for her precious feedback and advice on this research, and to Telmo Menezes, Jean-Philippe Cointet, Jean-Pierre Nadal, Sharon Peperkamp, \newtext{Nicolas Claidière} and Nicolas Baumard for useful suggestions and comments.
+
+{\new This work has also been partially supported by the French National Agency of Research (ANR) through the grant Algopol (ANR-12-CORD-0018).}
 

### 2016-06-24 brains-introduction.tex
feat/introduction: rework introduction to not give out all the results

--- 2016-05-30 brains-introduction.tex
@@
 
-\todo{
-\#14: Rework to
-(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
-(2) present current problem of testing theories to incorporate those discoveries,
-(3) present big data literature which can bring measurement tools,
-(4) introduce the question of cultural attractors,
-(5) to be studied through quotes on the web: it's the proposal of an online approach, we'll be describing and cross-checking with existing literature (vs.
-predicting, since there are too many possible things to predict)
+\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
+
+\begin{new}
+Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
+%the reciprocal influences of cognition and culture in meaningful ways.
+While it has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g.
+Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.
+
+These construe culture as an evolutionary process analogous and parallel to biological evolution (and especially the modern synthesis' account of it).
+That analogy can be traced a long way back in the 20th century and earlier, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\citeyearpar{dawkins_selfish_2006}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
+More recently, Dan Sperber has drawn on this principle to explicitly connect anthropology and cognitive science through the theory of \emph{Epidemiology of Representations} \citep{sperber_explaining_1996}, and the study of cultural evolution has been growing steadily since.
+
+The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
+\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
+\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
+\Citeauthor{kirby_cumulative_2008} \citeyearpar{kirby_cumulative_2008,cornish_systems_2013} have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
+Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.
+
+The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
+these are alternatively located in the mind ("mental representations" in Sperber's terminology), or in the outer world ("public representations") as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
+A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
+Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to e.g.
+memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.
+
+The theory makes an additional strong hypothesis, which this paper focuses on:
+as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called "cultural representations", because they characterize a given culture).
+This process should manifest itself as attractors (called "cultural attractors") in the dynamical system that models cultural evolution, that is:
+there should be areas of the representation space where any cognitive effect in transformation brings representations closer to a given stable asymptotic point.\footnote{
+Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
 }
 
-\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
-
-The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling; from social cognition to data mining.
-These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part, investigate phenomena pertaining to both cognitive science and social science, both at the individual and social levels.
-
-Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.
-A first stream is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution.
-It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citealp{Aunger00}, provides a solid overview), the development of evolutionary models of norms~\citep[see for instance][]{Ehrlich05} following the seminal work of \citet{Boyd85}; or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all kinds of cultural artifacts: texts, utterances, etc.).
-%\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
-The producer and the user of a mental representation are one and the same person.
-A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}''.}
-
-One of the core claims of this literature consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized.
-Sperber's cultural epidemiology classically opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
-Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
-Despite some recent modeling attempts~\citep[e.g.][]{Claidiere07}, the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.
-
-Another research stream deals with rather macroscopic studies of knowledge diffusion.
-Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace.
-The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by each individual).
-This research program nonetheless exhibits a particularly interesting empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
-While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
-Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace~\citep{Gruhl04} or the email network~\citep{Liben-Nowell25032008}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media~\citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues~\citep{Cointet09}.
-
-These latter studies are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis) and are relatively remote from cognitive science; for a significant part, they rely rather marginally on specific social science theories.
-They nonetheless show us the added value of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
-%and, %more broadly, large textual corpora \cam{google n-grams??}
-towards radically improving the empirical understanding of (individual-level) cultural evolution processes.
+This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations:
+quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
+One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
+This paper exemplifies a second approach, taking advantage of the ever-increasing avalanche of available digital footprints since the 2000's.
+Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
+For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in the email network~\citep{liben-nowell_tracing_2008};
+\citet{cointet_socio-semantic_2009} have described the reciprocal influence between the social network topology and the distribution of issues;
+\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} have studied the characteristics of particularly memorable quotes that circulate in those networks.
+We believe those works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.
 
 \bigskip
-Stepping back, we thus observe a gap between, on one side, empirical studies of diffusion dynamics in social systems and, on the other side, more theoretical works  focused on knowledge transformation processes.
-%--- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.
-Our research lies at the intersection of these two programs, aiming to shed light on micro-level information transformation by leveraging the empirical wealth of (\emph{in vivo}) social diffusion phenomena.
-More precisely, we hope to describe reformulation processes within a large distributed system such as blogspace; showing how some specific types and features of public representations may be altered by bloggers when they freely reproduce them.
 
-We focus on simple linguistic modifications, thereby connecting our research to the broader psycholinguistic literature.
-To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
-While these verbatim public representations should in theory not suffer any alterations when they are produced anew (as opposed to more elaborate expressions and opinions, not identified as quoted utterances), empirical observation shows that they are occasionally transformed.
-We will in particular exhibit a non-trivial process by which individual words in quotations are replaced.
-We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
-In a way using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes a first step towards building empirically realistic models of cultural evolution.
+To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
+These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
+authors spontaneously transform quotes, not only cropping them but also replacing words, when in fact they are implicitly required to copy them exactly.
+We can therefore assume that most transformations, especially the simple ones, are the result of automatic (i.e. hard to control) low-level cognitive biases of the authors.
+
+Our question is as follows: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases referred to in the theory of epidemiology of representations, do cultural attractors appear and how do cognitive biases participate in them?
+We chose to restrict our analysis to substitutions (one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
+As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
+This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
+}
+While this restricts the scope of our observations for the particular data set we use, the methodological point we also make is left intact.
+By characterizing words with 6 well-known features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
+This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
+While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
+This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, but with natural language out of the laboratory.
 
 The next section %(Sec.~\ref{sec:related})
-describes our hypotheses along with the relevant state-of-the-art on this psycholinguistic matter.
+describes our hypotheses along with a review of the psycholinguistics literature.
 %In Sec.~\ref{sec:protocol},
-Then, we detail the empirical protocol and the various assumptions that were made in order to deal with the available empirical material.
+Then, we describe the data set and detail the various assumptions that were made in order to analyze it.
 %Sec.~\ref{sec:results}
-We further describe the significant psycholinguistic biases observed during \emph{in vivo} quotation reformulation as well as their epidemiological setting, followed by a discussion and general guidelines for further work in the final section.%Sec.~\ref{sec:conclusion}.
+Next, we describe the measures built to observe the cognitive biases operating in quote transmission.
+Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.
+
+\end{new}
 

### 2016-06-24 brains-conclusion.tex
chore/todos: add some notes that were factored out of the introduction

@@
 We also chose exploratory vs. predictive to give a detailed view of what happens and because there's too many possible things to predict.
+
+ADDTHIS:
+By characterizing substitutions with 6 features on the disappearing and appearing words, we identify what makes a substitution more likely, and how a word changes when it is substituted.
+Consistent with known effects in linguistics, we observe that low-frequency words and words learned later in development are more susceptible to substitution than other words.
+Looking at the context those words appear in, we observe a marked effect for substitution of extreme words in a sentence (either very high-valued or very low-valued features compared to sentence average, except for word frequency).
+Focusing on how words are transformed, we see that the appearing words have significantly higher frequency and lower age-of-acquisition than synonyms of the disappearing word.
+Finally, the patterns we observe are also consistent with an attraction of each of the features towards a (feature-specific) asymptotic value.
+
+ADDTHIS:
+It is possible however, that these attractors appear due to an interaction between biases and sentence context, making it a contingency rather than a rule.
+This is not really dealt with (context, aside from relevance) by Sperber.
 

### 2016-06-28 brains-conclusion.tex
feat/related-work: major rework of the "related work" part

@@
 This is not really dealt with (context, aside from relevance) by Sperber.
+
+Attraction can also be defined on any number of dimensions.
+It can be on the structure, on anything, so saying there could be an attraction while not specifying the dimension is really meaningless.
+What's more important is to look at a specific dimension, and see if there's attraction on that one, as we did here for features.
+
+We could've done also on semantic grouping, predicting the new word based on semantic similarity (or on frequent dyads, i.e.
+collocation with previous word the same way Zaromb et al.
+2006 explain PLIs), and predicting the disappearing word based on the cost of doing such a substitution (lower cost -> higher prob of substitution).
+The point is, there's decades and many fields of psycholinguistics, and we can connect each of them with this question.
+
+All of this is possible with our software that we published.
+
+ADDTHIS:
+Taking context into account is more than what we did.
+For instance, building on Zaromb 2006, you could imagine that the substitutions appear because the word preceding the substituted one appears in another dyad a lot more than this one, triggering a substitution (Zaromb's associative vs.
+contextual retrieval processes in recall).
+
+ADDTHIS:
+Soooooo... following literature on word lists (Zaromb and DRM):
+- we could predict the new word in substitution? Take the strongest average association to words in the sentence.
+- we could predict substituted word? Take the word following the one that triggers the new word.
+Problems:
+- we probably won't find the exact word, but one similar to it (even Zaromb can't predict the exact word, they don't try, they just check it comes from the right list). How to evaluate that?
+- if computing LSA/LDA on the corpus (which probably isn't adapted because of the short sentence nature of the data -> topics = quote families), it's tautological unless you suppose substitutions have a negligible effect.
+Explain that to the reviewer.
+
+Again justify our approach w/ features by the shape of our data: few substitutions per cluster (avg.
+9), and substitutions on relatively few clusters overall -> opposite situation to a few lists repeated through 50 subjects where frequency of transition has meaning.
+Here for each word, it's nearly one shot.
+So we have to categorize words (by using low number of features -> known features best) to get frequencies.
+From there, you can predict many many things, so better describe.
 

### 2016-06-28 brains-introduction.tex
feat/related-work: major rework of the "related work" part

@@
 \begin{new}
+
+TOO GENERAL OR GIVES THE IMPRESSION OF TOO HIGH AMBITIONS. COMPARE TO OTHER INTROS TO JUST SAY WE'LL TALK ABOUT THIS.
+
 Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
 %the reciprocal influences of cognition and culture in meaningful ways.
-While it has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g.
+While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g.
 Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.
@@
 \Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
-\Citeauthor{kirby_cumulative_2008} \citeyearpar{kirby_cumulative_2008,cornish_systems_2013} have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
+\citeauthor{kirby_cumulative_2008}~(\citeyear{kirby_cumulative_2008,cornish_systems_2013}) have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
 Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.
@@
 This process should manifest itself as attractors (called "cultural attractors") in the dynamical system that models cultural evolution, that is:
-there should be areas of the representation space where any cognitive effect in transformation brings representations closer to a given stable asymptotic point.\footnote{
+there should be areas of the representation space where cognitive effects in transformations bring representation closer to a given stable asymptotic point.\footnote{
 Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
@@
 
-This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations:
-quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
+This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations as quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
 One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
@@
 
-\todo{
-\#15: add
-(1) sentence recall (Potter \& Lombardi),
-(2) false memories (Deese),
-(3) subjective organization (Tulving, Zaromb),
-(4) working memory and attention (Jefferies),
-(5) iterated learning (Kirby)
-}
-
 \todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
 
-%The relevant literature on \emph{public representation dynamics} features two main streams.
-%On one hand, we find studies of the macroscopic \emph{social diffusion} of public representations, describing for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
-%These studies are relatively independent from anthropology and cognition and are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
-%Without necessarily relying on specific social science theories, this research stream is of interest for its use of large textual corpora in studying cultural dynamics.
+The study of \newtext{cultural evolution on the part of cognitive science} emerged only recently.
+\newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
+The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
+Critics like \citet{ingold_trouble_2007}, however, have noted that such quantitative use of works from the social sciences approach risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
+}
+A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
+Transmission chains, in particular, have been used extensively to study the evolution of human language~\citep[see][for a review]{tamariz_cultural_2016}.
+Other recent examples include studies of the evolution of simple audio loops through consumer preference~\citep{maccallum_evolution_2012}, the emergence of structure in visual patterns transmitted by baboons~\citep{claidiere_cultural_2014}, and the amplification of risk perception through chains of casual conversation~\citep{moussaid_amplification_2015}.}
 
-The practical study of the transformation of public representations has emerged only recently.
-For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citealp{Claidiere07}, and \citealp{claidiere2014darwinian}, as well as a hybrid empirical-theoretical protocol in \citealp{maccallum2012evolution}).
-Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
-%They exhibit several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint.
-One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations.
-However, these studies address transformations by focusing on the properties of the source of the quotation (\hbox{e.g.} news outlet {vs.} blog), or the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
-One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations.
-They essentially examine the effect of some properties of the quotation source (\hbox{e.g.} news outlet {vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus).
-Some diffusion-transformation models have been proposed, yet the very cognitive features which may determine or, at least, influence these transformations, are overlooked; which may appear to be relatively unsatisfying from a cognitive viewpoint.
+\newtext{Research on online content points to a third approach to this question.}
+By investigating the transformations of quotations in a large corpus of US blog posts and online news stories initially collected and studied by \citet{leskovec_meme-tracking_2009}, \citet{simmons_memes_2011} and later \citet{omodei_multi-level_2012} show that even for quotations, a type of public representation that should change the least when transmitted on the Web, it is still possible to witness significant transformations.
+These studies focus on the influence of the quotation source (e.g.
+news outlet vs.
+blog) or of the surrounding public space (e.g.
+quotation frequency in the corpus), and suggest diffusion-transformation models to capture the dynamics of the population of quotations.
+But the cognitive features which may determine or, at least, influence these transformations, are overlooked.
+\newtext{Yet cognitive and linguistic features have been used in studies not involving transformation: \citet{danescu-niculescu-mizil_you_2012}, for instance, show that particularly memorable quotations (taken from movie scripts in this case) use more distinctive words and have more common syntax than less memorable quotations; they are also the quotes that adapt best to new contexts of use.
+One source of ideas to study the transformations of such quotes, then, might be the psycholinguistic literature studying word and sentence recall.}
 
-At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations for public representation evolution by studying the influence of word features on the ease of recall.
-This field is well developed and details the impact that classical psycholinguistic variables such as word frequency~\citep[see][for a review]{Yonelinas02}, age-of-acquisition~\citep{Zevin02}, number of phonemes or number of syllables~\citep[see for instance][]{Rey98,nick-diss}, have on this type of task.
+\begin{new}
 
-Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models~\citep[see for instance][]{collins1975spreading}.
-Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
-First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin
-alphabet.
-The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
-\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, that is the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
-A third psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
-\citet{nelson2013activation}, finally, show the importance of clustering coefficient in a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free assocation).
+\citet{potter_regeneration_1990} suggest that immediate recall of sentences is based on the retention of an unordered list of words which is then regenerated as a sentence at the moment of production.
+Priming recall with other words can lead to replacement in the recalled sentence if the primed words support the overall meaning of the sentence.
+Regenerated syntax can also be influenced by priming recall with another syntactic structure~\citep{potter_syntactic_1998}, or with verbs whose category constraints call for a different structure~\citep{lombardi_regeneration_1992}.
 
-On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
-On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
-On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{Yonelinas02}.
+Recall of word lists provides a situation that is easier to explore fully compared to recall of complete sentences, and has been extensively studied.
+In particular, the Deese, Roediger, and McDermott paradigm (introduced by \citealp{deese_prediction_1959}, and later popularized by \citealp{roediger_creating_1995}) has shown that it is possible to construct lists of words which reliably create the false memory of an external word related to those in the list.
+This is done by using lists of words produced by free association from the target intrusion word;
+the intruding recall then happens with probability nearly proportional to the average strength of semantic association between the intruding word and the words in the list.
+A sizable literature studies this type of task with varying complexities in the design of the lists, a good review of which is given by \citet{zaromb_temporal_2006}.
+One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which subjects recall a list~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}
+The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
+indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations (contributing to associative information retrieval) of subjects is at the center of many of these studies.
+
+These effects don't transpose simply to sentence recall however, as not only syntax but also effects of attention come into play for both retrieval and encoding.
+\citet{jefferies_automatic_2004}, for instance, show that attention is central to the encoding and retention of unrelated propositions, on top of more automatic syntactic and semantic processes.
+This involvement of executive resources also seems to contribute to the much greater memory span subjects exhibit for sentences compared to word lists~\citep[see][again, for more details]{jefferies_automatic_2004}.
+
+Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited when the data consists of a high number of measures over a smaller number of lists (in which case it makes sense to ask e.g.
+what proportion of intrusions come from prior lists).
+As is explained further down however, our data set is shaped the opposite way: a great number of sentences, with only very few to no measures at all on each sentence.
+}
+If a cognitive bias in the substitution of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.
+
+Lexical features, then, are obvious well-studied word measures that can be analyzed in aggregate.
+Indeed word frequency~\citep[see][for a review]{yonelinas_nature_2002}, age-of-acquisition~\citep{zevin_age_2002}, number of phonemes~\citep[see for instance][]{rey_phoneme_1998,nickels_dissociating_2004}, and phonological neighborhood density~\citep{garlock_age--acquisition_2001} to name a few, all have known effects on word recognition or production.
+More complex features based on word networks built from free association or phonological data have also been analyzed:
+\citet{nelson_how_2013} for instance, show the importance of clustering coefficient in such a network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free association).
+\citet{chan_network_2010} show that pictures are named faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network.
+\citet{griffiths_google_2007} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the alphabet.
+The authors show that there is a link between the ease of recall of words and their authority position (pagerank) in a language-wide semantic network built from external word association data~(\citealp{austerweil_human_2012}, further develop this tool to give a parsimonious account of the fact that related words are often retrieved together from memory).
+
+\end{new}
+
+On the whole, research on lexical features hints towards two antagonistic types of effects.
+On one hand, part of the literature shows that recall is easier for the least "awkward" words;
+those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
+On the other hand, when the task consists in recognizing a specific item in a list, "awkward" words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
 The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
-We also aim here at shedding some light on this debate, considering oddness as a dimension of the purported fitness of utterances.
+We also aim to shed some light on this debate, considering oddness as a dimension of the purported fitness of utterances.

### 2016-06-30 brains-introduction.tex
feat/methods: rewrite up to 'word-level measures' (excluded)
Also fix some typos and formatting things in the introduction (using
\hbox{} for "i.e.", "e.g.", "vs.").

@@
 %the reciprocal influences of cognition and culture in meaningful ways.
-While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g.
-Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.
+While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.
 
@@
 A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
-Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to e.g.
-memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.
+Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.
 
@@
 authors spontaneously transform quotes, not only cropping them but also replacing words, when in fact they are implicitly required to copy them exactly.
-We can therefore assume that most transformations, especially the simple ones, are the result of automatic (i.e. hard to control) low-level cognitive biases of the authors.
+We can therefore assume that most transformations, especially the simple ones, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.
 
@@
 The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
-Critics like \citet{ingold_trouble_2007}, however, have noted that such quantitative use of works from the social sciences approach risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
+Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
 }
@@
 By investigating the transformations of quotations in a large corpus of US blog posts and online news stories initially collected and studied by \citet{leskovec_meme-tracking_2009}, \citet{simmons_memes_2011} and later \citet{omodei_multi-level_2012} show that even for quotations, a type of public representation that should change the least when transmitted on the Web, it is still possible to witness significant transformations.
-These studies focus on the influence of the quotation source (e.g.
-news outlet vs.
-blog) or of the surrounding public space (e.g.
-quotation frequency in the corpus), and suggest diffusion-transformation models to capture the dynamics of the population of quotations.
+These studies focus on the influence of the quotation source (\hbox{e.g.} news outlet \hbox{vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), and suggest diffusion-transformation models to capture the dynamics of the population of quotations.
 But the cognitive features which may determine or, at least, influence these transformations, are overlooked.
@@
 
-Recall of word lists provides a situation that is easier to explore fully compared to recall of complete sentences, and has been extensively studied.
+Compared to full sentences, recall of word lists provides a situation that is easier to fully explore, and that has been extensively studied.
 In particular, the Deese, Roediger, and McDermott paradigm (introduced by \citealp{deese_prediction_1959}, and later popularized by \citealp{roediger_creating_1995}) has shown that it is possible to construct lists of words which reliably create the false memory of an external word related to those in the list.
@@
 A sizable literature studies this type of task with varying complexities in the design of the lists, a good review of which is given by \citet{zaromb_temporal_2006}.
-One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which subjects recall a list~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}
+One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which words are recalled~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}.
 The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
-indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations (contributing to associative information retrieval) of subjects is at the center of many of these studies.
+indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations of subjects (contributing to associative information retrieval) is at the center of many of these studies.
 
@@
 
-Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited when the data consists of a high number of measures over a smaller number of lists (in which case it makes sense to ask e.g.
-what proportion of intrusions come from prior lists).
+Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
+Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited to data consisting of a high number of measures over a smaller number of lists (in which case it makes sense to ask \hbox{e.g.} what proportion of intrusions come from prior lists).
 As is explained further down however, our data set is shaped the opposite way: a great number of sentences, with only very few to no measures at all on each sentence.
@@
 More complex features based on word networks built from free association or phonological data have also been analyzed:
-\citet{nelson_how_2013} for instance, show the importance of clustering coefficient in such a network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free association).
+\citet{nelson_how_2013} for instance, show the importance of clustering coefficient in such a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free association).
 \citet{chan_network_2010} show that pictures are named faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network.
@@
 On one hand, part of the literature shows that recall is easier for the least "awkward" words;
-those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
+those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in retrieval, that is in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
 On the other hand, when the task consists in recognizing a specific item in a list, "awkward" words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
-The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
+The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to rewrite a given utterance, is rather of the former or latter sort.
 We also aim to shed some light on this debate, considering oddness as a dimension of the purported fitness of utterances.

### 2016-06-30 brains-protocol.tex
chore/todos: rework some todos after rereading Cog. Sci. letter
feat/methods: rewrite up to 'word-level measures' (excluded)
Also fix some typos and formatting things in the introduction (using
\hbox{} for "i.e.", "e.g.", "vs.").

@@
 
-%In order to start bridging this gap, we set out to \emph{empirically} study public representation transformations at the microscopic level;
-%aiming to stay compatible with macroscopic-level studies of these public representations.
-Quotations appeared to be a perfect candidate to propose a first \emph{in vivo} measure of low-level cognitive bias in a reformulation task. % as public representations.
-First, they are usually cleanly delimited by quotation marks %(and often with HTML markup in web pages),
-which greatly facilitates their detection in text corpora.
-Second, they stem from a unique ``original'' version, and could ideally be traceable back to that version.
-Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.
-
-\todo{
-\#19: better explain choice of single-substitution case, referring to further down in (1) substitution detection and (2) susceptibility measure
-}
+Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
+First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
+Second, they stem from a unique original version, and are ideally traceable back to that version.
+Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping:
+not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.
 
 We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
-To keep the analysis palatable, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
-To quantify those substitutions, we decided to associate a number of features to each word, the variation of which we can statistically study.
+\newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
+Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions going on every day, with all the complexity and uncertainty of conditions that this entails.
+In this first study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations, as this would be akin to a drink from the firehose.
+Our goal, instead, is to show that such effects exist and are measurable even if they are part of a larger complexity, that is, we aim to prove first a relatively weak hypothesis before trying stronger ones.
+If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
+As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}
 
-The next subsections describe the dataset and measures we used to assess this cognitive bias.
-%Quotation evolution is therefore a perfect environment to measure cognition-induced transformations.%and relate those findings to macroscopic social dynamics.
+\newtext{With this goal of keeping} the analysis \newtext{tractable}, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see "Substitution model").}
+To quantify those substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
+
+The next subsections describe the data set and the measures we used to assess this cognitive bias.
 
@@
 
-\todo{\#17: go into more detail on how quotes are selected by Leskovec et al. (or explain there's no info if there's none)}
-
-We used a quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
-This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009~---~\citealp{Leskovec09-url}).\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
-Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person. For instance,
+We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
+This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 \newtext{(\citealp{leskovec_memetracker:_2009}; unfortunately, the original article, \citealp{leskovec_meme-tracking_2009}, does not provide further details on the source selection methodology).}
+\newtext{The authors automatically extracted quotations} from this corpus. Each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person; for instance:
 \begin{quote}
@@
 
-\todo{\#20: make clear we didn't do this, it's Leskovec et al.}
+\newtext{Then, they gathered quotations in a graph and connected each pair that differed by no more than one word or that shared at least ten consecutive words (they tested this procedure with a number of different parameters, see~\citealp{leskovec_meme-tracking_2009}, for more details).}
+We find for example the following variation of the above quote:
+\begin{quote}
+"these operations are \textbf{intended} to address funding pressures over quarter-end."
+\end{quote}
+\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
+This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been assigned to another set)}.
 
-Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words).
-We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
-A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see][for more details]{Leskovec09}.
-This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.
-
-\todo{\#17: precision/recall analysis on anti-spam and language}
-
-Manual inspection of this dataset revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
+Manual inspection of this data set revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
 Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
-To filter them out, we exclude all quotes having less than 5 words long or lasting more than 80 days %\seb{it is asymmetrical...}
- (as well as quotes not written in English).
-If an entire cluster still lasts more than 80 days after this screening (because of short-lived but unrelated quotes far apart in time), we also exclude it.
-% New filtering: 1051 substitutions out of 6172 mined. 45749 clusters. -- correct text with numbers and for new figures.
-We eventually keep 45,749 clusters (out of 71,568; i.e.
-63.9\%), containing a total of 127,778 unique quotes (out of 310,457; i.e.
-41.2\%) making up about 2.43m occurrences (out of 8.16m, i.e.
-29.8\%).\footnote{The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is expected for everyday language utterances.}
-%We contend that the use of the unfiltered dataset would strongly jeopardize results related to linguistic evolution or information diffusion,
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (such as ``the city is tired of me and the organization and I have run our course together''), we check that our approach essentially fulfills its goals by manually coding a random subsample of 100 excluded clusters: a solid 71\% appear to be entirely irrelevant to our analysis (everyday language rather than quotations), and all but one of the remaining clusters were of relevance to the protocol {set out below.}% in Sec.~\ref{sec:model}.
+To filter them out, we exclude quotes with less than 5 words or \newtext{whose occurrences span more} than 80 days (indicating causally unrelated occurrences), as well as quotes not written in English.
+\newtext{Clusters that are emptied by this procedure are therefore excluded.}
+If, after this screening, a cluster's occurrences still span more than 80 days (because of short-lived but unrelated quotes far apart in time), we also exclude it.
+\newtext{We eventually keep 50,427 clusters (out of 71,568; \hbox{i.e.} 70.5\%), containing a total of 141,324 unique quotes (out of 310,457; \hbox{i.e.} 45.5\%) making up about 2.60m occurrences (out of 7.67m; \hbox{i.e.} 33.9\%).\footnote{
+The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is to be expected for everyday language utterances.
+}
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is "the city is tired of me and the organization and I have run our course together"), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
+35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
+Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.\footnote{
+A similar analysis was made for language detection, which is part of the cluster filtering:
+out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
+Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
+}}
 
@@
 
-\subsubsection{Psycholinguistic indices}
+\subsubsection{Lexical features}
 
@@
 \begin{APAitemize}
-    \item \textbf{Word frequency}: the frequency at which words appear in our dataset, known to be relevant for both recognition and recall~\citep{gregg1976word},
+    \item \textbf{Word frequency}: the frequency at which words appear in our data set, known to be relevant for both recognition and recall~\citep{gregg1976word},
     \item \textbf{Age of Acquisition}: the average age at which words are learned~\citep[obtained from][]{Kuperman12}, known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
@@
 
-The final set of variables we consider, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.\footnote{Note that feature values stem from different datasets which do not always encode the same words.
+The final set of variables we consider, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.\footnote{Note that feature values stem from different data sets which do not always encode the same words.
 Indeed, we have data on frequency for about 22.6k words, on age of acquisition for 30.1k words, on number of phonemes for 123.4k words, number of synonyms 111.2k, and clustering coefficient 5.7k words.
-Quite often then, not all features are available for all words in our dataset; % are encoded in only a subset of the features we use (if any at all);
+Quite often then, not all features are available for all words in our data set; % are encoded in only a subset of the features we use (if any at all);
 however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
@@
 
-We finally need a substitution detection model, for the utterance data we use presents a challenge: quote-to-quote transformations, and much less substitutions, are not explicitly encoded in the dataset.
+We finally need a substitution detection model, for the utterance data we use presents a challenge: quote-to-quote transformations, and much less substitutions, are not explicitly encoded in the data set.
 More precisely, each set of quotations bears no explicit information about either the authoritative original quotation, or the source quotation(s) each author relied on when creating a new post and reproducing (and possibly altering) that source.
@@
 We therefore model the underlying quotation selection process by making a few additional assumptions. % which let us define quote-to-quote substitutions from the available data.
-The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, that is from a source external to the data collection perimeter.
+The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
 
@@
 It is however not feasible to try them all and, for the sake of simplicity, we decided to go with a sensible set of assumptions, and stick to them without trying alternative options.
-%These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
+%These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the data set.
 %\TB{We identified and implemented eleven other such models, and they all yielded essentially the same results.} \rk{T: I would prefer a bit more detail here.
@@
 \section{Results}\label{sec:results}
+
+\todo{\#16: add a quick description of what types of substitutions we see, then moving on to the observables since it's not what principally interests us.}
 
@@
 
-We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was randomly chosen from the whole pool of words available in the dataset for that feature.\footnote{For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the dataset of FA norms.}
-%\TB{randomly chosen from the whole pool of words in the dataset from which the feature is built (e.g.
+We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was randomly chosen from the whole pool of words available in the data set for that feature.\footnote{For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.}
+%\TB{randomly chosen from the whole pool of words in the data set from which the feature is built (e.g.
 for Clustering coefficient: randomly chosen in the pool of words in the Free Association norms)}.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins?
@@
 
-To make things concrete, here is an example substitution taking place in the dataset.
+To make things concrete, here is an example substitution taking place in the data set.
 At the end of January 2009, many media websites reported the following quote,
@@
 \end{quote}
-The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the dataset, and has a Clustering coefficient of $.24$.
-The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the dataset, and has a Clustering coefficient of $.05$.
+The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the data set, and has a Clustering coefficient of $.24$.
+The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the data set, and has a Clustering coefficient of $.05$.
 (Both words have four phonemes.)

### 2016-06-30 brains-protocol.tex
fix/methods: some better phrasings

@@
 \newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
-Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions going on every day, with all the complexity and uncertainty of conditions that this entails.
+Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
 In this first study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations, as this would be akin to a drink from the firehose.
-Our goal, instead, is to show that such effects exist and are measurable even if they are part of a larger complexity, that is, we aim to prove first a relatively weak hypothesis before trying stronger ones.
+Our goal, instead, is to show that such effects exist and are measurable, even if they are part of a larger complexity.
+That is, we aim to prove first a relatively weak hypothesis before trying stronger ones.
 If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
@@
 
-\newtext{With this goal of keeping} the analysis \newtext{tractable}, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 \newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see "Substitution model").}

### 2016-07-12 brains-protocol.tex
style/methods: minor tex presentation changes
feat/features: rewrite feature introduction and selection

@@
 This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 \newtext{(\citealp{leskovec_memetracker:_2009}; unfortunately, the original article, \citealp{leskovec_meme-tracking_2009}, does not provide further details on the source selection methodology).}
-\newtext{The authors automatically extracted quotations} from this corpus. Each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person; for instance:
+\newtext{The authors automatically extracted quotations} from this corpus.
+Each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person; for instance:
 \begin{quote}
-The Bank of England said, ``these operations are designed to address funding pressures over quarter-end.''
+The Bank of England said, "these operations are designed to address funding pressures over quarter-end."
 \end{quote}
 
-\newtext{Then, they gathered quotations in a graph and connected each pair that differed by no more than one word or that shared at least ten consecutive words (they tested this procedure with a number of different parameters, see~\citealp{leskovec_meme-tracking_2009}, for more details).}
+\newtext{Then, the authors gathered quotations in a graph and connected each pair that differed by no more than one word or that shared at least ten consecutive words (they tested this procedure with a number of different parameters, see~\citealp{leskovec_meme-tracking_2009}, for more details).}
 We find for example the following variation of the above quote:
@@
 \newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
-This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been assigned to another set)}.
+This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.
 
@@
 
-\todo{\#16:
-(1) update the set of features,
-(2) better explain feature selection: no PCA or all-feature regression (a predictive model doesn't differentiate correlated features), it's robust feature selection with domain knowledge
-(3) note clustering values have changed because of weights
-}
-
-We first introduce some of the most classical psycholinguistic measures on words.
+We first introduce some lexical measures on words.
 
 \begin{APAitemize}
-    \item \textbf{Word frequency}: the frequency at which words appear in our data set, known to be relevant for both recognition and recall~\citep{gregg1976word},
-    \item \textbf{Age of Acquisition}: the average age at which words are learned~\citep[obtained from][]{Kuperman12}, known to have different effects than word frequency~\citep{morrison1995roles,dewhurst1998separate},
-    \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary, \citealp{Weide98})\footnote{The CMU Pronouncing Dictionary is included in the NTLK package~\citep{Bird09}, the natural language processing toolkit we used for the analysis.} as a proxy to word production cost,
-    \item The average \textbf{Number of Synonyms} for all meanings of a word~\citep[obtained from][]{WordNet10} as an \emph{a priori} indicator of how easy it would be to replace a word.
+
+    \item \textbf{Word frequency}: the frequency at which words appear in our data set, known to be relevant for both recognition and recall~\citep{gregg_word_1976},
+
+    \item \textbf{Age of Acquisition}: the average age at which words are learned~\citep[obtained from][]{kuperman_age--acquisition_2012}, known to have different effects than word frequency~\citep{morrison_roles_1995,dewhurst_separate_1998},
+
+    \item \newtext{\textbf{Phonological} and \textbf{Orthographic Neighborhood Density}~\citep[obtained from][]{marian_clearpond:_2012}, also known to be relevant for word production~\citep{garlock_age--acquisition_2001},}
+
+    \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary, \citealp{weide_cmu_1998})\footnote{
+    The CMU Pronouncing Dictionary is included in the NTLK package~\citep{bird_nltk_2009}, the natural language processing toolkit we used for the analysis.
+    },
+    \newtext{as well as \textbf{Number of Letters}}, as a proxy to word production cost,
+
+    \item The average \textbf{Number of Synonyms} for all meanings of a word~\citep[obtained from][]{wordnet_princeton_2010} as an \emph{a priori} indicator of how easy it would be to replace a word.
+
 \end{APAitemize}
 
-%We also considered grammatical types within quotations by detection of \emph{Part-of-Speech} (POS) categories, using the Penn TreeBank Project typology~\citep{Santorini90} and thereby distinguishing verbs, nouns, adjectives and adverbs.
-The results were however extremely similar across the various categories, exhibiting no specific effect of words belonging to different POS categories.
-\rk{See \#8 for this fact-check}
-
-The number of synonyms is related to a notion of the word connectivity in a semantic network.
-To go a bit further in this direction, we appraise the possible role of network-based variables which have received special attention in the recent related literature, following the blooming interest in networks from many disciplines over the last decade.
+\newtext{We also consider grammatical types within quotations by detecting \emph{Part-of-Speech} (POS) categories with TreeTagger~\citep{schmid_probabilistic_1994}; we distinguish between verbs, nouns, adjectives, adverbs, and stopword-like words.}
 
@@
 
-\todo{\#20: FA don't encode reformulation explicitly, they did norms. We hypothesize it's what's at work.}
+\newtext{Aside from these raw features, the systemic dimension of vocabulary~\citep{cornish_systems_2013} has led authors to develop measures based on the full topology of networks built from free association data or phonological similarity.
+Several such measures have been shown to be involved in recall, recognition, and naming tasks~\citep{nelson_how_2013,chan_network_2010,griffiths_google_2007}.}
 
-%, we thus considered more recently studied variables based on semantic network properties.
-We relied on the ``free association'' (FA) norms collected by \citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations.
-FA norms record the words that come to mind when someone is presented with a given cue% (that is the ``free association'' task)
-. As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
-%Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
-% \rk{we're in fact using the unweighed version of the network. Why?}
-%Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
-Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
-This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyze is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.
+\newtext{To compute those features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
+As \citet{nelson_university_2004} explain, "free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect."
+\newtext{Similar to what \citet{griffiths_google_2007} did, we first considered the directed \newtext{weighted} network formed by association norms, where nodes are words and edges are directed from cue to target word, with a weight equal to the association strength (that is the probability of that target word being produced when this particular cue is presented).
+This network is of particular interest since it lets us define features that reflect the associations driving false memories in word lists~\citep{deese_prediction_1959}, a phenomenon which may be involved in the transformation of quotations.}
 
-Three standard network-based measures are to be used on the FA network% \rk{adapt once the weighing question (\#8) is settled}
-:
+We used three standard measures on the FA network:
 
 \begin{APAitemize}
-    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by \citet{Griffiths07}.
-    In the present case these two polysemy-related measures are quasi-perfectly correlated. %, and measure polysemy by taking into account the structure of the entire network.
+
+    \item \textbf{\newtext{Incoming} degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \textbf{Pagerank}~\citep{page_pagerank_1999}, which has already been used on the FA network by \citet{griffiths_google_2007}.
+    In the present case these two polysemy-related measures are quasi-perfectly correlated.\footnote{
+    \newtext{Note that in-degree does not take the weights of links into account, as it counts 1 for each incoming link.
+    Pagerank on the other hand, does take the weights into account.}
+    }
+    %, and measure polysemy by taking into account the structure of the entire network.
     %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
 Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
-    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{free:set}.
-    This quantity tells us if some words behave like unavoidable waypoints on association chains connecting one word to another.
-    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\citep{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual versus possible edges between a node's neighbors.}
-    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
+
+    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{freeman_set_1977}.
+    This quantity tells us if some words behave as unavoidable waypoints on association chains connecting one word to another.\footnote{
+    \newtext{For this measure, weights are interpreted as inverse cost:
+    the stronger a link, the easier it is to travel across it.
+    A stronger link will be favored over weaker links in the computation of the shortest path between two words.}
+    }
+
+    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\citep{watts_collective_1998}, computed on the undirected \newtext{weighted} version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual versus possible edges between a node's neighbors;
+    \newtext{this is poorly defined in the case of directed networks, which led us to ignore the direction of links in the network for this measure (if two words are connected in both directions, the weights of both links are added to make the final undirected link's weight).}
+    }
+    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a free association point of view).
+
 \end{APAitemize}
@@
 
-An important question arises concerning the possible correlations between all the variables we use.
+\newtext{Several of these features are strongly related and can be grouped together.
+To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
+The distributions of original and log-transformed features appears in the supplementary material, along with correlation scatter plots for each feature pair, summarized below with correlation values.
+}}
 
-%\begin{figure}[!th]
-%    \centering
-%    \includegraphics[width=\linewidth]{images/computed-figures/all-feature_correlations.png}
-%    \caption{Spearman correlations in the initial set of features}
-%    \label{fig:feature-corrs-initial}
-%\end{figure}
+\begin{figure}[!th]
+    \centering
+    \includegraphics[width=\linewidth]{images/computed-figures/all-feature_correlations.png}
+    \caption{Spearman correlations in the initial set of features}
+    \label{fig:feature-corrs-initial}
+\end{figure}
 
-The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($.8$).
-Our analysis showed clearer results with number of phonemes over number of syllables, which is consistent with \citet{nick-diss}, and we therefore chose to only present results for the former.
+\begin{new}
 
-Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
-Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Fig.~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.
+The pairwise correlations in the initial set of features appears in Fig.~\ref{fig:feature-corrs-initial}.
+By looking at absolute values, three subsets of highly correlated features can be easily identified:
+(a) number of letters, phonemes, and syllables with pairwise correlations greater than .75;
+(b) orthographic and phonological neighborhood densities, with a correlation of .8;
+(c) age of acquisition, betweenness, degree, and pagerank centralities, with absolute pairwise correlations at .41, .6, .59, .63, .61 and .85.
+Applying a feature agglomeration algorithm targeted at 6 groups refined this observation by producing identical (a) and (b) groups, a (c) group without betweenness centrality which was instead assigned to a group (d) with clustering coefficient, and the remaining features (frequency and number of synonyms) as singletons.\footnote{
+\newtext{Agglomerating into less than 6 groups merged groups (a) and (b), which we excluded to keep neighborhood densities in their own group;
+agglomerating into more than 6 groups separated age of acquisition from group (c), which we excluded given its high correlation values to the rest of group (c). 
+We used scikit-learn's FeatureAgglomeration class for this procedure~\citep{pedregosa_scikit-learn:_2011}.}
+}
 
-Frequency and number of synonyms both have relatively low levels of correlation to the other variables (excluding again the network properties); we therefore also keep them in the rest of the analysis.
-
-\bigskip
-Network centrality properties, on the other hand, are strongly dependent on one another.
-As mentioned earlier, degree centrality and pagerank have a very strong correlation ($.85$), and are also redundant with betweenness centrality (with correlation levels at $.75$ and $.68$ respectively).
-Furthermore, the three variables are also strongly related to age of acquisition, which leads us to keep the latter as the sole indicator for centrality.
-This may trigger a chicken-and-egg issue where a strong centrality may be due, or be the result, of an early age of acquisition; in any case, the age of acquisition seems to partially capture centrality-based network properties.
-
-Conversely, clustering coefficient exhibits low correlation levels with all the variables we kept (maximum absolute value $.38$), leading us to include it in the rest of the analysis.
-
-The final set of variables we consider, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.\footnote{Note that feature values stem from different data sets which do not always encode the same words.
-Indeed, we have data on frequency for about 22.6k words, on age of acquisition for 30.1k words, on number of phonemes for 123.4k words, number of synonyms 111.2k, and clustering coefficient 5.7k words.
-Quite often then, not all features are available for all words in our data set; % are encoded in only a subset of the features we use (if any at all);
+Since our data is about written transformations, number of letters and orthographic neighborhood density are the natural representatives of groups (a) and (b) respectively.
+Given the importance of age of acquisition in the lexical feature literature, we chose it to represent group (c).
+Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
+The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in Supplementary Material).\footnote{
+Note that feature values stem from different data sets which do not always encode the same words.
+Indeed, we have data on frequency for about \newtext{33.5k} words, on age of acquisition for 30.1k words, on clustering coefficient for 5.7k words, number of synonyms 111.2k, and \newtext{orthographic density 17.8k words}.
+Quite often then, not all features are available for all words in our data set;
 however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
 
-%\begin{figure}[!th]
-%    \centering
-%    \includegraphics[width=0.6412\linewidth]{images/computed-figures/paper-feature_correlations-v2.png}
-%    \caption{Spearman correlations in the filtered set of features}
-%    \label{fig:feature-corrs-filtered}
-%\end{figure}
+\end{new}
+
+\begin{figure}[!th]
+    \centering
+    \includegraphics[width=0.6412\linewidth]{images/computed-figures/paper-feature_correlations-v2.png}
+    \caption{Spearman correlations in the filtered set of features}
+    \label{fig:feature-corrs-filtered}
+\end{figure}
 
@@
 
-\todo{\#12: fix substitution counts}
+\todo{\#12: fix substitution counts and add precision/recall}
 
@@
 center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 1,051 valid substitutions.
+
+\todo{Note that our minor-spelling-changes exclusion does not affect orthographic nd: in a sample of 100 substitutions, no rejected substitutions are minor orthographic changes that would interest us (they're always different spelling of the same concept).}
 

### 2016-07-12 brains-protocol.tex
feat/substitution-model: add new graphics describing a few models

@@
 \subsection{Substitution model}
-%\subsubsection{Temporal binning}
 \label{sec:temporal-binning}\label{sec:model}
@@
 
-We finally need a substitution detection model, for the utterance data we use presents a challenge: quote-to-quote transformations, and much less substitutions, are not explicitly encoded in the data set.
+We finally need a substitution detection model, for the \newtext{quotation} data we use presents a challenge:
+quote-to-quote transformations and substitutions are not explicitly encoded in the data set.
 More precisely, each set of quotations bears no explicit information about either the authoritative original quotation, or the source quotation(s) each author relied on when creating a new post and reproducing (and possibly altering) that source.
-We thus face an inference problem where, given all quotations and their occurrence timestamps, we should estimate which was the originating quotation for each instance of each quotation.
+\newtext{In other words} we face an inference problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
 
-We therefore model the underlying quotation selection process by making a few additional assumptions. % which let us define quote-to-quote substitutions from the available data.
+We therefore model the underlying quotation selection process by making a few additional assumptions.
 The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
 
-Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in other blogs twice on the 20th and once on the 21st of January.
+Let us give an example:
+say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in a blog on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in other blogs twice on the 20th and once on the 21st of January.
 If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 20th based himself on $q$ as is shown in Fig.~\ref{fig:substitution-temporal-binning-a}.
@@
 Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place)
-or accurate reproductions of the previous occurrences of $q'$? (Options shown in Fig.~\ref{fig:substitution-temporal-binning-a}.)
+or accurate reproductions of the previous occurrences of $q'$?
+(Options shown in Fig.~\ref{fig:substitution-unmodelled}.)
 
-\begin{figure}[h]
+\begin{figure}[!h]
     \centering
@@
 	    \small
-	    \input{images/substitution-temporal-binning-a.pdf_tex}
-	    \label{fig:substitution-temporal-binning-a}
+	    \input{images/substitution-unmodelled.pdf_tex}
+	    \label{fig:substitution-unmodelled}
 	}
@@
 	    \small
-	    \input{images/substitution-q_max.pdf_tex}
-	    \label{fig:substitution-q_max}
+	    \input{images/substitution-Time=discrete-Source=majority-Past=last_bin-Durl=all.pdf_tex}
+	    \label{fig:substitution-Time=discrete-Source=majority-Past=last_bin-Durl=all}
+	}
+	\hfill \\
+	\subfloat[Binned quotation family with majority rule and excluded past]{
+	    \def\svgwidth{\linewidth}
+	    \small
+	    \input{images/substitution-Time=discrete-Source=majority-Past=last_bin-Durl=exclude_past.pdf_tex}
+	    \label{fig:substitution-Time=discrete-Source=majority-Past=last_bin-Durl=exclude_past}
+	}
+	\hfill \\
+	\subfloat[Binned quotation family with \emph{no} majority rule and excluded past extended to the beginning of the quotation family]{
+	    \def\svgwidth{\linewidth}
+	    \small
+	    \input{images/substitution-Time=discrete-Source=all-Past=all-Durl=exclude_past.pdf_tex}
+	    \label{fig:substitution-Time=discrete-Source=all-Past=all-Durl=exclude_past}
 	}

### 2016-07-13 brains-protocol.tex
feat/substitution-models: finish first rewrite of substitution models

@@
 
-\todo{\#19: show more clearly why this is more robust with single-substitution, justifying our choice of this case}
-
-\todo{\#17: explain all models and the work involved, with graphs and explanatory figures for each model in annex}
-
 We finally need a substitution detection model, for the \newtext{quotation} data we use presents a challenge:
@@
 We therefore model the underlying quotation selection process by making a few additional assumptions.
-The main issue is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
+\newtext{Given a particular occurrence of a quotation, the first issue is deciding whether} that occurrence is a strict copy of an earlier occurrence, or a substitution of another \newtext{quotation}, or maybe a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
+\newtext{The second issue is deciding which source originated such a substitution when several candidate sources are available.}
 
 Let us give an example:
-say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in a blog on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in other blogs twice on the 20th and once on the 21st of January.
-If $q$ was sufficiently prominent when $q'$ first appeared, we can safely assume that the first author of $q'$ on the 20th based himself on $q$ as is shown in Fig.~\ref{fig:substitution-temporal-binning-a}.
-But what about the second and third occurrences of $q'$, on the 20th and 21st?
-Should we consider them to be substitutions based on $q$ %(i.e. re-creations of $q'$ by a new instance of the substitution process that brought from $q$ to $q'$ in the first place)
-or accurate reproductions of the previous occurrences of $q'$?
+\newtext{say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in two different blogs on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in another blog on the 20th of January.
+The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
+And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$ as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
+Now say a third version, "These \textbf{allegations} are false and \textbf{incoherent}" ($q''$) also appears once on January 19 and once on January 20 after $q'$.
+$q$ and $q''$ differ by two substitutions, so we ignore the possibility that one was written based on the other.\footnote{
+NOTE WE TESTED THIS.
+}
+$q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
+Conversely, the occurrence of $q''$ on the 20th could be considered as a substitution from $q'$, or as a faithful copy from its initial occurrence on January 19.}
 (Options shown in Fig.~\ref{fig:substitution-unmodelled}.)
 
-\begin{figure}[!h]
+\begin{figure}[h]
     \centering
-    \subfloat[Possible paths from occurrence to occurrence]{
 	\def\svgwidth{\linewidth}
@@
 	\input{images/substitution-unmodelled.pdf_tex}
+	\caption{{\bf Possible paths from occurrence to occurrence.}
+	\newtext{$q$, $q'$ and $q''$ are three quotation variants belonging to the same cluster.
+	$q$ and $q''$ differ by two words, but $q'$ differs from both $q$ and $q''$ by a word.
+	The second occurrence of $q$ can safely be considered a faithful copy of the first, but the occurrences of $q'$ and $q''$ are uncertain:
+	while the first occurrence of $q'$ is most likely a substitution from $q'$, it could also stem from $q''$;
+	conversely, the second occurrence of $q''$ could also be a substitution from $q'$ instead of being a faithful copy of its first occurrence.}
+	}
 	\label{fig:substitution-unmodelled}
-	}
-	\hfill \\
-	\subfloat[Binned quotation family with majority rule]{
-	    \def\svgwidth{\linewidth}
+\end{figure}
+
+\newtext{One way to settle these questions is the following:}
+group quote occurrences into fixed bins spanning $\Delta t$ days (1 day in the implementation), each one representing a unit of time evolution;
+when a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote of the preceding bin $t$ (or a substring thereof) by only one word;
+if not, $q'$ is not considered to be an instance of substitution.
+\newtext{Fig.~\ref{fig:substitution-modelled-majority-all} shows the inferences made by such a model.
+The assumptions it embeds, however, are a subset of a much wider set of possibilities, each leading to alternative inferences.}
+
+\begin{new}
+
+We identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
+The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
+(1) \textbf{bin positions}, which can be discretized (aligning the end of a bin to midnight, as in the model presented above) or kept continuous (for each occurrence, use a bin that ends precisely at that occurrence);
+(2) \textbf{bin span}, which can be 24 hours (as in the model above) or can be extended to start at the very first occurrence in the quotation family.
+The other two parameters define rules on the selection of source and destination quotes of a substitution:
+(3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
+(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or not (as in the model above).
+A substitution model, then, is the given of a value for each of those parameters;
+it considers valid all the substitutions (and only those) where the source and destination follow the rules set out by the parameters.
+If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.
+
+\end{new}
+\begin{figure*}[h]
+	\centering
+	\begin{subfigure}{.49\textwidth}
+	    \def\svgwidth{\textwidth}
 	    \small
 	    \input{images/substitution-Time=discrete-Source=majority-Past=last_bin-Durl=all.pdf_tex}
-	    \label{fig:substitution-Time=discrete-Source=majority-Past=last_bin-Durl=all}
-	}
-	\hfill \\
-	\subfloat[Binned quotation family with majority rule and excluded past]{
-	    \def\svgwidth{\linewidth}
+	    \caption{Source must be majority in preceding bin, destination can be anything}
+	    \label{fig:substitution-modelled-majority-all}
+	\end{subfigure}
+	~
+	\begin{subfigure}{.49\textwidth}
+	    \def\svgwidth{\textwidth}
 	    \small
 	    \input{images/substitution-Time=discrete-Source=majority-Past=last_bin-Durl=exclude_past.pdf_tex}
-	    \label{fig:substitution-Time=discrete-Source=majority-Past=last_bin-Durl=exclude_past}
+	    \caption{Source must be majority in preceding bin, destination must not appear in preceding bin}
+	    \label{fig:substitution-modelled-majority-exclude_past}
+	\end{subfigure}
+
+	\begin{subfigure}{.49\textwidth}
+	    \def\svgwidth{\textwidth}
+	    \small
+	    \input{images/substitution-Time=discrete-Source=all-Past=last_bin-Durl=all.pdf_tex}
+	    \caption{Source can be anything, destination can be anything}
+	    \label{fig:substitution-modelled-all-all}
+	\end{subfigure}
+	~
+	\begin{subfigure}{.49\textwidth}
+	    \def\svgwidth{\textwidth}
+	    \small
+	    \input{images/substitution-Time=discrete-Source=all-Past=last_bin-Durl=exclude_past.pdf_tex}
+	    \caption{Source can be anything, destination must not appear in preceding bin}
+	    \label{fig:substitution-modelled-all-exclude_past}
+	\end{subfigure}
+	\caption{\newtext{
+	{\bf Substitution models.}
+	Substitutions inferred by four models in the situation introduced by Fig.~\ref{fig:substitution-unmodelled}.
+	Each of these models uses discretely positioned bins spanning 1 day (see the main text for a complete description of parameters).
+	In the top left panel~(a), $q$ holds the majority in the first bin and is considered the unique basis for $q'$ in bin~2.
+	$q'$ and $q''$ have equal maximum frequency in bin~2 however, so both are sources of substitutions towards bin~3.
+	In the top right panel~(b), quotes that appear in the preceding bin cannot be the target of a substitution; this removes two substitutions compared to panel~(a).
+	In the bottom left panel~(c), the majority constraint is lifted compared to panel~(a), making $q''$ in bin~1 a candidate source for $q'$ in bin~2.
+	In the bottom right panel~(d), the majority constraint is also lifted compared to panel~(a) (adding the same $q'' \rightarrow q'$ substitution as in panel~(c)), and the excluded-past constraint is added as in panel~(b) (removing two same substitutions from bin~2 to bin~3 as in panel~(b)).
+	If the bins were extended to the beginning of the quotation family, the exclude past constraint would also remove the $q' \rightarrow q$ substitution from bin~2 to bin~3.
+	In all four panels, a background rectangle or square indicates the quotation is the source of a substitution.
+	A thick border on that rectangle or square indicates the quotation was selected because it has maximum frequency.}
 	}
-	\hfill \\
-	\subfloat[Binned quotation family with \emph{no} majority rule and excluded past extended to the beginning of the quotation family]{
-	    \def\svgwidth{\linewidth}
-	    \small
-	    \input{images/substitution-Time=discrete-Source=all-Past=all-Durl=exclude_past.pdf_tex}
-	    \label{fig:substitution-Time=discrete-Source=all-Past=all-Durl=exclude_past}
+    \label{fig:substitution-modelled}
+\end{figure*}
+
+\begin{new}
+
+Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
+Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use discretely positioned bins spanning 1 day:
+later occurrences of $q'$ and $q''$ are counted as substitutions in Fig.~\ref{fig:substitution-modelled-majority-all} and Fig.~\ref{fig:substitution-modelled-all-all}, whereas in Fig.~\ref{fig:substitution-modelled-majority-exclude_past} and Fig.~\ref{fig:substitution-modelled-all-exclude_past} they are not.
+
+The results reported and discussed in the following sections are valid for all 16 models, and the graphics we present were produced by the model first introduced above.
+Finally, note that this inference procedure is one of the reasons we restricted our analysis to single-substitutions:
+looking for more complex transformations would
+(a) exponentially increase the number of candidate sources for a destination occurrence, which correspondingly reduces the confidence in inferences made,
+and (b) greatly increase the complexity of the transformation models used to make these inferences.\footnote{
+\newtext{We checked that this restriction does not bias the results discussed below by extending our protocol to two-substitution transformations.
+The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.
+\todo{check they're the same}.}
 }
-	\caption{{\bf Temporal binning of quotation families.} $q$ and $q'$ are two versions of a quotation belonging to the same cluster.
-In the bottom panel (b), $q'$ holds the majority in the 3rd bin and is considered the unique basis for the last occurrence of $q$ (in the 4th bin).
-This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Fig.~\ref{fig:substitution-q_max}, this seems to be the most likely scenario).
-Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence of $q$ in bin~3.}
-    \label{fig:substitution-temporal-binning}
-\end{figure}
 
-To settle this question we group quote occurrences into fixed bins spanning $\Delta t$ days (1 day in the implementation), each one representing a unit of time evolution.
-When a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote $q$ of the preceding bin $t$ (or a substring thereof) by only one word.
-If not, $q'$ is not considered to be an instance of substitution.
-%We assume here that quotes are based on the most frequent quote of the preceding bin (\hbox{vs.} all the other quotes in that bin).
-Note that these assumptions are admittedly a subset of a much wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, the criterion of the most frequent quote in the preceding bin may be replaced with the most frequent quote overall, or the oldest quote; time can be sliced into fixed bins as is done here, or kept fine-grained by using sliding bins.}
-It is however not feasible to try them all and, for the sake of simplicity, we decided to go with a sensible set of assumptions, and stick to them without trying alternative options.
-%These various flavors of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the data set.
-%\TB{We identified and implemented eleven other such models, and they all yielded essentially the same results.} \rk{T: I would prefer a bit more detail here.
-Maybe on supp.
-mats.?
-What does “essentially the same” mean?
-This could cause concern to reviewers.}
+\end{new}
 
@@
 
-\todo{\#12: fix substitution counts and add precision/recall}
-
-Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Fig.~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.
-In practice, from the 2.43m initial occurrences spread into 45,749 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 6,172 real substitutions obeying to this model.
+In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread into 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
 From these substitutions we remove those featuring stop words, minor spelling changes (e.g.
-center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 1,051 valid substitutions.
-
-\todo{Note that our minor-spelling-changes exclusion does not affect orthographic nd: in a sample of 100 substitutions, no rejected substitutions are minor orthographic changes that would interest us (they're always different spelling of the same concept).}
-
-%Let us illustrate this ``majority'' rule by going back to the example described in Section~\ref{sec:temporal-binning} and extending it, in Fig.~\ref{fig:substitution-q_max}.
+center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
+this eventually yields \newtext{6,318} valid substitutions (before merging substitutions that share the same destination).\footnote{
+\newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
+Precision was evaluated over a random subset of 100 \emph{kept} substitutions, showing a score of .87.
+Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
+out of the first 100 substitutions coded for recall, those with levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}
+}
 

### 2016-07-13 brains-protocol.tex
style/protocol: missing \newtext's and left-over self-notes

@@
 Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.\footnote{
-A similar analysis was made for language detection, which is part of the cluster filtering:
+\newtext{A similar analysis was made for language detection, which is part of the cluster filtering:
 out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
 Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
-}}
+}}}
 
@@
 To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
-The distributions of original and log-transformed features appears in the supplementary material, along with correlation scatter plots for each feature pair, summarized below with correlation values.
-}}
+\newtext{The distributions of original and log-transformed features appears in the supplementary material, along with scatter plots for each feature pair, summarized below with correlation values.
+}}}
 
@@
 Now say a third version, "These \textbf{allegations} are false and \textbf{incoherent}" ($q''$) also appears once on January 19 and once on January 20 after $q'$.
-$q$ and $q''$ differ by two substitutions, so we ignore the possibility that one was written based on the other.\footnote{
-NOTE WE TESTED THIS.
-}
+$q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (this below for further details).
 $q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.

### 2016-07-16 brains-introduction.tex
feat/features: rewrite feature introduction and selection
comments + various direct minor corrections

@@
 
+\todo{add examples}
+
 To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
@@
 }
-While this restricts the scope of our observations for the particular data set we use, the methodological point we also make is left intact.
-By characterizing words with 6 well-known features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
+While this limits the scope of our observations to the particular data set we use, the methodological point we also make is left intact.
+By characterizing words using 6 well-known features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
 This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
@@
 But the cognitive features which may determine or, at least, influence these transformations, are overlooked.
-\newtext{Yet cognitive and linguistic features have been used in studies not involving transformation: \citet{danescu-niculescu-mizil_you_2012}, for instance, show that particularly memorable quotations (taken from movie scripts in this case) use more distinctive words and have more common syntax than less memorable quotations; they are also the quotes that adapt best to new contexts of use.
+\newtext{On the other hand cognitive and linguistic features have been used in diffusion studies not involving transformation: \citet{danescu-niculescu-mizil_you_2012}, for instance, show that particularly memorable quotations (taken from movie scripts in this case) use more distinctive words and have more common syntax than less memorable quotations; they are also the quotes that adapt best to new contexts of use.
 One source of ideas to study the transformations of such quotes, then, might be the psycholinguistic literature studying word and sentence recall.}
@@
 
-These effects don't transpose simply to sentence recall however, as not only syntax but also effects of attention come into play for both retrieval and encoding.
+These effects do not transpose simply to sentence recall however, as not only syntax but also effects of attention come into play for both retrieval and encoding.
 \citet{jefferies_automatic_2004}, for instance, show that attention is central to the encoding and retention of unrelated propositions, on top of more automatic syntactic and semantic processes.
@@
 
-On the whole, research on lexical features hints towards two antagonistic types of effects.
+On the whole, research on lexical features hints towards two antagonistic types of effects \newtext{(also known as the "word-frequency paradox", \citealp{mandler_word-frequency_1982})}.
 On one hand, part of the literature shows that recall is easier for the least "awkward" words;

### 2016-07-16 brains-protocol.tex
comments + various direct minor corrections

@@
 \todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
+
+\todo{\sf [Cam: Perhaps useful to say here and first off that it's a text corpus made from online data -- otherwise it's a bit abrupt and one may not immediately remember (from the introduction) which quotations we are talking about.
+Something like ``We rely on a text corpus of public representations made of blog posts and focus on the evolution of quotations.'']}
 
@@
 Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
-In this first study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations, as this would be akin to a drink from the firehose.
+In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations\todo{, as this would be akin to a drink from the firehose.
+\sf [Cam: supprimer?
+potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc.
+outside the scope etc.
+further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion?
+ou bien il s'agit d'une critique directe d'un reviewer ?
+Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
 Our goal, instead, is to show that such effects exist and are measurable, even if they are part of a larger complexity.
-That is, we aim to prove first a relatively weak hypothesis before trying stronger ones.
+\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
 If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
@@
 We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
-This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 \newtext{(\citealp{leskovec_memetracker:_2009}; unfortunately, the original article, \citealp{leskovec_meme-tracking_2009}, does not provide further details on the source selection methodology).}
+This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 \newtext{(\citealp{leskovec_memetracker:_2009}; \todo{unfortunately, the original article, \citealp{leskovec_meme-tracking_2009}, does not provide further details on the source selection methodology) \sf [Cam: mettrait-on cette précision au mieux en footnote?
+Le ``unfortunately'' laisse planer une sale ombre sur la qualité du dataset, alors qu'il ne s'agit là que d'une question de périmètre (bcp moins pertinente pour les problèmes de qualité du dataset que d'autres phénomènes qui l'affectent assurément)]}.}
 \newtext{The authors automatically extracted quotations} from this corpus.
@@
 Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
-The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in Supplementary Material).\footnote{
+The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in Supplementary Material) \todo{[Cam: Appendix dans ce journal je pense (oder?)]}.\footnote{
 Note that feature values stem from different data sets which do not always encode the same words.
@@
 	In the bottom right panel~(d), the majority constraint is also lifted compared to panel~(a) (adding the same $q'' \rightarrow q'$ substitution as in panel~(c)), and the excluded-past constraint is added as in panel~(b) (removing two same substitutions from bin~2 to bin~3 as in panel~(b)).
-	If the bins were extended to the beginning of the quotation family, the exclude past constraint would also remove the $q' \rightarrow q$ substitution from bin~2 to bin~3.
+	If the bins were extended to the beginning of the quotation family, the excluded-past constraint would also remove the $q' \rightarrow q$ substitution from bin~2 to bin~3.
 	In all four panels, a background rectangle or square indicates the quotation is the source of a substitution.

### 2016-07-18 brains-protocol.tex
fix/protocol: fixes after Camilles comments

--- 2016-07-13 brains-protocol.tex
@@
 
-Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
+\newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
+\newtext{Indeed} quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
 First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
@@
 Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
-In this first study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations, as this would be akin to a drink from the firehose.
-Our goal, instead, is to show that such effects exist and are measurable, even if they are part of a larger complexity.
-That is, we aim to prove first a relatively weak hypothesis before trying stronger ones.
+%In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
+%\todo{, as this would be akin to a drink from the firehose.
+\sf [Cam: supprimer?
+potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc.
+outside the scope etc.
+further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion?
+ou bien il s'agit d'une critique directe d'un reviewer ?
+Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
+Our goal, therefore, is to show that such effects exist and are measurable even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
+%\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
 If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
@@
 We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
-This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 \newtext{(\citealp{leskovec_memetracker:_2009}; unfortunately, the original article, \citealp{leskovec_meme-tracking_2009}, does not provide further details on the source selection methodology).}
+This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 (\citealp{leskovec_memetracker:_2009}).\footnote{
+The original article~\citep{leskovec_meme-tracking_2009} does not provide further details on the source selection methodology.
+}
 \newtext{The authors automatically extracted quotations} from this corpus.
@@
 Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
-The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in Supplementary Material).\footnote{
+The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in the appendix).\footnote{
 Note that feature values stem from different data sets which do not always encode the same words.

### 2016-07-25 brains-introduction.tex
style/color: some more text should be bluegreen

@@
 there should be areas of the representation space where cognitive effects in transformations bring representation closer to a given stable asymptotic point.\footnote{
-Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
-}
+\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
+}}
 
@@
 We chose to restrict our analysis to substitutions (one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
-As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
+\newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
 This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
-}
+}}
 While this limits the scope of our observations to the particular data set we use, the methodological point we also make is left intact.
@@
 The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
-Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
-}
+\newtext{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
+}}
 A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
@@
 Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
-Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited to data consisting of a high number of measures over a smaller number of lists (in which case it makes sense to ask \hbox{e.g.} what proportion of intrusions come from prior lists).
+\newtext{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited to data consisting of a high number of measures over a smaller number of lists (in which case it makes sense to ask \hbox{e.g.} what proportion of intrusions come from prior lists).
 As is explained further down however, our data set is shaped the opposite way: a great number of sentences, with only very few to no measures at all on each sentence.
-}
+}}
 If a cognitive bias in the substitution of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.

### 2016-07-25 brains-protocol.tex
chore/todos+colors: add some todos, fix some colors

@@
 As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}
+\todo{this last sentence is unclear or not useful}
 
@@
 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
-Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.\footnote{
+Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.}\footnote{
 \newtext{A similar analysis was made for language detection, which is part of the cluster filtering:
@@
 Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
-}}}
+}}
 
@@
 \newtext{We checked that this restriction does not bias the results discussed below by extending our protocol to two-substitution transformations.
-The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.
-\todo{check they're the same}.}
+The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.}
 }
@@
 center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
-this eventually yields \newtext{6,318} valid substitutions (before merging substitutions that share the same destination).\footnote{
+this eventually yields \newtext{6,318} valid substitutions \newtext{(before merging substitutions that share the same destination)}.\footnote{
 \newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
@@
 \todo{\#16: add a quick description of what types of substitutions we see, then moving on to the observables since it's not what principally interests us.}
+
+\todo{order of presentation:
+- distance in substitution
+- susceptibility of POS
+- susceptibility of bins of absolute value
+- variation of bins of absolute value
+- susceptibility of quartiles in sentence
+- variation of quartiles in sentence (bins->bins or quartile->bins or quartile->quartile)
+- PCA of susceptibilities absolute, quartile, absolute+quartile
+}
 

### 2016-07-25 brains-protocol.tex
chore/todos: rework todos for results

@@
 
-\todo{\#16: add a quick description of what types of substitutions we see, then moving on to the observables since it's not what principally interests us.}
+\todo{Quick description of what types of substitutions we see (\#16)}
 
-\todo{order of presentation:
-- distance in substitution
-- susceptibility of POS
-- susceptibility of bins of absolute value
-- variation of bins of absolute value
-- susceptibility of quartiles in sentence
-- variation of quartiles in sentence (bins->bins or quartile->bins or quartile->quartile)
-- PCA of susceptibilities absolute, quartile, absolute+quartile
-}
+\todo{Mention distance in substitution}
 
@@
 
-\todo{\#19: explain if we're susceptible to bias from single-substitution}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
+    \caption{}
+    \label{fig:pos-susceptibilities}
+\end{figure}
 
@@
 
-Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$, that is% (only taking into account words that are substituted at least once)
-:
+Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$, that is:
 
@@
 
-%\begin{figure*}[!th]
-%    \centering
-%    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
-%    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-%    The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.}
-%    \label{fig:feature-susceptibilities}
-%\end{figure*}
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
+    \caption{\textbf{Substitution susceptibility:}% average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
+    %The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
+    }
+    \label{fig:feature-susceptibilities-global}
+\end{figure*}
+
 \medskip
 
-\todo{\#16:
-(1) show POS don't differ,
-(2) note that stopwords are excluded, which excludes POS open/closed discussion
-}
+\todo{Exemplify with POS (note stopwords excluded) (\#16)}
 
-\todo{\#16: explain we do this on quartiles, rework results discussion}
-
-\todo{\#16: explicit all significance claims, referring to the CIs}
-
-\todo{\#17: show other feature variations and other models in annex}
-
-\todo{\#15: relate to missed literature}
+\todo{Discuss results for global value, then (\#15) relate to missed literature}
 
@@
 To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-%\TG{TBF: [[We see that outlier values (the two darkest cells) are both due to cells where the number of items is only slightly above the significance threshold (11 items, while we decided to cut at 10).]]}
-%\TG{TBF: [[If we focus on cells based on at least several dozens of words]]}
 {Even if there are a few outlier cells, values tend to navigate around the mean value ($.16$) with little obvious regularity (except for a low number of synonyms, consistent with the unidimensional graph)}.
 {On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
-%\TG{TBF: [[This confirms the fact that the slight late increases observed for Age of acquisition and Number of phonemes are rather non-significant.]]}
-
-%\TB{This class of words corresponds to \todo{} [\TB{highly specialized words? domain-specific? a hard limit on too complicated words?}]}
-%Age of acquisition and Number of phonemes delineate two categories of words: both features have little to no significant effect on susceptibility for low- to mid-valued words, but very high-valued words behave differently and tend to be more substituted (clearly for high Number of phonemes, and only as a tendency for high Age of acquisition).
-%These high-valued words also have low occurrence frequencies, which corresponds to the slope change in the frequency curve between extreme low frequency words and mid- to high-frequency words.
-\rk{check this with a plot of freq vs.
-clustering and freq vs.
-AoA, or simply plot of frequency for the high-valued words for AoA and CC.}
 
@@
 
-% plus ou moins choisies uniformément (modulo les réserves exhibées en 4.1), on peut montrer en quoi elles sont modifiées Finally the results for Clustering coefficient and Number of synonyms, despite some tendencies, do not allow us to conclude either to an effect or no effect of those features on substitution susceptibility.
+\subsection{Variation}
 
-\subsection{Variation}
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_global.png}
+    \caption{\textbf{Feature variation upon substitution:} %$\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
+    %The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
+    %The intersection with $y = x$ marks the attractor value.
+    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.
+    }
+    \label{fig:feature-variations-global}
+\end{figure*}
 
@@
 
+\todo{Introduce both H0 and H00}
+
 We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was randomly chosen from the whole pool of words available in the data set for that feature.\footnote{For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.}
-%\TB{randomly chosen from the whole pool of words in the data set from which the feature is built (e.g.
-for Clustering coefficient: randomly chosen in the pool of words in the Free Association norms)}.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins?
-la formulation est un peu ambigüe.
-S: oui et non.
-Oui: c'est pas juste les voisins.
-Non: le pool dépend de la feature, c'est pas juste FA.
-Il y avait une erreur, j'ai reformulé pour être plus clair.}
 In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, that is an arrival word chosen among semantically plausible though still random words.
@@
 $$\nu_{\phi}^0 (f) = \left<\phi\right>$$
-%We thereby obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.
 
@@
 \end{seriate}
-%\begin{figure*}[!th]
-%    \centering
-%    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
-%    \caption{\textbf{Feature variation upon substitution:} $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
-%    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
-%    The intersection with $y = x$ marks the attractor value.
-%    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.
-%    }
-%    \label{fig:feature-variations}
-%\end{figure*}
 
@@
 
-\todo{\#16:
-(1) add H00 lines to account for semantic similarity (nothing more fancy since it involves a threshold),
-(2) rework results discussion
-}
-
-\todo{\#17: show other feature variations and other models in annex}
-
-\todo{\#17: explain binning}
-
-\todo{\#18: tone down claims about contractile: it's a possible hypothesis if this were the only process, but not observed with the mix of all other processes}
-
-\todo{\#15: relate to missed literature}
+\todo{Update results, explaining binning and toning down contractility, and relate to missed literature}
 
@@
 This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
-%\cam{je crois qu'on peut éviter l'inférence inverse en ne disant pas "confirms" (c'est exagéré en effet) mais en disant juste que c'est cohérent avec un processus mais pas l'autre; j'ai changé le wording}
-% , Mais est-ce que c'est pas une inférence inverse que de dire ceci?
-i.e.
-on sait qu'une tâche "familiarity/recognition" implique un biais en faveur des mots awkards, et qu'une tâche "recall" implique un biais en faveur des mots fréquents, mais on peut pas vraiment tirer de conclusions en utilisant la réciproque, si?}.
-%However, the detailed variations of the Word frequency curve, despite an inflection for high-frequency words, do not seem to follow any significant pattern.\cam{Je serais pour supprimer cette phrase purement factuelle et pas très clear-cut non plus}
 
@@
 
-\todo{\#16: add the taking-context-into-account parts, potentially moving the non-relative results discussions to here:
-(1) explain and add sentence-relative susceptibility by quartiles, which has a perfectly clear interpretation + results discussion,
-(2) explain and add sentence-relative variations + results discussion
-(3) link those results to non sentence-relative values
-}
+\subsection{Sentence context}
 
-\todo{\#17: show other feature variations and other models in annex}
+\todo{Explain in-sentence susceptibility and sentence-relative variation, relating to global value results}
 
-\todo{\#15: relate to missed literature}
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
+    \caption{}
+    \label{fig:feature-susceptibilities-in_sentence}
+\end{figure*}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
+    \caption{}
+    \label{fig:feature-variations-sentencerel}
+\end{figure*}
 

### 2016-07-26 brains-protocol.tex
feat/results: rephrase introduction paragraphs
And some other minor presentational fixes.

@@
 To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
-\newtext{The distributions of original and log-transformed features appears in the supplementary material, along with scatter plots for each feature pair, summarized below with correlation values.
+\newtext{The distributions of original and log-transformed features appears in the Supplementary Information, along with scatter plots for each feature pair, summarized below with correlation values.
 }}}
@@
 The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
-(1) \textbf{bin positions}, which can be discretized (aligning the end of a bin to midnight, as in the model presented above) or kept continuous (for each occurrence, use a bin that ends precisely at that occurrence);
+(1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
 (2) \textbf{bin span}, which can be 24 hours (as in the model above) or can be extended to start at the very first occurrence in the quotation family.
@@
 (3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
-(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or not (as in the model above).
+(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
 A substitution model, then, is the given of a value for each of those parameters;
@@
 	    \input{images/substitution-Time=discrete-Source=all-Past=last_bin-Durl=all.pdf_tex}
-	    \caption{Source can be anything, destination can be anything}
+	    \caption{Source can be anything, destination can be anything\\\hfill} % \\\hfill to add a line, aligning with the other panels which have two-line captions
 	    \label{fig:substitution-modelled-all-all}
@@
 
-\todo{Quick description of what types of substitutions we see (\#16)}
+\begin{new}
 
-\todo{Mention distance in substitution}
+While most of the substitutions we obtain with this procedure replace a word with another semantically related word, the two are rarely direct synonyms:
+only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
+\newtext{A similar behavior is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
+}}
+Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
+The following examples illustrate this behavior:
 
-We may now use this substitution model to formulate a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.
-To this end, we build two main observables for each word feature.
-First, we measure the susceptibility for words to be the target of a substitution in a quote, knowing that there has been a variation, in order to show which semantic features are the most likely to ``attract'' a substitution under this condition.
+\begin{APAitemize}
+	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
+	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
+	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
+\end{APAitemize}
+
+\end{new}
+
+\newtext{Our question concerns the low-level properties of these substitutions; to address it} we build the following two observables for each word feature.
+First, we measure \newtext{which word features are more or less substituted compared to how often they would be if the process were random, in order to capture} the susceptibility for words to be the target of a substitution in a quote.
 Second, we measure the change in word feature upon substitution, looking at the variation of a given feature between start and arrival words.
+\newtext{Since sentence context is also central to this process, we extend these two observables by applying them to feature values relative to the distribution of feature values of the sentence in which a word appears.}
 
-Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
-Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source~(``{\sc Ctrl-C}/{\sc Ctrl-V}'').
-Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~for instance drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
+Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, \newtext{that is} we do not take invariant quotations into account.
+Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it \newtext{we cannot} know whether there has been perfect human reformulation or simply digital copy-pasting of a source~(``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+\newtext{Moreover}, perfect human reformulation possibly involves different practices than those involved in alteration ---~for instance drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
 The two situations are different enough to be studied separately, and we focus here on the latter.

### 2016-07-26 brains-protocol.tex
feat/results: rewrite susceptibility results

@@
 
+We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
+\newtext{For a given group of words $g$, susceptibility is computed as the ratio of the number of times $s_g$ words of that group are substituted to the number of times $s_g^0$ words of that group would be substituted if substitutions fell randomly on substitutable words, that is:}\footnote{
+\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
+If, for instance, half the content words of a given quote are from group $g$, such a quote contributes .5 to the total $s_g^0$.}
+}
+$$\sigma_g = \frac{s_g}{s_g^0}$$
+
+In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
+\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (e.g.
+for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
+
 \begin{figure}[h]
@@
     \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
-    \caption{}
+    \caption{\newtext{\textbf{Substitution susceptibility for POS categories:} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
+    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts, the bottom panel shows $\sigma_{POS}$, the ratio of the two.
+    Confidence intervals are computed with the Goodman method for multinomial proportions.}
+    \todo{Split into 2 subfigures; fix labels; add H0 dashed line.}}
     \label{fig:pos-susceptibilities}
@@
 
-We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether that substitution operates on that word or on another one.
-Word substitution susceptibility is computed as the ratio of the number of times $s_w$ a word is substituted to the number of times $p_w$ that word appears in a substitutable position, that is $\sfrac{s_w}{p_w}$.
-{In other words, it measures how often a word $w$ actually gets substituted, compared to how often it could have been substituted (because it appears in quotes undergoing substitution)}.
+\medskip
 
-Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$, that is:
+\begin{new}
 
-$$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$
+Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
+A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Stopword-like}\footnote{
+\newtext{The \emph{Stopword-like} category gathers all the POS groups partially of completely affected by the stopword filter for substitutions: coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections.
+Note that the stopword filter equally affects $s$ and $s^0$, so while both numbers are very small for these categories (which led us to group them together) the reported susceptibility is not biased by this filtering.}
+}
+and \emph{Adverb} categories.
+Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for stopword-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
+none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
+While we acknowledge the low susceptibilities of adverbs and stopword-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
+it seems, then, that POS categories don't capture any strong bias in the selection of substitution targets.
 
-Put shortly, susceptibility focuses on the selection of start words involved in substitutions, measuring the effect of features at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
+\end{new}
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
-    \caption{\textbf{Substitution susceptibility:}% average susceptibility to substitution versus average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
-    %The heatmap on the lower-right shows the joint effect of Number of synonyms and Number of phonemes on susceptibility, averaged over the respective single-variable ranges, with sample size (word numbers) in parentheses.
+    \caption{\textbf{Substitution susceptibility for feature values:} susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
     }
@@
 
-\medskip
+\todo{Relate to missed literature (\#15)}
 
-\todo{Exemplify with POS (note stopwords excluded) (\#16)}
+\begin{new}
 
-\todo{Discuss results for global value, then (\#15) relate to missed literature}
+The results for word features presented in Fig.~\ref{fig:feature-susceptibilities-global}, on the other hand, show marked effects for several features.
+Word frequency, Age of acquisition, and Number of letters each exhibit significant susceptibility variations (Goodman goodness-of-fit with $p < .05$ in all substitution models, $p < .001$ in most) consistent with known effects of those features on recall.
+High-frequency words, much easier to recall, are substituted about half as much as they would be at random; conversely low-frequency words, harder to recall, are substituted about 50\% more than random.
+Age of acquisition and Number of letters show the opposite pattern, consistent with their negative correlation to word frequency ($-.4$ and $-.19$):
+words learned before 5 or 6 years old, or made of less than 5 letters, are substituted less than random, whereas words learned after 10 years old, or made of more than 8 letters, are substituted far more than random.
+Orthographic neighborhood density also shows a slight effect (significant at $p < .05$ in 15 of the 16 substitution models):
+words with very sparse neighborhoods are more substituted than random (which may seem counter-intuitive, but is likely because over 70\% of those words have 7 letters or more).
+Clustering coefficient shows no effect on susceptibility, and neither does Number of synonyms:
+in particular, words with many synonyms do not attract substitutions more than random (in fact, half the substitution models show they have a slight tendency to be substituted less than random).
 
-Results for this measure are gathered in Fig.~\ref{fig:feature-susceptibilities}.
-They first show an obvious strong effect of Word frequency: the more frequent a word, the less likely it is to attract substitutions.
-Indeed, susceptibility goes from $.33$ for low-frequency words down to nearly~0 for very high-frequency words.
-To make things clear, this value of $.33$ means that low-frequency words, when present in a quote undergoing a substitution, are the ones being substituted 33\% of the time on average.
+On the whole, the trends observed are consistent with known effects of word frequency, age of acquisition, and number of letters, indicating that the triggering of a substitution could behave quite similarly to word recall in standard tasks.
 
-The other features --- Age of acquisition, Number of phonemes, Clustering coefficient and Number of synonyms --- do not seem to exhibit any particularly significant effect on susceptibility.
-{If we set aside the values for low Number of phonemes}, for each of these features it is indeed possible to draw a constant line which always remains within the respective confidence intervals.
-If these variables have an effect, it is by no means as strong as it is for Word frequency.
-This is remarkably clear for Clustering coefficient and Age of acquisition, where susceptibility values remain within quite small intervals (respectively $[.13-.18]$ and $[.16-.20]$).
-{We may notice a slight effect for the lowest values of Number of synonyms and Number of phonemes, where the mean susceptibility is almost half as high as the average of the other values (respectively $.09$ vs.
-$.16$, and $.11$ vs.
-$.17$).}
-Keeping in mind the poor statistical significance of this effect, we could still wonder if the shortest words and words with fewest synonyms are significantly less susceptible to substitution.
-To further examine this phenomenon, we plotted the two-dimensional map of susceptibility values for these two features (see heatmap at the bottom right of Fig.~\ref{fig:feature-susceptibilities}).
-{Even if there are a few outlier cells, values tend to navigate around the mean value ($.16$) with little obvious regularity (except for a low number of synonyms, consistent with the unidimensional graph)}.
-{On the whole, this makes it relatively hard to draw any conclusion as regards the direction of an effect, except for the least populated value ranges (which as a result are also less significant).}
-
-All in all, apart from Word frequency and despite some local tendencies, {in general} these results do not allow us to conclude to a marked effect %or to no effect
-of the selected psycholinguistic features on substitution susceptibility.
-We may therefore globally assume that substitution targets are chosen in a more or less uniform way with respect to these features.
+\end{new}
 
@@
 
-We can thus show how words are modified once we know they are substituted, that is how their features are modified by said substitution.
-Considering a word~$\wstart$ substituted for~$\warrival$, we measure how the feature of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
-Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
+We \newtext{now examine} how words are modified \newtext{when} they are substituted, that is how their features change upon substitution.
+Considering a word~$\wstart$ substituted for~$\warrival$, we measure how a feature~\newtext{$\phi$} of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
+Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
+\todo{explain this for susceptibility too}}
 $$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
 
-Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or a repulsion) effect towards (respectively from) some values of each feature.
-In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamical systems.
+\newtext{We are interested in comparing} the value of $\nu_{\phi}(f)$ to $f$ \newtext{itself}, as \newtext{this} shows whether there is an attraction (or a repulsion) effect towards (respectively from) some values of each feature.
+In other words, plotting the $y=x$ line, we can see if substitutions tend to \newtext{attract words} towards some typical \newtext{feature value} or not --- \newtext{a standard procedure} in the study of dynamical systems.
 
-\todo{Introduce both H0 and H00}
-
-We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ was randomly chosen from the whole pool of words available in the data set for that feature.\footnote{For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.}
-In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, that is an arrival word chosen among semantically plausible though still random words.
-In this case $\warrival_{00}$ does depend on $\wstart$.
-Our conclusions hold under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
+We also introduce \newtext{two} null \newtext{hypotheses}, $\mathcal{H}_0$ \newtext{and $\mathcal{H}_{00}$}, to compare the actual variation of a word's feature to expected variations \newtext{under unbiased transformations}.
+\newtext{$\mathcal{H}_0$ models the situation where} the arrival word $\warrival$ \newtext{is} randomly chosen from the whole pool of words available in the data set for that feature.\footnote{
+For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.
+}
+In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore), the baseline variation under $\mathcal{H}_0$ may be rewritten as:
 $$\nu_{\phi}^0 (f) = \left<\phi\right>$$
+\newtext{$\mathcal{H}_{00}$ models the situation where the arrival word $\warrival$ is chosen \emph{among immediate synonyms of the start word} $\wstart$, that is an arrival word chosen among semantically plausible though still random words.
+In this case $\nu_{\phi}^{00}$ does depend on $f$:\footnote{
+\newtext{The actual implementation has an additional level of averaging since WordNet, used to get a word's synonyms, defines several meanings for a single given word.
+Therefore:
+$$\nu_{\phi}^{00} (f) = \left< \left< \left< \phi(\warrival) \right>_{\warrival \in syn(m)} \right>_{m \in meanings(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}
+}
+$$\nu_{\phi}^{00} (f) = \left< \left< \phi(\warrival) \right>_{\warrival \in syn(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}
 
@@
 \begin{seriate}
-\item the original feature (\hbox{vs.} $y=x$) and
-\item a random arrival (\hbox{vs.} $\nu_{\phi}^0$).
+\item the original feature (\hbox{vs.} $y=x$),
+\item a random arrival (\hbox{vs.} $\nu_{\phi}^0$), \newtext{and
+\item an unbiased semantically plausible arrival (\hbox{vs.} $\nu_{\phi}^{00}$)}.
 \end{seriate}
@@
 
-Results are gathered in Fig.~\ref{fig:feature-variations}.
-We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
-In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
+Results are gathered in Fig.~\ref{fig:feature-variations-global}.
+A first observation is that all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, and the slope of $\nu_{\phi}$ is smaller than 1 in absolute value, independently of the feature considered.
+\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the start word, the arrival word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
+\newtext{The observation becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
+it is always closer to the intersection with $y=x$, meaning that that intersection is an attractor point for the substitution process.
+If the slope of $\nu_{\phi}$ were greater than one (in absolute value), the arrival value would always be \emph{farther} from the intersection than the start value was, making it a \emph{repulsor} point.
+This is how the number of intersections with $y=x$, and the slope of $\nu_{\phi}$ at those intersections, characterize the behavior of substitutions.
+}}}
+In other words, beyond individual variation patterns, the substitution process exhibits a unique attractor \newtext{for each feature}.
+
 Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
@@
 The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
-%\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
-%\rk{Cam: On vire ou bien?}
 

### 2016-07-27 brains-protocol.tex
feat/results: more work on results in variation

@@
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
-\newtext{For a given group of words $g$, susceptibility is computed as the ratio of the number of times $s_g$ words of that group are substituted to the number of times $s_g^0$ words of that group would be substituted if substitutions fell randomly on substitutable words, that is:}\footnote{
+% utiliser "word-level measures", et/ou "For a given value of a given type of word-level measure (say, words of length 5), we define susceptibility as the ratio (...)"
+\newtext{For a given group of words $g$ \todo{[not clear this is not a given word -- maybe present by example instead of generally]}, susceptibility is computed as the ratio of the number of times $s_g$ words of that group are substituted to the number of times $s_g^0$ words of that group would be substituted if substitutions fell randomly on substitutable words, that is:}\footnote{
 \newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
@@
     Confidence intervals are computed with the Goodman method for multinomial proportions.}
-    \todo{Split into 2 subfigures; fix labels; add H0 dashed line.}}
+    \todo{Split into 2 subfigures; fix labels; use log-y for second pane; add H0 dashed line.}}
     \label{fig:pos-susceptibilities}
@@
     \caption{\textbf{Substitution susceptibility for feature values:} susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
+    \todo{Use log-y}
     }
@@
 
-Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{images/computed-figures/paper-variations_regression-globals_to_globalsvariation.png}
+    \caption{\newtext{\textbf{Coefficients of the regressions of feature variations:}
+	\todo{only aoa and clustering are impacted by frequency, the rest is only impacted by itself}
+	}}
+    \label{fig:feature-variations_regression-global}
+\end{figure}
+
+Second, the comparison with $\nu_{\phi}^0$ \newtext{and $\nu_{\phi}^{00}$} shows that there are two classes of attractors, depending on whether:
 \begin{APAenumerate}
-\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
-\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
+\item there is a triple intersection (of $y=x$, $\nu_{\phi}$, \newtext{and $\nu_{\phi}^0$ or $\nu_{\phi}^{00}$});
+\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$ \newtext{and $\nu_{\phi}^{00}$}.
 \end{APAenumerate}
-The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
+The first class (Number of synonyms and \newtext{Orthographic neighborhood density}) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$ \newtext{(for Number of synonyms) or $\nu_{\phi}^{00}$ (for Orthographic neighborhood density)}, and no uniform bias can be observed.
+The second class (comprising Word frequency, Age of acquisition, Clustering coefficient\newtext{, and Number of letters}) are features for which the substitution process has a clear bias, positive or negative, with respect to \newtext{both} the purely random situation ($\mathcal{H}_0$) \newtext{and the semantically plausible random situation ($\mathcal{H}_{00}$)}.
 
-On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative, with respect to the purely random situation ($\mathcal{H}_0$).
-
-Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words.
-This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
-
-Age of acquisition and Clustering coefficient, on the other hand, exhibit a clear negative bias for the substitution process.
-Both curves are significantly below their respective $\nu_{\phi}^0$ values, which is consistent with the literature on recall: words learned earlier and words with lower clustering coefficient are easier to produce than average~\citep{nelson2013activation,Zevin02}.
-Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; that is on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq .1$.
+Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$ \newtext{and $\nu_{\phi}^{00}$}, exhibits a strong bias towards more frequent words.
+This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones.%, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
+Age of acquisition, Clustering coefficient \newtext{and Number of letters}, on the other hand, exhibit a clear negative bias for the substitution process (except for high clustering values or very high number of letters).
+\newtext{The three} curves are significantly below their respective $\nu_{\phi}^0$ \newtext{and $\nu_{\phi}^{00}$ curves for most start values}, which is consistent with the literature on recall:
+words learned earlier, with lower clustering coefficient \newtext{or with fewer letters} are easier to produce than average~\citep{nelson_how_2013,zevin_age_2002,baddeley_word_1975}.
 
 To make things concrete, here is an example substitution taking place in the data set.
-At the end of January 2009, many media websites reported the following quote,
-
+\newtext{Around mid-November 2008, several media websites reported the following quote from Burmese poet Saw Wai (arrested for one of his poems),
 \begin{quote}
-    ``The massive economic upheaval being experienced across the globe is sparing no one in the consumer electronics world.''
+    ``Senior general Than Shwe is foolish with power.''
 \end{quote}
@@
 \begin{quote}
-    ``The massive economic upheaval being experienced across the \textbf{world} is sparing no one in the consumer electronics world.''
+    ``Senior general Than Shwe is \textbf{crazy} with power.''
 \end{quote}
-The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the data set, and has a Clustering coefficient of $.24$.
-The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the data set, and has a Clustering coefficient of $.05$.
-(Both words have four phonemes.)
+The word \emph{foolish} is acquired at an average of 8.94 years old, appears 675 times in the data set, has a Clustering coefficient of $8.2 \times 10^{-3}$ and is 7 letters long.
+The word it was replaced with, \emph{crazy}, is acquired on average at 5.22 years old, appears about 4.1k times in the data set, has a Clustering coefficient of $1.7 \times 10^{-3}$, and is 5 letters long.}
 Such a change, though minor in appearance, is a typical example of alteration along the lines shown by our results.
 
-%globe (age 6.5, phonemes 4.0, freq 8.18, cc -1.42, susceptibility 1.0)
-%world  [age 5.32, phonemes 4.0, freq 11.89, cc -2.95, susceptibility 0.272727272727]
+% foolish / crazy
+%  frequency: (6.5147126908725301, 8.3192299386323256)
+%  aoa: (8.94, 5.22)
+%  clustering: (-4.7983132551456364, -6.369204119494885)
+%  letters_count: (7, 5)
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
-    \caption{}
+    \caption{\todo{Bring back to 1, and use log-y}}
     \label{fig:feature-variations-sentencerel}

### 2016-07-28 brains-protocol.tex
feat/results: finish variation subsection

@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_global.png}
-    \caption{\textbf{Feature variation upon substitution:} %$\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
-    %The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
-    %The intersection with $y = x$ marks the attractor value.
-    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.
+    \caption{\textbf{Feature variation upon substitution:}
+    $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals based on Student's \emph{t}-distribution.
+    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias \newtext{compared to a purely random variation}.
+    \newtext{The position with respect to the dash-dotted line representing $\mathcal{H}_{00}$ ($\nu_{\phi}^{00}$) indicates the bias compared to a semantically plausible random variation obtained by choosing a random synonym of the disappearing word.}
+    The intersection with $y = x$ marks the attractor value.
+    The fact that all curves have slopes smaller than 1 \newtext{in absolute value} means that the substitution operation is contractile on average:
+    \newtext{it brings} each feature \newtext{closer to} its own specific asymptotic range.
     }
@@
 
-\todo{Update results, explaining binning and toning down contractility, and relate to missed literature}
+\todo{Relate to missed literature}
 
@@
 A first observation is that all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, and the slope of $\nu_{\phi}$ is smaller than 1 in absolute value, independently of the feature considered.
-\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the start word, the arrival word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
-\newtext{The observation becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
-it is always closer to the intersection with $y=x$, meaning that that intersection is an attractor point for the substitution process.
-If the slope of $\nu_{\phi}$ were greater than one (in absolute value), the arrival value would always be \emph{farther} from the intersection than the start value was, making it a \emph{repulsor} point.
-This is how the number of intersections with $y=x$, and the slope of $\nu_{\phi}$ at those intersections, characterize the behavior of substitutions.
+\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the disappearing word, the appearing word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
+\newtext{This reasoning is standard in the analysis of dynamical systems (where the same transformation is applied to the whole system over and over), and becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
+the arrival value is always closer to the intersection with $y=x$, meaning that that intersection is an attractor point for the substitution process.
+If the slope of $\nu_{\phi}$ were greater than one (in absolute value), the arrival value would always be \emph{farther} from the intersection than the start value was, making the intersection with $y=x$ a \emph{repulsor} point.
+This is how the number of intersections with $y=x$ and the slope of $\nu_{\phi}$ at those intersections characterize the behavior of substitutions.
 }}}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/computed-figures/paper-variations_regression-globals_to_globalsvariation.png}
-    \caption{\newtext{\textbf{Coefficients of the regressions of feature variations:}
-	\todo{only aoa and clustering are impacted by frequency, the rest is only impacted by itself}
+    \includegraphics[width=\linewidth]{images/computed-figures/paper-variations_regression-globals_to_globalsvariation-heatmap.png}
+    \caption{\newtext{\textbf{Feature variations regression coefficients}:
+    source feature values and feature variations were normalized to $[0; 1]$ to ensure the coefficients are comparable.
+    Significance levels for individual \emph{t}-tests against the hypothesis of a null coefficient are denoted by stars below the corresponding coefficient (*** for $p \leq .001$, ** for $p \leq .01$, * for $p \leq .05$, and nothing when $p > .05$).
+	Frequency has a slight effect on Age of acquisition and Clustering coefficient, with small coefficients compared to the respective diagonal ones.
+	Aside from those two, only diagonal values are significantly non null.
 	}}
@@
 Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$ \newtext{and $\nu_{\phi}^{00}$}, exhibits a strong bias towards more frequent words.
-This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones.%, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
+This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones.
 Age of acquisition, Clustering coefficient \newtext{and Number of letters}, on the other hand, exhibit a clear negative bias for the substitution process (except for high clustering values or very high number of letters).
@@
 words learned earlier, with lower clustering coefficient \newtext{or with fewer letters} are easier to produce than average~\citep{nelson_how_2013,zevin_age_2002,baddeley_word_1975}.
+\newtext{All these effects are significant at $p < .05$ (and more often $p < .001$) and were verified across the 16 substitution models.}
+
+\begin{new}
+
+To make sure our observations are not the product of correlations or interactions, we model the variations of the 6 features as a linear function of the start word's feature values:
+$$\bm{\phi}(\warrival) - \bm{\phi}(\wstart) = \bm{A} +  \bm{B} \cdot \bm{\phi}(\wstart)$$
+where $\bm{\phi}$ is the vector of all 6 features of a word, $\bm{A}$ is an intercept vector, and $\bm{B}$ is a $6 \times 6$ coefficients matrix.
+This regression achieves an overall $R^2$ of $.33$.
+The corresponding matrix of coefficients $\bm{B}$ is shown in Fig.~\ref{fig:feature-variations_regression-global}:
+aside from Age of acquisition and Clustering coefficient on which word frequency has a slight effect, the variation of all other features depends solely on the disappearing word's same feature.
+In other words there is little to no interaction between a disappearing word's features in determining the variations that that word will undergo when substituted.
+
+\end{new}
 
 To make things concrete, here is an example substitution taking place in the data set.
-\newtext{Around mid-November 2008, several media websites reported the following quote from Burmese poet Saw Wai (arrested for one of his poems),
+\newtext{Around mid-November 2008 several media websites reported the following quote from Burmese poet Saw Wai (arrested for one of his poems),
 \begin{quote}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
-    \caption{}
+    \caption{\todo{Bring back to 1, and use log-y}}
     \label{fig:feature-susceptibilities-in_sentence}
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
-    \caption{\todo{Bring back to 1, and use log-y}}
+    \caption{}
     \label{fig:feature-variations-sentencerel}

### 2016-07-28 brains-appendix.tex
feat/results: finish presentation of results

--- 2016-07-26 brains-protocol.tex
@@
-% !TEX root = brainscopycut.tex
-% ============================
-\section{Methods} % =========
-% ============================
-\label{sec:protocol}
-
-\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
-
-\newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
-\newtext{Indeed} quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
-First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
-Second, they stem from a unique original version, and are ideally traceable back to that version.
-Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping:
-not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.
-
-We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
-\newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
-Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
-%In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
-%\todo{, as this would be akin to a drink from the firehose.
-\sf [Cam: supprimer?
-potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc.
-outside the scope etc.
-further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion?
-ou bien il s'agit d'une critique directe d'un reviewer ?
-Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
-Our goal, therefore, is to show that such effects exist and are measurable even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
-%\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
-If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
-As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}
-\todo{this last sentence is unclear or not useful}
-
-To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
-\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see "Substitution model").}
-To quantify those substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
-
-The next subsections describe the data set and the measures we used to assess this cognitive bias.
-
-\subsection{\emph{In vivo} utterances}
-
-We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
-This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 (\citealp{leskovec_memetracker:_2009}).\footnote{
-The original article~\citep{leskovec_meme-tracking_2009} does not provide further details on the source selection methodology.
-}
-\newtext{The authors automatically extracted quotations} from this corpus.
-Each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person; for instance:
-\begin{quote}
-The Bank of England said, "these operations are designed to address funding pressures over quarter-end."
-\end{quote}
-
-\newtext{Then, the authors gathered quotations in a graph and connected each pair that differed by no more than one word or that shared at least ten consecutive words (they tested this procedure with a number of different parameters, see~\citealp{leskovec_meme-tracking_2009}, for more details).}
-We find for example the following variation of the above quote:
-\begin{quote}
-"these operations are \textbf{intended} to address funding pressures over quarter-end."
-\end{quote}
-\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
-This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.
-
-Manual inspection of this data set revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
-Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
-To filter them out, we exclude quotes with less than 5 words or \newtext{whose occurrences span more} than 80 days (indicating causally unrelated occurrences), as well as quotes not written in English.
-\newtext{Clusters that are emptied by this procedure are therefore excluded.}
-If, after this screening, a cluster's occurrences still span more than 80 days (because of short-lived but unrelated quotes far apart in time), we also exclude it.
-\newtext{We eventually keep 50,427 clusters (out of 71,568; \hbox{i.e.} 70.5\%), containing a total of 141,324 unique quotes (out of 310,457; \hbox{i.e.} 45.5\%) making up about 2.60m occurrences (out of 7.67m; \hbox{i.e.} 33.9\%).\footnote{
-The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is to be expected for everyday language utterances.
-}
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is "the city is tired of me and the organization and I have run our course together"), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
-35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
-Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.}\footnote{
-\newtext{A similar analysis was made for language detection, which is part of the cluster filtering:
-out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
-Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
-}}
-
-\subsection{Word-level measures}
-
-\subsubsection{Lexical features}
-
-We first introduce some lexical measures on words.
-
-\begin{APAitemize}
-
-    \item \textbf{Word frequency}: the frequency at which words appear in our data set, known to be relevant for both recognition and recall~\citep{gregg_word_1976},
-
-    \item \textbf{Age of Acquisition}: the average age at which words are learned~\citep[obtained from][]{kuperman_age--acquisition_2012}, known to have different effects than word frequency~\citep{morrison_roles_1995,dewhurst_separate_1998},
-
-    \item \newtext{\textbf{Phonological} and \textbf{Orthographic Neighborhood Density}~\citep[obtained from][]{marian_clearpond:_2012}, also known to be relevant for word production~\citep{garlock_age--acquisition_2001},}
-
-    \item The average \textbf{Number of Phonemes} and \textbf{Number of Syllables} for all pronunciations of a word (obtained from the Carnegie Mellon University Pronouncing Dictionary, \citealp{weide_cmu_1998})\footnote{
-    The CMU Pronouncing Dictionary is included in the NTLK package~\citep{bird_nltk_2009}, the natural language processing toolkit we used for the analysis.
-    },
-    \newtext{as well as \textbf{Number of Letters}}, as a proxy to word production cost,
-
-    \item The average \textbf{Number of Synonyms} for all meanings of a word~\citep[obtained from][]{wordnet_princeton_2010} as an \emph{a priori} indicator of how easy it would be to replace a word.
-
-\end{APAitemize}
-
-\newtext{We also consider grammatical types within quotations by detecting \emph{Part-of-Speech} (POS) categories with TreeTagger~\citep{schmid_probabilistic_1994}; we distinguish between verbs, nouns, adjectives, adverbs, and stopword-like words.}
-
-\medskip
-
-\newtext{Aside from these raw features, the systemic dimension of vocabulary~\citep{cornish_systems_2013} has led authors to develop measures based on the full topology of networks built from free association data or phonological similarity.
-Several such measures have been shown to be involved in recall, recognition, and naming tasks~\citep{nelson_how_2013,chan_network_2010,griffiths_google_2007}.}
-
-\newtext{To compute those features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
-As \citet{nelson_university_2004} explain, "free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect."
-\newtext{Similar to what \citet{griffiths_google_2007} did, we first considered the directed \newtext{weighted} network formed by association norms, where nodes are words and edges are directed from cue to target word, with a weight equal to the association strength (that is the probability of that target word being produced when this particular cue is presented).
-This network is of particular interest since it lets us define features that reflect the associations driving false memories in word lists~\citep{deese_prediction_1959}, a phenomenon which may be involved in the transformation of quotations.}
-
-We used three standard measures on the FA network:
-
-\begin{APAitemize}
-
-    \item \textbf{\newtext{Incoming} degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \textbf{Pagerank}~\citep{page_pagerank_1999}, which has already been used on the FA network by \citet{griffiths_google_2007}.
-    In the present case these two polysemy-related measures are quasi-perfectly correlated.\footnote{
-    \newtext{Note that in-degree does not take the weights of links into account, as it counts 1 for each incoming link.
-    Pagerank on the other hand, does take the weights into account.}
-    }
-    %, and measure polysemy by taking into account the structure of the entire network.
-    %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
-Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
-
-    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{freeman_set_1977}.
-    This quantity tells us if some words behave as unavoidable waypoints on association chains connecting one word to another.\footnote{
-    \newtext{For this measure, weights are interpreted as inverse cost:
-    the stronger a link, the easier it is to travel across it.
-    A stronger link will be favored over weaker links in the computation of the shortest path between two words.}
-    }
-
-    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\citep{watts_collective_1998}, computed on the undirected \newtext{weighted} version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual versus possible edges between a node's neighbors;
-    \newtext{this is poorly defined in the case of directed networks, which led us to ignore the direction of links in the network for this measure (if two words are connected in both directions, the weights of both links are added to make the final undirected link's weight).}
-    }
-    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a free association point of view).
-
-\end{APAitemize}
-
-\subsubsection{Variable correlations}
-
-\newtext{Several of these features are strongly related and can be grouped together.
-To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
-\newtext{The distributions of original and log-transformed features appears in the Supplementary Information, along with scatter plots for each feature pair, summarized below with correlation values.
-}}}
-
-\begin{figure}[!th]
-    \centering
-    \includegraphics[width=\linewidth]{images/computed-figures/all-feature_correlations.png}
-    \caption{Spearman correlations in the initial set of features}
-    \label{fig:feature-corrs-initial}
-\end{figure}
-
-\begin{new}
-
-The pairwise correlations in the initial set of features appears in Fig.~\ref{fig:feature-corrs-initial}.
-By looking at absolute values, three subsets of highly correlated features can be easily identified:
-(a) number of letters, phonemes, and syllables with pairwise correlations greater than .75;
-(b) orthographic and phonological neighborhood densities, with a correlation of .8;
-(c) age of acquisition, betweenness, degree, and pagerank centralities, with absolute pairwise correlations at .41, .6, .59, .63, .61 and .85.
-Applying a feature agglomeration algorithm targeted at 6 groups refined this observation by producing identical (a) and (b) groups, a (c) group without betweenness centrality which was instead assigned to a group (d) with clustering coefficient, and the remaining features (frequency and number of synonyms) as singletons.\footnote{
-\newtext{Agglomerating into less than 6 groups merged groups (a) and (b), which we excluded to keep neighborhood densities in their own group;
-agglomerating into more than 6 groups separated age of acquisition from group (c), which we excluded given its high correlation values to the rest of group (c). 
-We used scikit-learn's FeatureAgglomeration class for this procedure~\citep{pedregosa_scikit-learn:_2011}.}
-}
-
-Since our data is about written transformations, number of letters and orthographic neighborhood density are the natural representatives of groups (a) and (b) respectively.
-Given the importance of age of acquisition in the lexical feature literature, we chose it to represent group (c).
-Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
-The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in the appendix).\footnote{
-Note that feature values stem from different data sets which do not always encode the same words.
-Indeed, we have data on frequency for about \newtext{33.5k} words, on age of acquisition for 30.1k words, on clustering coefficient for 5.7k words, number of synonyms 111.2k, and \newtext{orthographic density 17.8k words}.
-Quite often then, not all features are available for all words in our data set;
-however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
-
-\end{new}
-
-\begin{figure}[!th]
-    \centering
-    \includegraphics[width=0.6412\linewidth]{images/computed-figures/paper-feature_correlations-v2.png}
-    \caption{Spearman correlations in the filtered set of features}
-    \label{fig:feature-corrs-filtered}
-\end{figure}
-
-\subsection{Substitution model}
-\label{sec:temporal-binning}\label{sec:model}
-
-We finally need a substitution detection model, for the \newtext{quotation} data we use presents a challenge:
-quote-to-quote transformations and substitutions are not explicitly encoded in the data set.
-More precisely, each set of quotations bears no explicit information about either the authoritative original quotation, or the source quotation(s) each author relied on when creating a new post and reproducing (and possibly altering) that source.
-\newtext{In other words} we face an inference problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
-
-We therefore model the underlying quotation selection process by making a few additional assumptions.
-\newtext{Given a particular occurrence of a quotation, the first issue is deciding whether} that occurrence is a strict copy of an earlier occurrence, or a substitution of another \newtext{quotation}, or maybe a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
-\newtext{The second issue is deciding which source originated such a substitution when several candidate sources are available.}
-
-Let us give an example:
-\newtext{say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in two different blogs on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in another blog on the 20th of January.
-The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
-And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$ as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
-Now say a third version, "These \textbf{allegations} are false and \textbf{incoherent}" ($q''$) also appears once on January 19 and once on January 20 after $q'$.
-$q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (this below for further details).
-$q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
-Conversely, the occurrence of $q''$ on the 20th could be considered as a substitution from $q'$, or as a faithful copy from its initial occurrence on January 19.}
-(Options shown in Fig.~\ref{fig:substitution-unmodelled}.)
-
-\begin{figure}[h]
-    \centering
-	\def\svgwidth{\linewidth}
-	\small
-	\input{images/substitution-unmodelled.pdf_tex}
-	\caption{{\bf Possible paths from occurrence to occurrence.}
-	\newtext{$q$, $q'$ and $q''$ are three quotation variants belonging to the same cluster.
-	$q$ and $q''$ differ by two words, but $q'$ differs from both $q$ and $q''$ by a word.
-	The second occurrence of $q$ can safely be considered a faithful copy of the first, but the occurrences of $q'$ and $q''$ are uncertain:
-	while the first occurrence of $q'$ is most likely a substitution from $q'$, it could also stem from $q''$;
-	conversely, the second occurrence of $q''$ could also be a substitution from $q'$ instead of being a faithful copy of its first occurrence.}
-	}
-	\label{fig:substitution-unmodelled}
-\end{figure}
-
-\newtext{One way to settle these questions is the following:}
-group quote occurrences into fixed bins spanning $\Delta t$ days (1 day in the implementation), each one representing a unit of time evolution;
-when a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote of the preceding bin $t$ (or a substring thereof) by only one word;
-if not, $q'$ is not considered to be an instance of substitution.
-\newtext{Fig.~\ref{fig:substitution-modelled-majority-all} shows the inferences made by such a model.
-The assumptions it embeds, however, are a subset of a much wider set of possibilities, each leading to alternative inferences.}
-
-\begin{new}
-
-We identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
-The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
-(1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
-(2) \textbf{bin span}, which can be 24 hours (as in the model above) or can be extended to start at the very first occurrence in the quotation family.
-The other two parameters define rules on the selection of source and destination quotes of a substitution:
-(3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
-(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
-A substitution model, then, is the given of a value for each of those parameters;
-it considers valid all the substitutions (and only those) where the source and destination follow the rules set out by the parameters.
-If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.
-
-\end{new}
-\begin{figure*}[h]
-	\centering
-	\begin{subfigure}{.49\textwidth}
-	    \def\svgwidth{\textwidth}
-	    \small
-	    \input{images/substitution-Time=discrete-Source=majority-Past=last_bin-Durl=all.pdf_tex}
-	    \caption{Source must be majority in preceding bin, destination can be anything}
-	    \label{fig:substitution-modelled-majority-all}
-	\end{subfigure}
-	~
-	\begin{subfigure}{.49\textwidth}
-	    \def\svgwidth{\textwidth}
-	    \small
-	    \input{images/substitution-Time=discrete-Source=majority-Past=last_bin-Durl=exclude_past.pdf_tex}
-	    \caption{Source must be majority in preceding bin, destination must not appear in preceding bin}
-	    \label{fig:substitution-modelled-majority-exclude_past}
-	\end{subfigure}
-
-	\begin{subfigure}{.49\textwidth}
-	    \def\svgwidth{\textwidth}
-	    \small
-	    \input{images/substitution-Time=discrete-Source=all-Past=last_bin-Durl=all.pdf_tex}
-	    \caption{Source can be anything, destination can be anything\\\hfill} % \\\hfill to add a line, aligning with the other panels which have two-line captions
-	    \label{fig:substitution-modelled-all-all}
-	\end{subfigure}
-	~
-	\begin{subfigure}{.49\textwidth}
-	    \def\svgwidth{\textwidth}
-	    \small
-	    \input{images/substitution-Time=discrete-Source=all-Past=last_bin-Durl=exclude_past.pdf_tex}
-	    \caption{Source can be anything, destination must not appear in preceding bin}
-	    \label{fig:substitution-modelled-all-exclude_past}
-	\end{subfigure}
-	\caption{\newtext{
-	{\bf Substitution models.}
-	Substitutions inferred by four models in the situation introduced by Fig.~\ref{fig:substitution-unmodelled}.
-	Each of these models uses discretely positioned bins spanning 1 day (see the main text for a complete description of parameters).
-	In the top left panel~(a), $q$ holds the majority in the first bin and is considered the unique basis for $q'$ in bin~2.
-	$q'$ and $q''$ have equal maximum frequency in bin~2 however, so both are sources of substitutions towards bin~3.
-	In the top right panel~(b), quotes that appear in the preceding bin cannot be the target of a substitution; this removes two substitutions compared to panel~(a).
-	In the bottom left panel~(c), the majority constraint is lifted compared to panel~(a), making $q''$ in bin~1 a candidate source for $q'$ in bin~2.
-	In the bottom right panel~(d), the majority constraint is also lifted compared to panel~(a) (adding the same $q'' \rightarrow q'$ substitution as in panel~(c)), and the excluded-past constraint is added as in panel~(b) (removing two same substitutions from bin~2 to bin~3 as in panel~(b)).
-	If the bins were extended to the beginning of the quotation family, the excluded-past constraint would also remove the $q' \rightarrow q$ substitution from bin~2 to bin~3.
-	In all four panels, a background rectangle or square indicates the quotation is the source of a substitution.
-	A thick border on that rectangle or square indicates the quotation was selected because it has maximum frequency.}
-	}
-    \label{fig:substitution-modelled}
-\end{figure*}
-
-\begin{new}
-
-Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
-Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use discretely positioned bins spanning 1 day:
-later occurrences of $q'$ and $q''$ are counted as substitutions in Fig.~\ref{fig:substitution-modelled-majority-all} and Fig.~\ref{fig:substitution-modelled-all-all}, whereas in Fig.~\ref{fig:substitution-modelled-majority-exclude_past} and Fig.~\ref{fig:substitution-modelled-all-exclude_past} they are not.
-
-The results reported and discussed in the following sections are valid for all 16 models, and the graphics we present were produced by the model first introduced above.
-Finally, note that this inference procedure is one of the reasons we restricted our analysis to single-substitutions:
-looking for more complex transformations would
-(a) exponentially increase the number of candidate sources for a destination occurrence, which correspondingly reduces the confidence in inferences made,
-and (b) greatly increase the complexity of the transformation models used to make these inferences.\footnote{
-\newtext{We checked that this restriction does not bias the results discussed below by extending our protocol to two-substitution transformations.
-The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.}
-}
-
-\end{new}
-
-\medskip
-
-In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread into 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
-From these substitutions we remove those featuring stop words, minor spelling changes (e.g.
-center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
-this eventually yields \newtext{6,318} valid substitutions \newtext{(before merging substitutions that share the same destination)}.\footnote{
-\newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
-Precision was evaluated over a random subset of 100 \emph{kept} substitutions, showing a score of .87.
-Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
-out of the first 100 substitutions coded for recall, those with levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}
-}
-
-\section{Results}\label{sec:results}
-
-\begin{new}
-
-While most of the substitutions we obtain with this procedure replace a word with another semantically related word, the two are rarely direct synonyms:
-only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
-\newtext{A similar behavior is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
-}}
-Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
-The following examples illustrate this behavior:
-
-\begin{APAitemize}
-	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
-	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
-	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
-\end{APAitemize}
-
-\end{new}
-
-\newtext{Our question concerns the low-level properties of these substitutions; to address it} we build the following two observables for each word feature.
-First, we measure \newtext{which word features are more or less substituted compared to how often they would be if the process were random, in order to capture} the susceptibility for words to be the target of a substitution in a quote.
-Second, we measure the change in word feature upon substitution, looking at the variation of a given feature between start and arrival words.
-\newtext{Since sentence context is also central to this process, we extend these two observables by applying them to feature values relative to the distribution of feature values of the sentence in which a word appears.}
-
-Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, \newtext{that is} we do not take invariant quotations into account.
-Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it \newtext{we cannot} know whether there has been perfect human reformulation or simply digital copy-pasting of a source~(``{\sc Ctrl-C}/{\sc Ctrl-V}'').
-\newtext{Moreover}, perfect human reformulation possibly involves different practices than those involved in alteration ---~for instance drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
-The two situations are different enough to be studied separately, and we focus here on the latter.
-
-\subsection{Susceptibility}
-
-We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
-\newtext{For a given group of words $g$, susceptibility is computed as the ratio of the number of times $s_g$ words of that group are substituted to the number of times $s_g^0$ words of that group would be substituted if substitutions fell randomly on substitutable words, that is:}\footnote{
-\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
-If, for instance, half the content words of a given quote are from group $g$, such a quote contributes .5 to the total $s_g^0$.}
-}
-$$\sigma_g = \frac{s_g}{s_g^0}$$
-
-In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
-\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (e.g.
-for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
-    \caption{\newtext{\textbf{Substitution susceptibility for POS categories:} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
-    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts, the bottom panel shows $\sigma_{POS}$, the ratio of the two.
-    Confidence intervals are computed with the Goodman method for multinomial proportions.}
-    \todo{Split into 2 subfigures; fix labels; add H0 dashed line.}}
-    \label{fig:pos-susceptibilities}
-\end{figure}
-
-\medskip
-
-\begin{new}
-
-Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
-A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Stopword-like}\footnote{
-\newtext{The \emph{Stopword-like} category gathers all the POS groups partially of completely affected by the stopword filter for substitutions: coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections.
-Note that the stopword filter equally affects $s$ and $s^0$, so while both numbers are very small for these categories (which led us to group them together) the reported susceptibility is not biased by this filtering.}
-}
-and \emph{Adverb} categories.
-Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for stopword-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
-none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
-While we acknowledge the low susceptibilities of adverbs and stopword-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
-it seems, then, that POS categories don't capture any strong bias in the selection of substitution targets.
-
-\end{new}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
-    \caption{\textbf{Substitution susceptibility for feature values:} susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
-    }
-    \label{fig:feature-susceptibilities-global}
-\end{figure*}
-
-\todo{Relate to missed literature (\#15)}
-
-\begin{new}
-
-The results for word features presented in Fig.~\ref{fig:feature-susceptibilities-global}, on the other hand, show marked effects for several features.
-Word frequency, Age of acquisition, and Number of letters each exhibit significant susceptibility variations (Goodman goodness-of-fit with $p < .05$ in all substitution models, $p < .001$ in most) consistent with known effects of those features on recall.
-High-frequency words, much easier to recall, are substituted about half as much as they would be at random; conversely low-frequency words, harder to recall, are substituted about 50\% more than random.
-Age of acquisition and Number of letters show the opposite pattern, consistent with their negative correlation to word frequency ($-.4$ and $-.19$):
-words learned before 5 or 6 years old, or made of less than 5 letters, are substituted less than random, whereas words learned after 10 years old, or made of more than 8 letters, are substituted far more than random.
-Orthographic neighborhood density also shows a slight effect (significant at $p < .05$ in 15 of the 16 substitution models):
-words with very sparse neighborhoods are more substituted than random (which may seem counter-intuitive, but is likely because over 70\% of those words have 7 letters or more).
-Clustering coefficient shows no effect on susceptibility, and neither does Number of synonyms:
-in particular, words with many synonyms do not attract substitutions more than random (in fact, half the substitution models show they have a slight tendency to be substituted less than random).
-
-On the whole, the trends observed are consistent with known effects of word frequency, age of acquisition, and number of letters, indicating that the triggering of a substitution could behave quite similarly to word recall in standard tasks.
-
-\end{new}
-
-\subsection{Variation}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_global.png}
-    \caption{\textbf{Feature variation upon substitution:} %$\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
-    %The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
-    %The intersection with $y = x$ marks the attractor value.
-    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.
-    }
-    \label{fig:feature-variations-global}
-\end{figure*}
-
-We \newtext{now examine} how words are modified \newtext{when} they are substituted, that is how their features change upon substitution.
-Considering a word~$\wstart$ substituted for~$\warrival$, we measure how a feature~\newtext{$\phi$} of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
-Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
-\todo{explain this for susceptibility too}}
-$$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
-
-\newtext{We are interested in comparing} the value of $\nu_{\phi}(f)$ to $f$ \newtext{itself}, as \newtext{this} shows whether there is an attraction (or a repulsion) effect towards (respectively from) some values of each feature.
-In other words, plotting the $y=x$ line, we can see if substitutions tend to \newtext{attract words} towards some typical \newtext{feature value} or not --- \newtext{a standard procedure} in the study of dynamical systems.
-
-We also introduce \newtext{two} null \newtext{hypotheses}, $\mathcal{H}_0$ \newtext{and $\mathcal{H}_{00}$}, to compare the actual variation of a word's feature to expected variations \newtext{under unbiased transformations}.
-\newtext{$\mathcal{H}_0$ models the situation where} the arrival word $\warrival$ \newtext{is} randomly chosen from the whole pool of words available in the data set for that feature.\footnote{
-For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.
-}
-In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore), the baseline variation under $\mathcal{H}_0$ may be rewritten as:
-$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
-\newtext{$\mathcal{H}_{00}$ models the situation where the arrival word $\warrival$ is chosen \emph{among immediate synonyms of the start word} $\wstart$, that is an arrival word chosen among semantically plausible though still random words.
-In this case $\nu_{\phi}^{00}$ does depend on $f$:\footnote{
-\newtext{The actual implementation has an additional level of averaging since WordNet, used to get a word's synonyms, defines several meanings for a single given word.
-Therefore:
-$$\nu_{\phi}^{00} (f) = \left< \left< \left< \phi(\warrival) \right>_{\warrival \in syn(m)} \right>_{m \in meanings(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}
-}
-$$\nu_{\phi}^{00} (f) = \left< \left< \phi(\warrival) \right>_{\warrival \in syn(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}
-
-This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to
-\begin{seriate}
-\item the original feature (\hbox{vs.} $y=x$),
-\item a random arrival (\hbox{vs.} $\nu_{\phi}^0$), \newtext{and
-\item an unbiased semantically plausible arrival (\hbox{vs.} $\nu_{\phi}^{00}$)}.
-\end{seriate}
-
-\medskip
-
-\todo{Update results, explaining binning and toning down contractility, and relate to missed literature}
-
-Results are gathered in Fig.~\ref{fig:feature-variations-global}.
-A first observation is that all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, and the slope of $\nu_{\phi}$ is smaller than 1 in absolute value, independently of the feature considered.
-\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the start word, the arrival word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
-\newtext{The observation becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
-it is always closer to the intersection with $y=x$, meaning that that intersection is an attractor point for the substitution process.
-If the slope of $\nu_{\phi}$ were greater than one (in absolute value), the arrival value would always be \emph{farther} from the intersection than the start value was, making it a \emph{repulsor} point.
-This is how the number of intersections with $y=x$, and the slope of $\nu_{\phi}$ at those intersections, characterize the behavior of substitutions.
-}}}
-In other words, beyond individual variation patterns, the substitution process exhibits a unique attractor \newtext{for each feature}.
-
-Second, the comparison with $\nu_{\phi}^0$ shows that there are two classes of attractors, depending on whether:
-\begin{APAenumerate}
-\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
-\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
-\end{APAenumerate}
-The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
-
-On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative, with respect to the purely random situation ($\mathcal{H}_0$).
-
-Word frequency, with $\nu_{\phi}$ always significantly above $\nu_{\phi}^0$, exhibits a strong bias towards more frequent words.
-This, in turn, is consistent with the hypothesis that substitution is a recall process, since common words are favored over awkward ones, while it goes against the idea that it could be a familiarity process, where awkward terms would be favored.
-
-Age of acquisition and Clustering coefficient, on the other hand, exhibit a clear negative bias for the substitution process.
-Both curves are significantly below their respective $\nu_{\phi}^0$ values, which is consistent with the literature on recall: words learned earlier and words with lower clustering coefficient are easier to produce than average~\citep{nelson2013activation,Zevin02}.
-Clustering coefficient has the additional particularity that, on average, the destination word does not depend on the start word; that is on average, substitutions will always produce words with a clustering coefficient around $\exp(-2.4) \simeq .1$.
-
-To make things concrete, here is an example substitution taking place in the data set.
-At the end of January 2009, many media websites reported the following quote,
-
-\begin{quote}
-    ``The massive economic upheaval being experienced across the globe is sparing no one in the consumer electronics world.''
-\end{quote}
-and a smaller number of media websites, and blogs, reported the following,
-\begin{quote}
-    ``The massive economic upheaval being experienced across the \textbf{world} is sparing no one in the consumer electronics world.''
-\end{quote}
-The word \emph{globe} is acquired at an average of 6.5 years old, appears about 3.5k times in the data set, and has a Clustering coefficient of $.24$.
-The word it was replaced with, \emph{world}, is acquired on average at 5.3 years old, appears about 146k times in the data set, and has a Clustering coefficient of $.05$.
-(Both words have four phonemes.)
-Such a change, though minor in appearance, is a typical example of alteration along the lines shown by our results.
-
-%globe (age 6.5, phonemes 4.0, freq 8.18, cc -1.42, susceptibility 1.0)
-%world  [age 5.32, phonemes 4.0, freq 11.89, cc -2.95, susceptibility 0.272727272727]
-
-\subsection{Sentence context}
-
-\todo{Explain in-sentence susceptibility and sentence-relative variation, relating to global value results}
+\section{Sentence-relative feature effects}
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
-    \caption{}
+    \caption{\todo{Bring back to 1, and use log-y}}
     \label{fig:feature-susceptibilities-in_sentence}
 \end{figure*}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
-    \caption{}
-    \label{fig:feature-variations-sentencerel}
-\end{figure*}
-
-\medskip
-We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
-Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e.
-$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age slightly below 8, and a Clustering coefficient of $.1$).

### 2016-07-28 brains-protocol.tex
feat/results: finish presentation of results

@@
 % utiliser "word-level measures", et/ou "For a given value of a given type of word-level measure (say, words of length 5), we define susceptibility as the ratio (...)"
-\newtext{For a given group of words $g$ \todo{[not clear this is not a given word -- maybe present by example instead of generally]}, susceptibility is computed as the ratio of the number of times $s_g$ words of that group are substituted to the number of times $s_g^0$ words of that group would be substituted if substitutions fell randomly on substitutable words, that is:}\footnote{
+\newtext{For a given group of words $g$ \todo{[not clear this is not a given word -- maybe present by example instead of generally]}, susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
+That is:}\footnote{
 \newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
@@
 
-\todo{Explain in-sentence susceptibility and sentence-relative variation, relating to global value results}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
-    \caption{\todo{Bring back to 1, and use log-y}}
-    \label{fig:feature-susceptibilities-in_sentence}
-\end{figure*}
-
 \begin{figure*}[!th]
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
-    \caption{}
+    \caption{\newtext{\textbf{Sentence-relative feature variation:}
+    $\nu_{\phi_r}$, average sentence-relative feature value of the appearing word as a function of the sentence-relative value of the disappearing word (binned by quartiles), with 95\% asymptotic confidence intervals based on Student's \emph{t}-distribution.
+    $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ are similarly converted to be sentence-relative.
+    Attraction, magnitude and direction of bias with respect to null hypotheses are similar to Fig.~\ref{fig:feature-variations-global}.
+    However, attractors are always positioned between sentence median ($y=0$) on one side and $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ on the other side.
+    Clustering coefficient, Number of synonyms and Orthographic neighborhood density are limit cases, with triple intersections with one of the null hypothesis curves.
+    }
+    \todo{fix labels}
+    }
     \label{fig:feature-variations-sentencerel}
@@
 
+\begin{new}
+
+The alterations we study are always made in a context, that is while the author is writing.
+We wish to ask, therefore, if taking that context into account can provide more insight into the substitution process.
+To do so we adapt the two observables presented above to capture some of the relationships between a word and the sentence it appears in.
+
+Let us start with the first one:
+given a feature $\phi$, we define the context-relative susceptibility to substitution in the following three steps.
+(1) For each quote in which a substitution appears, compute the distribution of $\phi$ values in that quote (excluding stopwords) and divide it into quartiles.
+(2) Count how many times each quartile (first, second, third or fourth) contains a word that is substituted.
+This procedure tells us, for $i \in \{1, 2, 3, 4\}$, how many times substitutions fall in the $i$-th quartile of each in-quote distribution of $\phi$;
+in other words it gives us the numerator $s_{q_i}$ for the computation of susceptibility, where $q_i$ represents the $i$-th quartile of the distributions of $\phi$ in the quotes.
+(3) Finally divide each quartile count by its corresponding $s_{q_i}^0$, that is the number of times the $i$-th quartile would receive substitutions if targets were picked at random;
+since the random situation would equally distribute a fourth of all substitutions to each quartile, we divide by the number of substitutions divided by 4.
+Taking for instance word frequency, this measure tells us if words that have high- or low-frequency \emph{compared to the quote they appear in} are more or less substituted than at random.
+
+Surprisingly the results for this measure are no different from the context-free measure, as can be seen in Fig.~\ref{fig:feature-susceptibilities-in_sentence} of the Appendix:
+low-frequency words compared to the sentence are substituted much more than higher-frequency words, words learned earlier than the rest of the sentence are substituted less than words learned later, shorter words less than longer words, and words with scarce neighborhoods slightly more than words with denser neighborhoods.
+Clustering coefficient and Number of synonyms are, here again and across all substitution models, not significantly different from $\mathcal{H}_0$:
+with or without context, they do not seem relevant to the selection of substitution targets.
+
+Feature variation is more easily extended to the context-relative case.
+To do so we consider all feature values relative to the median word feature in the sentence.
+That is, in all the equations of the previous section we replace $\phi(w)$ with:
+$$\phi_r(w) = \phi(w) - \operatorname{median}\left\lbrace \phi(w) | w \in \text{sentence} \right\rbrace$$
+$\nu_{\phi}$, $\nu_{\phi}^0$ and $\nu_{\phi}^{00}$ each transpose to $\nu_{\phi_r}$, $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ (note that $\nu_{\phi_r}^0$ now depends on $w$ since it is sentence-relative, whereas $\nu_{\phi}^0$ did not).
+
+The results for sentence-relative feature variations are gathered in Fig.~\ref{fig:feature-variations-sentencerel}.
+Here too, the behavior is strikingly similar to the context-free measure:
+a single attractor is visible for each feature, and the magnitude and direction of biases are near-identical to those for the previous measure.
+The values of the appearing words give an additional insight into the process, however:
+the attractor value of a feature, at the intersection of $\nu_{\phi_r}$ and $y=x$, is always between the sentence median, on one side, and $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ on the other side (for Number of synonyms it is a triple intersection with $\nu_{\phi_r}^0$; for Clustering coefficient and Orthographic neighborhood density, a triple intersection with $\nu_{\phi_r}^{00}$).
+Substitutions, therefore, seem to attract words closer to the sentence median than what a random process would do.
+This is true with respect to both null hypotheses (semantically plausible or not) for Frequency, Age of Acquisition and Number of letters, and true with respect to at least one of the two null hypotheses for the remaining features.
+
+\end{new}
+
 \medskip
-We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
-Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e.
-$\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age slightly below 8, and a Clustering coefficient of $.1$).
+
+\newtext{On the whole}, we observe a clear \newtext{attraction} pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
+\newtext{More awkward words along relevant features (less frequent, learned later, or with made of more letters), both globally and with respect to the sentence they appear in, are substituted more often than more common words.
+And across all features, words are brought generally closer to the sentence median than what a random process would do.}

### 2016-07-28 brains-protocol.tex
fix/results: phrasing

@@
 \newtext{On the whole}, we observe a clear \newtext{attraction} pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
-\newtext{More awkward words along relevant features (less frequent, learned later, or with made of more letters), both globally and with respect to the sentence they appear in, are substituted more often than more common words.
-And across all features, words are brought generally closer to the sentence median than what a random process would do.}
+\newtext{More awkward words along relevant features (less frequent, learned later, or made of more letters), both globally and with respect to the sentence they appear in, are substituted more often than what would happen if targets were picked randomly in the sentences;
+conversely, more common words are substituted less.
+Finally, across all features, substituted words are attracted towards a point closer to the sentence median than what a random process, semantically plausible or not, would do.}

### 2016-07-28 brains-appendix.tex
chore/all: typos, colors, todos

@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
-    \caption{\todo{Bring back to 1, and use log-y}}
+    \caption{\newtext{\textbf{Substitution susceptibility for in-quote feature quartiles:}
+    susceptibility to substitution versus quartile of the feature distribution in the originating quote, with 95\% asymptotic confidence intervals (Goodman-based multinomial)}.
+    \todo{Bring back to 1, and use log-y}}
     \label{fig:feature-susceptibilities-in_sentence}

### 2016-07-28 brains-conclusion.tex
chore/all: typos, colors, todos

@@
 \section{Discussion}
+
+- place back into the initial question: in our example, do reps converge?
+-> we don't know because we couldn't test (ref. sub models and complicated data), but there is attraction.
+  - Remind the details
+  - It's consistent with known effects (word frequency: easier recall for high freq, age of acquistion: Zevin, Dewhurst, Morrison, clustering coefficient: nelson, age of acquisition (corr pagerank): griffiths, orth.
+neighborhoos: Garlock), which is interesting, because it's applying cogsci tools on real life data
+  - it's also a first step towards analyzing together cogsci and culture (because cogsci on real life data), which is much of a challenge still, although we don't get to see any global effect on the corpus nor reciprocity, therefore (i.e.
+we don't loop the loop yet)
+  - there's a cost to this however: noise, uncertainty (hence single sub), data shaped oppositely to lab experiments.
+  - so naturally it's a far cry from the whole story.
+It's not even the main transformation at work, so we can't conclude on the evolution of the dataset.
+But we know something more for this simple case, and what it costs to reach that knowledge.
+
+- Now we also looked at context, but there's vast oceans to explore in this direction too. Indeed there's way more to do on this case:
+  - in lab exps you can predict the word if proper measures or well-designed lists
+  - you can ask what triggers the substitution with semantic similarity: low cost of switching to a more frequent dyad, given the sentence, following Zaromb
+  - you can see how previous context influences the semantic similarity (PLI, LSA/LDA). Note that these LSA/LDA methods are ill-suited to the data.
+  - with better data, building on our software, we could do that, and look at chains
+
+More broadly with Sperber:
+  - you can have a simplistic interpretation, or you can look for attraction on different dimensions, which a more rich (and useful) interpretation.
+  - kirby and co have looked at that on artificial languages, on non-meaningful patterns (claidiere), we could look at that on real language.
+  - What we see is consistent with lineage specificity (contraction around sentence median), and many things push towards that (context that influences the fitness, as in "language as a system" from kirby, previous history that changes bias, as in PLIs), but right now we don't know in detail what pressure creates that.
+It's likely the type of task influences that.
+  - What we see is not clear w.r.t.
+convergence however: (1) we don't observe it on our data, and (2) if there's lineage specificity there's no reason convergence should appear (on contrary).
+But, this is only half the story, since in reality there's also selection, i.e.
+not all chains go on forever, so you lose some diversity, and it changes the dynamics.
+  - finally, context is addressed with relevance in Sperber, but it's still all representational, as we did here. Others have said that it will never be enough. Ingold, Di Paolo.
 

### 2016-07-28 brains-introduction.tex
chore/all: typos, colors, todos

@@
 This process should manifest itself as attractors (called "cultural attractors") in the dynamical system that models cultural evolution, that is:
-there should be areas of the representation space where cognitive effects in transformations bring representation closer to a given stable asymptotic point.\footnote{
+there should be areas of the representation space where cognitive effects in transformations bring representations closer to a given stable asymptotic point.\footnote{
 \newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
@@
 }}
-While this limits the scope of our observations to the particular data set we use, the methodological point we also make is left intact.
-By characterizing words using 6 well-known features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
+While this limits the scope of our results to the particular data set we use, the methodological point we also make is left intact.
+By characterizing words using 6 well-studied features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
 This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
@@
 This is done by using lists of words produced by free association from the target intrusion word;
-the intruding recall then happens with probability nearly proportional to the average strength of semantic association between the intruding word and the words in the list.
+the intruding recall then happens with probability nearly proportional to the average semantic association strength between the intruding word and the words in the list.
 A sizable literature studies this type of task with varying complexities in the design of the lists, a good review of which is given by \citet{zaromb_temporal_2006}.
@@
 
-Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
+\end{new}
+
+\newtext{Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
 \newtext{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited to data consisting of a high number of measures over a smaller number of lists (in which case it makes sense to ask \hbox{e.g.} what proportion of intrusions come from prior lists).
@@
 }}
-If a cognitive bias in the substitution of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.
+If a cognitive bias in the} substitution \newtext{of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.}
+
+\begin{new}
 

### 2016-07-29 brains-conclusion.tex
feat/discussion-conclusion: finish all the way to conclusion

--- 2016-06-24 brains-conclusion.tex
@@
 
-\todo{Discuss related to introduction. Attractors, lineage with specification, what we couldn't observe, how it fits into Kirby.
-}
+\begin{new}
 
-\todo{\#15: relate to missed literature}
+We aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
+The data set we used imposed a few constraints on our analysis:
+first, it was necessary to infer source-destination links, an operation made more reliable when restricting the scope of transformations to very simple cases, which we did by focusing on word substitutions.
+Second, contrary to laboratory experiments which produce data made of many repeated measures on a small number of cases (\hbox{e.g.} a given list of words), we have a great number of different cases (one case per cluster in which substitutions are found, \hbox{i.e.} 698 cases), with very few measures on each of them (average 9, median 5).
+This rendered the prediction of individual words impractical: if we cannot compute a percentage of explained data for a given case, any approximate prediction will be heavily underestimated.
+This last factor, added to the potential for variation of external conditions when authors wrote the quotes, led us to use word features to analyze the transformations by aggregating over individual cases.
+
+By characterizing substitutions with 6 features on the disappearing word, we show that authors preferentially substitute words known for being harder to recall:
+most prominently words with low frequency~\citep{gregg_word_1976}, learned later~\citep{dewhurst_separate_1998}, or made up of more letters~\citep{nickels_dissociating_2004}, both globally and in comparison to the sentence they appear in.
+Further characterizing the substitutions by examining the variation of word features from disappearing to appearing words, we show:
+(1) that the operation is contractile on average, that is words are brought closer to an attractor point on each feature;
+(2) that authors produce words that are easier to remember than the average of synonyms of the disappearing word (a fact that is reflected in the position of the attraction point).
+
+We do not actually observe quotes converging on a global scale towards attractors in their various dimensions.
+Indeed the limits of the data set don't allow us to infer chains of substitutions, and substitutions themselves are not the only type of transformation at work in the data set.
+Nonetheless, these findings (1) bring light to this simple type of transformation, and (2) are consistent with known psycholinguistic effects, with the hypothesis of cultural attractors in representations from everyday life, and with the lineage specificity discussed in the iterated learning literature~\citep{claidiere_cultural_2014,cornish_systems_2013}.
+They are obtained by successfully applying knowledge from cognitive science to real-life complex data, a task that remains a challenge in the study of cultural evolution.
+We believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.
+
+In the simple case presented here, however, much remains to be explored.
+Since it is clear that observing cognitive biases in such data is now possible, questions addressed in controlled laboratory situations could be opened by further research.
+One question concerns the influence of the context surrounding a quote, be it in terms of other quotes preceding it temporally or of text surrounding it in a post.
+A first step could be the application of results from \citet{zaromb_temporal_2006} who have shown, in the simpler task of recall of random word lists, that the source of prior-list intrusions can be predicted based on the associations those preceding lists have formed:
+in our case, a substitution could be triggered and directed by associations formed by preceding context.
+A further step would be to follow \citet{cornish_systems_2013} have shown about reciprocal influences between context and transformations (in their case, with transmission chains of artificial content).
+Indeed substitutions, and more generally all transformations, also participate in creating the context for later quotes.
+One can ask, therefore, what are the reciprocal effects between, on one side, the corpus-level evolution of quotes through iterated transformations, and on the other side, a gradual change in the properties of transformations operated because of the evolution of surrounding context.
+Such interactions have been shown to underlie the lineage specificity observed in transmission chains~\citep{claidiere_cultural_2014}.
+Exploring how similar loop interactions happen real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
+In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge.
+
+\end{new}
+
+% A few more unredacted ideas:
+% More broadly with Sperber:
+%  - you can have a simplistic interpretation and expect simple convergence, or you can look for attraction on different dimensions, which is a richer (and more useful) interpretation.
+%  - note also that it can apply to long term or short term evolution.
+%  - What we see is not clear w.r.t.
+convergence: (1) we don't observe it on our data, and (2) if there's lineage specificity there's no reason convergence should appear (on contrary).
+But, this is only half the story, since in reality there's also selection, i.e.
+not all chains go on forever, so you lose some diversity, and it changes the dynamics.
+%  - finally, context is addressed with relevance in Sperber, but it's still all representational, as we did here. Others have said that it will never be enough. Ingold, Di Paolo.
+
+\section{Concluding remarks}\label{sec:conclusion}
+
+\newtext{The theory of Epidemiology of Representations proposes a unifying framework for the study of cultural evolution.
+One of its core claims, the existence of cultural attractors, has been both a challenge to test empirically and a fruitful line to pursue in the study of cultural evolution.}
+We aimed to contribute to \newtext{testing this hypothesis} by studying a simple everyday-life task where individuals are implicitly trying to reproduce quotations.
+To some extent, our work amounts to \newtext{an out-of-laboratory} experiment where we \newtext{examine the influence} of \newtext{well-known word features} on the accuracy of the reproduction of short sentences.
+\newtext{Our analysis of substitutions shows that words are attracted, in each dimension, to a feature-specific value.
+Furthermore, the features' known effects in psycholinguistic experiments are reflected in the biases of these attraction points, meaning that the evolution of such quotations can be partially explained by known low-level cognitive biases.
+We believe that such an approach, combining psycholinguistic knowledge and data mining tools, can be fruitfully developed to improve the study of cultural attractors and explore the reciprocal influences of cognition and culture.}
 
@@
 
-ADDTHIS:
-We also chose exploratory vs. predictive to give a detailed view of what happens and because there's too many possible things to predict.
-
-ADDTHIS:
-By characterizing substitutions with 6 features on the disappearing and appearing words, we identify what makes a substitution more likely, and how a word changes when it is substituted.
-Consistent with known effects in linguistics, we observe that low-frequency words and words learned later in development are more susceptible to substitution than other words.
-Looking at the context those words appear in, we observe a marked effect for substitution of extreme words in a sentence (either very high-valued or very low-valued features compared to sentence average, except for word frequency).
-Focusing on how words are transformed, we see that the appearing words have significantly higher frequency and lower age-of-acquisition than synonyms of the disappearing word.
-Finally, the patterns we observe are also consistent with an attraction of each of the features towards a (feature-specific) asymptotic value.
-
-ADDTHIS:
-It is possible however, that these attractors appear due to an interaction between biases and sentence context, making it a contingency rather than a rule.
-This is not really dealt with (context, aside from relevance) by Sperber.
+Let us conclude by noting that the question of short- and long-term cultural evolution, and the approaches to study them, are becoming increasingly relevant to other fields.
+In biology in particular, work on evo-devo and non-genetic inheritance has accumulated evidence that is poorly accounted for by the modern synthesis of biological evolution, and is creating a demand for new or extended  approaches to joint cultural and biological evolution~\citep[see][for instance]{gilbert_eco-evo-devo:_2015}.
+Such an approach has long been called upon by anthropologists like~\citet{ingold_beyond_2004,ingold_three_1999}, in line with Mauss' initial works~\citep{mauss_les_1936}, and the debate is not entirely foreign to the enactive-representational discussion in cognitive science.
+As regards our particular case, a most promising line of work on the evolution of language is being developed by \citet{cuffari_participatory_2014}.
+The study of cultural evolution will most likely benefit greatly from the growing interactions between these disciplines.
 
 \end{new}
-
-\section{Concluding remarks}\label{sec:conclusion}
-
-\todo{\#14:
-link to introduction discussion:
-(1) this can be a model system,
-(2) convergence can be looked for in any dimension, but that doesn't make a theory, so Epidemiology of Representations is all nice, but:
-(3a) taken simplistically it predicts obviously simplistic stuff (all quotes CV. to a single quote),
-(3b) taken more realistically (many dimensions in life) it gives some ideas, but it's not clear it's a core principle
-(3c) we need more controlled investigation to fuel the discussion and see how relevant it is.
-}
-
-\begin{new}
-
-ADDTHIS:
-On the other side an enactive proposition which anthropologists like Ingold, in line with Mauss' works, are calling for \CN, is being developed by Froese, Di Paolo, and De Jaegher among others \CNs.
-
-The question is also gaining relevance in other fields, as work in evo-devo and non-genetic inheritance is accumulating evidence not accounted for by the modern synthesis \CN;
-these discoveries are creating demand for new or extended approaches to life evolution that unify its different levels, as well as creative empirical methods to test the predictions these approaches make \CN.
-
-\end{new}
-
-We aimed to contribute to the empirical understanding of representation transformation processes %\marginnote{fixed after Telmo saying: Is this really about knowledge transformation of more about representation transformations and the insight they provide over underlaying cognitive processes?}
- by studying a simple task where individuals are \emph{implicitly} trying to reproduce textual content.
-To some extent, our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
-In more detail, we describe the joint properties of the substituted and substituting terms in the reformulation by individuals of a specific type of utterances (quotations).
-%--- in this sense we also diverge from psycholinguistic experiments that focus on ease of recall.
-
-\todo{\#18: tone down claims about contractile: it's a possible hypothesis if this were the only process, but not observed with the mix of all other processes}
-
-For each of the selected psycholinguistic variables, we demonstrate the existence of attractor values in the underlying variable spaces.
-More precisely, beyond the interpretation of our results for each variable, we notice that all variables remarkably exhibit a single attractor and are generally contractile --- as such, even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are susceptible to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
-
-%Rather, we exhibit the bias of substitution, and that in this respect it not only provides a  we provide a fine description of the bias but also corresponds to an  ``input-output'' reformulation couple describing the joint properties of (substituted$\rightarrow$substituting) terms.
-
-%We contend that our results provide some of the first bricks of an empirical \emph{fitness landscape} for the epidemiology of representations
-
-%(age of acquisition, number of phonemes, ), it also emphasizes the importance of the semantic network structure (Wordnet: Pagerank, degree, clustering[, distance?]),
-
-%\begin{itemize}
-%\item new in the sense that it does not focus
-%From a psycholingustic viewpoint,

### 2016-07-29 brains-introduction.tex
fix/todos: some minor rephrasings

@@
 
-\todo{add examples}
-
 To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
 These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
-authors spontaneously transform quotes, not only cropping them but also replacing words, when in fact they are implicitly required to copy them exactly.
-We can therefore assume that most transformations, especially the simple ones, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.
+authors spontaneously transform quotes, not only cropping them but also replacing words.
+For instance the quote ``we will not be scared of these cowards'' (a substring of a quote from former Pakistani President Asif Ali Zardari) is also found as ``we will not be \textbf{afraid} of these cowards''.
+More meaningful changes often happen too, such as the transformation of McCain's ``I admire Senator Obama and his accomplishments'' during the 2008 U.S.
+presidential campaign, into ``I \textbf{respect} Senator Obama and his accomplishments''.
+Since authors are implicitly required to copy quotes exactly, we can assume that most transformations, especially simple ones such as those shown above, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.
 

### 2016-07-29 brains-protocol.tex
chore/all: typos, colors, todos
feat/discussion-conclusion: finish all the way to conclusion
fix/todos: some minor rephrasings

@@
 If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
-As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}
-\todo{this last sentence is unclear or not useful}
+Aiming to exhibit such subtle biases in complex data is the main reason we chose to use aggregate measures that have a chance of smoothing out the possible variations of experimental conditions in the data set.}
 
@@
 % utiliser "word-level measures", et/ou "For a given value of a given type of word-level measure (say, words of length 5), we define susceptibility as the ratio (...)"
-\newtext{For a given group of words $g$ \todo{[not clear this is not a given word -- maybe present by example instead of generally]}, susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
+\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (e.g.
+words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
 That is:}\footnote{
 \newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
-If, for instance, half the content words of a given quote are from group $g$, such a quote contributes .5 to the total $s_g^0$.}
+If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.}
 }
@@
 \newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (e.g.
-for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
+for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.\footnote{
+\newtext{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
+That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
+}
 
@@
     \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
-    \caption{\textbf{Substitution susceptibility for feature values:} susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
+    \caption{\textbf{Substitution susceptibility for feature values:}
+    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
     \todo{Use log-y}
@@
 \end{figure*}
-
-\todo{Relate to missed literature (\#15)}
 
@@
 Considering a word~$\wstart$ substituted for~$\warrival$, we measure how a feature~\newtext{$\phi$} of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
-Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
-\todo{explain this for susceptibility too}}
+Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{
+\newtext{Similarly to what we do for susceptibility, we avoid possible autocorrelation effects by averaging start- and arrival-word features of substitutions from the same cluster into a single aggregate substitution per cluster.}
+}
 $$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$
@@
 
-\todo{Relate to missed literature}
-
 Results are gathered in Fig.~\ref{fig:feature-variations-global}.
@@
 In other words, beyond individual variation patterns, the substitution process exhibits a unique attractor \newtext{for each feature}.
+\newtext{Note that this is also true under $\mathcal{H}_0$ or $\mathcal{H}_{00}$ (both null hypothesis curves have single intersections with $y=x$ with slopes smaller than 1):
+the substitution process naturally leads to an attraction even under reasonable random conditions.}
 

### 2016-07-29 brains-protocol.tex
fix/protocol: we won't use Supplementary Information for now

@@
 out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
-Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.
-}}
+Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.}
+}
 
@@
 \newtext{Several of these features are strongly related and can be grouped together.
-To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).\footnote{
-\newtext{The distributions of original and log-transformed features appears in the Supplementary Information, along with scatter plots for each feature pair, summarized below with correlation values.
-}}}
+To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).}
+%\footnote{
+%\newtext{The distributions of original and log-transformed features appears in the Supplementary Information, %along with scatter plots for each feature pair, summarized below with correlation values.
+%}}
 

### 2016-07-29 brains-protocol.tex
fix/todos: fix small cosmetic changes, comment out an appendix-related sentence

@@
 Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
-The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in the appendix).\footnote{
+The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.% (the analysis on the complete set of features can be found in the appendix).
+\footnote{
 Note that feature values stem from different data sets which do not always encode the same words.
@@
     Confidence intervals are computed with the Goodman method for multinomial proportions.}
-    \todo{Split into 2 subfigures; fix labels; use log-y for second pane; add H0 dashed line.}}
+    }
     \label{fig:pos-susceptibilities}
@@
     susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
-    \todo{Use log-y}
     }
@@
     However, attractors are always positioned between sentence median ($y=0$) on one side and $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ on the other side.
-    Clustering coefficient, Number of synonyms and Orthographic neighborhood density are limit cases, with triple intersections with one of the null hypothesis curves.
-    }
-    \todo{fix labels}
+    Clustering coefficient, Number of synonyms and Orthographic neighborhood density are limit cases, with triple intersections with one of the null hypothesis curves.}
     }

### 2016-07-29 brains-introduction.tex
feat/start: rewrite abstract

@@
 
-\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
-
 \begin{new}
 
-TOO GENERAL OR GIVES THE IMPRESSION OF TOO HIGH AMBITIONS. COMPARE TO OTHER INTROS TO JUST SAY WE'LL TALK ABOUT THIS.
-
 Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
-%the reciprocal influences of cognition and culture in meaningful ways.
-While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.
+The question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), but today's discussion is mostly structured by proponents from cognitive science.
 

### 2016-07-29 brainscopycut.tex
feat/substitution-models: finish first rewrite of substitution models
feat/results: finish variation subsection
feat/results: finish presentation of results
feat/start: rewrite abstract

@@
 \usepackage{amsmath}
+\usepackage{bm}
 \usepackage{amsfonts}
@@
 \usepackage{graphicx}
-\usepackage{subfig}
+%\usepackage{subfig}
+%\usepackage{subfigure}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 191, D-10117 Berlin}
-\abstract{We describe reformulation processes within a large distributed system such as blogspace;
-showing how some specific features of public representations may be altered by bloggers when they freely reproduce them.
-To deal with robust and simple cultural representations, we focus on the evolution of quotations.
-In particular, we uncover some of the semantic and structural characteristics of individual words and the substitutions they undergo.
-Our work amounts to a large \emph{in vivo} experiment where we appraise the impact of classically-influent psycholinguistic variables in the accuracy of the reproduction.
-We show that all variables remarkably exhibit a single attractor and are generally contractile.
-Even though the observed convergence patterns only partially explain quotation evolution, we shed light on a class of phenomena which are prone to constitute a key element of a broader empirically-grounded, attractor-based theory of cultural evolution.
-\todo{Update after introduction rework.}}
+\abstract{
+We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations online.
+Such attractors have been suggested as a useful tool for the study of cultural evolution and its reciprocal influences with cognition, but their existence has been hard to test.
+Our study focuses on how quotations are transformed when they are copied from blog to blog or media website:
+by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
+In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
+Our results support the hypothesis that cultural attractors can arise from the combination of individual cognitive biases in interpretation and production of representations.
+}
 \authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
-\keywords{word production; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; \emph{in vivo} psycholinguistics}
+\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; \emph{in vivo} psycholinguistics}
 %\note{\TB{Note XXXX (version? date?)}}
@@
 %\clearpage
-%\input{brains-garbage}
+\appendix
+\input{brains-appendix}
 

### 2016-07-29 brainscopycut.tex
style/abstract: missing colours from previous commit

@@
 \abstract{
-We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations online.
+\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations online.
 Such attractors have been suggested as a useful tool for the study of cultural evolution and its reciprocal influences with cognition, but their existence has been hard to test.
@@
 In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
-Our results support the hypothesis that cultural attractors can arise from the combination of individual cognitive biases in interpretation and production of representations.
+Our results support the hypothesis that cultural attractors can arise from the combination of individual cognitive biases in interpretation and production of representations.}
 }

### 2016-07-29 brainscopycut.tex
cite software colophon

@@
 
+\begin{new}
+
+\subsection*{Software colophon}
+
+Finally, this paper was developed using Python's scientific computing ecosystem~\citep{millman_python_2011}.
+In particular, we directly used NumPy and SciPy~\citep{walt_numpy_2011}, Matplotlib~\citep{hunter_matplotlib:_2007}, Pandas~\citep{mckinney_data_2010}, scikit-learn~\citep{pedregosa_scikit-learn:_2011}, NetworkX~\citep{hagberg_exploring_2008}, NLTK~\citep{bird_nltk_2009}, IPython~\citep{perez_ipython:_2007}, and many other libraries from the Python ecosystem.
+The software and analyses written for the paper are documented and published under a Free Software license.
+They can be found at \url{https://github.com/wehlutyk/brainscopypaste}.
+
+\end{new}
+
 \bibliographystyle{apacite}

### 2016-07-29 brains-introduction.tex
remove done todos, move figures around
style: fix quotation marks, i.e.'s and e.g.'s

@@
 The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
-these are alternatively located in the mind ("mental representations" in Sperber's terminology), or in the outer world ("public representations") as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
+these are alternatively located in the mind (``mental representations'' in Sperber's terminology), or in the outer world (``public representations'') as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
 A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
@@
 The theory makes an additional strong hypothesis, which this paper focuses on:
-as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called "cultural representations", because they characterize a given culture).
-This process should manifest itself as attractors (called "cultural attractors") in the dynamical system that models cultural evolution, that is:
+as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called ``cultural representations'', because they characterize a given culture).
+This process should manifest itself as attractors (called ``cultural attractors'') in the dynamical system that models cultural evolution, that is:
 there should be areas of the representation space where cognitive effects in transformations bring representations closer to a given stable asymptotic point.\footnote{
@@
 \label{sec:related}
-
-\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
 
@@
 
-On the whole, research on lexical features hints towards two antagonistic types of effects \newtext{(also known as the "word-frequency paradox", \citealp{mandler_word-frequency_1982})}.
-On one hand, part of the literature shows that recall is easier for the least "awkward" words;
+On the whole, research on lexical features hints towards two antagonistic types of effects \newtext{(also known as the ``word-frequency paradox'', \citealp{mandler_word-frequency_1982})}.
+On one hand, part of the literature shows that recall is easier for the least ``awkward'' words;
 those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in retrieval, that is in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
-On the other hand, when the task consists in recognizing a specific item in a list, "awkward" words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
+On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
 The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to rewrite a given utterance, is rather of the former or latter sort.

### 2016-07-29 brains-protocol.tex
remove done todos, move figures around
style: fix quotation marks, i.e.'s and e.g.'s

@@
 \label{sec:protocol}
-
-\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}
 
@@
 To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
-\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see "Substitution model").}
+\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see ``Substitution model'').}
 To quantify those substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
@@
 \begin{quote}
-The Bank of England said, "these operations are designed to address funding pressures over quarter-end."
+The Bank of England said, ``these operations are designed to address funding pressures over quarter-end.''
 \end{quote}
@@
 \begin{quote}
-"these operations are \textbf{intended} to address funding pressures over quarter-end."
+``these operations are \textbf{intended} to address funding pressures over quarter-end.''
 \end{quote}
@@
 }
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is "the city is tired of me and the organization and I have run our course together"), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
@@
 \newtext{To compute those features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
-As \citet{nelson_university_2004} explain, "free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect."
+As \citet{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
 \newtext{Similar to what \citet{griffiths_google_2007} did, we first considered the directed \newtext{weighted} network formed by association norms, where nodes are words and edges are directed from cue to target word, with a weight equal to the association strength (that is the probability of that target word being produced when this particular cue is presented).
@@
 Let us give an example:
-\newtext{say the quotation "These accusations are false and \textbf{absurd}" ($q$) appears in two different blogs on January 19, and the slightly different quotation "These accusations are false and \textbf{incoherent}" ($q'$) appears in another blog on the 20th of January.
+\newtext{say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in two different blogs on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in another blog on the 20th of January.
 The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
 And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$ as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
-Now say a third version, "These \textbf{allegations} are false and \textbf{incoherent}" ($q''$) also appears once on January 19 and once on January 20 after $q'$.
+Now say a third version, ``These \textbf{allegations} are false and \textbf{incoherent}'' ($q''$) also appears once on January 19 and once on January 20 after $q'$.
 $q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (this below for further details).
@@
 In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread into 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
-From these substitutions we remove those featuring stop words, minor spelling changes (e.g.
-center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
+From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
 this eventually yields \newtext{6,318} valid substitutions \newtext{(before merging substitutions that share the same destination)}.\footnote{
@@
 
-We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
-% utiliser "word-level measures", et/ou "For a given value of a given type of word-level measure (say, words of length 5), we define susceptibility as the ratio (...)"
-\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (e.g.
-words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
-That is:}\footnote{
-\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
-If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.}
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
+    \caption{\textbf{Substitution susceptibility for feature values:}
+    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
     }
-$$\sigma_g = \frac{s_g}{s_g^0}$$
-
-In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
-\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (e.g.
-for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.\footnote{
-\newtext{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
-That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
-}
+    \label{fig:feature-susceptibilities-global}
+\end{figure*}
 
@@
 \end{figure}
+
+We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
+\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
+That is:}\footnote{
+\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
+If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.}
+}
+$$\sigma_g = \frac{s_g}{s_g^0}$$
+
+In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
+\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.\footnote{
+\newtext{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
+That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
+}
 
@@
 \end{new}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
-    \caption{\textbf{Substitution susceptibility for feature values:}
-    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
-    }
-    \label{fig:feature-susceptibilities-global}
-\end{figure*}
 

### 2016-07-30 brains-introduction.tex
beginning of introduction - end of conclusion

@@
 
-Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
+\tb{The reciprocal influence between cognition and culture has a long history in both social science and psychology.
+While this question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.}
+%Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
 The question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), but today's discussion is mostly structured by proponents from cognitive science.
-
-These construe culture as an evolutionary process analogous and parallel to biological evolution (and especially the modern synthesis' account of it).
-That analogy can be traced a long way back in the 20th century and earlier, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\citeyearpar{dawkins_selfish_2006}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
+These construe culture as an evolutionary process analogous and parallel to biological evolution \tbn{(and especially the modern synthesis' account of it)}.
+That analogy can \tb{also} be traced a long way back \tbn{in the 20th century and earlier}, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\tb{\citeyearpar{dawkins_selfish_1976}}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
 More recently, Dan Sperber has drawn on this principle to explicitly connect anthropology and cognitive science through the theory of \emph{Epidemiology of Representations} \citep{sperber_explaining_1996}, and the study of cultural evolution has been growing steadily since.

### 2016-07-30 brains-introduction.tex
passed over introduction, related work, methods

@@
 
-\tb{The reciprocal influence between cognition and culture has a long history in both social science and psychology.
-While this question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.}
+\tb{The reciprocal influence between cognition and culture has a long history in} both social science and psychology.
+While this \tb{question} has been the subject of intense debate in the social sciences in the 20th century\tbn{ (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980})}, today's discussion is mostly structured by proponents from cognitive science.
 %Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
@@
 
-The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
-\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
-\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
+The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how \tbn{the theory of} memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
+\Citet{mesoudi_multiple_2008} \tbn{have} discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
+\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, \tb{demonstrate} how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
 \citeauthor{kirby_cumulative_2008}~(\citeyear{kirby_cumulative_2008,cornish_systems_2013}) have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
@@
 A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
-Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.
+Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the \tb{above cited} studies are doing.
 
@@
 there should be areas of the representation space where cognitive effects in transformations bring representations closer to a given stable asymptotic point.\footnote{
-\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
+\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or \tb{confined to}) the target sub-area.
 }}
@@
 Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
-For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in the email network~\citep{liben-nowell_tracing_2008};
-\citet{cointet_socio-semantic_2009} have described the reciprocal influence between the social network topology and the distribution of issues;
-\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} have studied the characteristics of particularly memorable quotes that circulate in those networks.
-We believe those works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.
+For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in \tb{emails}~\citep{liben-nowell_tracing_2008};
+\citet{cointet_socio-semantic_2009} \tbn{have }described the reciprocal influence between the social network topology and the distribution of issues;
+\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} \tbn{have }studied the characteristics of particularly memorable quotes that circulate in those networks.
+We believe \tb{these} works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.
 
@@
 
-To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
+To show this we analyze \tb{how} quotes in blogs and media outlets are modified when they are copied from website to website.
 These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
@@
 
-Our question is as follows: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases referred to in the theory of epidemiology of representations, do cultural attractors appear and how do cognitive biases participate in them?
-We chose to restrict our analysis to substitutions (one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
+\tb{We thus ask the following question}: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases \tb{evoked} in the theory of epidemiology of representations, %do cultural attractors appear and
+how do cognitive biases participate in \tb{the appearance of cultural attractors}?
+We chose to restrict our analysis to substitutions (i.e., one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
 \newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
@@
 While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
-This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, but with natural language out of the laboratory.
+This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, \tb{yet} with natural language out of the laboratory.
 
@@
 %Sec.~\ref{sec:results}
-Next, we describe the measures built to observe the cognitive biases operating in quote transmission.
+Next, we \tb{introduce} the measures \tb{we} built to observe \tbn{the} cognitive biases operating in quote transmission.
 Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.
-
 \end{new}
@@
 \newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
-The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
-\newtext{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
-}}
+The first \tb{above-mentioned} method consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{\tbn{Critics like }\tb{\citet{ingold_trouble_2007}}\tbn{, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.}}
 A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
@@
 \citet{potter_regeneration_1990} suggest that immediate recall of sentences is based on the retention of an unordered list of words which is then regenerated as a sentence at the moment of production.
-Priming recall with other words can lead to replacement in the recalled sentence if the primed words support the overall meaning of the sentence.
+Priming recall with other words can lead to replacement in the recalled sentence if the primed words \tb{are consistent with} the overall meaning of the sentence.
 Regenerated syntax can also be influenced by priming recall with another syntactic structure~\citep{potter_syntactic_1998}, or with verbs whose category constraints call for a different structure~\citep{lombardi_regeneration_1992}.

### 2016-07-30 brains-protocol.tex
passed over introduction, related work, methods

@@
 \newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
-\newtext{Indeed} quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
+\newtext{\tbn{Indeed}}
+Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
 First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
@@
 \newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
-Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
+Indeed, the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
 %In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
@@
 
-To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting in the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting \tb{of} the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 \newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see ``Substitution model'').}
-To quantify those substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
+To quantify \tb{these} substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
 
@@
 \end{quote}
-\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
+\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, \tb{i.e.} sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.
@@
 }
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our approach fulfills its goals by coding a random sub-sample of 100 clusters:
-35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our \tb{filtering} approach fulfills its goals by coding a random sub-sample of 100 clusters: 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
 Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.}\footnote{
@@
 
-\newtext{To compute those features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
+\newtext{To compute \tb{these} features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
 As \citet{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
@@
 The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
-And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$ as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
+And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$, as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
 Now say a third version, ``These \textbf{allegations} are false and \textbf{incoherent}'' ($q''$) also appears once on January 19 and once on January 20 after $q'$.
-$q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (this below for further details).
+\tb{Here,} $q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (\tb{see} below for further details).
 $q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
@@
 (1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
-(2) \textbf{bin span}, which can be 24 hours (as in the model above) or can be extended to start at the very first occurrence in the quotation family.
-The other two parameters define rules on the selection of source and destination quotes of a substitution:
+(2) \textbf{bin span}, which can be \tb{$\Delta t=1$ day} %24 hours 
+(as in the model above) or can be extended to start at the very first occurrence in the quotation family.
+The other two parameters \tb{configure} rules on the selection of source and destination quotes of a substitution:
 (3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
@@
 Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
-Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use discretely positioned bins spanning 1 day:
+Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use \tb{discrete} %ly positioned 
+bins spanning 1 day:
 later occurrences of $q'$ and $q''$ are counted as substitutions in Fig.~\ref{fig:substitution-modelled-majority-all} and Fig.~\ref{fig:substitution-modelled-all-all}, whereas in Fig.~\ref{fig:substitution-modelled-majority-exclude_past} and Fig.~\ref{fig:substitution-modelled-all-exclude_past} they are not.
@@
 
-In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread into 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
+In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread \tb{over} 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
 From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
@@
 Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
-out of the first 100 substitutions coded for recall, those with levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}
+out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}
 }

### 2016-07-31 reviews/2016-07-response-to-reviewers.tex
initial LaTeX file for reviewer response / cover letter

\documentclass[a4paper,10pt]{article}
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{utopia}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage[authoryear,round,compress]{natbib}
\usepackage{multicol}
\usepackage[usenames,dvipsnames]{color}
\usepackage{url}
\usepackage{eurosym}
\usepackage{array}
\usepackage{tabularx}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{wrapfig}

\usepackage{float}
\usepackage{enumerate}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\usepackage[frenchb]{babel}
\usepackage{setspace}
\setstretch{1.05}

\newcommand{\itx}[1]{\begin{itemize}#1\end{itemize}}
\newcommand{\enx}[1]{\begin{enumerate}#1\end{enumerate}}
\newcommand{\x}{\item}

\usepackage[bookmarks, colorlinks, breaklinks, pdftitle={Research project},pdfauthor={Camille Roth}]{hyperref}  
\hypersetup{linkcolor= MidnightBlue,citecolor= MidnightBlue,filecolor=black,urlcolor= MidnightBlue}

\title{Manuscript CS15-80\medskip\\
{\Large\em``The semantic drift of quotations in blogspace: a case study in short-term cultural evolution''}
\bigskip\\Response to Reviewers
}
\author{Sébastien Lerique and Camille Roth}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

\makeatletter
\renewcommand{\subsection}{\clearpage\@startsection{section}{1}{0mm}
{\baselineskip}{\baselineskip\medskip\hrule\medskip}{\raggedright\Large\bf~\textcolor{MidnightBlue}}}
\makeatother

\newcommand{\critique}[1]{\begin{quote}#1\end{quote}}
\newcommand{\answer}[1]{{\setlength{\parindent}{0pt}\par\color{MidnightBlue} #1}}

\vspace{3cm}
Dear Todd Gureckis,

\medskip
We enclose an extensively revised version of the manuscript CS15-80 (initially titled ``How do we copy and paste? The semantic drift of quotations in blogspace'') based on the report you sent us on Jul 18, 2015.

\medskip
We hope that we have appropriately addressed the detailed remarks and very constructive questions made by you and the referees: to this effect, we also provide below a detailed response to each of these comments.  We additionally used a different color to emphasize the portions which were thoroughly rewritten.

\medskip
This comprehensive revision work led to a significantly longer manuscript than was originally submitted --- from around 6,300 words to about 9,000 words now, making it closer to a ``regular article'' than a ``brief report'': we hope that this will not be problematic at this stage.

\medskip
We look forward to reading your feedback on this new manuscript.

\medskip
Best wishes,

\medskip
Sébastien Lerique and Camille Roth

\subsection*{Editor}
\critique{
1. The main results of the paper are based on a lot of arbitrary decisions about how various quantities are defined.  I am interested if small differences in these choices radically changes the conclusions you would draw, and how many other approaches you tried before you arrived at the data patterns you present.  All the reviewers agreed on this point (asking for more clarification about feature selection, why certain choices were made in the analysis, etc…).  This ranges from what types of quotes were included, to the rules to detecting substitution errors, to the inclusion of certain features in your analyses.  The reader needs to be convinced that the few “strong” looking patterns you do find are not too dependent on these arbitrary choices.  I actually question if this is the best way to address this question (effect of memory and linguistic variables on copy errors) given the noisiness of the data and the difficult in attributing copying errors to any particular parent
source.

}

\answer{
Point to:
- work on the substitution models showing that the results are robust
- the rationalization of feature selection
- clarifications of choices, such a single-substitution restriction
- precision/recall analyses for all filters
}

\critique{
2. The conclusions about “contractile” processes seem pretty speculative (as almost all reviewers agreed).   Sure, this might be the theoretical implication of iterating this substitution process, but you do not present any further evidence of this type of contraction towards a fixed point.  I think if you want to make this a major conclusion of the paper there is more work to be done.  Also this idea has been explored, in my view, in a much more sophisticated way in the iterated learning literature (e.g., Kalish, Lewandowsky, Griffiths, Kirby… see my PDF comments for specific references).  You should really check these arguments and papers out because you really have an iterated learning task.
}

\answer{We agree the presentation was too much. Point to:
- the integration of the iterated learning literature in review and general discussion
- the fact that the constraints of the data don't let us explore chains, so we explained our claims more appropriately
}

\critique{3. The writing of this paper is of very low quality compared to what is expected at the journal.  I will email separately a very detailed editing of the paper via Adobe Acrobat where I tried to point these problems out.  However, it simply couldn’t be published without a nearly complete re-write or very heavy editing.
}

\answer{Thank you. We paid a lot of attention to the writing style in this new version, as can be seen in the large rewrites of many paragraphs.}

%\critique{
%The bottom line is that while I think this is a really interesting idea/question, this is not the best execution of this idea that I could imagine being published in the journal.  My gut reaction is that this paper needs a lot of work, additional analysis, and rewriting before it could be published and therefore should be rejected.  However, because the reviewers are unanimously more positive in their evaluation I will invited a revision for this in the hope that you can make it a pretty significant change and can benefit from the excellent and detailed reviewer comments.
%
%So you understand what is requested:  typically papers which are revised at Cognitive Science are returned with a detailed cover letter responding to each of the reviewer comments in detail, along with pointers to what has changed in the paper.  I would appreciate this type of cover letter if you choose to resubmit because it can help me and the reviewers process what changes you made.
%
%Sincerely,
%Todd Gureckis
%Associate editor
%}
%
%\answer{
%}

\subsection*{Reviewer \#1}

\critique{I think this is an interesting approach and dataset for better understanding how sentences are recalled and how paraphrases are constructed.  However, I think that the experiment design needs to be refined, and the project as a whole needs to be better connected to larger theoretical problems and better grounded within cognitive science.
}

\answer{Point to:
- the revamped introduction, review of related work, and discussion of results which connects our approach to the literature
}

\critique{
Starting with the grounding, the Introduction talks about theories of "knowledge transmission mechanisms" but doesn't provide examples aside from the quote-copying problem. The theories also don't re-appear in the results discussion, leaving the experimental results stranded and unconnected to any larger domain.  No discussion of sentence recall tasks within psycholinguistics is given - there isn't a huge literature on this, but there are at least a few papers that should inform this work, such as Potter \& Lombardi 1990 "Regeneration in the short-term recall of sentences". No discussion is made about whether and how much the substitutions change the quotes (e.g., is the substituted word a synonym, phonological-neighbor, or something else? - cf. Lauf et al 2013 "Analyzing Variation Patterns in Quotes Over Time"). Lastly, the meaning of the analysis (pg. 17-19) of this process as a dynamical system with "cultural attractors" is not fleshed out - why doesn't this predict that all quotes will eventually, through substitution, converge to the same string? (Does it?) As a result, the conclusion that psycholinguistic factors don't have a big impact on substitution likelihood is interesting but doesn't have clear theoretical repercussions, or any clear meaning outside the very specific problem of quote replication.  This needs to be substantially revised and expanded to make this work useful to the audience at large.
}

\answer{Point to:
- the new discussion of results
- the improved review of literature
- the use of H00 to compare to synonyms, and the small exploration of distances travelled by substitutions on the FA and synonym networks
}

\critique{ The experiment design also has some substantial flaws. First, some crucial factors were not considered - most notably, part of speech. I suspect, for instance, that pronouns and other closed-class words less often substituted than open-class words.  Phonological neighborhood density could also be important, though probably less so.  Second, no analysis was done that included all of the factors in, e.g., a multi-feature regression model.  While the features being considered are not strongly correlated, they are still likely to interact - especially if part-of-speech is added.  Third, the restriction to single-word substitutions introduces some significant confounds.  I'm not convinced that correctly and incorrectly copying a quote should be viewed as separate behaviors.  But more importantly, the $\sigma_\phi(f)$ calculation can be misleading under the single-substitution restriction:
}

\answer{
Point to:
- susceptibility of POS, showing that the effect (closed/open class words being different) expected is very small at best (closed class words would be stopword-likes in our graph)
- added phonological neighbourhood density as a feature
- the multi-feature regression of variation upon substitution, showing that variation of a feature is predicted mainly by the same source feature
- the expansion of substitution models to two-substitutions, giving the same results as single-substitutions. Notebooks for all models can be found at https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks , and the graphs produced by those notebooks can be found at https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb .
}

\critique{Consider a toy case where all quotes are two words, and there is a single binary feature $F$. $p(F=1) = 3/4$, and the true substitution probabilities, given the value of F, are $p(S|F=1)=1$, $p(S|F=0)=1/2$.  All the quotes where both words have feature $F=1$ will have two substitutions and be omitted from the dataset, as will half of the quotes with one $F=1$ \& one $F=0$ word.  Half of the quotes with two $F=0$ words will also be omitted for having no substitutions.  Due to these omissions, if you actually generate some test data and calculate the $\sigma_F$ values, the $\sigma_F(1)$ value is correctly estimated as 1, but the $\sigma_F(0)$ value is erroneously estimated as 1/8 (not 1/2), because the $F=0$ words are never substituted in the $F=1/F=0$ pairs, and these are three times more common than the $F=0/F=0$ pairs.  I haven't run more complicated cases, such where the length of the quotes vary, but this leads me to have limited confidence that the estimates in the Figure 4 are appropriate to draw conclusions from. I would want to see either a mathematical analysis with realistic values showing that the error is expected to be small in this data or an expansion to include multiple-substitution examples to avoid this problem.
}

\answer{This case is indeed valid, if there is such a feature (F=1 here) that:
(a) is extremly frequent,
(b) is extremely substituted,
(c) interacts strongly with another feature so that both are often substituted together.

The combination of (a) and (b) is an anti-pattern in our results (more frequent =less substituted), so we could in theory give a bound to the worst error that could have been committed.
However we extend our substitution models to two-substitutions, and observe that the results do not change.
}

\critique{ Some high-level suggestions:
\itx{
\x more discussion of what dynamical systems are and especially their interpretation in this task before their appearance in the results
\x needs a substantive discussion of the results}
}

\answer{Point to rewrite of introduction and the added discussion of results.}

\critique{\itx{
\x show that the phi value is capturing what it purports to capture
\x and that the factors aren't conflated with length
\x is the single-substitution biasing the data?
\x make sure that the reported measures are interpretable
\x if low-frequency words have 1/3 chance of being chosen as substitution (Fig. 4), what if there are four low-frequency words in a quote?
}
}

\answer{
Point to:
- our control of single-substitution bias
- the change in definition of phi. Indeed it was confusing to present it (and measure it) as a probability (which it was, but an aggregated global one, not an in-sentence one), we now measure it as the strength of a bias.
}

\critique{\itx{
\x improving statements of motivation and grounding within cognitive science
\x what are representation transformation processes?
\x give examples in the introduction
\x show how the quotation test case connects to more general representation transformation?}
}

\answer{
Point to improved introduction and added discussion. Especially the connection with iterated learning literature.
}

\critique{\itx{
\x what about the semantics of quote replication?
\x does part of speech influence substitution rate?
\x can you quantify the semantic distance between the original and substituted quote (e.g., get Mechanical Turkers to rate the semantic distance?)}
}

\answer{
Point to:
- the added POS analysis
- the short discussion of distances travelled upon substitution
}

\critique{\itx{
\x what happens when you run a multi-factor regression model? Do the features interact?
\x clarify from the beginning the predictions you are making so that the results will be interpretable 
}}

\answer{Point to the regression model we introduced.}

\critique{Some low-level points:\itx{
\x the in vitro/in vivo distinction is unfamiliar and ought to be defined in the text
}
}

\answer{We hope it is now clarified.}

\critique{\itx{
\x introducing eight measures and then removing three of them due to collinearity is presentationally confusing
\x was there a reason to expect that \# of synonyms and \# of phonemes would interact (Fig. 4)?]
}
}

\answer{Yes, because of the similar curve shapes they had.
We removed number of phonemes now (in favor of number of letters), and number of synonyms doesn't show any effect of susceptibility on the new, improved measure.
}

\critique{\itx{
\x how are you binning the continuous-valued features into categorical features for Fig 4 \& 5?
\x are there multiple y=x crossovers in Fig 5 if more bins are used?}
}

\answer{
Binning is now by quantiles (makes more sense for the confidence intervals), but the extended notebooks (https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks and https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb) also include graphs with the fixed-width bins we presented in the first version.
No multiple crossovers were ever observed.
}

\subsection*{Reviewer \#2}
\critique{This is a review of "How do we copy and paste? The semantic drift of quotations in blogspace." This paper attempts to discover the underlying psycholinguistic variables that is involved in blog authors making mistakes in the copying of quotations across time. Although I think the motivations and goals of this article are interesting and noteworthy, I think the actual analysis needs improvement in a number of areas.

Although I appreciate the brief writeup on intrusions in recall, the article fails to capture the amount of research done on this topic. Going back to Deese (1959; the precursor to modern experimental studies on false memory) and the subjective organization results of Tulving (1962) it has been well known that extra-list intrusions in free recall are often semantic in nature, with the intruding word being very similar in meaning to a studied item. A recent study that reinforces this is one by Zaromb, et al. (2006; JEP:LMC).
}

\answer{Thank you for these very useful pointers, we accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' section).}

\critique{ This leads into a larger issue: there is no direct word-word semantic similarity score included in the psycholinguistic variables in the analysis. Global measures such as clustering coefficients based on free association norms are included, but this isn't as appropriate, since as the authors find, the vast majority of words that are being substituted are low frequency words. Low frequency words are not as well represented in the Nelson norms, given that they are less likely to be produced, and are thus less likely to have an appropriate clustering coefficient from this dataset. Variables based on corpus analyses need to be included to provide an appropriate overview of the impacts of semantics on this task. Even so, I would suspect that global semantic network measures are not nearly as important as word-word metrics.
}

\answer{Point to:
- discussion of why we use word features (intro + related work), 
- why we don't do exact word prediction (related work showing the question is not even mastered for random word lists in the laboratory, so forget about sentences in out-of-laboratory data), 
- clustering coefficient is not so central to our results now
- the addition of H00 gives insight into what the new words are w.r.t. to synonyms of the disappearing word
}

\critique{I would guess that the substitutions that are made in quotations are overwhelmingly higher frequency, highly semantically similar words. As the example on pg. 19 states, the substitution was "world" for "globe," where a higher frequency word replaced a lower frequency, semantically related word. There are a number of different models that can be used to assess this type of similarity, with the classic model being latent semantic analysis (Landauer \& Dumais, 1997). There are a number of additional models that have been proposed, such as Topics (Griffiths, Steyvers, \& Tenenbaum, 2007, Psyc Review) or Beagle (Jones \& Mewhort, 2007, Psyc Review). These models differ in their implementational complexity, but there are tools available that allow for useful statistics to be acquired easily. A tool by Recchia \& Jones (2010, BRM) allows for pointwise mutual information (PMI) of different words to be retrieved from a given corpus very simply. A tool such as this would allow for word-word similarity to be included in the analysis.

I understand that word-word metrics are not easily introduced in the type of analysis that the authors are attempting to do, where the analysis of the contributions of different global psycholinguistic variables are being weighed against each other. I just think that type of analysis is inappropriate here. Since the quotations were selected on the basis of single substitutions, it seems to me that the appropriate level of analysis is what properties of the word that is being substituted is the most predictive of the substitution. I suspect that the combination of frequency and semantic similarity will be by far the most predictive variables, consistent with past results on recall.
}

\answer{
Point to:
- discussion of distance travelled by substitution on FA and synonyms networks
- our data is ill-shaped for the kind of topic modelling suggested (aside from the earlier point that even with experiments in the laboratory, such tools don't predict so well)
}

\critique{Beyond this, I think the authors are missing out on the most interesting aspect of the data that they have collected: how the surrounding context of the quotation used is changing across time. That is, the quote is being propagated, and the prior usage of that quote is likely predictive of how that information is spread. This is likely a dynamical process where multiple news sources are used to construct a new blog post, and being able to analyze this type of information propagation would be a very interesting study. I point to the semantic space models cited (and the work derived from them) as being very useful tools to analyze this type of question.
}

\answer{That would indeed be interesting, but not our goal in the paper.}

\subsection*{Reviewer \#3}
\critique{In this paper, the authors analyze a large data set of quotations taken from a variety of news sources to study the form of substitutions in quotes, to determine if the transformations follow expected patterns given psycholinguistic theory.  Specifically, the words that are likely to be substituted are found to be less frequent words, and the pattern of substitution is such that the substituting word is more frequent, acquired earlier, and has lower clustering in the free-association word network.  These patterns are consistent with expected cognitive biases.

There are a few issues I would want to see addressed.

Foremost is the method of detecting substitution.  I understand that, given the data set, there is no perfect way to identify what quote is a transformation of what quote, and that it is not feasible to test all possible methods AND that it is simpler to focus on one method for the purposes of the paper.  However, given that this decision of which method to use can directly influence the main conclusions, it behooves the authors to do some sort of robustness check around this method.  For example, one could imagine two distinct groups that only exactly copy from themselves and never copy from each other, where one always publishes earlier than the other.  If the earlier-publishing group tended to select variations of quotes that used lower-frequency words, and the later-publishing group tended to select variations of quotes that used higher-frequency words, this method would infer many more low-frequency to high-frequency substitutions than actually existed.  This may not be a likely scenario, but it demonstrates the benefit of a robustness check. Trying the same analysis with one or two different substitution models would strengthen the authors' claims.
}

\answer{Point to:
- the several substitution models implemented, tested, and checked}

\critique{ A second issue is the time spent on feature selection.  It is unclear why the authors do feature selection at all, as "the analysis is done on a per-feature basis."  It would again add to the robustness of their claims if they showed that all of the correlated features show the same pattern of susceptibility and bias in substitution.  In the same vein, it is not clear why the authors claim that age of acquisition, frequency, and number of synonyms have low levels of correlation "excluding the network properties".  Why would one ignore the correlations with the network features?
}

\answer{Point to:
- the improved substitution selection section
- the graphics for the other features, available in the notebooks (https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks) and figures (https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb), which we can put in a Supplementary Information if needed
}

\critique{
One smaller issue has to do with the filtering of the quotations.  The authors manually inspected the filtered quotes to see how many real event-related utterances are incorrectly filtered, but I would argue the precision of the filter is more important than the recall; how many of the kept quotes are not real event-related utterances?  Why not do manual coding of those and report that accuracy?
}

\answer{Yes thank you, this is done now.}

\critique{ Finally, while I understand the appeal of doing a dynamic systems analysis to find the convergence points of the substitutions, but it is not really valid here.  The analysis in this paper does not provide information on what happens with multiple substitutions, so the authors cannot really make claims about the convergence points given this methodology.
}

\answer{We better explained our claims without overstating our findings.}

\critique{Overall, the analysis of the patterns of substitution in a series of quotations to validate psycholinguistic theory is novel and advances the current science.  The results would be more solid with some additional robustness checks, but in general they are quite intuitive.  I hope in the future the authors find a dataset that provides certain information about the transmission of language from one person to another, as this would allow for stronger claims about causality and greater confidence in the pattern of substitutions, as well as a finer-grained analysis of repeated substitutions.
}

{\small
\bibliographystyle{../../authordate1}
\bibliography{../../mainbiblio-u8}
}

 \end{document}

### 2016-07-31 brains-introduction.tex
initial LaTeX file for reviewer response / cover letter

@@
 One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which words are recalled~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}.
-The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
-indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations of subjects (contributing to associative information retrieval) is at the center of many of these studies.
+The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006}\tb{.} Indeed\tb{,} the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations of subjects (contributing to associative information retrieval) is at the \tb{core} of many of these studies.
 

### 2016-07-31 brains-protocol.tex
initial LaTeX file for reviewer response / cover letter

@@
 
-\newtext{Our question concerns the low-level properties of these substitutions; to address it} we build the following two observables for each word feature.
+\newtext{Our question concerns the low-level properties of these substitutions: \tb{we ask (i) which words are targets of the substitutions and (ii) what change they are subjected to.
+To this end,}} we build the following two observables for each word feature.
 First, we measure \newtext{which word features are more or less substituted compared to how often they would be if the process were random, in order to capture} the susceptibility for words to be the target of a substitution in a quote.

### 2016-07-31 reviews/2016-07-response-to-reviewers.tex
intermediary point

@@
 
-\answer{Thank you for these very useful pointers, we accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' section).}
+\answer{Thank you for these very useful and, indeed, relevant pointers.
+We accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' section).}
 
@@
 There are a few issues I would want to see addressed.
+\bigskip
 
@@
 
+\answer{On a side note, we entirely agree with the relevance of pursuing this research program in these directions, as the \emph{in vitro} observation of information transmission and its evolution is part of an ongoing project (which nonetheless remains beyond the scope of the current paper).}
+
 {\small

### 2016-07-31 brains-protocol.tex
intermediary point

@@
 From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
-this eventually yields \newtext{6,318} valid substitutions \newtext{(before merging substitutions that share the same destination)}.\footnote{
+this eventually yields \newtext{6,318} valid substitutions\footnote{
 \newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
@@
 Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
-out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}
-}
+out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}} \newtext{\tbn{(before merging substitutions that share the same destination)}}.
 
@@
 
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
-    \caption{\textbf{Substitution susceptibility for feature values:}
-    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
-    }
-    \label{fig:feature-susceptibilities-global}
-\end{figure*}
-
 \begin{figure}[h]
@@
 
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
+    \caption{\textbf{Substitution susceptibility for feature values:}
+    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
+    }
+    \label{fig:feature-susceptibilities-global}
+\end{figure*}
+
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
@@
 \newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
-If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.}
+If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.} \tb{Further, t}o avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
+That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
 }
@@
 In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
-\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.\footnote{
-\newtext{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
-That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
-}
+\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions,
+\tb{\hbox{i.e.} it measures the preferential selection of some word properties given a substitution.}
+%\newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
 

### 2016-08-01 brains-conclusion.tex
manuscript revised

@@
 
-We aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
+We \tb{initially} aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
 The data set we used imposed a few constraints on our analysis:
@@
 They are obtained by successfully applying knowledge from cognitive science to real-life complex data, a task that remains a challenge in the study of cultural evolution.
-We believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.
+\tb{More broadly,} we believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.
 
@@
 Such interactions have been shown to underlie the lineage specificity observed in transmission chains~\citep{claidiere_cultural_2014}.
-Exploring how similar loop interactions happen real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
+Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
 In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge.
@@
 To some extent, our work amounts to \newtext{an out-of-laboratory} experiment where we \newtext{examine the influence} of \newtext{well-known word features} on the accuracy of the reproduction of short sentences.
-\newtext{Our analysis of substitutions shows that words are attracted, in each dimension, to a feature-specific value.
+\newtext{Our analysis of substitutions shows that words are attracted, in each dimension, to feature-specific values.
 Furthermore, the features' known effects in psycholinguistic experiments are reflected in the biases of these attraction points, meaning that the evolution of such quotations can be partially explained by known low-level cognitive biases.

### 2016-08-01 brains-protocol.tex
manuscript revised

@@
 
-While most of the substitutions we obtain with this procedure replace a word with another semantically related word, the two are rarely direct synonyms:
-only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
-\newtext{A similar behavior is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
-}}
-Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
-The following examples illustrate this behavior:
+\tb{Substitutions usually} replace a word with another semantically related word: the new word achieves a similar meaning in the context of its sentence, while still slightly changing the implications or the attitude expressed by the author.
+The following examples illustrate this \tb{phenomenon}:
 
@@
 \end{APAitemize}
+
+\tb{However, the substituted and subtituting words are rarely direct synonyms:} only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
+\newtext{A similar \tb{phenomenon} is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
+}}
+
+%Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
+%The following examples illustrate this behavior:
+
+%\begin{APAitemize}
+%	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
+%	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
+%	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
+%\end{APAitemize}
 
@@
     \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
-    \caption{\newtext{\textbf{Substitution susceptibility for POS categories:} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
-    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts, the bottom panel shows $\sigma_{POS}$, the ratio of the two.
+    \caption{\newtext{\textbf{\tb{POS-related results:}} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
+    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows \tb{the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts}.
     Confidence intervals are computed with the Goodman method for multinomial proportions.}
@@
 That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
-}
+
 $$\sigma_g = \frac{s_g}{s_g^0}$$
@@
 
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_global.png}
-    \caption{\textbf{Feature variation upon substitution:}
-    $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals based on Student's \emph{t}-distribution.
-    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias \newtext{compared to a purely random variation}.
-    \newtext{The position with respect to the dash-dotted line representing $\mathcal{H}_{00}$ ($\nu_{\phi}^{00}$) indicates the bias compared to a semantically plausible random variation obtained by choosing a random synonym of the disappearing word.}
-    The intersection with $y = x$ marks the attractor value.
-    The fact that all curves have slopes smaller than 1 \newtext{in absolute value} means that the substitution operation is contractile on average:
-    \newtext{it brings} each feature \newtext{closer to} its own specific asymptotic range.
-    }
-    \label{fig:feature-variations-global}
-\end{figure*}
-
 We \newtext{now examine} how words are modified \newtext{when} they are substituted, that is how their features change upon substitution.
@@
 $$\nu_{\phi}^0 (f) = \left<\phi\right>$$
-\newtext{$\mathcal{H}_{00}$ models the situation where the arrival word $\warrival$ is chosen \emph{among immediate synonyms of the start word} $\wstart$, that is an arrival word chosen among semantically plausible though still random words.
+\newtext{$\mathcal{H}_{00}$ models the situation where the arrival word $\warrival$ is chosen \emph{among immediate synonyms of the start word} $\wstart$, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
 In this case $\nu_{\phi}^{00}$ does depend on $f$:\footnote{
@@
 \end{seriate}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_global.png}
+    \caption{\textbf{Feature variation upon substitution:}
+    $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals based on Student's \emph{t}-distribution.
+    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias \newtext{compared to a purely random variation}.
+    \newtext{The position with respect to the dash-dotted line representing $\mathcal{H}_{00}$ ($\nu_{\phi}^{00}$) indicates the bias compared to a semantically plausible random variation obtained by choosing a random synonym of the disappearing word.}
+    The intersection with $y = x$ marks the attractor value.
+    The fact that all curves have slopes smaller than 1 \newtext{in absolute value} means that the substitution operation is contractile on average:
+    \newtext{it brings} each feature \newtext{closer to} its own specific asymptotic range.
+    }
+    \label{fig:feature-variations-global}
+\end{figure*}
 

### 2016-08-01 brainscopycut.tex
beginning of introduction - end of conclusion
manuscript revised

@@
 \newcommand{\tb}[1]{\textcolor{blue}{#1}}
+\newcommand{\tbn}[1]{\tb{\sout{#1}}}
 \newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
@@
 \abstract{
-\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations online.
-Such attractors have been suggested as a useful tool for the study of cultural evolution and its reciprocal influences with cognition, but their existence has been hard to test.
-Our study focuses on how quotations are transformed when they are copied from blog to blog or media website:
-by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
+\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations.
+Such attractors have been \tb{proposed} as \tb{a concept useful for} understanding cultural evolution %and its reciprocal influences 
+\tb{in relation} with \tb{individual} cognition, but their existence has been hard to test.
+\tb{We empirically} focus on \tb{the online transformation of} quotations when they are copied from blog to blog or media website: by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
 In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
-Our results support the hypothesis that cultural attractors can arise from the combination of individual cognitive biases in interpretation and production of representations.}
+Our results support the hypothesis that cultural attractors can \tb{result} from the combination of individual cognitive biases in the interpretation and reproduction of representations.}
 }

### 2016-08-06 reviews/2016-07-response-to-reviewers.tex
first comments
minor before leaving

@@
 \newcommand{\critique}[1]{\begin{quote}#1\end{quote}}
-\newcommand{\answer}[1]{{\setlength{\parindent}{0pt}\par\color{MidnightBlue} #1}}
+\newcommand{\answer}[1]{{\setlength{\parindent}{0pt}\par\color{blue} #1}}
+\newcommand{\answerok}[1]{{\setlength{\parindent}{0pt}\par\color{MidnightBlue} #1}}
 
@@
 \medskip
-We look forward to reading your feedback on this new manuscript.
+We look forward to reading your feedback on this new manuscript and apologize in advance for this long revision round.
 
@@
 
-\answer{We hope it is now clarified.}
+\answerok{In light of this potential unfamiliarity, we preferred to entirely avoid referring explicitly to this dichotomy (for instance, we now speak of ``corpus-based utterances'' instead of ``in vivo'').}
 

### 2016-08-06 brainscopycut.tex
minor before leaving

@@
 \authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
-\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; \emph{in vivo} psycholinguistics}
+\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; psycholinguistics}
 %\note{\TB{Note XXXX (version? date?)}}

### 2016-08-06 reviews/2016-07-response-to-reviewers.tex
further cover letter corrections, to be cont'd later today

@@
 \newcommand{\answerok}[1]{{\setlength{\parindent}{0pt}\par\color{MidnightBlue} #1}}
+\newcommand{\unclear}[1]{\textcolor{red}{\sf [C: #1]}}
 
@@
 
-\answer{
-Point to:
-- the added POS analysis
-- the short discussion of distances travelled upon substitution
+\answerok{We added a discussion on the semantics of substituted quotes.
+We first aimed at quantifying the semantic distance by estimating the distance ``traveled'' by the substituted word in the WordNet hyperonym-hyperonym network as well as the FA network.
+Substituted and substituting words are rarely synonyms, yet substitutions usually replace a word with a semantically-related word which achieves a similar meaning in the given context \unclear{voir ma question sur le manuscrit principal, ceci dit / d'ailleurs il faut faire gaffe car H00 concerne les synonymes, qui finalement ne sont pas si efficaces}.
+
+We also examined the effect of part-of-speech (POS): Figure 5 details the results on five POS categories.
+A Goodman-based multinomial goodness-of-fit shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
 }
+%\answer{ Point to: - the added POS analysis - the short discussion of distances travelled upon substitution }
 
@@
 \x what happens when you run a multi-factor regression model? Do the features interact?
+}}
+
+\answerok{Thank you for this question, we decided to introduce a simple regression model to check feature interaction, see the discussion related to Fig.
+8 (in a nutshell: we show that features generally do not interact).}
+
+\critique{\itx{
 \x clarify from the beginning the predictions you are making so that the results will be interpretable 
@@
 
-\answer{Point to the regression model we introduced.}
+\unclear{Voir le point {\bf (7)} de mon email}
 
@@
 \critique{\itx{
-\x introducing eight measures and then removing three of them due to collinearity is presentationally confusing
+\x introducing eight measures and then removing three of them due to collinearity is presentationally confusing}}
+
+\unclear{ah oui que répond-on là? qqch comme ça?:}
+\answer{We agree that introducing (now) twelve measures before eventually keeping six of them could complicate the flow of the section on word-level measures.
+However, we found it more natural to first propose a comprehensive, literature-based list of candidate features before addressing the quantitative analysis with a reduced yet sufficient set of metrics --- especially now that this section has been significantly expanded.
+We initially favored this solution over first presenting the final list of candidate features and justifying afterwards that they are anyway correlated with a broader list of traditional variables.
+(NB: An alternative option could consist in presenting the contents of Figs.
+1 and 2 in the same figure, if this could be less confusing.)}
+
+\critique{
+\itx
+{
 \x was there a reason to expect that \# of synonyms and \# of phonemes would interact (Fig. 4)?]
@@
 
-\answer{Yes, because of the similar curve shapes they had.
+\answer{
+Yes, because of the similar curve shapes they had.
 We removed number of phonemes now (in favor of number of letters), and number of synonyms doesn't show any effect of susceptibility on the new, improved measure.
 }
+\unclear{ne serait-ce pas plutôt la raison suivante?\\Except for word frequency, we observed that synonyms and phonems were the only two word features to exhibit an effect on susceptibility (and q quite similar effect for that matter).
+We thus reasoned that they were good candidates to look for and illustrate potential joint effects.\\However, the revised version now features a much braoder framework to statistically assess joint effects (see Fig.
+8), we think it favorably replaces the previous variable-to-variable correlation check (NB: we also replaced the number of phonemes with the number of letters, which makes more sense in our context; letters and synonyms do not seem to interact).}
 
@@
 
-\answer{
+\answer{\unclear{comment était fait le binning avant, d'ailleurs?
+j'aurais eu envie de dire "integer-based" mais à la vue de l'ancien PDF ça ne colle pas avec les graphiques, et j'ai l'impression qu'on n'en parlait pas non plus -- auquel cas la réponse devrait aussi faire mention de l'avant, qui était probablement l'objectif de la question du reviewer}
 Binning is now by quantiles (makes more sense for the confidence intervals), but the extended notebooks (https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks and https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb) also include graphs with the fixed-width bins we presented in the first version.

### 2016-08-06 brains-protocol.tex
minor before leaving
further cover letter corrections, to be cont'd later today

@@
 
-\subsection{\emph{In vivo} utterances}
+\subsection{\tb{Corpus-based} utterances}
 
@@
 
-\tb{However, the substituted and subtituting words are rarely direct synonyms:} only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
+\tb{However, the substituted and subtituting words are not so often direct synonyms:} only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
 \newtext{A similar \tb{phenomenon} is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
@@
 While we acknowledge the low susceptibilities of adverbs and stopword-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
-it seems, then, that POS categories don't capture any strong bias in the selection of substitution targets.
+it seems, then, that POS categories do not capture any strong bias in the selection of substitution targets.
 
@@
     \caption{\newtext{\textbf{Feature variations regression coefficients}:
-    source feature values and feature variations were normalized to $[0; 1]$ to ensure the coefficients are comparable.
+    source feature values \tb{(columns)} and feature variations \tb{(rows)} were normalized to $[0; 1]$ to ensure the coefficients are comparable.
     Significance levels for individual \emph{t}-tests against the hypothesis of a null coefficient are denoted by stars below the corresponding coefficient (*** for $p \leq .001$, ** for $p \leq .01$, * for $p \leq .05$, and nothing when $p > .05$).

### 2016-08-06 reviews/2016-07-response-to-reviewers.tex
further cover letter corrections, passing over to Seb

@@
 
-\answer{We agree the presentation was too much. Point to:
-- the integration of the iterated learning literature in review and general discussion
-- the fact that the constraints of the data don't let us explore chains, so we explained our claims more appropriately
-}
+\answer{We agree the presentation was too much. \unclear{too much?} Point to:
+\itx{
+\x the integration of the iterated learning literature in review and general discussion
+\x the fact that the constraints of the data don't let us explore chains, so we explained our claims more appropriately
+}}
 
@@
 
-\answer{Thank you. We paid a lot of attention to the writing style in this new version, as can be seen in the large rewrites of many paragraphs.}
+\answerok{Thank you for this very constructive and extensive feedback.
+We paid signification attention to the writing style in this new version, and large portions of the document have been rewritten and expanded --- all updates are marked in dark blue.}
 
@@
 \critique{I think this is an interesting approach and dataset for better understanding how sentences are recalled and how paraphrases are constructed.
-However, I think that the experiment design needs to be refined, and the project as a whole needs to be better connected to larger theoretical problems and better grounded within cognitive science.
-}
+However, I think that the experiment design needs to be refined, and the project as a whole needs to be better connected to larger theoretical problems and better grounded within cognitive science.}
 
-\answer{Point to:
-- the revamped introduction, review of related work, and discussion of results which connects our approach to the literature
+\answerok{We have been careful to introduce many more connections to the relevant literature: the introduction has been entirely rewritten and is followed by a more thorough presentation of the related work in cognitive science, while the concluding sections (discussion \& concluding remarks) follow your recommendation of connecting our results with the literature.
 }
@@
 I'm not convinced that correctly and incorrectly copying a quote should be viewed as separate behaviors.
-But more importantly, the $\sigma_\phi(f)$ calculation can be misleading under the single-substitution restriction:
 }
 
-\answer{
-Point to:
-- susceptibility of POS, showing that the effect (closed/open class words being different) expected is very small at best (closed class words would be stopword-likes in our graph)
-- added phonological neighbourhood density as a feature
-- the multi-feature regression of variation upon substitution, showing that variation of a feature is predicted mainly by the same source feature
-- the expansion of substitution models to two-substitutions, giving the same results as single-substitutions.
-Notebooks for all models can be found at https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks , and the graphs produced by those notebooks can be found at https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb .
+\answerok{
+We addressed all these points in the revised version:
+\itx{
+\x We checked the susceptibility of part-of-speech (POS), \unclear{je n'ai pas bien compris ton: ``showing that the effect (closed/open class words being different) expected is very small at best (closed class words would be stopword-likes in our graph)''  (et c'est très légèrement lié à mon point {\bf (5)} dans l'email) -- est-ce qu'il s'agit de l'effet des stopwords-like (== closed class words, exactement?) sur nos résultats (auquel cas, oui, ça a l'air de ne pas arriver souvent, en même temps on se focalise sur les substitutions) ou bien de la susceptibilité des stopwords-like par rapport aux autres (auquel cas elle semble effectivement très inférieure)}
+\x Following this suggestion, we added phonological neighborhood density as a word feature.
+We show that it forms a subgroup of strongly correlated variables, together with orthographic neighborhood density, and choose the latter as a more natural representative of this subgroup given the task under study.
+\x We introduced a multi-feature regression model to estimate variation upon substitution, showing that the magnitude of the variation of a feature is mainly predicted by the value of that very feature, generally with negligible interaction with the other features (see also below and figure 8).
+\x We expanded substitution models to account for \emph{two-substitutions} \unclear{explain here what is a ``two-substitution'', maybe?}, which yield the same results as single substitutions. 
+(Note: %computation notebooks for all models can be found at https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks , and the 
+graphs produced for all models may be found at <https://huit.re/aEH77Gc3>, and may be commented further upon request.)
 }
 
-\critique{Consider a toy case where all quotes are two words, and there is a single binary feature $F$.
+%We also examined the effect of part-of-speech (POS): Figure 5 details the results on five POS categories.
+A Goodman-based multinomial goodness-of-fit shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
+
+}
+
+\critique{But more importantly, the $\sigma_\phi(f)$ calculation can be misleading under the single-substitution restriction: Consider a toy case where all quotes are two words, and there is a single binary feature $F$.
 $p(F=1) = 3/4$, and the true substitution probabilities, given the value of F, are $p(S|F=1)=1$, $p(S|F=0)=1/2$.
@@
 \itx{
-\x more discussion of what dynamical systems are and especially their interpretation in this task before their appearance in the results
+\x more discussion of what dynamical systems are and especially their interpretation in this task before their appearance in the results}
+}
+
+\answerok{We clarified the notion of dynamical systems when rewriting the introduction and in the ``Variation'' section (including footnote \#22).}
+
+\critique{\itx{
 \x needs a substantive discussion of the results}
@@
 
-\answer{Point to rewrite of introduction and the added discussion of results.}
+\answerok{An entirely new ``Discussion'' section is now devoted to this task.}
 
@@
 
-\answer{
-Point to improved introduction and added discussion. Especially the connection with iterated learning literature.
+\answerok{
+These points should be addressed by the improved introduction.
+In particular regarding the last question, the new ``Discussion'' section additionally makes a connection with iterated learning literature.
 }
@@
 
-\unclear{ah oui que répond-on là? qqch comme ça?:}
-\answer{We agree that introducing (now) twelve measures before eventually keeping six of them could complicate the flow of the section on word-level measures.
+\unclear{ah oui que répond-on là? qqch comme ça? je te laisse valider:}
+\answerok{We agree that introducing twelve\footnote{In the revised version we have 12 measures, instead of 8 in the initial submission.} measures before eventually keeping six of them could complicate the flow of the section on word-level measures.
 However, we found it more natural to first propose a comprehensive, literature-based list of candidate features before addressing the quantitative analysis with a reduced yet sufficient set of metrics --- especially now that this section has been significantly expanded.
@@
 We thus reasoned that they were good candidates to look for and illustrate potential joint effects.\\However, the revised version now features a much braoder framework to statistically assess joint effects (see Fig.
-8), we think it favorably replaces the previous variable-to-variable correlation check (NB: we also replaced the number of phonemes with the number of letters, which makes more sense in our context; letters and synonyms do not seem to interact).}
+8), we think it favorably replaces the previous variable-to-variable correlation check (NB: we also replaced the number of phonemes with the number of letters [which makes more sense in our context, see the end of the ``Word-level features'' subsection], number of letters and synonyms do not seem to interact).}
 
@@
 
-\answer{Thank you for these very useful and, indeed, relevant pointers.
+\answerok{Thank you for these very useful and, indeed, relevant pointers.
 We accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' section).}
@@
 
-\answer{That would indeed be interesting, but not our goal in the paper.}
+\answer{That would indeed be interesting, but not our goal in the paper.\unclear{``merci vieux mais c'est hors-sujet'' ;) Pourrait-on lui dire plus diplomatiquement:} \textcolor{red}{``This would indeed be a very interesting project and prospect, but the data we collected do not enable the study of the surrounding context (in particular, we do neither know the prior usage nor the way information sources may be used) so that this goal remains beyond the scope of our paper.''}}
 
@@
 
-\answer{Point to:
-- the several substitution models implemented, tested, and checked}
+\answerok{There are indeed very many possible models, yet we agree that a robustness check is also very sensible.
+Substitution detection relies on a number of assumptions and we identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
+We thus implemented, tested and checked these models, and we can report that our results remain valid for all 16 models (for the sake of clarity, we do not specifically report on these models, though).
+\\
+
+This leads to a much more detailed ``Substitution Model'' subsection which discusses these options along with the reference model we used throughout the paper.
+}
 
@@
 
-\answer{Yes thank you, this is done now.}
+\answerok{This is an important empirical point which we did not address systematically in the first submission.
+We now have done this manual coding step and are glad to be able to report a (satisfying) accuracy in the relevant cases.}
 
-\critique{ Finally, while I understand the appeal of doing a dynamic systems analysis to find the convergence points of the substitutions, but it is not really valid here.
+\critique{Finally, while I understand the appeal of doing a dynamic systems analysis to find the convergence points of the substitutions, but it is not really valid here.
 The analysis in this paper does not provide information on what happens with multiple substitutions, so the authors cannot really make claims about the convergence points given this methodology.
@@
 
-\answer{We better explained our claims without overstating our findings.}
+\answerok{In this revised version, we better explained our claims and their limits (see for instance the second paragraph of the ``Discussion'' section, without overstating our findings.}
 
@@
 
-\answer{On a side note, we entirely agree with the relevance of pursuing this research program in these directions, as the \emph{in vitro} observation of information transmission and its evolution is part of an ongoing project (which nonetheless remains beyond the scope of the current paper).}
+\answerok{\textcolor{red}{[C: j'ai ajouté ça, je te laisse confirmer si ce teasing te semble approprié]}\\ On a side note, we entirely agree with the relevance of pursuing this research program in these directions, as the \emph{in vitro} observation of information transmission and its evolution is part of an ongoing project (which nonetheless remains beyond the scope of the current paper).}
 

### 2016-08-06 brains-conclusion.tex
further cover letter corrections, passing over to Seb

@@
 We do not actually observe quotes converging on a global scale towards attractors in their various dimensions.
-Indeed the limits of the data set don't allow us to infer chains of substitutions, and substitutions themselves are not the only type of transformation at work in the data set.
+Indeed the limits of the data set do not allow us to infer chains of substitutions, and substitutions themselves are not the only type of transformation at work in the data set.
 Nonetheless, these findings (1) bring light to this simple type of transformation, and (2) are consistent with known psycholinguistic effects, with the hypothesis of cultural attractors in representations from everyday life, and with the lineage specificity discussed in the iterated learning literature~\citep{claidiere_cultural_2014,cornish_systems_2013}.

### 2016-08-06 brains-protocol.tex
further cover letter corrections, passing over to Seb

@@
 }
+\end{new}
 
+\begin{new}
 Since our data is about written transformations, number of letters and orthographic neighborhood density are the natural representatives of groups (a) and (b) respectively.
@@
 however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
-
 \end{new}

### 2016-08-06 reviews/2016-07-response-to-reviewers.tex
further cover letter corrections, passing over to Seb

@@
 
-\medskip
+\bigskip
 We enclose an extensively revised version of the manuscript CS15-80 (initially titled ``How do we copy and paste?
@@
 
-\medskip
+\bigskip
 Best wishes,
 
-\medskip
+\bigskip
 Sébastien Lerique and Camille Roth
@@
 
-\answer{
-Point to:
-- work on the substitution models showing that the results are robust
-- the rationalization of feature selection
-- clarifications of choices, such a single-substitution restriction
-- precision/recall analyses for all filters
+\answerok{
+We acknowledge these issues and worked extensively to address them.  In particular (further details may also be found in the responses to reviewers): 
+\itx{
+\x We developed a variety of substitution models to show that the results are robust
+\x We rationalized feature selection, adding some variables and much detail to this section
+\x In general, we clarified further the assumptions we made, such as the restriction to single substitutions
+\x We also introduced manually-coded precision \& recall analyses for all filters, which justifies our choices.
+}
 }

### 2016-08-10 brains-conclusion.tex
Corrections after Camille's manuscript and emails

@@
 
-We \tb{initially} aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
+We initially aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
 The data set we used imposed a few constraints on our analysis:
@@
 They are obtained by successfully applying knowledge from cognitive science to real-life complex data, a task that remains a challenge in the study of cultural evolution.
-\tb{More broadly,} we believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.
+More broadly, we believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.
 
@@
 in our case, a substitution could be triggered and directed by associations formed by preceding context.
-A further step would be to follow \citet{cornish_systems_2013} have shown about reciprocal influences between context and transformations (in their case, with transmission chains of artificial content).
+A further step would be to follow what \citet{cornish_systems_2013} have shown about reciprocal influences between context and transformations (in their case, with transmission chains of artificial content).
 Indeed substitutions, and more generally all transformations, also participate in creating the context for later quotes.
@@
 Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
-In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge.
+In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge, and \tb{help assess the potential causal effect they have on subsequent evolution}.
 
@@
 Such an approach has long been called upon by anthropologists like~\citet{ingold_beyond_2004,ingold_three_1999}, in line with Mauss' initial works~\citep{mauss_les_1936}, and the debate is not entirely foreign to the enactive-representational discussion in cognitive science.
-As regards our particular case, a most promising line of work on the evolution of language is being developed by \citet{cuffari_participatory_2014}.
+%As regards our particular case, a most promising line of work on the evolution of language is being developed by \citet{cuffari_participatory_2014}.
 The study of cultural evolution will most likely benefit greatly from the growing interactions between these disciplines.

### 2016-08-10 brains-introduction.tex
Corrections after Camille's manuscript and emails

--- 2016-07-30 brains-introduction.tex
@@
 
-\tb{The reciprocal influence between cognition and culture has a long history in both social science and psychology.
-While this question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.}
 %Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
 The question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), but today's discussion is mostly structured by proponents from cognitive science.
-These construe culture as an evolutionary process analogous and parallel to biological evolution \tbn{(and especially the modern synthesis' account of it)}.
-That analogy can \tb{also} be traced a long way back \tbn{in the 20th century and earlier}, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\tb{\citeyearpar{dawkins_selfish_1976}}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
+The reciprocal influence between cognition and culture has a long history in both social science and psychology.
+While this question has been the subject of intense debate in the social sciences in the 20th century, today's discussion is mostly structured by proponents from cognitive science\tb{, who} construe culture as an evolutionary process analogous and parallel to biological evolution.
+That analogy can be traced a long way back, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\citeyearpar{dawkins_selfish_1976}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
 More recently, Dan Sperber has drawn on this principle to explicitly connect anthropology and cognitive science through the theory of \emph{Epidemiology of Representations} \citep{sperber_explaining_1996}, and the study of cultural evolution has been growing steadily since.
 
-The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
-\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
-\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
+The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
+\Citet{mesoudi_multiple_2008} discuss the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
+\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, demonstrate how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
 \citeauthor{kirby_cumulative_2008}~(\citeyear{kirby_cumulative_2008,cornish_systems_2013}) have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
@@
 A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
-Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.
+Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the above-cited studies are doing.
 
@@
 there should be areas of the representation space where cognitive effects in transformations bring representations closer to a given stable asymptotic point.\footnote{
-\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
+\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or confined to) the target sub-area.
 }}
@@
 Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
-For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in the email network~\citep{liben-nowell_tracing_2008};
-\citet{cointet_socio-semantic_2009} have described the reciprocal influence between the social network topology and the distribution of issues;
-\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} have studied the characteristics of particularly memorable quotes that circulate in those networks.
-We believe those works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.
+For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in emails~\citep{liben-nowell_tracing_2008};
+\citet{cointet_socio-semantic_2009} described the reciprocal influence between the social network topology and the distribution of issues;
+\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} studied the characteristics of particularly memorable quotes that circulate in those networks.
+We believe these works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.
 
@@
 
-To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
+To show this we analyze how quotes in blogs and media outlets are modified when they are copied from website to website.
 These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
@@
 
-Our question is as follows: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases referred to in the theory of epidemiology of representations, do cultural attractors appear and how do cognitive biases participate in them?
-We chose to restrict our analysis to substitutions (one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
+We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, \tb{do cultural attractors appear and how do cognitive biases participate in them?}
+%how do cognitive biases participate in \tb{the appearance of cultural attractors}?
+We chose to restrict our analysis to substitutions (i.e., one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
 \newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
@@
 While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
-This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, but with natural language out of the laboratory.
+This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, yet with natural language out of the laboratory.
 
@@
 %Sec.~\ref{sec:results}
-Next, we describe the measures built to observe the cognitive biases operating in quote transmission.
+Next, we introduce the measures we built to observe cognitive biases operating in quote transmission.
 Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.
-
 \end{new}
@@
 \newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
-The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
-\newtext{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
+The first above-mentioned method consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
+\tb{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the importance of the originating cultural systems in the interpretation of such data.
+%have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
 }}
@@
 \citet{potter_regeneration_1990} suggest that immediate recall of sentences is based on the retention of an unordered list of words which is then regenerated as a sentence at the moment of production.
-Priming recall with other words can lead to replacement in the recalled sentence if the primed words support the overall meaning of the sentence.
+Priming recall with other words can lead to replacement in the recalled sentence if the primed words are consistent with the overall meaning of the sentence.
 Regenerated syntax can also be influenced by priming recall with another syntactic structure~\citep{potter_syntactic_1998}, or with verbs whose category constraints call for a different structure~\citep{lombardi_regeneration_1992}.
@@
 One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which words are recalled~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}.
-The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
-indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations of subjects (contributing to associative information retrieval) is at the center of many of these studies.
+The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006}.
+Indeed, the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations of subjects (contributing to associative information retrieval) is at the core of many of these studies.
 
@@
 }}
-If a cognitive bias in the} substitution \newtext{of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.}
+If a cognitive bias in the substitution of words manifests itself with simple measures, then it} \newtext{will be worth applying predictive models of the substitution process in further research.}
 

### 2016-08-10 brains-protocol.tex
Corrections after Camille's manuscript and emails

@@
 \newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
-\newtext{\tbn{Indeed}}
 Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
@@
 
-To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting \tb{of} the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
+To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting of the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
 \newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see ``Substitution model'').}
-To quantify \tb{these} substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
+To quantify these substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
 
@@
 
-\subsection{\tb{Corpus-based} utterances}
+\subsection{Corpus-based utterances}
 
@@
 \end{quote}
-\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, \tb{i.e.} sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
+\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is
+% Le APA style guide force "that is" au lieu de "i.e." hors de parenthèses
+sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.
@@
 }
-Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our \tb{filtering} approach fulfills its goals by coding a random sub-sample of 100 clusters: 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
+Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our filtering approach fulfills its goals by coding a random sub-sample of 100 clusters: 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
 Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.}\footnote{
@@
 
-\newtext{To compute \tb{these} features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
+\newtext{To compute these features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
 As \citet{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
@@
 Now say a third version, ``These \textbf{allegations} are false and \textbf{incoherent}'' ($q''$) also appears once on January 19 and once on January 20 after $q'$.
-\tb{Here,} $q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (\tb{see} below for further details).
+Here, $q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (see below for further details).
 $q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
@@
 (1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
-(2) \textbf{bin span}, which can be \tb{$\Delta t=1$ day} %24 hours 
+(2) \textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
 (as in the model above) or can be extended to start at the very first occurrence in the quotation family.
-The other two parameters \tb{configure} rules on the selection of source and destination quotes of a substitution:
+The other two parameters configure rules on the selection of source and destination quotes of a substitution:
 (3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
@@
 Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
-Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use \tb{discrete} %ly positioned 
+Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use discrete %ly positioned 
 bins spanning 1 day:
@@
 
-In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread \tb{over} 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
+In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread over 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
 From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
-this eventually yields \newtext{6,318} valid substitutions\footnote{
+this eventually yields \newtext{6,177} valid substitutions.\footnote{
 \newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
@@
 Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
-out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}} \newtext{\tbn{(before merging substitutions that share the same destination)}}.
+out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}}
 
@@
 
-\tb{Substitutions usually} replace a word with another semantically related word: the new word achieves a similar meaning in the context of its sentence, while still slightly changing the implications or the attitude expressed by the author.
-The following examples illustrate this \tb{phenomenon}:
-
+Substitutions usually replace a word with another semantically related word:
+\tb{manual observation of a random subset of 100 substitutions shows that, compared to the word it replaces,} the new word often achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
+The following examples illustrate this phenomenon:\footnote{
+\newtext{However, the substituted and substituting words are not so often direct synonyms:
+only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.
+A similar phenomenon is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
+}}
 \begin{APAitemize}
@@
 \end{APAitemize}
-
-\tb{However, the substituted and subtituting words are not so often direct synonyms:} only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
-\newtext{A similar \tb{phenomenon} is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
-}}
 
@@
 
-\newtext{Our question concerns the low-level properties of these substitutions: \tb{we ask (i) which words are targets of the substitutions and (ii) what change they are subjected to.
-To this end,}} we build the following two observables for each word feature.
+\newtext{Our question concerns the low-level properties of these substitutions: we ask (i) which words are targets of the substitutions and (ii) what change they are subjected to.
+To this end,} we build the following two observables for each word feature.
 First, we measure \newtext{which word features are more or less substituted compared to how often they would be if the process were random, in order to capture} the susceptibility for words to be the target of a substitution in a quote.
@@
     \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
-    \caption{\newtext{\textbf{\tb{POS-related results:}} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
-    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows \tb{the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts}.
+    \caption{\newtext{\textbf{POS-related results:}
+    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories \tb{essentially made of stopwords} (see main text for details).
+    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
     Confidence intervals are computed with the Goodman method for multinomial proportions.}
@@
 \newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
-If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.} \tb{Further, t}o avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
-That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
+If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.
+Further, to avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
+\tb{That is, each quote cluster has a maximum contribution of 1, computed as the average contribution of all substitutions in that cluster.}
+}}
 
@@
 \newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions,
-\tb{\hbox{i.e.} it measures the preferential selection of some word properties given a substitution.}
+\hbox{i.e.} \newtext{it measures the preferential selection of some word properties \tb{for} substitution.}
 %\newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
@@
 A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Stopword-like}\footnote{
-\newtext{The \emph{Stopword-like} category gathers all the POS groups partially of completely affected by the stopword filter for substitutions: coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections.
-Note that the stopword filter equally affects $s$ and $s^0$, so while both numbers are very small for these categories (which led us to group them together) the reported susceptibility is not biased by this filtering.}
+\newtext{The \emph{Stopword-like} category gathers all the POS groups \tb{essentially made of stopwords (coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections), which are affected by our decision to exclude stopwords from the analysis.
+Indeed substitutions falling on stopwords are filtered out (affecting $s$), and stopwords are never counted as substitutable in the computation of $s^0$.
+Both numbers are subsequently very small for these categories (which led us to group them together) but the reported susceptibility is left unbiased.}}
 }
@@
     \caption{\newtext{\textbf{Feature variations regression coefficients}:
-    source feature values \tb{(columns)} and feature variations \tb{(rows)} were normalized to $[0; 1]$ to ensure the coefficients are comparable.
+    source feature values (columns) and feature variations (rows) were normalized to $[0; 1]$ to ensure the coefficients are comparable.
     Significance levels for individual \emph{t}-tests against the hypothesis of a null coefficient are denoted by stars below the corresponding coefficient (*** for $p \leq .001$, ** for $p \leq .01$, * for $p \leq .05$, and nothing when $p > .05$).

### 2016-08-10 brainscopycut.tex
Corrections after Camille's manuscript and emails

@@
 \abstract{
-\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations.
-Such attractors have been \tb{proposed} as \tb{a concept useful for} understanding cultural evolution %and its reciprocal influences 
-\tb{in relation} with \tb{individual} cognition, but their existence has been hard to test.
-\tb{We empirically} focus on \tb{the online transformation of} quotations when they are copied from blog to blog or media website: by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
+\newtext{We present an empirical case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations.
+Such attractors have been proposed as a useful concept for understanding cultural evolution in relation with individual cognition, but their existence has been hard to test.
+We focus on the transformation of quotations when they are copied from blog to blog or media website:
+by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
 In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
-Our results support the hypothesis that cultural attractors can \tb{result} from the combination of individual cognitive biases in the interpretation and reproduction of representations.}
+Our results support the hypothesis that cultural attractors can result from the combination of individual cognitive biases in the interpretation and reproduction of representations.}
 }

### 2016-08-10 reviews/2016-07-response-to-reviewers.tex
Fill in the gaps for the cover letter

@@
 \newcommand{\answer}[1]{{\setlength{\parindent}{0pt}\par\color{blue} #1}}
+\newcommand{\tb}[1]{\textcolor{blue}{#1}}
 \newcommand{\answerok}[1]{{\setlength{\parindent}{0pt}\par\color{MidnightBlue} #1}}
@@
 \bigskip
-We enclose an extensively revised version of the manuscript CS15-80 (initially titled ``How do we copy and paste?
+We enclose an extensively revised version of the manuscript CS15-80 (initially entitled ``How do we copy and paste?
 The semantic drift of quotations in blogspace'') based on the report you sent us on Jul 18, 2015.
@@
 We hope that we have appropriately addressed the detailed remarks and very constructive questions made by you and the referees: to this effect, we also provide below a detailed response to each of these comments.
-We additionally used a different color to emphasize the portions which were thoroughly rewritten.
+We additionally used a different color in the manuscript to emphasize the portions which were thoroughly rewritten.
 
@@
 The reader needs to be convinced that the few “strong” looking patterns you do find are not too dependent on these arbitrary choices.
-I actually question if this is the best way to address this question (effect of memory and linguistic variables on copy errors) given the noisiness of the data and the difficult in attributing copying errors to any particular parent
+I actually question if this is the best way to address this question (effect of memory and linguistic variables on copy errors) given the noisiness of the data and the difficulty in attributing copying errors to any particular parent
 source.
@@
 
-\answer{We agree the presentation was too much. \unclear{too much?} Point to:
-\itx{
-\x the integration of the iterated learning literature in review and general discussion
-\x the fact that the constraints of the data don't let us explore chains, so we explained our claims more appropriately
-}}
+\answer{We agree that the initial manuscript overstated some results.
+We believe the new manuscript states its claims more appropriately, and makes a better distinction between actual conclusions and open questions that could not be addressed given the data (in particular, we attempted to explore substitution chains in earlier stages of this work, but were quickly limited by the constraints of the data).
+
+We did indeed overlook several developments in the iterated learning literature, and are grateful for the pointer.
+The works we thought were most relevant have been integrated to the introductory sections, and their implications for our work is discussed in the newly ``Discussion'' section.
+}
 
@@
 
-\answer{Point to:
-- the new discussion of results
-- the improved review of literature
-- the use of H00 to compare to synonyms, and the small exploration of distances travelled by substitutions on the FA and synonym networks
+\answer{We thoroughly reworked the general grounding and flow of the manuscript. In particular:
+\itx{
+\x The introductory parts provide a clearer presentation of our question and how it relates to existing literature (including the areas we overlooked) and hypotheses
+\x The new ``Discussion'' section, which was indeed missing, properly reconnects our results to the related works introduced at the beginning of the paper.
+}
+
+Finally, several changes in the ``Results'' section (more examples, a short discussion of the distance traveled by substitutions on the synonym network in footnote \#16, and the addition of the second null hypothesis $\mathcal{H}_{00}$ in the ``Variation'' subsection) should better describe the type of changes undergone by quotes.
 }
@@
 First, some crucial factors were not considered - most notably, part of speech.
-I suspect, for instance, that pronouns and other closed-class words less often substituted than open-class words.
+I suspect, for instance, that pronouns and other closed-class words are less often substituted than open-class words.
 Phonological neighborhood density could also be important, though probably less so.
@@
 \itx{
-\x We checked the susceptibility of part-of-speech (POS), \unclear{je n'ai pas bien compris ton: ``showing that the effect (closed/open class words being different) expected is very small at best (closed class words would be stopword-likes in our graph)''  (et c'est très légèrement lié à mon point {\bf (5)} dans l'email) -- est-ce qu'il s'agit de l'effet des stopwords-like (== closed class words, exactement?) sur nos résultats (auquel cas, oui, ça a l'air de ne pas arriver souvent, en même temps on se focalise sur les substitutions) ou bien de la susceptibilité des stopwords-like par rapport aux autres (auquel cas elle semble effectivement très inférieure)}
+\x We checked and discussed the susceptibility of part-of-speech (POS).
+\tb{While the open-/closed-class question is precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), if we note that the manuscript's \emph{Stopword-like} category is mainly made of closed-class words, the data indicates that they are indeed less substituted than open-class words.
+As the manuscript discusses however, this concerns less than 7\% of the substitutions we consider, and has likely no effect on our results.}
 \x Following this suggestion, we added phonological neighborhood density as a word feature.
-We show that it forms a subgroup of strongly correlated variables, together with orthographic neighborhood density, and choose the latter as a more natural representative of this subgroup given the task under study.
-\x We introduced a multi-feature regression model to estimate variation upon substitution, showing that the magnitude of the variation of a feature is mainly predicted by the value of that very feature, generally with negligible interaction with the other features (see also below and figure 8).
-\x We expanded substitution models to account for \emph{two-substitutions} \unclear{explain here what is a ``two-substitution'', maybe?}, which yield the same results as single substitutions. 
-(Note: %computation notebooks for all models can be found at https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks , and the 
-graphs produced for all models may be found at <https://huit.re/aEH77Gc3>, and may be commented further upon request.)
+We show that it forms a subgroup of strongly correlated variables, together with orthographic neighborhood density, and choose the latter as a more natural representative of this subgroup given the task studied.
+\x We introduced a multi-feature regression model to estimate variation upon substitution, showing that the magnitude of the variation of a feature is mainly predicted by the \tb{start word's value for} that very feature, generally with negligible interaction with the other features (see also below and figure 8).
+\x We expanded substitution models to \tb{also detect quotes undergoing up to two substitutions}, which yielded the same results as single substitutions \tb{(see below)}. 
 }
@@
 
-\answer{This case is indeed valid, if there is such a feature (F=1 here) that:
-(a) is extremly frequent,
-(b) is extremely substituted,
-(c) interacts strongly with another feature so that both are often substituted together.
+\answer{The case presented here is valid and could indeed hide a bias in our measures.
+Rather than analyzing all possible such cases to bound the error, we extended substitution models to detect quotes undergoing up to two substitutions, and applied the same analysis to these mined substitutions.
+The results were unchanged (note: graphs produced for all models may be found at <https://huit.re/aEH77Gc3>, and may be commented further upon request).
+For example, the extended version of the model presented in the paper mines 70k substitutions, 9.2k of which are kept after filtering (the lowered acceptance rate indicates that more invalid substitutions are generated when allowing up to two substitutions per quote).
 
-The combination of (a) and (b) is an anti-pattern in our results (more frequent =less substituted), so we could in theory give a bound to the worst error that could have been committed.
-However we extend our substitution models to two-substitutions, and observe that the results do not change.
+We did not expand to more than two substitutions, as more than 20\% of quotes are 5 to 9 words long (with a minimum of 5 words for quotes in the data set).
+For such quotes, three or more substitutions means changing at least a third of the sentence, which makes any attempt to link it to a similar variant pointless.
 }
@@
 
-\answer{
-Point to:
-- our control of single-substitution bias
-- the change in definition of phi.
-Indeed it was confusing to present it (and measure it) as a probability (which it was, but an aggregated global one, not an in-sentence one), we now measure it as the strength of a bias.
+\answer{Fig.
+4 of the previous manuscript represented probabilities \emph{not knowing which sentence the word appears in} (i.e.
+marginal probabilities), which cannot be used to ask questions given a particular sentence.
+
+Since such a measure is indeed not very useful for practical interpretation, we refined the definition of susceptibility to make it more straightforward.
+It is now defined relative to a null hypothesis and measures the strength of a bias for a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly); it is presented in Figs.
+5 \& 6.
+
+Finally, we extended our measures to capture part of the context of the sentence substitutions occur in (see the ``Sentence context'' subsection).
+Sentence-relative susceptibility, in particular, gives a straightforward answer to the question of which words are more likely to attract substitution in a given sentence.
 }
@@
 These points should be addressed by the improved introduction.
-In particular regarding the last question, the new ``Discussion'' section additionally makes a connection with iterated learning literature.
+In particular regarding the last question, the new ``Discussion'' section additionally makes a connection with the iterated learning literature.
 }
@@
 
-\answerok{We added a discussion on the semantics of substituted quotes.
-We first aimed at quantifying the semantic distance by estimating the distance ``traveled'' by the substituted word in the WordNet hyperonym-hyperonym network as well as the FA network.
-Substituted and substituting words are rarely synonyms, yet substitutions usually replace a word with a semantically-related word which achieves a similar meaning in the given context \unclear{voir ma question sur le manuscrit principal, ceci dit / d'ailleurs il faut faire gaffe car H00 concerne les synonymes, qui finalement ne sont pas si efficaces}.
+\answerok{We added a \tb{short} discussion on the semantics of substituted quotes.
+We first aimed at quantifying the semantic distance by estimating the distance ``traveled'' by the substituted word in the WordNet hypernym-hyponym network as well as the FA network.
+\tb{This showed that substituted and substituting words are rarely direct synonyms, yet they usually achieve a similar meaning in the context of their sentence.
+To further grasp the relationship of substitutions to synonyms we introduced a second null hypothesis in the ``Variation'' section, which compares the actual variation to what would happen if a random synonym of the start word was picked instead.}
 
-We also examined the effect of part-of-speech (POS): Figure 5 details the results on five POS categories.
-A Goodman-based multinomial goodness-of-fit shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
+Finally, we also examined the effect of part-of-speech (POS): Fig.
+5 details the results on five POS categories, \tb{and the subsequent discussion} shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
 }
-%\answer{ Point to: - the added POS analysis - the short discussion of distances travelled upon substitution }
 
@@
 
-\answerok{Thank you for this question, we decided to introduce a simple regression model to check feature interaction, see the discussion related to Fig.
-8 (in a nutshell: we show that features generally do not interact).}
+\answerok{Thank you for this question.
+We introduced a simple regression model to check for feature interactions, \tb{which is presented in Fig.
+8 with a corresponding discussion} (in a nutshell: we show that features generally do not interact).}
 
@@
 
-\unclear{Voir le point {\bf (7)} de mon email}
+\answer{We hope the improved introduction is clearer as to what we predict and what is left open due to the exploratory nature of our study.}
 
@@
 
-\unclear{ah oui que répond-on là? qqch comme ça? je te laisse valider:}
 \answerok{We agree that introducing twelve\footnote{In the revised version we have 12 measures, instead of 8 in the initial submission.} measures before eventually keeping six of them could complicate the flow of the section on word-level measures.
-However, we found it more natural to first propose a comprehensive, literature-based list of candidate features before addressing the quantitative analysis with a reduced yet sufficient set of metrics --- especially now that this section has been significantly expanded.
+However, we found it more natural to first propose a comprehensive list of candidate features based on the literature before addressing the quantitative analysis with a reduced yet sufficient set of metrics --- especially now that this section has been significantly expanded.
 We initially favored this solution over first presenting the final list of candidate features and justifying afterwards that they are anyway correlated with a broader list of traditional variables.
-(NB: An alternative option could consist in presenting the contents of Figs.
-1 and 2 in the same figure, if this could be less confusing.)}
+(Note: An alternative option could consist in presenting the contents of Figs.
+1 and 2 in the same figure, if this would be less confusing.)}
 
@@
 
-\answer{
-Yes, because of the similar curve shapes they had.
-We removed number of phonemes now (in favor of number of letters), and number of synonyms doesn't show any effect of susceptibility on the new, improved measure.
-}
-\unclear{ne serait-ce pas plutôt la raison suivante?\\Except for word frequency, we observed that synonyms and phonems were the only two word features to exhibit an effect on susceptibility (and q quite similar effect for that matter).
-We thus reasoned that they were good candidates to look for and illustrate potential joint effects.\\However, the revised version now features a much braoder framework to statistically assess joint effects (see Fig.
-8), we think it favorably replaces the previous variable-to-variable correlation check (NB: we also replaced the number of phonemes with the number of letters [which makes more sense in our context, see the end of the ``Word-level features'' subsection], number of letters and synonyms do not seem to interact).}
+\answerok{Except for word frequency, we observed that synonyms and phonemes were the only two word features to exhibit an effect on susceptibility (and a quite similar effect for that matter).
+We thus reasoned that they might exhibit a joint effect.
+
+However, the revised version now features a much broader framework to statistically assess joint effects (see Fig.
+8); we think it favorably replaces the previous variable-to-variable correlation check.
+(Note: we also replaced the number of phonemes with the number of letters -- which makes more sense in our context, see the end of the ``Word-level features'' subsection --, which does not seem to interact with number of synonyms).}
 
@@
 
-\answer{\unclear{comment était fait le binning avant, d'ailleurs?
-j'aurais eu envie de dire "integer-based" mais à la vue de l'ancien PDF ça ne colle pas avec les graphiques, et j'ai l'impression qu'on n'en parlait pas non plus -- auquel cas la réponse devrait aussi faire mention de l'avant, qui était probablement l'objectif de la question du reviewer}
-Binning is now by quantiles (makes more sense for the confidence intervals), but the extended notebooks (https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks and https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb) also include graphs with the fixed-width bins we presented in the first version.
-No multiple crossovers were ever observed.
+\answer{The data in Figs.
+4 \& 5 was binned into 5 integer-based bins, and the graphs were cropped whenever the confidence intervals were unreasonably large (which gave the impression of some graphs having only 4 bins).
+
+The data in the new Figs.
+6, 7 \& 9 is now binned into 4 quartile-based bins, which makes more sense for the confidence intervals.
+This decision was reached after exploring several other binning configurations: none of these substantially changed the results, but opting for quartile-based bins does improve the overall readability.
+Notably, no multiple crossovers were ever observed.
 }
@@
 Even so, I would suspect that global semantic network measures are not nearly as important as word-word metrics.
-}
 
-\answer{Point to:
-- discussion of why we use word features (intro + related work), 
-- why we don't do exact word prediction (related work showing the question is not even mastered for random word lists in the laboratory, so forget about sentences in out-of-laboratory data), 
-- clustering coefficient is not so central to our results now
-- the addition of H00 gives insight into what the new words are w.r.t. to synonyms of the disappearing word
-}
-
-\critique{I would guess that the substitutions that are made in quotations are overwhelmingly higher frequency, highly semantically similar words.
+I would guess that the substitutions that are made in quotations are overwhelmingly higher frequency, highly semantically similar words.
 As the example on pg.
@@
 A tool by Recchia \& Jones (2010, BRM) allows for pointwise mutual information (PMI) of different words to be retrieved from a given corpus very simply.
-A tool such as this would allow for word-word similarity to be included in the analysis.
+A tool such as this would allow for word-word similarity to be included in the analysis.}
 
-I understand that word-word metrics are not easily introduced in the type of analysis that the authors are attempting to do, where the analysis of the contributions of different global psycholinguistic variables are being weighed against each other.
+\answer{We understand that the analysis we present is not the approach usually taken for similar laboratory data.
+The rewritten introductory sections of this revision now explain why we chose to use a feature-based analysis over the prediction of individual words (see below for more about this).
+
+Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (footnote \#16).
+We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is discussed in that section.
+
+Finally, we introduced a new null hypothesis ($\mathcal{H}_{00}$) in the ``Variation'' subsection, which gives insight into how an appearing word relates to the synonyms of the word it replaces.
+}
+
+\critique{I understand that word-word metrics are not easily introduced in the type of analysis that the authors are attempting to do, where the analysis of the contributions of different global psycholinguistic variables are being weighed against each other.
 I just think that type of analysis is inappropriate here.
@@
 
-\answer{
-Point to:
-- discussion of distance travelled by substitution on FA and synonyms networks
-- our data is ill-shaped for the kind of topic modelling suggested (aside from the earlier point that even with experiments in the laboratory, such tools don't predict so well)
+\answer{While we agree that predicting the actual disappearing and appearing words would be an extremely interesting result, we believe it is several steps away from what is currently possible.
+As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
+Zaromb et al.
+(2006), for instance, predict the list from which prior-list intrusions come, but not the exact intruding word.
+To our knowledge, the prediction of intrusions in full sentences (either manually designed or not) has not been tackled.
+Finally, as explained in footnote \#4 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.
 }
@@
 
-\answer{That would indeed be interesting, but not our goal in the paper.\unclear{``merci vieux mais c'est hors-sujet'' ;) Pourrait-on lui dire plus diplomatiquement:} \textcolor{red}{``This would indeed be a very interesting project and prospect, but the data we collected do not enable the study of the surrounding context (in particular, we do neither know the prior usage nor the way information sources may be used) so that this goal remains beyond the scope of our paper.''}}
+\answer{This would indeed be a very interesting project and prospect, but the data we collected do not enable the study of the surrounding context (in particular, we know neither the prior usage nor the way information sources may be used) so that this goal remains beyond the scope of our paper.
+
+We did however extend our measures to capture part of the context of the sentence substitutions occur in (see the ``Sentence context'' subsection).}
 
@@
 
-\answer{Point to:
-- the improved substitution selection section
-- the graphics for the other features, available in the notebooks (https://github.com/wehlutyk/brainscopypaste/tree/master/data/notebooks) and figures (https://github.com/wehlutyk/brainscopypaste/tree/master/data/figures/susceptibility.ipynb), which we can put in a Supplementary Information if needed
+\answer{The goal of our feature selection is to render the results more intelligible.
+This section was indeed poorly written, and has been thoroughly revised to rationalize the process.
+With few exceptions (notably number of phonemes and number of syllables less biased than number of letters), agglomerated features exhibit very similar behaviors across all measures (graphs produced for all features may be found at <https://huit.re/aEH77Gc3>, and could be included in a Supplementary Information bundle if requested).
 }
@@
 \answerok{This is an important empirical point which we did not address systematically in the first submission.
-We now have done this manual coding step and are glad to be able to report a (satisfying) accuracy in the relevant cases.}
+We have now done this manual coding step, \tb{and improved filters wherever precision or recall were not good enough.
+Accuracy is now assessed and satisfying in the relevant cases.}}
 
@@
 
-\answerok{In this revised version, we better explained our claims and their limits (see for instance the second paragraph of the ``Discussion'' section, without overstating our findings.}
+\answerok{In this revised version, we better explained our claims and their limits (see for instance the second paragraph of the ``Discussion'' section), without overstating our findings.}
 
@@
 
-\answerok{\textcolor{red}{[C: j'ai ajouté ça, je te laisse confirmer si ce teasing te semble approprié]}\\ On a side note, we entirely agree with the relevance of pursuing this research program in these directions, as the \emph{in vitro} observation of information transmission and its evolution is part of an ongoing project (which nonetheless remains beyond the scope of the current paper).}
+\answerok{On a side note, we entirely agree with the relevance of pursuing this research program in these directions, as the \emph{in vitro} observation of information transmission and its evolution is part of an ongoing project (which nonetheless remains beyond the scope of the current paper).}
+
+\subsection*{Final remarks}
+
+\tb{A few additional numbers have changed throughout the manuscript:
+\itx{
+\x Clustering coefficient values have changed, because Free Association link weights are now taken into account in the computation (i.e.
+coefficients are computed on the undirected weighted graph instead of the undirected unweighted graph).
+\x An update in the language detection module led to a small change in the results of cluster filtering: there are a few more clusters and quotes than in the previous version.
+As a result of this, the number of words coded by Word Frequency has also changed (since frequency of words is computed from the set of quotes we keep after cluster filtering).
+\x The discovery of three bugs, and the improvement of substitution filtering, led us to mine many more substitutions than previously.
+All in all the software is now much more reliable, as unit tests cover nearly all of it.
+}
+}
 

### 2016-08-20 reviews/2016-07-response-to-reviewers.tex
pre-final round of modifications manuscript + cover

@@
 
-\answer{We agree that the initial manuscript overstated some results.
+\answerok{We agree that the initial manuscript overstated some results.
 We believe the new manuscript states its claims more appropriately, and makes a better distinction between actual conclusions and open questions that could not be addressed given the data (in particular, we attempted to explore substitution chains in earlier stages of this work, but were quickly limited by the constraints of the data).
 
-We did indeed overlook several developments in the iterated learning literature, and are grateful for the pointer.
-The works we thought were most relevant have been integrated to the introductory sections, and their implications for our work is discussed in the newly ``Discussion'' section.
+We \tb{had also overlooked} several developments in the iterated learning literature, and are grateful for the \tb{various} pointers.
+The works we thought were most relevant have been integrated to the introductory sections, and their implications for our work are discussed in the new ``Discussion'' section.
 }
@@
 
-\answer{We thoroughly reworked the general grounding and flow of the manuscript. In particular:
+\answerok{We thoroughly reworked the general grounding and flow of the manuscript. In particular:
 \itx{
-\x The introductory parts provide a clearer presentation of our question and how it relates to existing literature (including the areas we overlooked) and hypotheses
-\x The new ``Discussion'' section, which was indeed missing, properly reconnects our results to the related works introduced at the beginning of the paper.
+\x \tb{We believe that} the introductory parts provide a clearer presentation of our question and how it relates to existing literature (including the areas we overlooked) and hypotheses
+\x The new ``Discussion'' section, which was indeed missing, \tb{should} properly reconnect our results to the related works introduced at the beginning of the paper.
 }
 
-Finally, several changes in the ``Results'' section (more examples, a short discussion of the distance traveled by substitutions on the synonym network in footnote \#16, and the addition of the second null hypothesis $\mathcal{H}_{00}$ in the ``Variation'' subsection) should better describe the type of changes undergone by quotes.
+Finally, several modifications in the ``Results'' section (more examples, a short discussion of the distance traveled by substitutions on the synonym network in footnote \#16, and the addition of the second null hypothesis $\mathcal{H}_{00}$ in the ``Variation'' subsection) should better describe the type of changes undergone by quotes.
 }
@@
 \x We checked and discussed the susceptibility of part-of-speech (POS).
-\tb{While the open-/closed-class question is precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), if we note that the manuscript's \emph{Stopword-like} category is mainly made of closed-class words, the data indicates that they are indeed less substituted than open-class words.
-As the manuscript discusses however, this concerns less than 7\% of the substitutions we consider, and has likely no effect on our results.}
+While the open-/closed-class question is \tb{largely} precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), if we note that the manuscript's \tb{\emph{Stopword-like}} category is mainly made of \tb{closed-class words}, the data indicates that they are indeed less substituted than open-class words.
+As the manuscript discusses however, this concerns less than 7\% of the substitutions we consider, and has likely no effect on our results.
 \x Following this suggestion, we added phonological neighborhood density as a word feature.
 We show that it forms a subgroup of strongly correlated variables, together with orthographic neighborhood density, and choose the latter as a more natural representative of this subgroup given the task studied.
-\x We introduced a multi-feature regression model to estimate variation upon substitution, showing that the magnitude of the variation of a feature is mainly predicted by the \tb{start word's value for} that very feature, generally with negligible interaction with the other features (see also below and figure 8).
-\x We expanded substitution models to \tb{also detect quotes undergoing up to two substitutions}, which yielded the same results as single substitutions \tb{(see below)}. 
+\x We introduced a multi-feature regression model to estimate variation upon substitution, showing that the magnitude of the variation of a feature is mainly predicted by the {start word's value for} that very feature, generally with negligible interaction with the other features (see also below and figure 8).
+\x We expanded substitution models to {also detect quotes undergoing up to two substitutions}, which yielded the same results as single substitutions {(see below)}. 
 }
@@
 A Goodman-based multinomial goodness-of-fit shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
-
 }
@@
 
-\answer{The case presented here is valid and could indeed hide a bias in our measures.
+\answerok{The case presented here is valid and could indeed hide a bias in our measures.
 Rather than analyzing all possible such cases to bound the error, we extended substitution models to detect quotes undergoing up to two substitutions, and applied the same analysis to these mined substitutions.
@@
 
-\answer{Fig.
-4 of the previous manuscript represented probabilities \emph{not knowing which sentence the word appears in} (i.e.
-marginal probabilities), which cannot be used to ask questions given a particular sentence.
+\answerok{\tb{The figure 4 in the previous manuscript represented probabilities \emph{not knowing which sentence the word appears in} (i.e.
+marginal probabilities), which cannot be used to ask questions given a particular sentence}.
 
-Since such a measure is indeed not very useful for practical interpretation, we refined the definition of susceptibility to make it more straightforward.
-It is now defined relative to a null hypothesis and measures the strength of a bias for a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly); it is presented in Figs.
+\tb{Since it appears indeed that such a measure is relatively difficult to connect to} a practical interpretation, we redefined the definition of susceptibility to make it more straightforward.
+\tb{We now define it relatively} to a null hypothesis\tb{: it} measures the strength of a bias \tb{affecting} a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly).
+\tb{We present the corresponding results} in Figs.
 5 \& 6.
 
-Finally, we extended our measures to capture part of the context of the sentence substitutions occur in (see the ``Sentence context'' subsection).
-Sentence-relative susceptibility, in particular, gives a straightforward answer to the question of which words are more likely to attract substitution in a given sentence.
+Finally, we also extended our measures to capture part of the \tb{sentence-level context in which substitutions occur} (see the ``Sentence context'' subsection).
+Sentence-relative susceptibility, in particular, gives a straightforward answer to the question of which words are more likely to attract substitution \tb{with}in a given sentence.
 }
@@
 
-\answerok{We added a \tb{short} discussion on the semantics of substituted quotes.
+\answerok{We added a short discussion on the semantics of substituted quotes.
 We first aimed at quantifying the semantic distance by estimating the distance ``traveled'' by the substituted word in the WordNet hypernym-hyponym network as well as the FA network.
-\tb{This showed that substituted and substituting words are rarely direct synonyms, yet they usually achieve a similar meaning in the context of their sentence.
-To further grasp the relationship of substitutions to synonyms we introduced a second null hypothesis in the ``Variation'' section, which compares the actual variation to what would happen if a random synonym of the start word was picked instead.}
+This showed that substituted and substituting words are rarely direct synonyms, yet they usually achieve a similar meaning in the context of their sentence.
+To further grasp the relationship of substitutions to synonyms we introduced a second null hypothesis in the ``Variation'' section, which compares the actual variation to what would happen if a random synonym of the start word was picked instead.
 
 Finally, we also examined the effect of part-of-speech (POS): Fig.
-5 details the results on five POS categories, \tb{and the subsequent discussion} shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
+5 details the results on five POS categories, and the subsequent discussion shows that \tb{POS categories do not capture any significant bias} in the selection of substitution targets, which is why we largely keep them outside of our analysis.
 }
@@
 \answerok{Thank you for this question.
-We introduced a simple regression model to check for feature interactions, \tb{which is presented in Fig.
-8 with a corresponding discussion} (in a nutshell: we show that features generally do not interact).}
+We introduced a simple regression model to check for feature interactions, which is presented in Fig.
+8 with a corresponding discussion (in a nutshell: we show that features generally do not interact).}
 
@@
 
-\answer{We hope the improved introduction is clearer as to what we predict and what is left open due to the exploratory nature of our study.}
+\answerok{We hope the improved introduction is clearer as to what we \tb{expect} and what is left open due to the exploratory nature of our study.}
 
@@
 
-\answer{The data in Figs.
-4 \& 5 was binned into 5 integer-based bins, and the graphs were cropped whenever the confidence intervals were unreasonably large (which gave the impression of some graphs having only 4 bins).
+\answerok{Data in Figs.
+4 \& 5 was binned into 5 integer-based bins, and the graphs were cropped whenever \tb{the data was unreasonably scarce and, accordingly,} confidence intervals were unreasonably large (which gave the impression of some graphs having only 4 bins).
 
-The data in the new Figs.
+\tb{Notwithstanding,} data in the new Figs.
 6, 7 \& 9 is now binned into 4 quartile-based bins, which makes more sense for the confidence intervals.
@@
 
-\answer{We understand that the analysis we present is not the approach usually taken for similar laboratory data.
-The rewritten introductory sections of this revision now explain why we chose to use a feature-based analysis over the prediction of individual words (see below for more about this).
-
-Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (footnote \#16).
-We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is discussed in that section.
+\answerok{The rewritten introductory sections of this revision now \tb{aim at explaining} why we chose to use a feature-based analysis over the prediction of individual words (see below for more details about this).
+Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (\tb{footnote \#15}).
+We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is being discussed in that section.
 
 Finally, we introduced a new null hypothesis ($\mathcal{H}_{00}$) in the ``Variation'' subsection, which gives insight into how an appearing word relates to the synonyms of the word it replaces.
+\tb{While we understand that the analysis we present differs from typical approaches usually taken for similar laboratory data, we hope that these various additions contribute to alleviate concerns regarding word-word similarity.}
 }
@@
 
-\answer{While we agree that predicting the actual disappearing and appearing words would be an extremely interesting result, we believe it is several steps away from what is currently possible.
-As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
+\answer{We agree that predicting the actual disappearing and appearing words would be an extremely interesting result, \tb{yet we believe that this prospect stands} several steps away from what is currently possible.
+\tb{As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
 Zaromb et al.
@@
 Finally, as explained in footnote \#4 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.
-}
+}}
 
@@
 
-\answer{This would indeed be a very interesting project and prospect, but the data we collected do not enable the study of the surrounding context (in particular, we know neither the prior usage nor the way information sources may be used) so that this goal remains beyond the scope of our paper.
+\answerok{This would indeed be a very interesting project and prospect, but the data we collected do not enable the study of the surrounding context (in particular, we know neither the prior usage nor the way information sources may be used) so that this goal remains beyond the scope of our paper.
 
@@
 
-\answer{The goal of our feature selection is to render the results more intelligible.
+\answerok{\tb{Our feature selection step aims at making} the results more intelligible.
 This section was indeed poorly written, and has been thoroughly revised to rationalize the process.
-With few exceptions (notably number of phonemes and number of syllables less biased than number of letters), agglomerated features exhibit very similar behaviors across all measures (graphs produced for all features may be found at <https://huit.re/aEH77Gc3>, and could be included in a Supplementary Information bundle if requested).
+\tb{In particular, we identify (six) groups of strongly correlated features (both manually by visual inspection and automatically with the help of an agglomeration algorithm, yielding identical results) and justify the choice of our six prototype features for each group.
+Features belonging to the same group generally} exhibit very similar behaviors across all measures \tb{(with few exceptions, notably number of phonemes and number of syllables less biased than number of letters)}.
+Graphs produced for all features may be found at <https://huit.re/aEH77Gc3>, and could be included in a Supplementary Information bundle if requested.
 }
@@
 \answerok{This is an important empirical point which we did not address systematically in the first submission.
-We have now done this manual coding step, \tb{and improved filters wherever precision or recall were not good enough.
-Accuracy is now assessed and satisfying in the relevant cases.}}
+We have now done this manual coding step, and improved filters wherever precision or recall were not good enough.
+\tb{We now assess filtering accuracy systematically and it is} satisfying in the relevant cases.}
 
@@
 
-\tb{A few additional numbers have changed throughout the manuscript:
+\answerok{A few additional numbers have changed in this revised manuscript:
 \itx{
@@
 As a result of this, the number of words coded by Word Frequency has also changed (since frequency of words is computed from the set of quotes we keep after cluster filtering).
-\x The discovery of three bugs, and the improvement of substitution filtering, led us to mine many more substitutions than previously.
-All in all the software is now much more reliable, as unit tests cover nearly all of it.
+\x \tb{The improvement of substitution filtering}, led us to mine many more substitutions than previously.
+All in all \tb{raw data processing} is now much more reliable, as unit tests cover nearly all of it.
 }

### 2016-08-20 brains-conclusion.tex
pre-final round of modifications manuscript + cover

@@
 Such interactions have been shown to underlie the lineage specificity observed in transmission chains~\citep{claidiere_cultural_2014}.
-Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
-In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge, and \tb{help assess the potential causal effect they have on subsequent evolution}.
+Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the \tb{co}evolution of both cultural content and the ways in which it is transformed.
+In our particular case, such insight could shed some light on how the feature attractors \tb{examined} in this paper actually emerge, and help assess their potential \tb{role on this coevolution}.
 

### 2016-08-20 brains-introduction.tex
pre-final round of modifications manuscript + cover

@@
 The reciprocal influence between cognition and culture has a long history in both social science and psychology.
-While this question has been the subject of intense debate in the social sciences in the 20th century, today's discussion is mostly structured by proponents from cognitive science\tb{, who} construe culture as an evolutionary process analogous and parallel to biological evolution.
+While this question has been the subject of intense debate in the social sciences in the 20th century, today's discussion is mostly structured by proponents from cognitive science, who construe culture as an evolutionary process analogous and parallel to biological evolution.
 That analogy can be traced a long way back, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\citeyearpar{dawkins_selfish_1976}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
@@
 \newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
-The first above-mentioned method consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
-\tb{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the importance of the originating cultural systems in the interpretation of such data.
+The first above-mentioned method consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).
+%\footnote{{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the importance of the originating cultural systems in the interpretation of such data.
 %have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
-}}
+%}}
 A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.

### 2016-08-20 brains-protocol.tex
pre-final round of modifications manuscript + cover

@@
 \newtext{We checked that this restriction does not bias the results discussed below by extending our protocol to two-substitution transformations.
-The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.}
+\tbn{The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.}}
 }
@@
 Substitutions usually replace a word with another semantically related word:
-\tb{manual observation of a random subset of 100 substitutions shows that, compared to the word it replaces,} the new word often achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
+manual observation of a random subset of 100 substitutions shows that, compared to the word it replaces, the new word often achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
 The following examples illustrate this phenomenon:\footnote{
@@
     \caption{\newtext{\textbf{POS-related results:}
-    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories \tb{essentially made of stopwords} (see main text for details).
+    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \tb{\emph{S} means \emph{Stopword-like}} %together all the categories \tb{essentially made of stopwords}
+    (see main text for details).
     The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
@@
 Further, to avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
-\tb{That is, each quote cluster has a maximum contribution of 1, computed as the average contribution of all substitutions in that cluster.}
+That is, each quote cluster has a maximum contribution of 1, computed as the average contribution of all substitutions in that cluster.
 }}
@@
 \newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions,
-\hbox{i.e.} \newtext{it measures the preferential selection of some word properties \tb{for} substitution.}
+\hbox{i.e.} \newtext{it measures the preferential selection of some word properties for substitution.}
 %\newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
@@
 
-Surprisingly the results for this measure are no different from the context-free measure, as can be seen in Fig.~\ref{fig:feature-susceptibilities-in_sentence} of the Appendix:
+Surprisingly the results for this measure are no different from the context-free measure (as can be seen in Fig.~\ref{fig:feature-susceptibilities-in_sentence} of the Appendix):
 low-frequency words compared to the sentence are substituted much more than higher-frequency words, words learned earlier than the rest of the sentence are substituted less than words learned later, shorter words less than longer words, and words with scarce neighborhoods slightly more than words with denser neighborhoods.

### 2016-08-23 reviews/2016-07-response-to-reviewers.tex
Fixes after Camille's comments and email

@@
 
-We \tb{had also overlooked} several developments in the iterated learning literature, and are grateful for the \tb{various} pointers.
+We had also overlooked several developments in the iterated learning literature, and are grateful for the various pointers.
 The works we thought were most relevant have been integrated to the introductory sections, and their implications for our work are discussed in the new ``Discussion'' section.
@@
 \itx{
-\x \tb{We believe that} the introductory parts provide a clearer presentation of our question and how it relates to existing literature (including the areas we overlooked) and hypotheses
-\x The new ``Discussion'' section, which was indeed missing, \tb{should} properly reconnect our results to the related works introduced at the beginning of the paper.
+\x We believe that the introductory parts provide a clearer presentation of our question and how it relates to existing literature (including the areas we overlooked) and hypotheses
+\x The new ``Discussion'' section, which was indeed missing, should properly reconnect our results to the related works introduced at the beginning of the paper.
 }
 
-Finally, several modifications in the ``Results'' section (more examples, a short discussion of the distance traveled by substitutions on the synonym network in footnote \#16, and the addition of the second null hypothesis $\mathcal{H}_{00}$ in the ``Variation'' subsection) should better describe the type of changes undergone by quotes.
+Finally, several modifications in the ``Results'' section (more examples, a short discussion of the distance traveled by substitutions on the synonym network in footnote \#15, and the addition of the second null hypothesis $\mathcal{H}_{00}$ in the ``Variation'' subsection) should better describe the type of changes undergone by quotes.
 }
@@
 \x We checked and discussed the susceptibility of part-of-speech (POS).
-While the open-/closed-class question is \tb{largely} precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), if we note that the manuscript's \tb{\emph{Stopword-like}} category is mainly made of \tb{closed-class words}, the data indicates that they are indeed less substituted than open-class words.
+While \tb{a rigorous treatment of} the open-/closed-class question is largely precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), \tb{the data does indeed indicate that closed class words are less substituted than open class words (the \emph{Closed class-like} category is mainly made of \tb{closed class words})}.
 As the manuscript discusses however, this concerns less than 7\% of the substitutions we consider, and has likely no effect on our results.
@@
 
-\answerok{We clarified the notion of dynamical systems when rewriting the introduction and in the ``Variation'' section (including footnote \#22).}
+\answerok{We clarified the notion of dynamical systems when rewriting the introduction and in the ``Variation'' section (including footnote \#21).}
 
@@
 
-\answerok{\tb{The figure 4 in the previous manuscript represented probabilities \emph{not knowing which sentence the word appears in} (i.e.
-marginal probabilities), which cannot be used to ask questions given a particular sentence}.
+\answerok{\tb{The measure we had proposed in the first version of the manuscript (for information, this was the subject of Fig.
+4 in that manuscript) considered probabilities of substitutions as an aggregate quantity (and not at the individual sentence level).
+It was difficult to use it to understand what happens when substitution occurs in a given sentence.}
 
-\tb{Since it appears indeed that such a measure is relatively difficult to connect to} a practical interpretation, we redefined the definition of susceptibility to make it more straightforward.
-\tb{We now define it relatively} to a null hypothesis\tb{: it} measures the strength of a bias \tb{affecting} a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly).
-\tb{We present the corresponding results} in Figs.
+\tb{We therefore reworked the notion of susceptibility to make it more straightforward}.
+We now define it relatively to a null hypothesis: it measures the strength of a bias affecting a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly).
+We present the corresponding results in Figs.
 5 \& 6.
 
-Finally, we also extended our measures to capture part of the \tb{sentence-level context in which substitutions occur} (see the ``Sentence context'' subsection).
-Sentence-relative susceptibility, in particular, gives a straightforward answer to the question of which words are more likely to attract substitution \tb{with}in a given sentence.
+Finally, we also extended our measures to capture part of the sentence-level context in which substitutions occur (see the ``Sentence context'' subsection).
+Sentence-relative susceptibility, in particular, gives a straightforward answer to the question of which words are more likely to attract substitution within a given sentence.
 }
@@
 Finally, we also examined the effect of part-of-speech (POS): Fig.
-5 details the results on five POS categories, and the subsequent discussion shows that \tb{POS categories do not capture any significant bias} in the selection of substitution targets, which is why we largely keep them outside of our analysis.
+5 details the results on five POS categories, \tb{with a discussion of the corresponding effects in the main text, explaining} why we largely keep POS outside of our analysis.
 }
@@
 
-\answerok{We hope the improved introduction is clearer as to what we \tb{expect} and what is left open due to the exploratory nature of our study.}
+\answerok{We hope the improved introduction is clearer as to what we expect and what is left open due to the exploratory nature of our study.}
 
@@
 \answerok{Data in Figs.
-4 \& 5 was binned into 5 integer-based bins, and the graphs were cropped whenever \tb{the data was unreasonably scarce and, accordingly,} confidence intervals were unreasonably large (which gave the impression of some graphs having only 4 bins).
+4 \& 5 was binned into 5 integer-based bins, and the graphs were cropped whenever the data was unreasonably scarce and, accordingly, confidence intervals were unreasonably large (this cropping gave the impression of some graphs having only 4 bins).
 
-\tb{Notwithstanding,} data in the new Figs.
-6, 7 \& 9 is now binned into 4 quartile-based bins, which makes more sense for the confidence intervals.
+Notwithstanding, data in the new Figs.
+6, 7, 9 \& 10 is now binned into 4 quartile-based bins, which makes more sense for the confidence intervals.
 This decision was reached after exploring several other binning configurations: none of these substantially changed the results, but opting for quartile-based bins does improve the overall readability.
@@
 A tool by Recchia \& Jones (2010, BRM) allows for pointwise mutual information (PMI) of different words to be retrieved from a given corpus very simply.
-A tool such as this would allow for word-word similarity to be included in the analysis.}
+A tool such as this would allow for word-word similarity to be included in the analysis.
 
-\answerok{The rewritten introductory sections of this revision now \tb{aim at explaining} why we chose to use a feature-based analysis over the prediction of individual words (see below for more details about this).
-Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (\tb{footnote \#15}).
-We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is being discussed in that section.
-
-Finally, we introduced a new null hypothesis ($\mathcal{H}_{00}$) in the ``Variation'' subsection, which gives insight into how an appearing word relates to the synonyms of the word it replaces.
-\tb{While we understand that the analysis we present differs from typical approaches usually taken for similar laboratory data, we hope that these various additions contribute to alleviate concerns regarding word-word similarity.}
-}
-
-\critique{I understand that word-word metrics are not easily introduced in the type of analysis that the authors are attempting to do, where the analysis of the contributions of different global psycholinguistic variables are being weighed against each other.
+I understand that word-word metrics are not easily introduced in the type of analysis that the authors are attempting to do, where the analysis of the contributions of different global psycholinguistic variables are being weighed against each other.
 I just think that type of analysis is inappropriate here.
@@
 
-\answer{We agree that predicting the actual disappearing and appearing words would be an extremely interesting result, \tb{yet we believe that this prospect stands} several steps away from what is currently possible.
-\tb{As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
+\answerok{The rewritten introductory sections of this revision now aim at explaining why we chose to use a feature-based analysis over the prediction of individual words (see below for more details about this).
+Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (footnote \#15).
+We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is being discussed in that section.
+
+\tb{Regarding prediction, we would like to distinguish two questions.
+First, prediction of the actual disappearing and appearing words, which would be an extremely interesting result, but a prospect we believe to stand several steps away from what is currently possible.
+As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
 Zaromb et al.
@@
 To our knowledge, the prediction of intrusions in full sentences (either manually designed or not) has not been tackled.
-Finally, as explained in footnote \#4 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.
-}}
+Finally, as explained in footnote \#3 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.}
+
+\tb{These reasons led us to focus on the second question: prediction of properties of the disappearing and appearing words.
+Indeed, Figs.
+7 \& 10 provide average predictions of appearing word features given the properties of a disappearing word.
+We show that there are little to no interactions between word features (see Fig.
+8), such that word frequency is not sufficient to predict variation of other features.
+Finally, we introduced a new null hypothesis ($\mathcal{H}_{00}$) in the ``Variation'' subsection which gives insight into how an appearing word relates to the synonyms of the word it replaces, and shows that appearing words are much more frequent than the average frequency of synonyms of the disappearing word.}
+
+While we understand that the analysis we present differs from typical approaches usually taken for similar laboratory data, we hope that these various additions contribute to alleviate concerns regarding word-word similarity.
+}
+
+%\answer{We agree that predicting the actual disappearing and appearing words would be an extremely interesting result, \tb{yet we believe that this prospect stands} several steps away from what is currently possible.
+%\tb{As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
+Zaromb et al.
+(2006), for instance, predict the list from which prior-list intrusions come, but not the exact intruding word.
+To our knowledge, the prediction of intrusions in full sentences (either manually designed or not) has not been tackled.
+Finally, as explained in footnote \#3 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.
+%}}
 
@@
 
-\answerok{\tb{Our feature selection step aims at making} the results more intelligible.
+\answerok{Our feature selection step aims at making the results more intelligible.
 This section was indeed poorly written, and has been thoroughly revised to rationalize the process.
-\tb{In particular, we identify (six) groups of strongly correlated features (both manually by visual inspection and automatically with the help of an agglomeration algorithm, yielding identical results) and justify the choice of our six prototype features for each group.
-Features belonging to the same group generally} exhibit very similar behaviors across all measures \tb{(with few exceptions, notably number of phonemes and number of syllables less biased than number of letters)}.
+In particular, we identify (six) groups of strongly correlated features (both manually by visual inspection and automatically with the help of an agglomeration algorithm, yielding identical results) and justify the choice of our six prototype features for each group.
+Features belonging to the same group generally exhibit very similar behaviors across all measures (with few exceptions, notably number of phonemes and number of syllables \tb{exhibiting less cognitive bias} than number of letters).
 Graphs produced for all features may be found at <https://huit.re/aEH77Gc3>, and could be included in a Supplementary Information bundle if requested.
@@
 We have now done this manual coding step, and improved filters wherever precision or recall were not good enough.
-\tb{We now assess filtering accuracy systematically and it is} satisfying in the relevant cases.}
+We now assess filtering accuracy systematically and it is satisfying in the relevant cases.}
 
@@
 As a result of this, the number of words coded by Word Frequency has also changed (since frequency of words is computed from the set of quotes we keep after cluster filtering).
-\x \tb{The improvement of substitution filtering}, led us to mine many more substitutions than previously.
-All in all \tb{raw data processing} is now much more reliable, as unit tests cover nearly all of it.
+\x The improvement of substitution filtering led us to mine many more substitutions than previously. All in all raw data processing is now much more reliable, as unit tests cover nearly all of it.
 }

### 2016-08-23 brains-conclusion.tex
Fixes after Camille's comments and email

@@
 Such interactions have been shown to underlie the lineage specificity observed in transmission chains~\citep{claidiere_cultural_2014}.
-Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the \tb{co}evolution of both cultural content and the ways in which it is transformed.
-In our particular case, such insight could shed some light on how the feature attractors \tb{examined} in this paper actually emerge, and help assess their potential \tb{role on this coevolution}.
+Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the coevolution of cultural content and the ways in which it is transformed.
+In our particular case, such insight could shed some light on how the feature attractors examined in this paper actually emerge, and help assess their potential role on this coevolution.
 

### 2016-08-23 brains-introduction.tex
Fixes after Camille's comments and email

@@
 
-We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, \tb{do cultural attractors appear and how do cognitive biases participate in them?}
+We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, \tb{do cultural attractors appear, and if so how do cognitive biases participate in them?}
 %how do cognitive biases participate in \tb{the appearance of cultural attractors}?

### 2016-08-23 brains-protocol.tex
Fixes after Camille's comments and email

@@
 
-\newtext{We also consider grammatical types within quotations by detecting \emph{Part-of-Speech} (POS) categories with TreeTagger~\citep{schmid_probabilistic_1994}; we distinguish between verbs, nouns, adjectives, adverbs, and stopword-like words.}
+\newtext{We also consider grammatical types within quotations by detecting \emph{Part-of-Speech} (POS) categories with TreeTagger~\citep{schmid_probabilistic_1994}; we distinguish between verbs, nouns, adjectives, adverbs, and closed class-like words.}
 
@@
 Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
-The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.% (the analysis on the complete set of features can be found in the appendix).
+The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered}.
 \footnote{
@@
 \end{new}
+
 \begin{figure*}[h]
@@
 and (b) greatly increase the complexity of the transformation models used to make these inferences.\footnote{
-\newtext{We checked that this restriction does not bias the results discussed below by extending our protocol to two-substitution transformations.
-\tbn{The results and graphics for the 16 additional models involved are available in the code repository for this paper: \url{https://github.com/wehlutyk/brainscopypaste}.}}
+\newtext{We checked that this restriction does not bias the results discussed below by extending our protocol to two-substitution transformations;
+the results were unchanged.}
 }
@@
     \caption{\newtext{\textbf{POS-related results:}
-    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \tb{\emph{S} means \emph{Stopword-like}} %together all the categories \tb{essentially made of stopwords}
+    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \tb{\emph{C} means \emph{Closed class-like}}
     (see main text for details).
@@
 Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
-A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Stopword-like}\footnote{
-\newtext{The \emph{Stopword-like} category gathers all the POS groups \tb{essentially made of stopwords (coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections), which are affected by our decision to exclude stopwords from the analysis.
-Indeed substitutions falling on stopwords are filtered out (affecting $s$), and stopwords are never counted as substitutable in the computation of $s^0$.
-Both numbers are subsequently very small for these categories (which led us to group them together) but the reported susceptibility is left unbiased.}}
+A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Closed class-like}\footnote{
+\newtext{The \emph{Closed class-like} category gathers all the POS groups \tb{representing closed class words (coordinating conjunctions, prepositions, subordinating conjunctions, modals and possessive endings).
+These groups, essentially made of stopwords, feature very low counts for both $s$ (substitutions falling on stopwords are filtered out) and $s^0$ (stopwords are never counted as substitutable).
+While the susceptibility reported for the remaining words is left unbiased (as $s$ and $s^0$ are equally affected), they represent a very small portion of all substitutions, which led us to group them together.
+Finally, we added to this meta-category the few POS groups that cover words entirely excluded from the analysis (foreign words, punctuation symbols and interjections), only sporadically present because of tagging fluctuations;
+hence the name \emph{Closed class-like}.
+}}
 }
 and \emph{Adverb} categories.
-Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for stopword-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
+Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for closed class-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
 none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
-While we acknowledge the low susceptibilities of adverbs and stopword-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
+While we acknowledge the low susceptibilities of adverbs and closed class-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
 it seems, then, that POS categories do not capture any strong bias in the selection of substitution targets.
@@
 
+%\begin{figure*}[!th]
+%	\centering
+%	\begin{subfigure}{.49\textwidth}
+%	    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
+%	    \caption{\newtext{\textbf{Sentence-relative feature variation:}
+%	    $\nu_{\phi_r}$, average sentence-relative feature value of the appearing word as a function of the sentence-relative value of the disappearing word (binned by quartiles), with 95\% asymptotic confidence intervals based on Student's \emph{t}-distribution.
+%	    $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ are similarly converted to be sentence-relative.
+%	    Attraction, magnitude and direction of bias with respect to null hypotheses are similar to Fig.~\ref{fig:feature-variations-global}.
+%	    However, attractors are always positioned between sentence median ($y=0$) on one side and $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ on the other side.
+%	    Clustering coefficient, Number of synonyms and Orthographic neighborhood density are limit cases, with triple intersections with one of the null hypothesis curves.}
+%	    }
+%	    \label{fig:feature-variations-sentencerel}
+%	\end{subfigure}
+%	~
+%	\begin{subfigure}{.49\textwidth}
+%	    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
+%	    \caption{\newtext{\textbf{Substitution susceptibility for in-quote feature quartiles:}
+%	    susceptibility to substitution versus quartile of the feature distribution in the originating quote, with 95\% asymptotic confidence intervals (Goodman-based multinomial)}.
+%	    }
+%	    \label{fig:feature-susceptibilities-in_sentence}
+%	\end{subfigure}
+%\end{figure*}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
+    \caption{\newtext{\textbf{Substitution susceptibility for in-quote feature quartiles:}
+    susceptibility to substitution versus quartile of the feature distribution in the originating quote, with 95\% asymptotic confidence intervals (Goodman-based multinomial)}.
+    }
+    \label{fig:feature-susceptibilities-in_sentence}
+\end{figure}
+
 \begin{figure*}[!th]
@@
 
-Surprisingly the results for this measure are no different from the context-free measure (as can be seen in Fig.~\ref{fig:feature-susceptibilities-in_sentence} of the Appendix):
+Surprisingly the results for this measure are no different from the context-free measure, as can be seen in Fig.~\ref{fig:feature-susceptibilities-in_sentence}:
 low-frequency words compared to the sentence are substituted much more than higher-frequency words, words learned earlier than the rest of the sentence are substituted less than words learned later, shorter words less than longer words, and words with scarce neighborhoods slightly more than words with denser neighborhoods.

### 2016-08-23 reviews/2016-07-response-to-reviewers.tex
Editorial changes after full proof-read (after a long rest)

@@
 All the reviewers agreed on this point (asking for more clarification about feature selection, why certain choices were made in the analysis, etc…).
-This ranges from what types of quotes were included, to the rules to detecting substitution errors, to the inclusion of certain features in your analyses.
+This ranges from what types of quotes were included, to the rules for detecting substitution errors, to the inclusion of certain features in your analyses.
 The reader needs to be convinced that the few “strong” looking patterns you do find are not too dependent on these arbitrary choices.
@@
 \x We rationalized feature selection, adding some variables and much detail to this section
-\x In general, we clarified further the assumptions we made, such as the restriction to single substitutions
-\x We also introduced manually-coded precision \& recall analyses for all filters, which justifies our choices.
+\x In general, we clarified further the assumptions and decisions made, such as the restriction to single substitutions
+\x We also introduced manually-coded precision \& recall analyses for all filters, justifying relevant choices.
 }
@@
 \answerok{Thank you for this very constructive and extensive feedback.
-We paid signification attention to the writing style in this new version, and large portions of the document have been rewritten and expanded --- all updates are marked in dark blue.}
+We paid significant attention to the writing style in this new version, and large portions of the document have been rewritten and expanded --- all updates are marked in dark blue.}
 
@@
 
-\answerok{We have been careful to introduce many more connections to the relevant literature: the introduction has been entirely rewritten and is followed by a more thorough presentation of the related work in cognitive science, while the concluding sections (discussion \& concluding remarks) follow your recommendation of connecting our results with the literature.
+\answerok{We have been careful to introduce many more connections to the relevant works in cognitive science: the introduction has been entirely rewritten and is followed by a more thorough presentation of the related work, while the concluding sections (discussion \& concluding remarks) follow your recommendation of connecting our results with the literature.
 }
@@
 
-\critique{ The experiment design also has some substantial flaws.
+\critique{The experiment design also has some substantial flaws.
 First, some crucial factors were not considered - most notably, part of speech.
@@
 \answerok{
-We addressed all these points in the revised version:
+We addressed these points in the revised version:
 \itx{
@@
 
-\answerok{We clarified the notion of dynamical systems when rewriting the introduction and in the ``Variation'' section (including footnote \#21).}
+\answerok{The rewritten ``Variation'' subsection (including footnote \#21) should now clarify this notion.}
 
@@
 This showed that substituted and substituting words are rarely direct synonyms, yet they usually achieve a similar meaning in the context of their sentence.
-To further grasp the relationship of substitutions to synonyms we introduced a second null hypothesis in the ``Variation'' section, which compares the actual variation to what would happen if a random synonym of the start word was picked instead.
+To further grasp the relationship of substitutions to synonyms we introduced a second null hypothesis in the ``Variation'' subsection, which compares the actual variation to what would happen if a random synonym of the start word were picked instead.
 
@@
 
-\answerok{Thank you for this question.
-We introduced a simple regression model to check for feature interactions, which is presented in Fig.
+\answerok{We introduced a simple regression model to check for feature interactions, which is presented in Fig.
 8 with a corresponding discussion (in a nutshell: we show that features generally do not interact).}
@@
 We initially favored this solution over first presenting the final list of candidate features and justifying afterwards that they are anyway correlated with a broader list of traditional variables.
-(Note: An alternative option could consist in presenting the contents of Figs.
+(Note: an alternative option could consist in presenting the contents of Figs.
 1 and 2 in the same figure, if this would be less confusing.)}
@@
 
-However, the revised version now features a much broader framework to statistically assess joint effects (see Fig.
+However, the revised version now features a broader analysis assessing joint effects (see Fig.
 8); we think it favorably replaces the previous variable-to-variable correlation check.
-(Note: we also replaced the number of phonemes with the number of letters -- which makes more sense in our context, see the end of the ``Word-level features'' subsection --, which does not seem to interact with number of synonyms).}
+(Note: we also replaced number of phonemes with number of letters, which makes more sense in our context -- see the end of the ``Word-level measures'' subsection.
+The latter does not seem to interact with number of synonyms.)}
 
@@
 \answerok{Thank you for these very useful and, indeed, relevant pointers.
-We accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' section).}
+We accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' and ``Discussion'' sections).}
 
@@
 \answerok{The rewritten introductory sections of this revision now aim at explaining why we chose to use a feature-based analysis over the prediction of individual words (see below for more details about this).
-Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (footnote \#15).
+Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (and in footnote \#15).
 We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is being discussed in that section.
@@
 \answerok{This is an important empirical point which we did not address systematically in the first submission.
-We have now done this manual coding step, and improved filters wherever precision or recall were not good enough.
+We have now done this manual coding step, and have improved filters wherever precision or recall (whichever was most relevant) were not good enough.
 We now assess filtering accuracy systematically and it is satisfying in the relevant cases.}

### 2016-08-23 brains-conclusion.tex
Editorial changes after full proof-read (after a long rest)

@@
 
-We initially aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
+We initially aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes gradually transformed by low-level biases.
 The data set we used imposed a few constraints on our analysis:
-first, it was necessary to infer source-destination links, an operation made more reliable when restricting the scope of transformations to very simple cases, which we did by focusing on word substitutions.
+first, it was necessary to infer source-destination links, an operation made more reliable when restricting the scope of transformations to very simple cases, which we did by focusing on single word substitutions.
 Second, contrary to laboratory experiments which produce data made of many repeated measures on a small number of cases (\hbox{e.g.} a given list of words), we have a great number of different cases (one case per cluster in which substitutions are found, \hbox{i.e.} 698 cases), with very few measures on each of them (average 9, median 5).
@@
 Further characterizing the substitutions by examining the variation of word features from disappearing to appearing words, we show:
-(1) that the operation is contractile on average, that is words are brought closer to an attractor point on each feature;
-(2) that authors produce words that are easier to remember than the average of synonyms of the disappearing word (a fact that is reflected in the position of the attraction point).
+(a) that the operation is contractile on average, that is words are brought closer to an attractor point on each feature;
+(b) that authors produce words that are easier to remember than the average of synonyms of the disappearing word (a fact that is reflected in the position of the attraction point).
 
@@
 Indeed the limits of the data set do not allow us to infer chains of substitutions, and substitutions themselves are not the only type of transformation at work in the data set.
-Nonetheless, these findings (1) bring light to this simple type of transformation, and (2) are consistent with known psycholinguistic effects, with the hypothesis of cultural attractors in representations from everyday life, and with the lineage specificity discussed in the iterated learning literature~\citep{claidiere_cultural_2014,cornish_systems_2013}.
+Nonetheless, these findings (a) bring light to this simple type of transformation, and (b) are consistent with known psycholinguistic effects, with the hypothesis of cultural attractors in representations from everyday life, and with the lineage specificity discussed in the iterated learning literature~\citep{claidiere_cultural_2014,cornish_systems_2013}.
 They are obtained by successfully applying knowledge from cognitive science to real-life complex data, a task that remains a challenge in the study of cultural evolution.
@@
 We aimed to contribute to \newtext{testing this hypothesis} by studying a simple everyday-life task where individuals are implicitly trying to reproduce quotations.
-To some extent, our work amounts to \newtext{an out-of-laboratory} experiment where we \newtext{examine the influence} of \newtext{well-known word features} on the accuracy of the reproduction of short sentences.
+To some extent, our work amounts to \newtext{an out-of-laboratory} experiment where we \newtext{examine the influence} of \newtext{well-known word features} on the accuracy of reproduction of short sentences.
 \newtext{Our analysis of substitutions shows that words are attracted, in each dimension, to feature-specific values.
 Furthermore, the features' known effects in psycholinguistic experiments are reflected in the biases of these attraction points, meaning that the evolution of such quotations can be partially explained by known low-level cognitive biases.
-We believe that such an approach, combining psycholinguistic knowledge and data mining tools, can be fruitfully developed to improve the study of cultural attractors and explore the reciprocal influences of cognition and culture.}
+We believe that such an approach, which combines psycholinguistic knowledge and data mining tools, can be fruitfully developed to improve the study of cultural attractors and explore the reciprocal influences of cognition and culture.}
 
@@
 In biology in particular, work on evo-devo and non-genetic inheritance has accumulated evidence that is poorly accounted for by the modern synthesis of biological evolution, and is creating a demand for new or extended  approaches to joint cultural and biological evolution~\citep[see][for instance]{gilbert_eco-evo-devo:_2015}.
-Such an approach has long been called upon by anthropologists like~\citet{ingold_beyond_2004,ingold_three_1999}, in line with Mauss' initial works~\citep{mauss_les_1936}, and the debate is not entirely foreign to the enactive-representational discussion in cognitive science.
+Such an approach has long been called upon by anthropologists like~\citet{ingold_beyond_2004,ingold_three_1999}, in line with Mauss' initial works~\citep{mauss_les_1936}, and the question is not entirely foreign to the enactive-representational debate in cognitive science.
 %As regards our particular case, a most promising line of work on the evolution of language is being developed by \citet{cuffari_participatory_2014}.

### 2016-08-23 brains-introduction.tex
Editorial changes after full proof-read (after a long rest)

@@
 
-The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
+The theory of Epidemiology of Representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
 these are alternatively located in the mind (``mental representations'' in Sperber's terminology), or in the outer world (``public representations'') as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
@@
 For instance the quote ``we will not be scared of these cowards'' (a substring of a quote from former Pakistani President Asif Ali Zardari) is also found as ``we will not be \textbf{afraid} of these cowards''.
-More meaningful changes often happen too, such as the transformation of McCain's ``I admire Senator Obama and his accomplishments'' during the 2008 U.S.
-presidential campaign, into ``I \textbf{respect} Senator Obama and his accomplishments''.
+More meaningful changes often happen too, such as the transformation of McCain's ``I admire Senator Obama and his accomplishments'' during the 2008 US presidential campaign, into ``I \textbf{respect} Senator Obama and his accomplishments''.
 Since authors are implicitly required to copy quotes exactly, we can assume that most transformations, especially simple ones such as those shown above, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.
@@
 
-Compared to full sentences, recall of word lists provides a situation that is easier to fully explore, and that has been extensively studied.
+Compared to full sentences, recall of word lists provides a situation that is easier to fully explore and has been extensively studied.
 In particular, the Deese, Roediger, and McDermott paradigm (introduced by \citealp{deese_prediction_1959}, and later popularized by \citealp{roediger_creating_1995}) has shown that it is possible to construct lists of words which reliably create the false memory of an external word related to those in the list.
@@
 
-\newtext{Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
+\newtext{Given this complexity we decided to focus on more aggregate measures, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
 \newtext{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited to data consisting of a high number of measures over a smaller number of lists (in which case it makes sense to ask \hbox{e.g.} what proportion of intrusions come from prior lists).

### 2016-08-23 brains-protocol.tex
Editorial changes after full proof-read (after a long rest)

@@
 Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
-First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
+First, they are usually cleanly delimited by quotation marks, which greatly facilitates their detection in text corpora.
 Second, they stem from a unique original version, and are ideally traceable back to that version.
@@
 Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
-Our goal, therefore, is to show that such effects exist and are measurable even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
+Our goal, therefore, is to show that cognitive biases have measurable effects in this setting even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
 %\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
@@
 \newtext{To compute these features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
-As \citet{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
+As \citeauthor{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
 \newtext{Similar to what \citet{griffiths_google_2007} did, we first considered the directed \newtext{weighted} network formed by association norms, where nodes are words and edges are directed from cue to target word, with a weight equal to the association strength (that is the probability of that target word being produced when this particular cue is presented).
@@
     }
-    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a free association point of view).
+    This tells us if a word belongs more or less to a group of equivalent words (from a free association point of view).
 
@@
 \newtext{Several of these features are strongly related and can be grouped together.
-To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (a few words valued orders of magnitude higher than the vast majority of other words).}
+To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (\hbox{i.e.} a few words valued orders of magnitude higher than the vast majority of other words).}
 %\footnote{
@@
 (b) orthographic and phonological neighborhood densities, with a correlation of .8;
-(c) age of acquisition, betweenness, degree, and pagerank centralities, with absolute pairwise correlations at .41, .6, .59, .63, .61 and .85.
+(c) age of acquisition, betweenness, degree, and pagerank centralities, with absolute pairwise correlations at .41, .59, .6, .61, .63 and .85.
 Applying a feature agglomeration algorithm targeted at 6 groups refined this observation by producing identical (a) and (b) groups, a (c) group without betweenness centrality which was instead assigned to a group (d) with clustering coefficient, and the remaining features (frequency and number of synonyms) as singletons.\footnote{
@@
 Indeed, we have data on frequency for about \newtext{33.5k} words, on age of acquisition for 30.1k words, on clustering coefficient for 5.7k words, number of synonyms 111.2k, and \newtext{orthographic density 17.8k words}.
-Quite often then, not all features are available for all words in our data set;
+Quite often then, not all features are available for a given word in our data set;
 however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
@@
 We therefore model the underlying quotation selection process by making a few additional assumptions.
-\newtext{Given a particular occurrence of a quotation, the first issue is deciding whether} that occurrence is a strict copy of an earlier occurrence, or a substitution of another \newtext{quotation}, or maybe a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
+\newtext{Given a particular occurrence of a quotation, the first issue is deciding whether} that occurrence is a strict copy of an earlier occurrence, or a substitution from another \newtext{quotation}, or maybe a substitution or copy from quotes appearing outside the data set, that is from a source external to the data collection perimeter.
 \newtext{The second issue is deciding which source originated such a substitution when several candidate sources are available.}
@@
 The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
-(1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
-(2) \textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
+(1)~\textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
+(2)~\textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
 (as in the model above) or can be extended to start at the very first occurrence in the quotation family.
 The other two parameters configure rules on the selection of source and destination quotes of a substitution:
-(3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
-(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
+(3)~\textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
+(4)~\textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
 A substitution model, then, is the given of a value for each of those parameters;
@@
 	Substitutions inferred by four models in the situation introduced by Fig.~\ref{fig:substitution-unmodelled}.
-	Each of these models uses discretely positioned bins spanning 1 day (see the main text for a complete description of parameters).
+	Each of these models uses bins spanning 1 day aligned to midnight (see the main text for a complete description of parameters).
 	In the top left panel~(a), $q$ holds the majority in the first bin and is considered the unique basis for $q'$ in bin~2.
@@
 	In the bottom left panel~(c), the majority constraint is lifted compared to panel~(a), making $q''$ in bin~1 a candidate source for $q'$ in bin~2.
-	In the bottom right panel~(d), the majority constraint is also lifted compared to panel~(a) (adding the same $q'' \rightarrow q'$ substitution as in panel~(c)), and the excluded-past constraint is added as in panel~(b) (removing two same substitutions from bin~2 to bin~3 as in panel~(b)).
+	In the bottom right panel~(d), the majority constraint is also lifted compared to panel~(a) (adding the same $q'' \rightarrow q'$ substitution as in panel~(c)), and the excluded-past constraint is added as in panel~(b) (removing the two same substitutions from bin~2 to bin~3 as in panel~(b)).
 	If the bins were extended to the beginning of the quotation family, the excluded-past constraint would also remove the $q' \rightarrow q$ substitution from bin~2 to bin~3.
@@
 Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
-Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use discrete %ly positioned 
-bins spanning 1 day:
+Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use bins spanning 1 day aligned to midnight:
 later occurrences of $q'$ and $q''$ are counted as substitutions in Fig.~\ref{fig:substitution-modelled-majority-all} and Fig.~\ref{fig:substitution-modelled-all-all}, whereas in Fig.~\ref{fig:substitution-modelled-majority-exclude_past} and Fig.~\ref{fig:substitution-modelled-all-exclude_past} they are not.
@@
 In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread over 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
-From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
+From these substitutions we remove those featuring stopwords, minor spelling changes (\hbox{e.g.} center/centre), abbreviations (\hbox{e.g.} November/Nov or Senator/Sen), spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
 this eventually yields \newtext{6,177} valid substitutions.\footnote{
@@
 Substitutions usually replace a word with another semantically related word:
-manual observation of a random subset of 100 substitutions shows that, compared to the word it replaces, the new word often achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
-The following examples illustrate this phenomenon:\footnote{
+manual observation of a random subset of 100 substitutions shows that, compared to the word it replaces, the new word often achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.\footnote{
 \newtext{However, the substituted and substituting words are not so often direct synonyms:
@@
 }}
+The following examples illustrate this phenomenon:
+
 \begin{APAitemize}
@@
 
-\newtext{Our question concerns the low-level properties of these substitutions: we ask (i) which words are targets of the substitutions and (ii) what change they are subjected to.
+\newtext{Our question concerns the low-level properties of these substitutions: we ask (a) which words are targets of the substitutions and (b) what change these words are subjected to.
 To this end,} we build the following two observables for each word feature.
@@
 Second, we measure the change in word feature upon substitution, looking at the variation of a given feature between start and arrival words.
-\newtext{Since sentence context is also central to this process, we extend these two observables by applying them to feature values relative to the distribution of feature values of the sentence in which a word appears.}
+\newtext{Since sentence context is also central to this process, we extend these two observables by applying them to feature values relative to the distribution of feature values in the sentence in which a word appears.}
 
 Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, \newtext{that is} we do not take invariant quotations into account.
-Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it \newtext{we cannot} know whether there has been perfect human reformulation or simply digital copy-pasting of a source~(``{\sc Ctrl-C}/{\sc Ctrl-V}'').
+Indeed, in the former case we know there has been a human reformulation, whereas in the latter case \newtext{we cannot} know whether there has been perfect human reformulation or simply digital copy-pasting of a source~(``{\sc Ctrl-C}/{\sc Ctrl-V}'').
 \newtext{Moreover}, perfect human reformulation possibly involves different practices than those involved in alteration ---~for instance drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
@@
     \caption{\newtext{\textbf{POS-related results:}
-    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \tb{\emph{C} means \emph{Closed class-like}}
-    (see main text for details).
+    categories are simplified from the TreeTagger tag set: \tb{\emph{C} means \emph{Closed class-like}} (see main text for details), \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, and \emph{V} means verb.
     The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
-    Confidence intervals are computed with the Goodman method for multinomial proportions.}
+    Confidence intervals are computed with the \citet{goodman_simultaneous_1965} method for multinomial proportions.}
     }
@@
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
-\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
-That is:}\footnote{
+\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.\footnote{
 \newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
@@
 }}
+That is:}
 
@@
 In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
-\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions,
+\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a; b]$, $g = \{w | \phi(w) \in [a; b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions,
 \hbox{i.e.} \newtext{it measures the preferential selection of some word properties for substitution.}
@@
 Orthographic neighborhood density also shows a slight effect (significant at $p < .05$ in 15 of the 16 substitution models):
-words with very sparse neighborhoods are more substituted than random (which may seem counter-intuitive, but is likely because over 70\% of those words have 7 letters or more).
+words with very sparse neighborhoods are more substituted than random (which may seem counter-intuitive, but is probably because over 70\% of those words have 7 letters or more).
 Clustering coefficient shows no effect on susceptibility, and neither does Number of synonyms:
@@
 A first observation is that all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, and the slope of $\nu_{\phi}$ is smaller than 1 in absolute value, independently of the feature considered.
-\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the disappearing word, the appearing word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
+\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the disappearing word, the appearing word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection of $\nu_{\phi}$ with $y=x$.\footnote{
 \newtext{This reasoning is standard in the analysis of dynamical systems (where the same transformation is applied to the whole system over and over), and becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
@@
 words learned earlier, with lower clustering coefficient \newtext{or with fewer letters} are easier to produce than average~\citep{nelson_how_2013,zevin_age_2002,baddeley_word_1975}.
-\newtext{All these effects are significant at $p < .05$ (and more often $p < .001$) and were verified across the 16 substitution models.}
+\newtext{All these effects are significant with two-tailed \emph{t}-tests at $p < .05$ (and more often $p < .001$) and were verified across the 16 substitution models.}
 
@@
 Let us start with the first one:
-given a feature $\phi$, we define the context-relative susceptibility to substitution in the following three steps.
+given a feature $\phi$, we define the context-relative susceptibility to substitution with the following three steps.
 (1) For each quote in which a substitution appears, compute the distribution of $\phi$ values in that quote (excluding stopwords) and divide it into quartiles.
 (2) Count how many times each quartile (first, second, third or fourth) contains a word that is substituted.
-This procedure tells us, for $i \in \{1, 2, 3, 4\}$, how many times substitutions fall in the $i$-th quartile of each in-quote distribution of $\phi$;
+This procedure tells us, for $i \in \{1; 2; 3; 4\}$, how many times substitutions fall in the $i$-th quartile of each in-quote distribution of $\phi$;
 in other words it gives us the numerator $s_{q_i}$ for the computation of susceptibility, where $q_i$ represents the $i$-th quartile of the distributions of $\phi$ in the quotes.
@@
 To do so we consider all feature values relative to the median word feature in the sentence.
-That is, in all the equations of the previous section we replace $\phi(w)$ with:
+That is, in all the equations of the previous subsection we replace $\phi(w)$ with:
 $$\phi_r(w) = \phi(w) - \operatorname{median}\left\lbrace \phi(w) | w \in \text{sentence} \right\rbrace$$

### 2016-08-23 brains-introduction.tex
okey-dokey manuscrit

@@
 
-We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, \tb{do cultural attractors appear, and if so how do cognitive biases participate in them?}
+We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, do cultural attractors appear, and if so how do cognitive biases participate in them?
 %how do cognitive biases participate in \tb{the appearance of cultural attractors}?

### 2016-08-23 brains-protocol.tex
Move figures around to bring the closer to their citing text
okey-dokey manuscrit

@@
 
-\newtext{One way to settle these questions is the following:}
-group quote occurrences into fixed bins spanning $\Delta t$ days (1 day in the implementation), each one representing a unit of time evolution;
-when a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote of the preceding bin $t$ (or a substring thereof) by only one word;
-if not, $q'$ is not considered to be an instance of substitution.
-\newtext{Fig.~\ref{fig:substitution-modelled-majority-all} shows the inferences made by such a model.
-The assumptions it embeds, however, are a subset of a much wider set of possibilities, each leading to alternative inferences.}
-
-\begin{new}
-
-We identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
-The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
-(1)~\textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
-(2)~\textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
-(as in the model above) or can be extended to start at the very first occurrence in the quotation family.
-The other two parameters configure rules on the selection of source and destination quotes of a substitution:
-(3)~\textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
-(4)~\textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
-A substitution model, then, is the given of a value for each of those parameters;
-it considers valid all the substitutions (and only those) where the source and destination follow the rules set out by the parameters.
-If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.
-
-\end{new}
-
 \begin{figure*}[h]
@@
 
+\newtext{One way to settle these questions is the following:}
+group quote occurrences into fixed bins spanning $\Delta t$ days (1 day in the implementation), each one representing a unit of time evolution;
+when a quotation $q'$ appears in bin $t+1$, it is counted as a substitution if it differs from the most frequent quote of the preceding bin $t$ (or a substring thereof) by only one word;
+if not, $q'$ is not considered to be an instance of substitution.
+\newtext{Fig.~\ref{fig:substitution-modelled-majority-all} shows the inferences made by such a model.
+The assumptions it embeds, however, are a subset of a much wider set of possibilities, each leading to alternative inferences.}
+
 \begin{new}
+
+We identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
+The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
+(1)~\textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
+(2)~\textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
+(as in the model above) or can be extended to start at the very first occurrence in the quotation family.
+The other two parameters configure rules on the selection of source and destination quotes of a substitution:
+(3)~\textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
+(4)~\textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
+A substitution model, then, is the given of a value for each of those parameters;
+it considers valid all the substitutions (and only those) where the source and destination follow the rules set out by the parameters.
+If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
-    \caption{\newtext{\textbf{POS-related results:}
-    categories are simplified from the TreeTagger tag set: \tb{\emph{C} means \emph{Closed class-like}} (see main text for details), \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, and \emph{V} means verb.
-    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
-    Confidence intervals are computed with the \citet{goodman_simultaneous_1965} method for multinomial proportions.}
-    }
-    \label{fig:pos-susceptibilities}
-\end{figure}
-
-\begin{figure*}[!th]
-    \centering
-    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
-    \caption{\textbf{Substitution susceptibility for feature values:}
-    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
-    }
-    \label{fig:feature-susceptibilities-global}
-\end{figure*}
-
 We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
+    \caption{\newtext{\textbf{POS-related results:}
+    categories are simplified from the TreeTagger tag set: \emph{C} means \emph{Closed class-like} (see main text for details), \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, and \emph{V} means verb.
+    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
+    Confidence intervals are computed with the \citet{goodman_simultaneous_1965} method for multinomial proportions.}
+    }
+    \label{fig:pos-susceptibilities}
+\end{figure}
+
 \begin{new}
@@
 A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Closed class-like}\footnote{
-\newtext{The \emph{Closed class-like} category gathers all the POS groups \tb{representing closed class words (coordinating conjunctions, prepositions, subordinating conjunctions, modals and possessive endings).
+\newtext{The \emph{Closed class-like} category gathers all the POS groups representing closed class words (coordinating conjunctions, prepositions, subordinating conjunctions, modals and possessive endings).
 These groups, essentially made of stopwords, feature very low counts for both $s$ (substitutions falling on stopwords are filtered out) and $s^0$ (stopwords are never counted as substitutable).
@@
 hence the name \emph{Closed class-like}.
-}}
+}
 }
@@
 \end{new}
+
+\begin{figure*}[!th]
+    \centering
+    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-quantilebins_global.png}
+    \caption{\textbf{Substitution susceptibility for feature values:}
+    susceptibility to substitution versus feature value of a candidate word for substitution \newtext{(binned by quartiles)}, with 95\% asymptotic confidence intervals \newtext{(Goodman-based multinomial)}.
+    }
+    \label{fig:feature-susceptibilities-global}
+\end{figure*}
 

### 2016-08-23 reviews/2016-07-response-to-reviewers.tex
okey-dokey cover letter

@@
 \x We checked and discussed the susceptibility of part-of-speech (POS).
-While \tb{a rigorous treatment of} the open-/closed-class question is largely precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), \tb{the data does indeed indicate that closed class words are less substituted than open class words (the \emph{Closed class-like} category is mainly made of \tb{closed class words})}.
+While {a rigorous treatment of} the open-/closed-class question is largely precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), {the data does indeed indicate that closed class words are less substituted than open class words (the \emph{Closed class-like} category is mainly made of {closed class words})}.
 As the manuscript discusses however, this concerns less than 7\% of the substitutions we consider, and has likely no effect on our results.
@@
 
-\answerok{\tb{The measure we had proposed in the first version of the manuscript (for information, this was the subject of Fig.
+\answerok{{The measure we had proposed in the first version of the manuscript (for information, this was the subject of Fig.
 4 in that manuscript) considered probabilities of substitutions as an aggregate quantity (and not at the individual sentence level).
@@
 
-\tb{We therefore reworked the notion of susceptibility to make it more straightforward}.
+{We therefore reworked the notion of susceptibility to make it more straightforward}.
 We now define it relatively to a null hypothesis: it measures the strength of a bias affecting a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly).
@@
 Finally, we also examined the effect of part-of-speech (POS): Fig.
-5 details the results on five POS categories, \tb{with a discussion of the corresponding effects in the main text, explaining} why we largely keep POS outside of our analysis.
+5 details the results on five POS categories, {with a discussion of the corresponding effects in the main text, explaining} why we largely keep POS outside of our analysis.
 }
@@
 
-\tb{Regarding prediction, we would like to distinguish two questions.
+{Regarding prediction, we would like to distinguish two questions.
 First, prediction of the actual disappearing and appearing words, which would be an extremely interesting result, but a prospect we believe to stand several steps away from what is currently possible.
@@
 
-\tb{These reasons led us to focus on the second question: prediction of properties of the disappearing and appearing words.
+{These reasons led us to focus on the second question: prediction of properties of the disappearing and appearing words.
 Indeed, Figs.
@@
 In particular, we identify (six) groups of strongly correlated features (both manually by visual inspection and automatically with the help of an agglomeration algorithm, yielding identical results) and justify the choice of our six prototype features for each group.
-Features belonging to the same group generally exhibit very similar behaviors across all measures (with few exceptions, notably number of phonemes and number of syllables \tb{exhibiting less cognitive bias} than number of letters).
+Features belonging to the same group generally exhibit very similar behaviors across all measures (with few exceptions, notably number of phonemes and number of syllables {exhibiting less cognitive bias} than number of letters).
 Graphs produced for all features may be found at <https://huit.re/aEH77Gc3>, and could be included in a Supplementary Information bundle if requested.

### 2016-08-25 brains-protocol.tex
Remove spaces between panesl for manuscript formatting
fix/protocol: fix two typos

@@
 	\newtext{$q$, $q'$ and $q''$ are three quotation variants belonging to the same cluster.
-	$q$ and $q''$ differ by two words, but $q'$ differs from both $q$ and $q''$ by a word.
+	$q$ and $q''$ differ by two words, but $q'$ differs from both $q$ and $q''$ by one word.
 	The second occurrence of $q$ can safely be considered a faithful copy of the first, but the occurrences of $q'$ and $q''$ are uncertain:
-	while the first occurrence of $q'$ is most likely a substitution from $q'$, it could also stem from $q''$;
+	while the first occurrence of $q'$ is most likely a substitution from $q$, it could also stem from $q''$;
 	conversely, the second occurrence of $q''$ could also be a substitution from $q'$ instead of being a faithful copy of its first occurrence.}
@@
 	\end{subfigure}
-	~
 	\begin{subfigure}{.49\textwidth}

### 2017-02-02 brains-protocol.tex
Make final corrections from reviewer 3

@@
 %\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
-If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
+If this is confirmed, we will have successfully tested fundamental cognitive biases with out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
 Aiming to exhibit such subtle biases in complex data is the main reason we chose to use aggregate measures that have a chance of smoothing out the possible variations of experimental conditions in the data set.}
@@
 To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting of the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
-\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see ``Substitution model'').}
+\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained in the ``Substitution model'' section.}
 To quantify these substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.
@@
 (2)~\textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
-(as in the model above) or can be extended to start at the very first occurrence in the quotation family.
+(as in the model above) or can be extended up to the very first occurrence in the quotation family.
 The other two parameters configure rules on the selection of source and destination quotes of a substitution:
@@
 In this case $\nu_{\phi}^{00}$ does depend on $f$:\footnote{
-\newtext{The actual implementation has an additional level of averaging since WordNet, used to get a word's synonyms, defines several meanings for a single given word.
+\newtext{The actual implementation has an additional level of averaging since WordNet, used to get a word's synonyms, defines several meanings for a single given word, which we have no means of disambiguating.
 Therefore:

### 2017-02-03 brainscopycut.tex
Use manuscript formatting for submission
Back to journal formatting after submission
Remove revision colours after submission
formatting: fix title page (keywords ang corresponding author)

@@
 % General document style ------------------------------------------------------------------
-\documentclass[jou]{apa6}
+\documentclass[man]{apa6}
 %\documentclass[man]{apa6}
@@
 \newcommand{\tbn}[1]{\tb{\sout{#1}}}
-\newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
+%\newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
+\newcommand{\tg}[1]{{#1}}
 \newcommand{\newtext}[1]{\tg{#1}}
-\newenvironment{new}{\par\color{MidnightBlue}}{\par}
+%\newenvironment{new}{\par\color{MidnightBlue}}{\par}
+\newenvironment{new}{\par}{\par}
 \newcommand{\add}[1]{\tb{ADD: #1}}
@@
 }
-\authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
+\authornote{\emph{Keywords:} word recall; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; psycholinguistics.\\\emph{Corresponding author:} Sébastien Lerique, CAMS, EHESS, 190 av.
+de France, F-75013 Paris.
+Email: sebastien.lerique@normalesup.org.}
 \keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; psycholinguistics}
@@
 
-%\clearpage
-\appendix
-\input{brains-appendix}
-
 \end{document}

### 2017-02-03 brains-protocol.tex
formatting: add figure placement indications for manuscript format

@@
 
+\indicatemanfigure{fig:feature-corrs-initial}
 \begin{figure}[!th]
@@
 
+\indicatemanfigure{fig:feature-corrs-filtered}
 \begin{figure}[!th]
@@
 
+\indicatemanfigure{fig:substitution-unmodelled}
 \begin{figure}[h]
@@
 	    \input{images/substitution-Time=discrete-Source=all-Past=last_bin-Durl=all.pdf_tex}
-	    \caption{Source can be anything, destination can be anything\\\hfill} % \\\hfill to add a line, aligning with the other panels which have two-line captions
+	    \caption{Source can be anything, destination can be anything\ifapamode{}{\\\hfill}{}} % apamode man/jou/doc. \\\hfill to add a line, aligning with the other panels which have two-line captions
 	    \label{fig:substitution-modelled-all-all}
@@
 If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.
+
+\indicatemanfigure{fig:substitution-modelled}
 
@@
 
+\indicatemanfigure{fig:pos-susceptibilities}
 \begin{figure}[h]
@@
 
+\indicatemanfigure{fig:feature-susceptibilities-global}
 \begin{figure*}[!th]
@@
 
+\indicatemanfigure{fig:feature-variations-global}
 \begin{figure*}[!th]
@@
 \end{new}
+
+\indicatemanfigure{fig:feature-variations_regression-global}
 
@@
 
+\indicatemanfigure{fig:feature-susceptibilities-in_sentence}
+
 Surprisingly the results for this measure are no different from the context-free measure, as can be seen in Fig.~\ref{fig:feature-susceptibilities-in_sentence}:
@@
 $\nu_{\phi}$, $\nu_{\phi}^0$ and $\nu_{\phi}^{00}$ each transpose to $\nu_{\phi_r}$, $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ (note that $\nu_{\phi_r}^0$ now depends on $w$ since it is sentence-relative, whereas $\nu_{\phi}^0$ did not).
+
+\indicatemanfigure{fig:feature-variations-sentencerel}
 

### 2017-02-03 brains-introduction.tex
formatting: clean up source tex from irrelevant comments

@@
 
-%Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
-The question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), but today's discussion is mostly structured by proponents from cognitive science.
 The reciprocal influence between cognition and culture has a long history in both social science and psychology.
@@
 We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, do cultural attractors appear, and if so how do cognitive biases participate in them?
-%how do cognitive biases participate in \tb{the appearance of cultural attractors}?
 We chose to restrict our analysis to substitutions (i.e., one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
@@
 
-The next section %(Sec.~\ref{sec:related})
-describes our hypotheses along with a review of the psycholinguistics literature.
-%In Sec.~\ref{sec:protocol},
+The next section describes our hypotheses along with a review of the psycholinguistics literature.
 Then, we describe the data set and detail the various assumptions that were made in order to analyze it.
-%Sec.~\ref{sec:results}
 Next, we introduce the measures we built to observe cognitive biases operating in quote transmission.
-Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.
+Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.
 \end{new}
@@
 The first above-mentioned method consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).
-%\footnote{{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the importance of the originating cultural systems in the interpretation of such data.
-%have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
-%}}
 A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.

### 2017-02-03 brains-protocol.tex
formatting: clean up source tex from irrelevant comments

@@
 Indeed, the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
-%In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
-%\todo{, as this would be akin to a drink from the firehose.
-\sf [Cam: supprimer?
-potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc.
-outside the scope etc.
-further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion?
-ou bien il s'agit d'une critique directe d'un reviewer ?
-Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
 Our goal, therefore, is to show that cognitive biases have measurable effects in this setting even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
-%\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
 If this is confirmed, we will have successfully tested fundamental cognitive biases with out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
@@
 \end{quote}
-\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is
-% Le APA style guide force "that is" au lieu de "i.e." hors de parenthèses
-sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
+\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
 This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.
@@
     }
-    %, and measure polysemy by taking into account the structure of the entire network.
-    %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on.
-Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
 
@@
 To make correlation values as well as future comparisons more reliable, we log-transformed features that have marked exponential distributions (\hbox{i.e.} a few words valued orders of magnitude higher than the vast majority of other words).}
-%\footnote{
-%\newtext{The distributions of original and log-transformed features appears in the Supplementary Information, %along with scatter plots for each feature pair, summarized below with correlation values.
-%}}
 
@@
 (1)~\textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
-(2)~\textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
-(as in the model above) or can be extended up to the very first occurrence in the quotation family.
+(2)~\textbf{bin span}, which can be $\Delta t=1$ day (as in the model above) or can be extended up to the very first occurrence in the quotation family.
 The other two parameters configure rules on the selection of source and destination quotes of a substitution:
@@
 
-%Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
-%The following examples illustrate this behavior:
-
-%\begin{APAitemize}
-%	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
-%	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
-%	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
-%\end{APAitemize}
-
 \end{new}
@@
 \hbox{i.e.} \newtext{it measures the preferential selection of some word properties for substitution.}
-%\newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.
 
@@
 
-%\begin{figure*}[!th]
-%	\centering
-%	\begin{subfigure}{.49\textwidth}
-%	    \includegraphics[width=\textwidth]{images/computed-figures/paper-variations-quantilebins_sentencerel.png}
-%	    \caption{\newtext{\textbf{Sentence-relative feature variation:}
-%	    $\nu_{\phi_r}$, average sentence-relative feature value of the appearing word as a function of the sentence-relative value of the disappearing word (binned by quartiles), with 95\% asymptotic confidence intervals based on Student's \emph{t}-distribution.
-%	    $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ are similarly converted to be sentence-relative.
-%	    Attraction, magnitude and direction of bias with respect to null hypotheses are similar to Fig.~\ref{fig:feature-variations-global}.
-%	    However, attractors are always positioned between sentence median ($y=0$) on one side and $\nu_{\phi_r}^0$ and $\nu_{\phi_r}^{00}$ on the other side.
-%	    Clustering coefficient, Number of synonyms and Orthographic neighborhood density are limit cases, with triple intersections with one of the null hypothesis curves.}
-%	    }
-%	    \label{fig:feature-variations-sentencerel}
-%	\end{subfigure}
-%	~
-%	\begin{subfigure}{.49\textwidth}
-%	    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
-%	    \caption{\newtext{\textbf{Substitution susceptibility for in-quote feature quartiles:}
-%	    susceptibility to substitution versus quartile of the feature distribution in the originating quote, with 95\% asymptotic confidence intervals (Goodman-based multinomial)}.
-%	    }
-%	    \label{fig:feature-susceptibilities-in_sentence}
-%	\end{subfigure}
-%\end{figure*}
-
 \begin{figure}[h]

### 2017-02-03 brains-conclusion.tex
formatting: clean up source tex from irrelevant comments
formatting: fix all references

@@
 
-% A few more unredacted ideas:
-% More broadly with Sperber:
-%  - you can have a simplistic interpretation and expect simple convergence, or you can look for attraction on different dimensions, which is a richer (and more useful) interpretation.
-%  - note also that it can apply to long term or short term evolution.
-%  - What we see is not clear w.r.t.
-convergence: (1) we don't observe it on our data, and (2) if there's lineage specificity there's no reason convergence should appear (on contrary).
-But, this is only half the story, since in reality there's also selection, i.e.
-not all chains go on forever, so you lose some diversity, and it changes the dynamics.
-%  - finally, context is addressed with relevance in Sperber, but it's still all representational, as we did here. Others have said that it will never be enough. Ingold, Di Paolo.
-
 \section{Concluding remarks}\label{sec:conclusion}
@@
 In biology in particular, work on evo-devo and non-genetic inheritance has accumulated evidence that is poorly accounted for by the modern synthesis of biological evolution, and is creating a demand for new or extended  approaches to joint cultural and biological evolution~\citep[see][for instance]{gilbert_eco-evo-devo:_2015}.
-Such an approach has long been called upon by anthropologists like~\citet{ingold_beyond_2004,ingold_three_1999}, in line with Mauss' initial works~\citep{mauss_les_1936}, and the question is not entirely foreign to the enactive-representational debate in cognitive science.
-%As regards our particular case, a most promising line of work on the evolution of language is being developed by \citet{cuffari_participatory_2014}.
+Such an approach has long been called upon by anthropologists like~\citet{ingold_beyond_2004,ingold_complementarity_1998}, in line with Mauss' initial works~\citep{mauss_les_1936}, and the question is not entirely foreign to the enactive-representational debate in cognitive science.
 The study of cultural evolution will most likely benefit greatly from the growing interactions between these disciplines.

### 2017-02-03 brains-protocol.tex
formatting: fix all references

@@
 We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
-This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 (\citealp{leskovec_memetracker:_2009}).\footnote{
+This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 (\citealp{leskovec_meme-tracking_2009}).\footnote{
 The original article~\citep{leskovec_meme-tracking_2009} does not provide further details on the source selection methodology.

### 2017-02-04 reviews/2017-02-response-to-reviewers.tex
submission: add response cover letter

--- 2016-08-23 reviews/2016-07-response-to-reviewers.tex
@@
 \bigskip
-We enclose an extensively revised version of the manuscript CS15-80 (initially entitled ``How do we copy and paste?
-The semantic drift of quotations in blogspace'') based on the report you sent us on Jul 18, 2015.
+Thank you for the vivifying report and reviewer comments you sent us for our revision to CS15-80.
 
 \medskip
-We hope that we have appropriately addressed the detailed remarks and very constructive questions made by you and the referees: to this effect, we also provide below a detailed response to each of these comments.
-We additionally used a different color in the manuscript to emphasize the portions which were thoroughly rewritten.
-
-\medskip
-This comprehensive revision work led to a significantly longer manuscript than was originally submitted --- from around 6,300 words to about 9,000 words now, making it closer to a ``regular article'' than a ``brief report'': we hope that this will not be problematic at this stage.
-
-\medskip
-We look forward to reading your feedback on this new manuscript and apologize in advance for this long revision round.
+Attached is the final version of the manuscript, updated with the suggestions from reviewer \#3 as well as a few minor stylistic and bibliographic fixes for production.
 
@@
 
-\subsection*{Editor}
-\critique{
-1.
-The main results of the paper are based on a lot of arbitrary decisions about how various quantities are defined.
-I am interested if small differences in these choices radically changes the conclusions you would draw, and how many other approaches you tried before you arrived at the data patterns you present.
-All the reviewers agreed on this point (asking for more clarification about feature selection, why certain choices were made in the analysis, etc…).
-This ranges from what types of quotes were included, to the rules for detecting substitution errors, to the inclusion of certain features in your analyses.
-The reader needs to be convinced that the few “strong” looking patterns you do find are not too dependent on these arbitrary choices.
-I actually question if this is the best way to address this question (effect of memory and linguistic variables on copy errors) given the noisiness of the data and the difficulty in attributing copying errors to any particular parent
-source.
-
-}
-
-\answerok{
-We acknowledge these issues and worked extensively to address them.  In particular (further details may also be found in the responses to reviewers): 
-\itx{
-\x We developed a variety of substitution models to show that the results are robust
-\x We rationalized feature selection, adding some variables and much detail to this section
-\x In general, we clarified further the assumptions and decisions made, such as the restriction to single substitutions
-\x We also introduced manually-coded precision \& recall analyses for all filters, justifying relevant choices.
-}
-}
-
-\critique{
-2.
-The conclusions about “contractile” processes seem pretty speculative (as almost all reviewers agreed).
-Sure, this might be the theoretical implication of iterating this substitution process, but you do not present any further evidence of this type of contraction towards a fixed point.
-I think if you want to make this a major conclusion of the paper there is more work to be done.
-Also this idea has been explored, in my view, in a much more sophisticated way in the iterated learning literature (e.g., Kalish, Lewandowsky, Griffiths, Kirby… see my PDF comments for specific references).
-You should really check these arguments and papers out because you really have an iterated learning task.
-}
-
-\answerok{We agree that the initial manuscript overstated some results.
-We believe the new manuscript states its claims more appropriately, and makes a better distinction between actual conclusions and open questions that could not be addressed given the data (in particular, we attempted to explore substitution chains in earlier stages of this work, but were quickly limited by the constraints of the data).
-
-We had also overlooked several developments in the iterated learning literature, and are grateful for the various pointers.
-The works we thought were most relevant have been integrated to the introductory sections, and their implications for our work are discussed in the new ``Discussion'' section.
-}
-
-\critique{3.
-The writing of this paper is of very low quality compared to what is expected at the journal.
-I will email separately a very detailed editing of the paper via Adobe Acrobat where I tried to point these problems out.
-However, it simply couldn’t be published without a nearly complete re-write or very heavy editing.
-}
-
-\answerok{Thank you for this very constructive and extensive feedback.
-We paid significant attention to the writing style in this new version, and large portions of the document have been rewritten and expanded --- all updates are marked in dark blue.}
-
-%\critique{
-%The bottom line is that while I think this is a really interesting idea/question, this is not the best execution of this idea that I could imagine being published in the journal.
-My gut reaction is that this paper needs a lot of work, additional analysis, and rewriting before it could be published and therefore should be rejected.
-However, because the reviewers are unanimously more positive in their evaluation I will invited a revision for this in the hope that you can make it a pretty significant change and can benefit from the excellent and detailed reviewer comments.
-%
-%So you understand what is requested:  typically papers which are revised at Cognitive Science are returned with a detailed cover letter responding to each of the reviewer comments in detail, along with pointers to what has changed in the paper.
-I would appreciate this type of cover letter if you choose to resubmit because it can help me and the reviewers process what changes you made.
-%
-%Sincerely,
-%Todd Gureckis
-%Associate editor
-%}
-%
-%\answer{
-%}
-
-\subsection*{Reviewer \#1}
-
-\critique{I think this is an interesting approach and dataset for better understanding how sentences are recalled and how paraphrases are constructed.
-However, I think that the experiment design needs to be refined, and the project as a whole needs to be better connected to larger theoretical problems and better grounded within cognitive science.}
-
-\answerok{We have been careful to introduce many more connections to the relevant works in cognitive science: the introduction has been entirely rewritten and is followed by a more thorough presentation of the related work, while the concluding sections (discussion \& concluding remarks) follow your recommendation of connecting our results with the literature.
-}
-
-\critique{
-Starting with the grounding, the Introduction talks about theories of "knowledge transmission mechanisms" but doesn't provide examples aside from the quote-copying problem.
-The theories also don't re-appear in the results discussion, leaving the experimental results stranded and unconnected to any larger domain.
-No discussion of sentence recall tasks within psycholinguistics is given - there isn't a huge literature on this, but there are at least a few papers that should inform this work, such as Potter \& Lombardi 1990 "Regeneration in the short-term recall of sentences".
-No discussion is made about whether and how much the substitutions change the quotes (e.g., is the substituted word a synonym, phonological-neighbor, or something else?
-- cf.
-Lauf et al 2013 "Analyzing Variation Patterns in Quotes Over Time").
-Lastly, the meaning of the analysis (pg.
-17-19) of this process as a dynamical system with "cultural attractors" is not fleshed out - why doesn't this predict that all quotes will eventually, through substitution, converge to the same string?
-(Does it?) As a result, the conclusion that psycholinguistic factors don't have a big impact on substitution likelihood is interesting but doesn't have clear theoretical repercussions, or any clear meaning outside the very specific problem of quote replication.
-This needs to be substantially revised and expanded to make this work useful to the audience at large.
-}
-
-\answerok{We thoroughly reworked the general grounding and flow of the manuscript. In particular:
-\itx{
-\x We believe that the introductory parts provide a clearer presentation of our question and how it relates to existing literature (including the areas we overlooked) and hypotheses
-\x The new ``Discussion'' section, which was indeed missing, should properly reconnect our results to the related works introduced at the beginning of the paper.
-}
-
-Finally, several modifications in the ``Results'' section (more examples, a short discussion of the distance traveled by substitutions on the synonym network in footnote \#15, and the addition of the second null hypothesis $\mathcal{H}_{00}$ in the ``Variation'' subsection) should better describe the type of changes undergone by quotes.
-}
-
-\critique{The experiment design also has some substantial flaws.
-First, some crucial factors were not considered - most notably, part of speech.
-I suspect, for instance, that pronouns and other closed-class words are less often substituted than open-class words.
-Phonological neighborhood density could also be important, though probably less so.
-Second, no analysis was done that included all of the factors in, e.g., a multi-feature regression model.
-While the features being considered are not strongly correlated, they are still likely to interact - especially if part-of-speech is added.
-Third, the restriction to single-word substitutions introduces some significant confounds.
-I'm not convinced that correctly and incorrectly copying a quote should be viewed as separate behaviors.
-}
-
-\answerok{
-We addressed these points in the revised version:
-\itx{
-\x We checked and discussed the susceptibility of part-of-speech (POS).
-While {a rigorous treatment of} the open-/closed-class question is largely precluded by our exclusion of stopwords from the analysis (which led us to not discuss that point in the manuscript), {the data does indeed indicate that closed class words are less substituted than open class words (the \emph{Closed class-like} category is mainly made of {closed class words})}.
-As the manuscript discusses however, this concerns less than 7\% of the substitutions we consider, and has likely no effect on our results.
-\x Following this suggestion, we added phonological neighborhood density as a word feature.
-We show that it forms a subgroup of strongly correlated variables, together with orthographic neighborhood density, and choose the latter as a more natural representative of this subgroup given the task studied.
-\x We introduced a multi-feature regression model to estimate variation upon substitution, showing that the magnitude of the variation of a feature is mainly predicted by the {start word's value for} that very feature, generally with negligible interaction with the other features (see also below and figure 8).
-\x We expanded substitution models to {also detect quotes undergoing up to two substitutions}, which yielded the same results as single substitutions {(see below)}. 
-}
-
-%We also examined the effect of part-of-speech (POS): Figure 5 details the results on five POS categories.
-A Goodman-based multinomial goodness-of-fit shows that POS categories do not capture any significant bias in the selection of substitution targets, which is why we largely keep them outside of our analysis.
-}
-
-\critique{But more importantly, the $\sigma_\phi(f)$ calculation can be misleading under the single-substitution restriction: Consider a toy case where all quotes are two words, and there is a single binary feature $F$.
-$p(F=1) = 3/4$, and the true substitution probabilities, given the value of F, are $p(S|F=1)=1$, $p(S|F=0)=1/2$.
-All the quotes where both words have feature $F=1$ will have two substitutions and be omitted from the dataset, as will half of the quotes with one $F=1$ \& one $F=0$ word.
-Half of the quotes with two $F=0$ words will also be omitted for having no substitutions.
-Due to these omissions, if you actually generate some test data and calculate the $\sigma_F$ values, the $\sigma_F(1)$ value is correctly estimated as 1, but the $\sigma_F(0)$ value is erroneously estimated as 1/8 (not 1/2), because the $F=0$ words are never substituted in the $F=1/F=0$ pairs, and these are three times more common than the $F=0/F=0$ pairs.
-I haven't run more complicated cases, such where the length of the quotes vary, but this leads me to have limited confidence that the estimates in the Figure 4 are appropriate to draw conclusions from.
-I would want to see either a mathematical analysis with realistic values showing that the error is expected to be small in this data or an expansion to include multiple-substitution examples to avoid this problem.
-}
-
-\answerok{The case presented here is valid and could indeed hide a bias in our measures.
-Rather than analyzing all possible such cases to bound the error, we extended substitution models to detect quotes undergoing up to two substitutions, and applied the same analysis to these mined substitutions.
-The results were unchanged (note: graphs produced for all models may be found at <https://huit.re/aEH77Gc3>, and may be commented further upon request).
-For example, the extended version of the model presented in the paper mines 70k substitutions, 9.2k of which are kept after filtering (the lowered acceptance rate indicates that more invalid substitutions are generated when allowing up to two substitutions per quote).
-
-We did not expand to more than two substitutions, as more than 20\% of quotes are 5 to 9 words long (with a minimum of 5 words for quotes in the data set).
-For such quotes, three or more substitutions means changing at least a third of the sentence, which makes any attempt to link it to a similar variant pointless.
-}
-
-\critique{ Some high-level suggestions:
-\itx{
-\x more discussion of what dynamical systems are and especially their interpretation in this task before their appearance in the results}
-}
-
-\answerok{The rewritten ``Variation'' subsection (including footnote \#21) should now clarify this notion.}
-
-\critique{\itx{
-\x needs a substantive discussion of the results}
-}
-
-\answerok{An entirely new ``Discussion'' section is now devoted to this task.}
-
-\critique{\itx{
-\x show that the phi value is capturing what it purports to capture
-\x and that the factors aren't conflated with length
-\x is the single-substitution biasing the data?
-\x make sure that the reported measures are interpretable
-\x if low-frequency words have 1/3 chance of being chosen as substitution (Fig. 4), what if there are four low-frequency words in a quote?
-}
-}
-
-\answerok{{The measure we had proposed in the first version of the manuscript (for information, this was the subject of Fig.
-4 in that manuscript) considered probabilities of substitutions as an aggregate quantity (and not at the individual sentence level).
-It was difficult to use it to understand what happens when substitution occurs in a given sentence.}
-
-{We therefore reworked the notion of susceptibility to make it more straightforward}.
-We now define it relatively to a null hypothesis: it measures the strength of a bias affecting a given feature value (i.e., how much a feature value is substituted compared to what it would be if targets were picked randomly).
-We present the corresponding results in Figs.
-5 \& 6.
-
-Finally, we also extended our measures to capture part of the sentence-level context in which substitutions occur (see the ``Sentence context'' subsection).
-Sentence-relative susceptibility, in particular, gives a straightforward answer to the question of which words are more likely to attract substitution within a given sentence.
-}
-
-\critique{\itx{
-\x improving statements of motivation and grounding within cognitive science
-\x what are representation transformation processes?
-\x give examples in the introduction
-\x show how the quotation test case connects to more general representation transformation?}
-}
-
-\answerok{
-These points should be addressed by the improved introduction.
-In particular regarding the last question, the new ``Discussion'' section additionally makes a connection with the iterated learning literature.
-}
-
-\critique{\itx{
-\x what about the semantics of quote replication?
-\x does part of speech influence substitution rate?
-\x can you quantify the semantic distance between the original and substituted quote (e.g., get Mechanical Turkers to rate the semantic distance?)}
-}
-
-\answerok{We added a short discussion on the semantics of substituted quotes.
-We first aimed at quantifying the semantic distance by estimating the distance ``traveled'' by the substituted word in the WordNet hypernym-hyponym network as well as the FA network.
-This showed that substituted and substituting words are rarely direct synonyms, yet they usually achieve a similar meaning in the context of their sentence.
-To further grasp the relationship of substitutions to synonyms we introduced a second null hypothesis in the ``Variation'' subsection, which compares the actual variation to what would happen if a random synonym of the start word were picked instead.
-
-Finally, we also examined the effect of part-of-speech (POS): Fig.
-5 details the results on five POS categories, {with a discussion of the corresponding effects in the main text, explaining} why we largely keep POS outside of our analysis.
-}
-
-\critique{\itx{
-\x what happens when you run a multi-factor regression model? Do the features interact?
-}}
-
-\answerok{We introduced a simple regression model to check for feature interactions, which is presented in Fig.
-8 with a corresponding discussion (in a nutshell: we show that features generally do not interact).}
-
-\critique{\itx{
-\x clarify from the beginning the predictions you are making so that the results will be interpretable 
-}}
-
-\answerok{We hope the improved introduction is clearer as to what we expect and what is left open due to the exploratory nature of our study.}
-
-\critique{Some low-level points:\itx{
-\x the in vitro/in vivo distinction is unfamiliar and ought to be defined in the text
-}
-}
-
-\answerok{In light of this potential unfamiliarity, we preferred to entirely avoid referring explicitly to this dichotomy (for instance, we now speak of ``corpus-based utterances'' instead of ``in vivo'').}
-
-\critique{\itx{
-\x introducing eight measures and then removing three of them due to collinearity is presentationally confusing}}
-
-\answerok{We agree that introducing twelve\footnote{In the revised version we have 12 measures, instead of 8 in the initial submission.} measures before eventually keeping six of them could complicate the flow of the section on word-level measures.
-However, we found it more natural to first propose a comprehensive list of candidate features based on the literature before addressing the quantitative analysis with a reduced yet sufficient set of metrics --- especially now that this section has been significantly expanded.
-We initially favored this solution over first presenting the final list of candidate features and justifying afterwards that they are anyway correlated with a broader list of traditional variables.
-(Note: an alternative option could consist in presenting the contents of Figs.
-1 and 2 in the same figure, if this would be less confusing.)}
-
-\critique{
-\itx
-{
-\x was there a reason to expect that \# of synonyms and \# of phonemes would interact (Fig. 4)?]
-}
-}
-
-\answerok{Except for word frequency, we observed that synonyms and phonemes were the only two word features to exhibit an effect on susceptibility (and a quite similar effect for that matter).
-We thus reasoned that they might exhibit a joint effect.
-
-However, the revised version now features a broader analysis assessing joint effects (see Fig.
-8); we think it favorably replaces the previous variable-to-variable correlation check.
-(Note: we also replaced number of phonemes with number of letters, which makes more sense in our context -- see the end of the ``Word-level measures'' subsection.
-The latter does not seem to interact with number of synonyms.)}
-
-\critique{\itx{
-\x how are you binning the continuous-valued features into categorical features for Fig 4 \& 5?
-\x are there multiple y=x crossovers in Fig 5 if more bins are used?}
-}
-
-\answerok{Data in Figs.
-4 \& 5 was binned into 5 integer-based bins, and the graphs were cropped whenever the data was unreasonably scarce and, accordingly, confidence intervals were unreasonably large (this cropping gave the impression of some graphs having only 4 bins).
-
-Notwithstanding, data in the new Figs.
-6, 7, 9 \& 10 is now binned into 4 quartile-based bins, which makes more sense for the confidence intervals.
-This decision was reached after exploring several other binning configurations: none of these substantially changed the results, but opting for quartile-based bins does improve the overall readability.
-Notably, no multiple crossovers were ever observed.
-}
-
-\subsection*{Reviewer \#2}
-\critique{This is a review of "How do we copy and paste?
-The semantic drift of quotations in blogspace." This paper attempts to discover the underlying psycholinguistic variables that is involved in blog authors making mistakes in the copying of quotations across time.
-Although I think the motivations and goals of this article are interesting and noteworthy, I think the actual analysis needs improvement in a number of areas.
-
-Although I appreciate the brief writeup on intrusions in recall, the article fails to capture the amount of research done on this topic.
-Going back to Deese (1959; the precursor to modern experimental studies on false memory) and the subjective organization results of Tulving (1962) it has been well known that extra-list intrusions in free recall are often semantic in nature, with the intruding word being very similar in meaning to a studied item.
-A recent study that reinforces this is one by Zaromb, et al.
-(2006; JEP:LMC).
-}
-
-\answerok{Thank you for these very useful and, indeed, relevant pointers.
-We accordingly devoted a significant portion of the literature review to the issue of recall (see the ``Related work'' and ``Discussion'' sections).}
-
-\critique{ This leads into a larger issue: there is no direct word-word semantic similarity score included in the psycholinguistic variables in the analysis.
-Global measures such as clustering coefficients based on free association norms are included, but this isn't as appropriate, since as the authors find, the vast majority of words that are being substituted are low frequency words.
-Low frequency words are not as well represented in the Nelson norms, given that they are less likely to be produced, and are thus less likely to have an appropriate clustering coefficient from this dataset.
-Variables based on corpus analyses need to be included to provide an appropriate overview of the impacts of semantics on this task.
-Even so, I would suspect that global semantic network measures are not nearly as important as word-word metrics.
-
-I would guess that the substitutions that are made in quotations are overwhelmingly higher frequency, highly semantically similar words.
-As the example on pg.
-19 states, the substitution was "world" for "globe," where a higher frequency word replaced a lower frequency, semantically related word.
-There are a number of different models that can be used to assess this type of similarity, with the classic model being latent semantic analysis (Landauer \& Dumais, 1997).
-There are a number of additional models that have been proposed, such as Topics (Griffiths, Steyvers, \& Tenenbaum, 2007, Psyc Review) or Beagle (Jones \& Mewhort, 2007, Psyc Review).
-These models differ in their implementational complexity, but there are tools available that allow for useful statistics to be acquired easily.
-A tool by Recchia \& Jones (2010, BRM) allows for pointwise mutual information (PMI) of different words to be retrieved from a given corpus very simply.
-A tool such as this would allow for word-word similarity to be included in the analysis.
-
-I understand that word-word metrics are not easily introduced in the type of analysis that the authors are attempting to do, where the analysis of the contributions of different global psycholinguistic variables are being weighed against each other.
-I just think that type of analysis is inappropriate here.
-Since the quotations were selected on the basis of single substitutions, it seems to me that the appropriate level of analysis is what properties of the word that is being substituted is the most predictive of the substitution.
-I suspect that the combination of frequency and semantic similarity will be by far the most predictive variables, consistent with past results on recall.
-}
-
-\answerok{The rewritten introductory sections of this revision now aim at explaining why we chose to use a feature-based analysis over the prediction of individual words (see below for more details about this).
-Word-word similarity is now also shortly discussed in the beginning of the ``Results'' section (and in footnote \#15).
-We explored a number of other similarity measures provided by WordNet (based on paths in the WordNet network, or on similarity of Information Content extracted from external corpora), but none of them provided more insight than what is being discussed in that section.
-
-{Regarding prediction, we would like to distinguish two questions.
-First, prediction of the actual disappearing and appearing words, which would be an extremely interesting result, but a prospect we believe to stand several steps away from what is currently possible.
-As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
-Zaromb et al.
-(2006), for instance, predict the list from which prior-list intrusions come, but not the exact intruding word.
-To our knowledge, the prediction of intrusions in full sentences (either manually designed or not) has not been tackled.
-Finally, as explained in footnote \#3 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.}
-
-{These reasons led us to focus on the second question: prediction of properties of the disappearing and appearing words.
-Indeed, Figs.
-7 \& 10 provide average predictions of appearing word features given the properties of a disappearing word.
-We show that there are little to no interactions between word features (see Fig.
-8), such that word frequency is not sufficient to predict variation of other features.
-Finally, we introduced a new null hypothesis ($\mathcal{H}_{00}$) in the ``Variation'' subsection which gives insight into how an appearing word relates to the synonyms of the word it replaces, and shows that appearing words are much more frequent than the average frequency of synonyms of the disappearing word.}
-
-While we understand that the analysis we present differs from typical approaches usually taken for similar laboratory data, we hope that these various additions contribute to alleviate concerns regarding word-word similarity.
-}
-
-%\answer{We agree that predicting the actual disappearing and appearing words would be an extremely interesting result, \tb{yet we believe that this prospect stands} several steps away from what is currently possible.
-%\tb{As the revised introductory parts now allude to, the prediction of word intrusions is well mastered in the case of manually designed word lists but is not as exact in the case of random lists of words.
-Zaromb et al.
-(2006), for instance, predict the list from which prior-list intrusions come, but not the exact intruding word.
-To our knowledge, the prediction of intrusions in full sentences (either manually designed or not) has not been tackled.
-Finally, as explained in footnote \#3 and the ``Discussion'' section, our data set is ill-shaped to measure the accuracy of predictions we could make.
-%}}
-
-\critique{Beyond this, I think the authors are missing out on the most interesting aspect of the data that they have collected: how the surrounding context of the quotation used is changing across time.
-That is, the quote is being propagated, and the prior usage of that quote is likely predictive of how that information is spread.
-This is likely a dynamical process where multiple news sources are used to construct a new blog post, and being able to analyze this type of information propagation would be a very interesting study.
-I point to the semantic space models cited (and the work derived from them) as being very useful tools to analyze this type of question.
-}
-
-\answerok{This would indeed be a very interesting project and prospect, but the data we collected do not enable the study of the surrounding context (in particular, we know neither the prior usage nor the way information sources may be used) so that this goal remains beyond the scope of our paper.
-
-We did however extend our measures to capture part of the context of the sentence substitutions occur in (see the ``Sentence context'' subsection).}
-
-\subsection*{Reviewer \#3}
-\critique{In this paper, the authors analyze a large data set of quotations taken from a variety of news sources to study the form of substitutions in quotes, to determine if the transformations follow expected patterns given psycholinguistic theory.
-Specifically, the words that are likely to be substituted are found to be less frequent words, and the pattern of substitution is such that the substituting word is more frequent, acquired earlier, and has lower clustering in the free-association word network.
-These patterns are consistent with expected cognitive biases.
-
-There are a few issues I would want to see addressed.
-\bigskip
-
-Foremost is the method of detecting substitution.
-I understand that, given the data set, there is no perfect way to identify what quote is a transformation of what quote, and that it is not feasible to test all possible methods AND that it is simpler to focus on one method for the purposes of the paper.
-However, given that this decision of which method to use can directly influence the main conclusions, it behooves the authors to do some sort of robustness check around this method.
-For example, one could imagine two distinct groups that only exactly copy from themselves and never copy from each other, where one always publishes earlier than the other.
-If the earlier-publishing group tended to select variations of quotes that used lower-frequency words, and the later-publishing group tended to select variations of quotes that used higher-frequency words, this method would infer many more low-frequency to high-frequency substitutions than actually existed.
-This may not be a likely scenario, but it demonstrates the benefit of a robustness check.
-Trying the same analysis with one or two different substitution models would strengthen the authors' claims.
-}
-
-\answerok{There are indeed very many possible models, yet we agree that a robustness check is also very sensible.
-Substitution detection relies on a number of assumptions and we identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
-We thus implemented, tested and checked these models, and we can report that our results remain valid for all 16 models (for the sake of clarity, we do not specifically report on these models, though).
-\\
-
-This leads to a much more detailed ``Substitution Model'' subsection which discusses these options along with the reference model we used throughout the paper.
-}
-
-\critique{ A second issue is the time spent on feature selection.
-It is unclear why the authors do feature selection at all, as "the analysis is done on a per-feature basis."  It would again add to the robustness of their claims if they showed that all of the correlated features show the same pattern of susceptibility and bias in substitution.
-In the same vein, it is not clear why the authors claim that age of acquisition, frequency, and number of synonyms have low levels of correlation "excluding the network properties".
-Why would one ignore the correlations with the network features?
-}
-
-\answerok{Our feature selection step aims at making the results more intelligible.
-This section was indeed poorly written, and has been thoroughly revised to rationalize the process.
-In particular, we identify (six) groups of strongly correlated features (both manually by visual inspection and automatically with the help of an agglomeration algorithm, yielding identical results) and justify the choice of our six prototype features for each group.
-Features belonging to the same group generally exhibit very similar behaviors across all measures (with few exceptions, notably number of phonemes and number of syllables {exhibiting less cognitive bias} than number of letters).
-Graphs produced for all features may be found at <https://huit.re/aEH77Gc3>, and could be included in a Supplementary Information bundle if requested.
-}
-
-\critique{
-One smaller issue has to do with the filtering of the quotations.
-The authors manually inspected the filtered quotes to see how many real event-related utterances are incorrectly filtered, but I would argue the precision of the filter is more important than the recall; how many of the kept quotes are not real event-related utterances?
-Why not do manual coding of those and report that accuracy?
-}
-
-\answerok{This is an important empirical point which we did not address systematically in the first submission.
-We have now done this manual coding step, and have improved filters wherever precision or recall (whichever was most relevant) were not good enough.
-We now assess filtering accuracy systematically and it is satisfying in the relevant cases.}
-
-\critique{Finally, while I understand the appeal of doing a dynamic systems analysis to find the convergence points of the substitutions, but it is not really valid here.
-The analysis in this paper does not provide information on what happens with multiple substitutions, so the authors cannot really make claims about the convergence points given this methodology.
-}
-
-\answerok{In this revised version, we better explained our claims and their limits (see for instance the second paragraph of the ``Discussion'' section), without overstating our findings.}
-
-\critique{Overall, the analysis of the patterns of substitution in a series of quotations to validate psycholinguistic theory is novel and advances the current science.
-The results would be more solid with some additional robustness checks, but in general they are quite intuitive.
-I hope in the future the authors find a dataset that provides certain information about the transmission of language from one person to another, as this would allow for stronger claims about causality and greater confidence in the pattern of substitutions, as well as a finer-grained analysis of repeated substitutions.
-}
-
-\answerok{On a side note, we entirely agree with the relevance of pursuing this research program in these directions, as the \emph{in vitro} observation of information transmission and its evolution is part of an ongoing project (which nonetheless remains beyond the scope of the current paper).}
-
-\subsection*{Final remarks}
-
-\answerok{A few additional numbers have changed in this revised manuscript:
-\itx{
-\x Clustering coefficient values have changed, because Free Association link weights are now taken into account in the computation (i.e.
-coefficients are computed on the undirected weighted graph instead of the undirected unweighted graph).
-\x An update in the language detection module led to a small change in the results of cluster filtering: there are a few more clusters and quotes than in the previous version.
-As a result of this, the number of words coded by Word Frequency has also changed (since frequency of words is computed from the set of quotes we keep after cluster filtering).
-\x The improvement of substitution filtering led us to mine many more substitutions than previously. All in all raw data processing is now much more reliable, as unit tests cover nearly all of it.
-}
-}
-
-{\small
-\bibliographystyle{../../authordate1}
-\bibliography{../../mainbiblio-u8}
-}
-
 \end{document}

### 2017-02-04 brainscopycut.tex
formatting: fix section titles to look as requested for submission
formatting: better subsubsection spacing
formatting: add figure placement indications for manuscript format
formatting: clean up source tex from irrelevant comments
formatting: fix all references
affiliation

@@
 % General document style ------------------------------------------------------------------
+%\documentclass[jou]{apa6}
 \documentclass[man]{apa6}
-%\documentclass[man]{apa6}
 \usepackage[utf8]{inputenc}
@@
 \usepackage[usenames,dvipsnames]{color}
-
-%\acmVolume{VV}
-%\acmNumber{N}
-%\acmYear{YYYY}
-%\acmMonth{Month}
-%\acmArticleNum{XXX}
-%\acmdoi{10.1145/XXXXXXX.YYYYYYY}
-
-%\markboth{S. Lerique and C. Roth}{How do we copy and paste? The semantic drift of quotations in blogspace}
 
@@
 % Formatting bibliography
-%\usepackage{natbib}
-\usepackage[natbibapa,nodoi]{apacite}
-%\setcitestyle{authoryear,round,semicolon}
+\usepackage[natbibapa]{apacite}
 
@@
 \usepackage[group-separator={,}]{siunitx}
-%\usepackage{stmaryrd}
 \usepackage{xfrac}
 \usepackage{mathtools}
-%\newcommand{\defeq}{\overset{\underset{\mathrm{def}}{}}{=}}
-
-% Strikethrough (with the \sout comment)
-\usepackage[normalem]{ulem}
 
@@
 \usepackage{graphicx}
-%\usepackage{subfig}
-%\usepackage{subfigure}
 \usepackage{caption}
@@
 \newcommand{\tbn}[1]{\tb{\sout{#1}}}
-%\newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
 \newcommand{\tg}[1]{{#1}}
 \newcommand{\newtext}[1]{\tg{#1}}
-%\newenvironment{new}{\par\color{MidnightBlue}}{\par}
 \newenvironment{new}{\par}{\par}
@@
 
+% Fix subsubsection to look as requested for submission
+\renewcommand{\typesectitle}[1]{#1\hfill\\ \vskip0.2\baselineskip \hskip-1em}
+
+% Add figure placement indications in apamode "man"
+\newcommand{\indicatemanfigure}[1]{
+\ifapamode{%man
+  \begin{center}
+    -------------------------- Insert Figure \ref{#1} about here --------------------------
+  \end{center}
+}{%jou
+}{%doc
+}}
+
 % Start with the real content -------------------------------------------------------------
@@
 \twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av.
-de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr.
-191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr.
+de France, F-75013 Paris\\and\\Centre Marc Bloch Berlin e.V., UMIFRE 14 CNRS/MAEDI, Friedrichstr.
+191, D-10117 Berlin}{Sciences Po, médialab, Paris, France\\and\\Centre Marc Bloch Berlin e.V., UMIFRE 14 CNRS/MAEDI, Friedrichstr.
 191, D-10117 Berlin}
@@
 Email: sebastien.lerique@normalesup.org.}
-\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; psycholinguistics}
-%\note{\TB{Note XXXX (version? date?)}}
 
@@
 \maketitle
+
+% Force section numbering
+\setcounter{secnumdepth}{3}
 
@@
 
+% Remove section numbering for references
+\setcounter{secnumdepth}{0}
+
 \bibliographystyle{apacite}
